### 2021-11-05 src/ms.tex
Add files via upload

%% Beginning of file 'sample631.tex'
%%
%% Modified 2021 March
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.31 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls and other external packages such as
%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
%% AASTeX v6+ still uses v4.1. All of these external packages should 
%% already be present in the modern TeX distributions but not always.
%% For example, revtex4.1 seems to be missing in the linux version of
%% TexLive 2020. One should be able to get all packages from www.ctan.org.
%% In particular, revtex v4.1 can be found at 
%% https://www.ctan.org/pkg/revtex4-1.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
%\documentclass[linenumbers,preprint2]{aastex631}
\documentclass[twocolumn]{aastex631}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%   tighten      : Makes baselineskip slightly smaller, only works with 
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for 
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much 
%%                  longer author list but may be desirable for short 
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{longtable}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

\newcommand{\rhk}{$R'_\mathrm{HK}$\xspace}
\newcommand{\shk}{$S_\mathrm{HK}$\xspace}
\newcommand{\caiihk}{Ca II H\&K \xspace}

\newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}

\newcommand{\kms}{km~s$^{-1}$}
\newcommand{\ms}{m~s$^{-1}$}
\newcommand{\gcc}{g~cm$^{-3}$}
\newcommand{\masyr}{mas~yr$^{-1}$}
\newcommand{\err}{\textit{$\pm$}}

\newcommand{\Mstar}{\ensuremath{M_{\star}}\xspace}
\newcommand{\Rstar}{\ensuremath{R_{\star}}\xspace} 
\newcommand{\Lstar}{\ensuremath{L_{\star}}\xspace} 
\newcommand{\teff}{\ensuremath{T_{\mathrm{eff}}}\xspace}  
\newcommand{\logg}{\ensuremath{\log g}\xspace} 
\newcommand{\feh}{[Fe/H]\xspace}
\newcommand{\vsini}{\ensuremath{v \sin i}\xspace} 
\newcommand{\prot}{\ensuremath{P_\mathrm{rot}}\xspace}
\newcommand{\rvar}{\ensuremath{R_\mathrm{var}}\xspace}
\newcommand{\vtan}{\ensuremath{v_\mathrm{tan}}\xspace}
\newcommand{\vb}{\ensuremath{v_\mathrm{b}}\xspace}

\newcommand{\mearth}{$M_\oplus$\xspace}
\newcommand{\rearth}{$R_\oplus$\xspace}

\newcommand{\msun}{$M_\odot$\xspace}
\newcommand{\rsun}{$R_\odot$\xspace}
\newcommand{\lsun}{$L_\odot$\xspace}
\newcommand{\rhosun}{$\rho_\odot$\xspace}
\newcommand{\mstar}{$M_*$\xspace}
\newcommand{\rstar}{$R_*$\xspace}
\newcommand{\lstar}{$L_*$\xspace}
\newcommand{\rhostar}{$\rho_*$\xspace}

\newcommand{\logage}{$\text{log(age)}$\xspace}
\newcommand{\rp}{$R_P$\xspace}
\newcommand{\rprstar}{$R_P/R_\star$\xspace}

\newcommand{\logrp}{$\log_{10}(R_P/R_\oplus)$\xspace}
\newcommand{\deltaage}{$\Delta \log_{10}(\mathrm{age~yr}^{-1})$\xspace}
\newcommand{\deltamstar}{$\Delta M_*$\xspace}
\newcommand{\deltateff}{$\Delta T_\mathrm{eff}$\xspace}

\newcommand{\mjup}{$M_\mathrm{Jup}$\xspace}
\newcommand{\galex}{\textit{GALEX}\xspace}
\newcommand{\gaia}{\textit{Gaia}\xspace}
\newcommand{\kepler}{\textit{Kepler}\xspace}
\newcommand{\spitzer}{\textit{Spitzer}}
\newcommand{\hipparcos}{\textit{Hipparcos}}
\newcommand{\ktwo}{\textit{K2}\xspace}
\newcommand{\tess}{\textit{TESS}\xspace}
\newcommand{\emcee}{\texttt{emcee}\xspace}
\newcommand{\stardate}{\texttt{stardate}\xspace}
\newcommand{\python}{\textsc{python}\xspace}

\newcommand{\ymg}{[Y/Mg]\xspace}
\newcommand{\yal}{[Y/Al]\xspace}
\newcommand{\alphafe}{[$\alpha$/Fe]\xspace}
\newcommand{\dexgyr}{dex~Gyr$^{-1}$\xspace}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the argument.
%\submitjournal{PSJ}

%% For manuscript that include authors in collaborations, AASTeX v6.31
%% builds on the \collaboration command to allow greater freedom to 
%% keep the traditional author+affiliation information but only show
%% subsets. The \collaboration command now must appear AFTER the group
%% of authors in the collaboration and it takes TWO arguments. The last
%% is still the collaboration identifier. The text given in this
%% argument is what will be shown in the manuscript. The first argument
%% is the number of author above the \collaboration command to show with
%% the collaboration text. If there are authors that are not part of any
%% collaboration the \nocollaboration command is used. This command takes
%% one argument which is also the number of authors above to show. A
%% dashed line is shown to indicate no collaboration. This example manuscript
%% shows how these commands work to display specific set of authors 
%% on the front page.
%%
%% For manuscript without any need to use \collaboration the 
%% \AuthorCollaborationLimit command from v6.2 can still be used to 
%% show a subset of authors.
%
%\AuthorCollaborationLimit=2
%
%% will only show Schwarz & Muench on the front page of the manuscript
%% (assuming the \collaboration and \nocollaboration commands are
%% commented out).
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{The Rossby Ridge}
\shortauthors{David et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
%% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{figures/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

% \title{The Rossby Ridge: a Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages}
\title{A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking}

%\\or\\Weakened Magnetic Braking Supported by Rotation Rates of Kepler Exoplanet Hosts}

% Ridge
% Edge
% Ledge
% Cornice

\correspondingauthor{Trevor J. David}
\email{tdavid@flatironinstitute.org}

\author[0000-0001-6534-6246]{Trevor J.\ David}
\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
\affil{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th Street, New York, NY 10024, USA}

\author[0000-0003-4540-5661]{Ruth Angus}
\affil{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th Street, New York, NY 10024, USA}
\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
\affil{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY, USA}

\author[0000-0002-2792-134X]{Jason L.~Curtis}
\affil{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY, USA}

\author[0000-0002-4284-8638]{Jennifer van Saders}
\affil{Institute for Astronomy, University of Hawai’i, Honolulu, HI, USA}

\author[0000-0002-3011-4784]{Gabriella Contardo}
\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}

\author{Other authors TBD}

%Gaby Contardo
%Lucy Lu
%Joel Zinn
%Adrian Price-Whelan

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters derived from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a long-period pile-up in the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding 5800~K. Stars in the pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period pile-up is well-described by a single-parameter model of constant Rossby number, with a critical value of $\mathrm{Ro_{crit}} \sim$\textbf{XYZ}. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which stellar spin-down stalls once stars reach a critical Rossby number. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than $\sim$1--2~Gyr. \textbf{We additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.}
\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Stellar rotation (1629) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- F dwarf stars(516)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}
Solar-type and low-mass stars ($M\lesssim1.3$~\msun) lose mass and angular momentum through magnetized winds \citep{Parker1958, WeberDavis1967, Mestel1968, Kawaler1988}. Consequently, stellar rotation rates are observed to decline with age. \citet{Skumanich1972} presented the first attempt to calibrate this age-rotation relationship using the rotation periods of Sun-like stars in open clusters with independently determined ages, finding a $P_\mathrm{rot} \propto t^{1/2}$ scaling, where $t$ is stellar age. In the intervening decades, observational determinations of stellar rotation periods among open cluster members revealed how stellar spin rates evolve in more detail, leading to the calibration of the so-called gyrochronology method \citep{Barnes2003, Barnes2007, Barnes2010, MamajekHillenbrand2008, Meibom2009}.

The arrival of continuous, high-precision, long-baseline photometry from NASA's Kepler space telescope \citep{Borucki2010} provided a watershed moment for stellar rotation studies, yielding period detections for tens of thousands of stars \citep[e.g.][]{Reinhold2013, McQuillan2014, Santos2021} and allowing for gyrochronology to be extended to older ages \citep[e.g.][]{Meibom2011, Meibom2015}. NASA's subsequent K2 \citep{Howell2014} and TESS \citep{Ricker2015} missions propelled the field of stellar rotation further still, providing an exquisitely detailed picture of how spin rates evolve for stars with a broad range of masses and ages in stellar associations \citep[e.g.][]{Douglas2016, Douglas2017, Douglas2019, Rebull2016, Rebull2017, Rebull2018, Rebull2020, Curtis2019a, Curtis2019b, Curtis2020}. New and evermore precise data is becoming available at a rate that is outpacing efforts to re-calibrate gyrochronology, which is necessary to capture the complex relationship between a star's spin and its age.

%Nevertheless, it is by now clear that the simple $t^{1/2}$ scaling of rotation rates with age breaks down in certain regimes of stellar mass and age. Efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}. More recently, low-mass stars were shown to experience a temporary epoch of stalled spin-down which is not accounted for by standard gyrochronology models \citep{Curtis2019a}. \citet{Spada2020} and \citet{Curtis2020} proposed that this epoch of apparent stalling may arise from rotational coupling of the core and envelope, in which redistribution of angular momentum from the interior to the surface temporarily offsets the effects of magnetic braking. Regardless of the true mechanism, rotation periods in this part of parameter space have diminished utility in predicting stellar ages.

For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}. This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with masses $\gtrsim$~1\msun experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}. Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep{vanSaders2019}. Those authors also predicted a pile-up of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the Kepler Input Catalog \citep[KIC,][]{Brown2011}. While \citet{vanSaders2019} favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.

More recently, \citet{Hall2021} used the asteroseismic rotation rates of Kepler dwarfs, with a different selection bias from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model. While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which varies with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.

%Similar to the apparent stalling of low-mass stars mentioned above,

%If this is the case, a time-dependent (or, more accurately, Rossby-dependent) magnetic field morphology might explain both the anomalously rapid rotation of old field stars and the existence of a detection edge, as the amplitude of variability should also be connected to the field morphology.

%period-detection algorithms are less efficient for stars with (i) smaller amplitude variations, and (ii) quasi-periodic or stochastic variations. The 
%number, size, contrast, and relative orientations of spots on the surface of stars

Here we examine the rotation period distribution of main-sequence stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of a long-period pile-up for F- and early G-type stars. \textbf{In this work we refer to this feature as the long-period pile-up, the Rossby ridge, or simply the ridge.} We argue that several aspects of this pile-up feature are consistent with expectations from the WMB model. We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, and present our conclusions in \S\ref{sec:conclusions}.

\section{Sample Selection} \label{sec:sample}
Below, we describe the samples utilized in this work. All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data.

%We consider two samples in this work. The first sample is derived from the CKS sample, which itself is a subset of Kepler exoplanet candidate host stars that were spectroscopically characterized in a homogeneous manner. The second sample arises from a cross-match of the Kepler field with the LAMOST spectroscopic survey \textbf{(citation)}.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{surveys.pdf}
    \caption{The \teff-\prot plane using rotation periods from \citet{McQuillan2014} or, in the case of the CKS sample, \citet{McQuillan2013}, with \teff originating from the source denoted at top. The \citet{McQuillan2014} \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars. The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the differentially rotating surface.}
    \label{fig:surveys}
\end{figure*}

\subsection{California--Kepler Survey} \label{sec:cks}
The CKS project gathered high-resolution spectroscopy for 1305 Kepler planet host stars \citep{Petigura2017}. CKS spectra were acquired with the Keck/HIRES spectrograph \citep{Vogt1994} and spectroscopic parameters were determined by averaging parameters from the SpecMatch pipeline \citep{Petigura2015} and SME@XSEDE, a Python implementation of the Spectroscopy Made Easy pipeline \citep{Valenti1996}. The internal errors on \teff from the CKS catalog are 60~K.

We compiled rotation periods for these stars from a variety of literature sources including \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018}. For each star in the sample we then visually inspected the Kepler light curve folded on all available literature periods, as well as the first harmonics and sub-harmonics of those periods, and recorded our preferred period along with a reliability flag. Our procedure is explained in detail in \S2.1 of \citet{David2021}, and rotation period vetting sheets for each Kepler Object of Interest (KOI) are publicly available through Zenodo.\footnote{\url{http://10.0.20.161/zenodo.4645437}} We note that the vast majority of stars in the CKS sample host small planets ($R_P < 4$~\rearth) and as such it is not expected that the host stars have experienced tidal spin-up from the planets.

In addition to the original CKS catalog, we also cross-matched our sample with the catalogs of \citet{Brewer2018} and \citet{Martinez2019}, both of which presented spectroscopic parameters for CKS stars based on independent analysis of the same spectra. The \citet{Brewer2018} study, referred to here as SPOCS, also published elemental abundances and ages from isochrone fitting for the CKS sample.

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ridge.pdf}
    \caption{The \teff-\prot plane for the CKS sample. Point colors are scaled to the CKS ages determined from isochrone fitting. The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements. The black trapezoid indicates the approximate area of the ridge. The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
    \label{fig:ridge}
\end{figure*}
\fi

%Points on contours
%Different color for 
%, pixel

\subsection{LAMOST--Kepler} \label{sec:lamost}
The LAMOST--Kepler project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}. We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} rotation period catalog, which published rotation periods for $>$34000 Kepler targets. The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015}.

We matched 10844 LAMOST targets to 10650 unique Kepler IDs. For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude. We additionally selected sources with 3000~K$< \teff <$~8000~K, 3$<$~\logg~$<$5 dex, and -2~$<$[Fe/H]$<$~2 dex. The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K). There is negligible overlap (only 3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.

%\textbf{Jason: can you provide any relevant background info and describe the rationale behind the cuts you made to ensure the cross-match was relatively good?}

\textbf{We additionally cross-matched the LAMOST--Kepler catalog with the CKS catalog...Note, LAMOST catalog contains duplicates from multiple visits. Describe how we handle these.}

\subsection{APOGEE--Kepler}
The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE,][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV][]{Blanton2017}. The spectroscopic analysis pipeline for SDSS DR16 is described in \citet{Jonsson2020}. We used Gaia DR2 source IDs \citep{Gaia2016, Gaia2018} to cross-match Megan Bedell's Gaia--Kepler catalog\footnote{\url{https://gaia-kepler.fun/}} with the APOGEE DR16 catalog \citep{Ahumada2020}. Kepler IDs were then used to cross-match this table with the \citet{McQuillan2014} catalog. While the current overlap between Kepler targets and APOGEE is small compared to the LAMOST catalog, APOGEE DR17 will contain more dwarf stars and provide a better resource for studies such as ours. The focus of this work are overdensities in the \teff--\prot plane, and as these appear to be less prominent when using  APOGEE DR16 temperatures (Figure~\ref{fig:surveys}) we conclude that LAMOST and CKS provide more accurate estimates of \teff and do not analyze the APOGEE sample further.

\subsection{The Sun}
To place the Sun in the context of the long-period pile-up, we use up-to-date estimates of the Sun's effective temperature, rotation period, and age. Following the IAU 2015 Resolution B3 \citep{Mamajek2015}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K. For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25$~d with an upper error that encompasses the period of $\approx$~36~d measured near the poles from differential rotation studies \citep[][and references therein]{Thompson2003}.  The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep{Bahcall1995}.

%\prot = 26.09~d, with minimum and maximum periods of $P_\mathrm{min}=24.5$ and $P_\mathrm{max}=28.5$, from \citet{Donahue1996}. Measurements of differential rotation on the solar surface have found an equatorial period of approximately 25 days, increasing to about 36 days at the poles \citep[][and references therein]{Thompson2003}.

%\textbf{Measurements of differential rotation on the solar surface have found (SOHO/MDI results and possibly GONG...see e.g. Ulrich 1988)...}

\section{Analysis}
\label{sec:analysis}

\subsection{Initial observations}
A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}. Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}).

The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:ridge}). As we show in \S\ref{subsec:ages}, stars along the long-period edge have a range of ages. Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale. Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge. These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.

Interestingly, in both the CKS and LAMOST samples, there appears to be a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}. %\textbf{What to make of this? We can fit a model to the period distribution to measure exactly where the break is and test its significance.}

In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K. This cluster of points has a similar slope in the \teff--\prot plane as the ridge, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined. After inspecting the light curves of stars on the ridge and above the ridge in the temperature range 6100~K~$<$~ \teff~$<6300$~K we did not notice any obvious trends or differences between the two samples. The periods of the stars in the cluster above the ridge appear to be just as reliable as those on the ridge. Interestingly, a similar clustering of points is not observed in the LAMOST--Kepler sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.

\subsection{Constant Rossby model}
\label{subsec:rossby}
In the WMB model of \citet{vanSaders2016, vanSaders2019}, a star spins down until it reaches a critical Rossby number, at which point magnetic braking is stalled. Since Rossby number is highly sensitive to temperature through its dependence on the convective turnover timescale ($\mathrm{Ro} = P_\mathrm{rot}/\tau_\mathrm{cz}$, where $\tau_\mathrm{cz}$ is the convective turnover timescale), this critical threshold corresponds to different rotation periods for stars of different \teff, leading to a pile-up in the \teff-\prot plane. Using a small sample of Kepler targets with rotation periods determined from brightness modulations, \citet{vanSaders2016} proposed this threshold happens at a critical Rossby number of $\mathrm{Ro_{crit}} \sim \mathrm{Ro_\odot}$. From a sample of Kepler targets with asteroseismic rotation rates, \citet{Hall2021} found a similar critical Rossby number of $\mathrm{Ro_{crit}} = 1.97$.

We tested the hypothesis that the long-period pile-up observed in the LAMOST--Kepler sample is due to WMB by fitting constant Rossby models to the \prot boundary. For a given \teff, this model predicts \prot as:

\begin{equation}
    \prot (\mathrm{Ro}, \teff) = \mathrm{Ro} \times \tau_\mathrm{cz}(\teff),
\end{equation}

where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}:

\begin{multline}
\tau_\mathrm{cz}(\teff) = 314.24\exp \left [ -\left (\frac{T_\mathrm{eff}}{1952.5 \mathrm{K}}  \right ) - \left (\frac{T_\mathrm{eff}}{6250 \mathrm{K}}  \right )^{18} \right ] \\+ 0.002.
\end{multline}

To obtain a reasonable description of the long-period edge we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K. We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long period edge. Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K. To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling. The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values. We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane. We show these curves in relation to the full LAMOST--Kepler sample and to constant Rossby models in Figure~\ref{fig:pctl}.

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{pctl-curves.pdf}
    \caption{At top, The LAMOST--Kepler sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}. At bottom, the same curves are shown in relation to fiducial constant Rossby models.}
    \label{fig:pctl}
\end{figure}
\fi

%We computed constant Rossby models across the range of \teff bins using the equation for the convective turnover timescale presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}. This equation is valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K. %\textbf{Note: we can also try Landin 2010 and Barnes Kim 2010 formulas for convective turnover timescales.}

We performed an initial Levenberg-Marquardt non-linear least-squares fit of a constant Rossby model to the long-period edge with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class to optimize the following likelihood:

\begin{multline}
    \ln{p} (y | T_\mathrm{eff}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
\end{multline}

where

\begin{equation}
    s_n^2 = \sigma^2 + f^2 P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff})^2,
\end{equation}

and $y_n$ is the value of a \prot percentile curve in the $n$th \teff bin. This is a Gaussian likelihood where the variance is underestimated by some fraction, $f$. We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K. We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro and $f$ with ranges of (0.1,10) and (0,10), respectively. Convergence was assessed by ensuring the chain length was at least 100 times longer than the chain autocorrelation lengths for each parameter. A similar analysis was performed for the 10th percentile curve, restricted in the range of 4500~K$\lesssim \teff \lesssim$~5800~K where a constant Rossby model provides a reasonable fit.

The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$5000--6250~K, with fractional residuals $\lesssim5\%$ over this range. Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fits.pdf}
    \caption{Constant Rossby model fits to the 90th and 10th \prot percentile curves (top). Residuals from the median models from the MCMC sampling are shown in the lower panel.}
    \label{fig:fits}
\end{figure}
\fi

\begin{deluxetable}{lccc}
\tabletypesize{\scriptsize}
\tablecolumns{4}
\tablewidth{0pt}
\tablecaption{Results of constant Rossby model fits.\label{tab:mcmc}}
\tablehead{
\colhead{Sample} & 
\colhead{Curve} & 
\colhead{Ro} & 
\colhead{f}
}
\startdata
LAMOST--Kepler & 10th \prot pctl. & 0.355 $\pm$ 0.003 & 0.048 $\pm$ 0.007 \\
LAMOST--Kepler & 90th \prot pctl. & 1.467 $\pm$ 0.007 & 0.034 $\pm$ 0.004 \\
CKS & ridge & 
\enddata
%\tablecomments{}
\end{deluxetable}

\subsection{Comparison with asteroseismic rotation rates}
\label{subsec:asteroseismic}

\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies. We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the LAMOST--Kepler sample while such an offset is either absent or not as apparent relative to the CKS sample. Such an offset could be due to differing spectroscopic temperature scales between the two studies. To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019}. We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average. Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{asteroseismic.pdf}
    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.}
    \label{fig:asteroseismic}
\end{figure}
\fi

\subsection{Comparison with theory}
\label{subsec:models}

We compared the \teff--\prot distribution of the CKS and LAMOST--Kepler samples with the theoretical predictions of \citet{vanSaders2019}. Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function. The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}. Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars. \textbf{The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values, at fixed \teff, when compared to the observations. This suggests that, if the WMB model is correct, the critical Rossby number is lower than previously assumed.}

To quantify the degree of agreement between the theoretical models and observations we computed the 10th and 90th percentile \prot ranges of the standard and WMB models in overlapping \teff bins, analogous to how the upper and lower boundaries of the observed \prot distribution were found in \S\ref{subsec:rossby}. We computed the $\chi^2$ values between the observed upper edge and the 90th percentile ranges of the standard and WMB models, finding the WMB model is preferred with a $\Delta \chi^2 = 154$. Moreover, the WMB model better reproduces the slope of the observed long-period edge between 5300--6000~K (Figure~\ref{fig:model-pctl}). While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}.

Notably, both models fail to reproduce the observed short-period edge.

%\textbf{Some caveats about how the observations and models may not have exactly the same logg distributions? Jen, do you have any guidance here? I think you mentioned being able to get additional parameters for the model populations.}

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=0.49\linewidth]{std-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{std-model-lamost.pdf}
    \includegraphics[width=0.49\linewidth]{wmb-model-lamost.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.}
    \label{fig:models}
\end{figure*}
\fi

\iffalse
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{model-pctl.pdf}
    \caption{Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--Kepler observations.}
    \label{fig:model-pctl}
\end{figure}
\fi

% \subsection{Piecewise polynomial model}
% We fit a piecewise linear relation to the CKS stars on the ridge with $5800~K<$~\teff~$<6400~K$. We performed an initial Levenberg-Marquardt least-squares fit with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class. We then performed Markov chain Monte Carlo sampling (MCMC) with the \texttt{emcee} package \citep{emcee2013, emcee2019}.

% \textbf{Can we measure where the overdensity vanishes by fitting multi-component gaussians to the prot distributions?}

\subsection{Ages of stars on the ridge}
\label{subsec:ages}
The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge. In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs. Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting. In both cases, there appears to be an age gradient such that hotter stars are younger on average. However, we note that such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars. We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages.

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence. As a result, an age-gradient is expected along the pile-up as stars of different mass depart the feature at different age. Though we have not measured this gradient, it can be deduced that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.

%\textbf{In this study we have allowed for the ridge to have a width, and since stars are not expected to completely stop spinning down in the WMB model, we expect there to be an age-gradient on the ridge along the period axis.}

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{ages.pdf}
    \caption{Above, H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b). Below, the \teff-age plane for CKS stars along the ridge using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
    \label{fig:ages}
\end{figure*}
\fi

%\section{Discussion} \label{sec:discussion}

\section{Discussion} \label{sec:discussion}
\textbf{Some discussion of subgiants, some discussion of the other epoch of stalling noted in Curtis and how the two are not to be confused.}

\section{Conclusions} \label{sec:conclusions}

Our primary conclusions are summarized as follows:

\begin{enumerate}
    \item We observe an overdensity in the \teff-\prot plane at the long-period edge for main-sequence stars with \teff~$\gtrsim$5800~K. We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates. Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim1$~\msun.

    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$5000-6250~K. A pile-up of stars with a constant Rossby number and a variety of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.

    \item The pile-up is populated by stars with a wide range of isochrone ages ($\sim$1--6~Gyr). Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile-up onto the ridge on a timescale of 1--3~Gyr, compatible with the predictions of \citet{vanSaders2019}. 
    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, \textbf{appears to have evolved away from the pile-up. Or did it? Not clear}

    \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane. This overdensity, which we recover through kernel density estimation, is not as prominent as the long-period pile-up. \textbf{The pile-up on the short-period edge can also be fit with a constant Rossby model, though over a somewhat different range of \teff. }

    \item \textbf{Some statement about where the pile-up ends. At what Teff does density drop off by factor of X.}

    \item We observe a change in slope to the long-period edge of F-type stars at $\teff~=~6250~\pm~32$~K, coinciding with the Kraft break. \textbf{The statistical significance of the change in slope across this break is $\sim 3\sigma$. Note...this may just be due to a subgiant branch crossing through the main sequence Prot distribution}

    % \item We compared our data to the standard and weakened magnetic braking models presented in \citet{vanSaders2019}, \textbf{finding...}

    \item The existence of this pile-up, or ridge, limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the Rossby ridge. For example, a Sun-like star may spend several Gyr evolving across the ridge.

    \item We find tentative evidence for an age-gradient along the ridge, such that hotter stars on the ridge are younger on average. Relatedly, the age dispersion along the ridge is non-uniform as a function of temperature, with hotter stars showing a smaller dispersion. This observation suggests that hotter stars reside on the ridge for a shorter period of time relative to cooler stars. However, a more careful analysis is required to conclusively show this effect is not due to the intrinsic age gradient expected among a sample of main sequence stars with different masses.

    \item Authors using the method of gyro-tagging, which uses rotation periods to identify coeval stellar associations, should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.

\end{enumerate}

The code and data tables required to reproduce the figures and analysis presented here are publicly available on GitHub.\footnote{\url{https://github.com/trevordavid/rossby-ridge}}

\newpage 
\appendix 
\section{Additional figures}

% \begin{figure*}
%     \centering
%     \includegraphics[width=\linewidth]{rossby-rvar.pdf}
%     \caption{The relationship between Rossby number and \rvar for the LAMOST--Kepler sample. \rvar values are sourced from \citet{Lu2021}.}
%     \label{fig:rossby-rvar}
% \end{figure*}

% \begin{figure*}
%     \centering
%     \includegraphics[width=\linewidth]{hrd.pdf}
%     \caption{H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS and LAMOST--Kepler samples. \textbf{There seems to be a main cluster of orange points in the right panel, approximately tracing an isochrone or range of isochrones, and then a gap between that cluster and the orange points at upper right of the black box. Are these stars that have peeled off the MS, or are there multiple bursts of star formation?}}
%     \label{fig:hrd}
% \end{figure*}

% \begin{figure*}
%     \centering
%     \includegraphics[width=\linewidth]{abundance.pdf}
%     \caption{Distribution of isochrone ages, [Y/Al], and [Y/Mg] ratios for CKS stars on the ridge and in an equivalent range of Teff/logg (control). The ridge stars are younger and tend to have moderately higher [Y/Al] and [Y/Mg], also indicating they are younger than average.}
%     \label{fig:my_label}
% \end{figure*}

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{scatter-color.pdf}
    \caption{LAMOST--Kepler sample in \teff--\prot plane color coded by \logg (left) and [Fe/H] (right). \textbf{This plot shows overlap between the ridge and lower gravity subgiants evolving through it (we think).}}
    \label{fig:my_label}
\end{figure*}
\fi

\iffalse
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{tp-comparison.pdf}
    \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
    \label{fig:tp-comparison}
\end{figure*}
\fi

\begin{acknowledgments}
We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog. It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute and the Stars Group at the American Museum of Natural History for helpful discussions. This work made use of the gaia-kepler.fun crossmatch database created by Megan Bedell. This paper includes data collected by the Kepler mission and obtained from the MAST data archive at the Space Telescope Science Institute (STScI). Funding for the Kepler mission is provided by the NASA Science Mission Directorate. STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555. Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences. Funding for the project has been provided by the National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS-IV acknowledges support and resources from the Center for High Performance Computing  at the University of Utah. The SDSS website is www.sdss.org. SDSS-IV is managed by the Astrophysical Research Consortium for the Participating Institutions of the SDSS Collaboration including the Brazilian Participation Group, the Carnegie Institution for Science, Carnegie Mellon University, Center for Astrophysics | Harvard \& Smithsonian, the Chilean Participation Group, the French Participation Group, Instituto de Astrof\'isica de Canarias, The Johns Hopkins University, Kavli Institute for the Physics and Mathematics of the Universe (IPMU) / University of Tokyo, the Korean Participation Group, Lawrence Berkeley National Laboratory, Leibniz Institut f\"ur Astrophysik Potsdam (AIP),  Max-Planck-Institut f\"ur Astronomie (MPIA Heidelberg), Max-Planck-Institut f\"ur Astrophysik (MPA Garching), Max-Planck-Institut f\"ur Extraterrestrische Physik (MPE), National Astronomical Observatories of China, New Mexico State University, New York University, University of Notre Dame, Observat\'ario Nacional / MCTI, The Ohio State University, Pennsylvania State University, Shanghai Astronomical Observatory, United Kingdom Participation Group, Universidad Nacional Aut\'onoma de M\'exico, University of Arizona, University of Colorado Boulder, University of Oxford, University of Portsmouth, University of Utah, University of Virginia, University of Washington, University of Wisconsin, Vanderbilt University, and Yale University. This research has made use of NASA's Astrophysics Data System Bibliographic Services.

\end{acknowledgments}

\bibliography{main}{}
\bibliographystyle{aasjournal}

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

### 2021-11-07 src/ms.tex
Updating manuscript

@@
 %% using aastex version 6.3
-%\documentclass[linenumbers,preprint2]{aastex631}
-\documentclass[twocolumn]{aastex631}
+\documentclass[linenumbers,tighten,trackchanges,twocolumn]{aastex631}
 
@@
 %% Reintroduced the \received and \accepted commands from AASTeX v5.2
-%\received{March 1, 2021}
-%\revised{April 1, 2021}
+\received{\today}
+%\revised{\today}
 %\accepted{\today}
@@
 
-%\\or\\Weakened Magnetic Braking Supported by Rotation Rates of Kepler Exoplanet Hosts}
-
-% Ridge
-% Edge
-% Ledge
-% Cornice
-
 \correspondingauthor{Trevor J. David}
@@
 
-%Gaby Contardo
 %Lucy Lu
@@
 We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters derived from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a long-period pile-up in the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding 5800~K.
-Stars in the pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
+Stars in the long-period pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
 In the effective temperature range of 5000--6250~K the long-period pile-up is well-described by a single-parameter model of constant Rossby number, with a critical value of $\mathrm{Ro_{crit}} \sim$\textbf{XYZ}.
@@
 
-%% Keywords should appear after the \end{abstract} command. 
-%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
-%% https://astrothesaurus.org
-%% You will be asked to selected these concepts during the submission process
-%% but this old "keyword" functionality is maintained in case authors want
-%% to include these concepts in their preprints.
-\keywords{Stellar rotation (1629) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- F dwarf stars(516)}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
+\keywords{Stellar rotation (1629) --- Solar analogs (1941) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- Stellar winds(1636)}
 
@@
 
-%Similar to the apparent stalling of low-mass stars mentioned above,
-
-%If this is the case, a time-dependent (or, more accurately, Rossby-dependent) magnetic field morphology might explain both the anomalously rapid rotation of old field stars and the existence of a detection edge, as the amplitude of variability should also be connected to the field morphology.
-
-%period-detection algorithms are less efficient for stars with (i) smaller amplitude variations, and (ii) quasi-periodic or stochastic variations. The 
-%number, size, contrast, and relative orientations of spots on the surface of stars
-
 Here we examine the rotation period distribution of main-sequence stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of a long-period pile-up for F- and early G-type stars.
@@
 All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data.
-
-%We consider two samples in this work.
-The first sample is derived from the CKS sample, which itself is a subset of Kepler exoplanet candidate host stars that were spectroscopically characterized in a homogeneous manner.
-The second sample arises from a cross-match of the Kepler field with the LAMOST spectroscopic survey \textbf{(citation)}.
+For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys.
 
@@
 
-\iffalse
 \begin{figure*}
@@
 \end{figure*}
-\fi
-
-%Points on contours
-%Different color for 
-%, pixel
 
@@
 The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K).
-There is negligible overlap (only 3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
-
-%\textbf{Jason: can you provide any relevant background info and describe the rationale behind the cuts you made to ensure the cross-match was relatively good?}
+There is negligible overlap (3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
 
@@
 While the current overlap between Kepler targets and APOGEE is small compared to the LAMOST catalog, APOGEE DR17 will contain more dwarf stars and provide a better resource for studies such as ours.
-The focus of this work are overdensities in the \teff--\prot plane, and as these appear to be less prominent when using  APOGEE DR16 temperatures (Figure~\ref{fig:surveys}) we conclude that LAMOST and CKS provide more accurate estimates of \teff and do not analyze the APOGEE sample further.
+The focus of this work are overdensities in the \teff--\prot plane, and as these appear to be less prominent when using  APOGEE DR16 temperatures (Figure~\ref{fig:surveys}) we conclude that LAMOST and CKS provide more precise estimates of \teff and do not analyze the APOGEE sample further.
 
@@
 
-%\textbf{Measurements of differential rotation on the solar surface have found (SOHO/MDI results and possibly GONG...see e.g. Ulrich 1988)...}
-
 \section{Analysis}
@@
 The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:ridge}).
-As we show in \S\ref{subsec:ages}, stars along the long-period edge have a range of ages.
+As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
 Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale.
@@
 
-Interestingly, in both the CKS and LAMOST samples, there appears to be a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
-%\textbf{What to make of this?
-We can fit a model to the period distribution to measure exactly where the break is and test its significance.}
+In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
+A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
+\textbf{Some interpretation.}
 
@@
 
-\iffalse
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{pctl-curves.pdf}
-    \caption{At top, The LAMOST--Kepler sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}.
+    \caption{At top, The LAMOST--McQuillan sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}.
 At bottom, the same curves are shown in relation to fiducial constant Rossby models.}
@@
 \end{figure}
-\fi
 
@@
 
-The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
+The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
 Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.
 
-\iffalse
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{asteroseismic.pdf}
-    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.}
+    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
+\textbf{Note about teff scales and shifts}.}
     \label{fig:asteroseismic}
 \end{figure}
-\fi
 
@@
 
-Notably, both models fail to reproduce the observed short-period edge.
+\textbf{Notably, both models fail to reproduce the observed short-period edge.}
 
-%\textbf{Some caveats about how the observations and models may not have exactly the same logg distributions?
-Jen, do you have any guidance here?
-I think you mentioned being able to get additional parameters for the model populations.}
-
-\iffalse
 \begin{figure*}
@@
 \end{figure*}
-\fi
 
-\iffalse
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{model-pctl.pdf}
-    \caption{Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--Kepler observations.}
+    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
+Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
     \label{fig:model-pctl}
 \end{figure}
-\fi
 
@@
 We then performed Markov chain Monte Carlo sampling (MCMC) with the \texttt{emcee} package \citep{emcee2013, emcee2019}.
-
-% \textbf{Can we measure where the overdensity vanishes by fitting multi-component gaussians to the prot distributions?}
 
@@
 We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages.
+This observation could be due to cooler stars populating the ridge for longer periods of time (relative to hotter stars), the lower precision of isochrone ages for cooler stars, or some combination of the two effects.
 
@@
 In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence.
-As a result, an age-gradient is expected along the pile-up as stars of different mass depart the feature at different age.
-Though we have not measured this gradient, it can be deduced that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.
+As a result, an age-gradient is expected along the pile-up as stars of different mass depart the pile-up at different ages.
+Though we have not measured this gradient, the data suggest that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.
 
-%\textbf{In this study we have allowed for the ridge to have a width, and since stars are not expected to completely stop spinning down in the WMB model, we expect there to be an age-gradient on the ridge along the period axis.}
-
-\iffalse
 \begin{figure*}
@@
 \end{figure*}
-\fi
-
-%\section{Discussion} \label{sec:discussion}
 
@@
 \begin{enumerate}
-    \item We observe an overdensity in the \teff-\prot plane at the long-period edge for main-sequence stars with \teff~$\gtrsim$5800~K.
+    \item We observe an overdensity at the long-period edge of the \teff-\prot distribution of Kepler main-sequence stars with \teff~$\gtrsim$~5800~K.
 We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates.
@@
 
-    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$5000-6250~K.
+    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$~5000-6250~K.
 A pile-up of stars with a constant Rossby number and a variety of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
@@
 Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile-up onto the ridge on a timescale of 1--3~Gyr, compatible with the predictions of \citet{vanSaders2019}.
-    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, \textbf{appears to have evolved away from the pile-up.
-Or did it?
-Not clear}
+    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, may have evolved away from the pile-up.
+\textbf{However, this statement depends on the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample.
+For example, stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}
+
+    %Ruth: include some discussion of what this says about the Sun.
+Not just because it implies the Sun's dynamo might be different from McQuillan stars because it's been through WMB, but also how did the Sun get there?
+Its Rossby number is higher than the stars on the ridge...
+Can expansion alone explain the ~6 days of rotation period lengthening it has gone through since it passed through the ridge around 2.5 Gyr ago?
 
@@
 Note...this may just be due to a subgiant branch crossing through the main sequence Prot distribution}
-
-    % \item We compared our data to the standard and weakened magnetic braking models presented in \citet{vanSaders2019}, \textbf{finding...}
 
@@
 \end{figure*}
-\fi
 
-\iffalse
 \begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{tp-comparison.pdf}
+    \includegraphics[width=\linewidth]{comparison.pdf}
     \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
@@
 \end{figure*}
-\fi
 
 \begin{acknowledgments}
-We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog.
+We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog and Rodrigo Luger for assistance with the \texttt{showyourwork} package.
 It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute and the Stars Group at the American Museum of Natural History for helpful discussions.

### 2021-12-08 src/ms.tex
Initial commit

@@
 \usepackage{xspace}
+\usepackage{xcolor}
 \usepackage{amsmath}
 \usepackage{longtable}
+
+\newcommand{\taucz}{$\tau_\mathrm{cz}$\xspace}
+\newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
+\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
+\newcommand{\lamostkep}{LAMOST--Kepler\xspace}
+\newcommand{\lamostmcq}{LAMOST--McQuillan\xspace}
+\newcommand{\teffmin}{5800~K\xspace}
+\newcommand{\agerange}{$\sim$2--6~Gyr\xspace}
+\newcommand{\agemin}{$\sim$2--3~Gyr\xspace}
 
@@
 
-\newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
-
-\newcommand{\kms}{km~s$^{-1}$}
-\newcommand{\ms}{m~s$^{-1}$}
-\newcommand{\gcc}{g~cm$^{-3}$}
-\newcommand{\masyr}{mas~yr$^{-1}$}
-\newcommand{\err}{\textit{$\pm$}}
-
 \newcommand{\Mstar}{\ensuremath{M_{\star}}\xspace}
@@
 \newcommand{\vsini}{\ensuremath{v \sin i}\xspace} 
-\newcommand{\prot}{\ensuremath{P_\mathrm{rot}}\xspace}
-\newcommand{\rvar}{\ensuremath{R_\mathrm{var}}\xspace}
-\newcommand{\vtan}{\ensuremath{v_\mathrm{tan}}\xspace}
-\newcommand{\vb}{\ensuremath{v_\mathrm{b}}\xspace}
-
-\newcommand{\mearth}{$M_\oplus$\xspace}
-\newcommand{\rearth}{$R_\oplus$\xspace}
-
 \newcommand{\msun}{$M_\odot$\xspace}
@@
 \newcommand{\lsun}{$L_\odot$\xspace}
+\newcommand{\tsun}{$T_\mathrm{eff,\odot}$\xspace}
 \newcommand{\rhosun}{$\rho_\odot$\xspace}
@@
 \newcommand{\rhostar}{$\rho_*$\xspace}
+\newcommand{\logage}{$\text{log(age)}$\xspace}
+\newcommand{\mearth}{$M_\oplus$\xspace}
+\newcommand{\rearth}{$R_\oplus$\xspace}
+\newcommand{\prot}{\ensuremath{P_\mathrm{rot}}\xspace}
+\newcommand{\rvar}{\ensuremath{R_\mathrm{var}}\xspace}
+\newcommand{\rper}{\ensuremath{R_\mathrm{per}}\xspace}
 
-\newcommand{\logage}{$\text{log(age)}$\xspace}
-\newcommand{\rp}{$R_P$\xspace}
-\newcommand{\rprstar}{$R_P/R_\star$\xspace}
-
-\newcommand{\logrp}{$\log_{10}(R_P/R_\oplus)$\xspace}
-\newcommand{\deltaage}{$\Delta \log_{10}(\mathrm{age~yr}^{-1})$\xspace}
-\newcommand{\deltamstar}{$\Delta M_*$\xspace}
-\newcommand{\deltateff}{$\Delta T_\mathrm{eff}$\xspace}
-
-\newcommand{\mjup}{$M_\mathrm{Jup}$\xspace}
-\newcommand{\galex}{\textit{GALEX}\xspace}
 \newcommand{\gaia}{\textit{Gaia}\xspace}
@@
 \newcommand{\spitzer}{\textit{Spitzer}}
-\newcommand{\hipparcos}{\textit{Hipparcos}}
-\newcommand{\ktwo}{\textit{K2}\xspace}
 \newcommand{\tess}{\textit{TESS}\xspace}
@@
 
-\newcommand{\ymg}{[Y/Mg]\xspace}
-\newcommand{\yal}{[Y/Al]\xspace}
-\newcommand{\alphafe}{[$\alpha$/Fe]\xspace}
-\newcommand{\dexgyr}{dex~Gyr$^{-1}$\xspace}
+\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
+\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th Street, New York, NY 10024, USA}
+\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY, USA}
+\newcommand{\ifa}{Institute for Astronomy, University of Hawai’i, Honolulu, HI, USA}
 
@@
 % \title{The Rossby Ridge: a Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages}
-\title{A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking}
+%\title{A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking}
+%\title{Two Pile-ups in the Temperature--Period Distribution of Main-Sequence Stars:\\Possible Evidence for Weakened Magnetic Braking and Core-Envelope Coupling}
+\title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}
 
@@
 \author[0000-0001-6534-6246]{Trevor J.\ David}
-\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
-\affil{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th Street, New York, NY 10024, USA}
+\affil{\cca}
+\affil{\amnh}
 
 \author[0000-0003-4540-5661]{Ruth Angus}
-\affil{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th Street, New York, NY 10024, USA}
-\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
-\affil{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY, USA}
+\affil{\amnh}
+\affil{\cca}
+\affil{\columbia}
 
 \author[0000-0002-2792-134X]{Jason L.~Curtis}
-\affil{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY, USA}
+\affil{\columbia}
 
-\author[0000-0002-4284-8638]{Jennifer van Saders}
-\affil{Institute for Astronomy, University of Hawai’i, Honolulu, HI, USA}
+\author[0000-0002-4284-8638]{Jennifer L.~van Saders}
+\affil{\ifa}
 
 \author[0000-0002-3011-4784]{Gabriella Contardo}
-\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}
+\affil{\cca}
 
-\author{Other authors TBD}
+\author[0000-0003-4769-3273]{Yuxi Lu}
+\affil{\columbia}
+\affil{\amnh}
 
-%Lucy Lu
-%Joel Zinn
-%Adrian Price-Whelan
+\author[0000-0002-7550-7151]{Joel C.~Zinn}
+\affil{\amnh}
 
 \begin{abstract}
-We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters derived from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a long-period pile-up in the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding 5800~K.
-Stars in the long-period pile-up are found to have a wide range of ages ($\sim$2--6~Gyr), meaning that, along the ridge, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
-In the effective temperature range of 5000--6250~K the long-period pile-up is well-described by a single-parameter model of constant Rossby number, with a critical value of $\mathrm{Ro_{crit}} \sim$\textbf{XYZ}.
-The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which stellar spin-down stalls once stars reach a critical Rossby number.
-These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than $\sim$1--2~Gyr.
-\textbf{We additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.}
+We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin.
+Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
+In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
+The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
+These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin.
+Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.
+{\color{blue}The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.}
 \end{abstract}
 
-\keywords{Stellar rotation (1629) --- Solar analogs (1941) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- Stellar winds(1636)}
+\keywords{Stellar rotation (1629) --- Solar analogs (1941) --- Stellar evolution (1599) --- Stellar magnetic fields (1610) --- Stellar winds (1636)}
 
@@
 
-Here we examine the rotation period distribution of main-sequence stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of a long-period pile-up for F- and early G-type stars.
-\textbf{In this work we refer to this feature as the long-period pile-up, the Rossby ridge, or simply the ridge.} We argue that several aspects of this pile-up feature are consistent with expectations from the WMB model.
-We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, and present our conclusions in \S\ref{sec:conclusions}.
+Here we examine the rotation period distribution of main-sequence stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pile-ups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
+We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, discuss some implications of these results in \S\ref{sec:discussion}, and present our conclusions in \S\ref{sec:conclusions}.
+
+%We argue that several aspects of the long-period pile-up feature are consistent with expectations from the WMB model.
+The short-period pile-up, by comparison, may be due a temporary epoch of stalled wind-driven losses due to core-envelope coupling.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{surveys.pdf}
+    \includegraphics[width=\textwidth]{./figures/surveys.pdf}
     \caption{The \teff-\prot plane using rotation periods from \citet{McQuillan2014} or, in the case of the CKS sample, \citet{McQuillan2013}, with \teff originating from the source denoted at top.
@@
 The internal errors on \teff from the CKS catalog are 60~K.
+The metallicity distribution of the CKS sample is centered near solar, with a mean and standard deviation of +0.03~dex and 0.18~dex, respectively.
 
@@
 For each star in the sample we then visually inspected the Kepler light curve folded on all available literature periods, as well as the first harmonics and sub-harmonics of those periods, and recorded our preferred period along with a reliability flag.
-Our procedure is explained in detail in \S2.1 of \citet{David2021}, and rotation period vetting sheets for each Kepler Object of Interest (KOI) are publicly available through Zenodo.\footnote{\url{http://10.0.20.161/zenodo.4645437}} We note that the vast majority of stars in the CKS sample host small planets ($R_P < 4$~\rearth) and as such it is not expected that the host stars have experienced tidal spin-up from the planets.
+Our procedure is explained in detail in \S2.1 of \citet{David2021}, and rotation period vetting sheets for each Kepler Object of Interest (KOI) are publicly available through Zenodo.\footnote{\url{http://10.0.20.161/zenodo.4645437}} The vast majority of stars in the CKS sample host small planets ($R_P < 4$~\rearth) and as such it is not expected that the host stars have experienced tidal spin-up from the planets.
 
@@
 The \citet{Brewer2018} study, referred to here as SPOCS, also published elemental abundances and ages from isochrone fitting for the CKS sample.
+We additionally cross-matched the CKS catalog with the LAMOST--Kepler catalog \citep{Xiang2019} which is described further in \S\ref{subsec:lamost}.
+We compare the \teff--\prot distributions of the CKS sample using \teff and rotation periods from a variety of sources in Appendix~\ref{app:teffprot}.
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{ridge.pdf}
-    \caption{The \teff-\prot plane for the CKS sample.
-Point colors are scaled to the CKS ages determined from isochrone fitting.
-The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements.
-The black trapezoid indicates the approximate area of the ridge.
-The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
-    \label{fig:ridge}
-\end{figure*}
-
-\subsection{LAMOST--Kepler} \label{sec:lamost}
+\subsection{LAMOST--Kepler} \label{subsec:lamost}
 The LAMOST--Kepler project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
@@
 For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
-We additionally selected sources with 3000~K$< \teff <$~8000~K, 3$<$~\logg~$<$5 dex, and -2~$<$[Fe/H]$<$~2 dex.
+We additionally selected sources with 3000~K~$<~\teff~<$~8000~K, 3~$<$~\logg~$<$~5~dex, and -2~$<$~[Fe/H]~$<$~2~dex.
 The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K).
-There is negligible overlap (3 stars) between our LAMOST--Kepler samples and the CKS samples since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
-
-\textbf{We additionally cross-matched the LAMOST--Kepler catalog with the CKS catalog...Note, LAMOST catalog contains duplicates from multiple visits. Describe how we handle these.}
+The metallicity distribution of the LAMOST--McQuillan sample is centered near solar, with a mean and standard deviation of -0.1~dex and 0.26~dex, respectively.
+There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
 
 \subsection{APOGEE--Kepler}
-The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE,][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV][]{Blanton2017}.
+The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV;][]{Blanton2017}.
 The spectroscopic analysis pipeline for SDSS DR16 is described in \citet{Jonsson2020}.
@@
 For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25$~d with an upper error that encompasses the period of $\approx$~36~d measured near the poles from differential rotation studies \citep[][and references therein]{Thompson2003}.
-The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep{Bahcall1995}.
-
-%\prot = 26.09~d, with minimum and maximum periods of $P_\mathrm{min}=24.5$ and $P_\mathrm{max}=28.5$, from \citet{Donahue1996}.
-Measurements of differential rotation on the solar surface have found an equatorial period of approximately 25 days, increasing to about 36 days at the poles \citep[][and references therein]{Thompson2003}.
+The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
 
@@
 Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}).
+{\color{blue} A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
+As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.}
 
-The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:ridge}).
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/kde.pdf}
+    \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic \citet{Hall2021} samples, from left to right.
+Empirical cluster sequences from \citet{Curtis2020} are shown by the dark grey lines.
+The orange dashed lines show constant Rossby curves of fiducial values (see \S\ref{subsec:rossby}).
+The short-period pile-up can be observed in the LAMOST--McQuillan sample for \teff~$\gtrsim$~5500~K.
+The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from its differentially rotating surface.}
+    \label{fig:kde}
+\end{figure*}
+
+The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
@@
 In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
-A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
-\textbf{Some interpretation.}
+A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
+{\color{blue}This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.}
 
-In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K.
-This cluster of points has a similar slope in the \teff--\prot plane as the ridge, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
+In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in the top panels of Figure~\ref{fig:models}).
+This cluster of points has a similar slope in the \teff--\prot plane as the long-period pile-up, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
 After inspecting the light curves of stars on the ridge and above the ridge in the temperature range 6100~K~$<$~ \teff~$<6300$~K we did not notice any obvious trends or differences between the two samples.
@@
 
-\subsection{Constant Rossby model}
-\label{subsec:rossby}
-In the WMB model of \citet{vanSaders2016, vanSaders2019}, a star spins down until it reaches a critical Rossby number, at which point magnetic braking is stalled.
-Since Rossby number is highly sensitive to temperature through its dependence on the convective turnover timescale ($\mathrm{Ro} = P_\mathrm{rot}/\tau_\mathrm{cz}$, where $\tau_\mathrm{cz}$ is the convective turnover timescale), this critical threshold corresponds to different rotation periods for stars of different \teff, leading to a pile-up in the \teff-\prot plane.
-Using a small sample of Kepler targets with rotation periods determined from brightness modulations, \citet{vanSaders2016} proposed this threshold happens at a critical Rossby number of $\mathrm{Ro_{crit}} \sim \mathrm{Ro_\odot}$.
-From a sample of Kepler targets with asteroseismic rotation rates, \citet{Hall2021} found a similar critical Rossby number of $\mathrm{Ro_{crit}} = 1.97$.
+\subsection{Comparison with theory}
+\label{subsec:models}
 
-We tested the hypothesis that the long-period pile-up observed in the LAMOST--Kepler sample is due to WMB by fitting constant Rossby models to the \prot boundary.
-For a given \teff, this model predicts \prot as:
+We compared the \teff--\prot distribution of the CKS and LAMOST--Kepler samples with the theoretical predictions of \citet{vanSaders2019}.
+Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
+The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
+Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
+The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations.
 
-\begin{equation}
-    \prot (\mathrm{Ro}, \teff) = \mathrm{Ro} \times \tau_\mathrm{cz}(\teff),
-\end{equation}
+To quantify the degree of agreement between the theoretical models and observations we computed the 10th and 90th percentile \prot ranges of the standard and WMB models in overlapping \teff bins, analogous to how the upper and lower boundaries of the observed \prot distribution were found in \S\ref{subsec:rossby}.
+We computed the $\chi^2$ values between the observed upper edge and the 90th percentile ranges of the standard and WMB models, finding the WMB model is preferred with a $\Delta \chi^2 = 154$.
+Moreover, the WMB model better reproduces the slope of the observed long-period edge between 5300--6000~K (Figure~\ref{fig:percentiles}).
 
-where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}:
+While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
+{\color{blue} We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.}
 
-\begin{multline}
-\tau_\mathrm{cz}(\teff) = 314.24\exp \left [ -\left (\frac{T_\mathrm{eff}}{1952.5 \mathrm{K}}  \right ) - \left (\frac{T_\mathrm{eff}}{6250 \mathrm{K}}  \right )^{18} \right ] \\+ 0.002.
-\end{multline}
+{\color{blue} Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
+In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
+the low-mass stars would be rotating more slowly than the model predictions).
+Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
+A higher normalization factor would cause the low-mass stars to spin down more at fixed age.}
 
-To obtain a reasonable description of the long-period edge we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
-We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long period edge.
-Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K.
-To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling.
-The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values.
-We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
-We show these curves in relation to the full LAMOST--Kepler sample and to constant Rossby models in Figure~\ref{fig:pctl}.
+{\color{blue} Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models.
+At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
+The core and envelope at these times are thus assumed to be decoupled.
+However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015}.
+When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
+This hypothesis is discussed further in \S\ref{subsec:shortperiod}.}
+
+\begin{figure*}
+    \centering 
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
+    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.
+The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.}
+    \label{fig:models}
+\end{figure*}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{pctl-curves.pdf}
-    \caption{At top, The LAMOST--McQuillan sample and the 10th and 90th \prot percentiles, computed as described in \S\ref{subsec:rossby}.
-At bottom, the same curves are shown in relation to fiducial constant Rossby models.}
-    \label{fig:pctl}
+    \includegraphics[width=\linewidth]{./figures/percentiles.pdf}
+    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
+Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
+    \label{fig:percentiles}
 \end{figure}
-
-%We computed constant Rossby models across the range of \teff bins using the equation for the convective turnover timescale presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}.
-This equation is valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K.
-%\textbf{Note: we can also try Landin 2010 and Barnes Kim 2010 formulas for convective turnover timescales.}
-
-We performed an initial Levenberg-Marquardt non-linear least-squares fit of a constant Rossby model to the long-period edge with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class to optimize the following likelihood:
-
-\begin{multline}
-    \ln{p} (y | T_\mathrm{eff}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
-\end{multline}
-
-where
-
-\begin{equation}
-    s_n^2 = \sigma^2 + f^2 P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff})^2,
-\end{equation}
-
-and $y_n$ is the value of a \prot percentile curve in the $n$th \teff bin.
-This is a Gaussian likelihood where the variance is underestimated by some fraction, $f$.
-We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
-We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro and $f$ with ranges of (0.1,10) and (0,10), respectively.
-Convergence was assessed by ensuring the chain length was at least 100 times longer than the chain autocorrelation lengths for each parameter.
-A similar analysis was performed for the 10th percentile curve, restricted in the range of 4500~K$\lesssim \teff \lesssim$~5800~K where a constant Rossby model provides a reasonable fit.
-
-The constant Rossby model provides a reasonably good description of the long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
-Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fits.pdf}
-    \caption{Constant Rossby model fits to the 90th and 10th \prot percentile curves (top). Residuals from the median models from the MCMC sampling are shown in the lower panel.}
-    \label{fig:fits}
-\end{figure}
-
-\begin{deluxetable}{lccc}
-\tabletypesize{\scriptsize}
-\tablecolumns{4}
-\tablewidth{0pt}
-\tablecaption{Results of constant Rossby model fits.\label{tab:mcmc}}
-\tablehead{
-\colhead{Sample} & 
-\colhead{Curve} & 
-\colhead{Ro} & 
-\colhead{f}
-}
-\startdata
-LAMOST--Kepler & 10th \prot pctl. & 0.355 $\pm$ 0.003 & 0.048 $\pm$ 0.007 \\
-LAMOST--Kepler & 90th \prot pctl. & 1.467 $\pm$ 0.007 & 0.034 $\pm$ 0.004 \\
-CKS & ridge & 
-\enddata
-%\tablecomments{}
-\end{deluxetable}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{asteroseismic.pdf}
+    \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
     \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
-\textbf{Note about teff scales and shifts}.}
+Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
+Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
-\end{figure}
-
-\subsection{Comparison with theory}
-\label{subsec:models}
-
-We compared the \teff--\prot distribution of the CKS and LAMOST--Kepler samples with the theoretical predictions of \citet{vanSaders2019}.
-Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
-Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-\textbf{The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values, at fixed \teff, when compared to the observations.
-This suggests that, if the WMB model is correct, the critical Rossby number is lower than previously assumed.}
-
-To quantify the degree of agreement between the theoretical models and observations we computed the 10th and 90th percentile \prot ranges of the standard and WMB models in overlapping \teff bins, analogous to how the upper and lower boundaries of the observed \prot distribution were found in \S\ref{subsec:rossby}.
-We computed the $\chi^2$ values between the observed upper edge and the 90th percentile ranges of the standard and WMB models, finding the WMB model is preferred with a $\Delta \chi^2 = 154$.
-Moreover, the WMB model better reproduces the slope of the observed long-period edge between 5300--6000~K (Figure~\ref{fig:model-pctl}).
-While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}.
-
-\textbf{Notably, both models fail to reproduce the observed short-period edge.}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.49\linewidth]{std-model-cks.pdf}
-    \includegraphics[width=0.49\linewidth]{wmb-model-cks.pdf}
-    \includegraphics[width=0.49\linewidth]{std-model-lamost.pdf}
-    \includegraphics[width=0.49\linewidth]{wmb-model-lamost.pdf}
-    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.}
-    \label{fig:models}
-\end{figure*}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{model-pctl.pdf}
-    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
-Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
-    \label{fig:model-pctl}
 \end{figure}
@@
 
+\subsection{Constant Rossby model}
+\label{subsec:rossby}
+In the WMB model of \citet{vanSaders2016, vanSaders2019}, a star spins down until it reaches a critical Rossby number, at which point magnetic braking ceases.
+Since Rossby number is highly sensitive to temperature through its dependence on the convective turnover timescale ($\mathrm{Ro} = P_\mathrm{rot}/\tau_\mathrm{cz}$, where $\tau_\mathrm{cz}$ is the convective turnover timescale), this critical threshold corresponds to different rotation periods for stars of different \teff, leading to a pile-up in the \teff-\prot plane.
+Using a small sample of Kepler targets with rotation periods determined from brightness modulations, \citet{vanSaders2016} proposed this threshold happens at a critical Rossby number of $\mathrm{Ro_{crit}} \sim \mathrm{Ro_\odot}$.
+
+%{\color{blue} From a sample of Kepler targets with asteroseismic rotation rates, \citet{Hall2021} found a good match to observations when adopting a critical Rossby number of $\mathrm{Ro_{crit}} = 1.97$.}
+
+We tested the hypothesis that the long-period pile-up observed in the LAMOST--Kepler sample is due to WMB by fitting constant Rossby models to the \prot boundary.
+For a given \teff, this model predicts \prot as:
+
+\begin{equation} \label{eq:1}
+    \prot (\mathrm{Ro}, \teff) = \mathrm{Ro} \times \tau_\mathrm{cz}(\teff),
+\end{equation}
+
+where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}:
+
+\begin{multline} \label{eq:2}
+\tau_\mathrm{cz}(\teff) = 314.24\exp \left [ -\left (\frac{T_\mathrm{eff}}{1952.5 \mathrm{K}}  \right ) - \left (\frac{T_\mathrm{eff}}{6250 \mathrm{K}}  \right )^{18} \right ] \\+ 0.002.
+\end{multline}
+
+To obtain a reasonable description of the long-period edge we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
+We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
+Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K.
+To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling.
+The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values.
+We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
+We show these curves in relation to the full LAMOST--Kepler sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
+
+%We computed constant Rossby models across the range of \teff bins using the equation for the convective turnover timescale presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}.
+This equation is valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K.
+%\textbf{Note: we can also try Landin 2010 and Barnes Kim 2010 formulas for convective turnover timescales.}
+
+We performed an initial Levenberg-Marquardt non-linear least-squares fit of a constant Rossby model to the long-period edge with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class to optimize the following likelihood:
+
+\begin{multline} \label{eq:3}
+    \ln{p} (y | T_\mathrm{eff}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
+\end{multline}
+
+where
+
+\begin{equation} \label{eq:4}
+    s_n^2 = \sigma^2 + f^2 P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys})^2,
+\end{equation}
+
+and $y_n$ is the value of a \prot percentile curve in the $n$th \teff bin.
+This is a Gaussian likelihood where the variance is underestimated by some fraction, $f$.
+Here $T_\mathrm{sys}$ is a constant to allow for a systematic offset between the data and the models used to calibrate the \taucz relation.
+We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
+We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro $f$, and $T_\mathrm{sys}$ with ranges of (0.1,10), (0,10), and (-1000~K, 1000~K), respectively.
+Convergence was assessed by ensuring the chain length was at least 100 times longer than the chain autocorrelation lengths for each parameter.
+A similar analysis was performed for the 10th percentile curve, restricted in the range of 4500~K$\lesssim \teff \lesssim$~5800~K where a constant Rossby model provides a reasonable fit.
+
+{\color{blue}  We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set.
+For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d.
+To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
+We note that this selection is particular to the CKS \teff scale and is not general.
+For the CKS sample, we additionally required $\logg~>~4$ and 5850~K~$<~\teff~<$~6500~K.
+In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.}
+
+The constant Rossby model provides a reasonably good description of the \lamostmcq long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
+Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, possibly because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.
+
+{\color{blue} We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation.
+Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 may be compatible with all of the data sets if modest \teff systematics are present.
+Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff})}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/mcmc.pdf}
+    \caption{\textit{Top:} Gaussian kernel density estimation of the LAMOST--McQuillan sample and the 10th and 90th \prot percentiles (which and black points, respectively), computed as described in \S\ref{subsec:rossby}.
+\textit{Middle:} The same 10th and 90th percentile curves shown in relation to constant Rossby curves.
+\textit{Bottom:} Residuals from the median models resulting from the MCMC sampling.
+The fits depicted above do not include a \teff offset.}
+    \label{fig:mcmc}
+\end{figure}
+
+\begin{deluxetable*}{lllll}
+\tabletypesize{\scriptsize}
+\tablecolumns{5}
+\tablewidth{0pt}
+\tablecaption{Results of constant Rossby model fits.\label{tab:mcmc}}
+\tablehead{
+\colhead{Sample} & 
+\colhead{\teff range} & 
+\colhead{Ro} & 
+\colhead{f} & 
+\colhead{$T_\mathrm{sys}$ (K)} 
+}
+\startdata
+\textit{Short-period pile-up} \\
+\hline
+\lamostmcq 10th \prot pctl. & 4500~K--5800~K & 0.367 $\pm$ 0.009 & 0.008 $\pm$ 0.006 &  44 $\pm$ 36 \\
+\hline 
+\textit{Long-period pile-up} \\
+\hline
+\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.37 $\pm$ 0.02 & 0.005 $\pm$ 0.004 & 59 $\pm$ 15 \\
+\lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.41 $\pm$ 0.09 & 0.01 $\pm$ 0.01 & 42 $\pm$ 37 \\
+CKS (CKS \teff) & 5800~K--6250~K & 1.75 $\pm$ 0.07 & 0.05 $\pm$ 0.01 & -58 $\pm$ 16 \\ 
+CKS (SPOCS \teff) & 5800~K--6250~K & 1.9 $\pm$ 0.1 & 0.09 $\pm$ 0.01 & -123 $\pm$ 20 \\
+CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.1 & 0.17 $\pm$ 0.02 & 26 $\pm$ 38 \\
+\citet{Hall2021} main-sequence & 5000~K--6250~K & 1.7 $\pm$ 0.2 & 0.30 $\pm$ 0.04 & -5 $\pm$ 64 \\
+\citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & -29 $\pm$ 79 \\
+All samples & 5000~K--6250~K & 1.47 $\pm$ 0.05 & 0.132 $\pm$ 0.008 & 2 $\pm$ 23 (LAMOST) \\
+ & & & & 10 $\pm$ 15 (CKS) \\
+ & & & & 61 $\pm$ 21  (Hall) \\
+All samples & 5800~K--6250~K & 1.67 $\pm$ 0.08 & 0.116 $\pm$ 0.009 & -42 $\pm$ 24 (LAMOST) \\
+&  &  &  & -44 $\pm$ 20 (CKS) \\
+&  &  &  & 3 $\pm$ 23 (Hall) \\ 
+\enddata
+%\tablecomments{}
+\end{deluxetable*}
+
 \subsection{Ages of stars on the ridge}
@@
 The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge.
-In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs.
+In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up.
 Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting.
 In both cases, there appears to be an age gradient such that hotter stars are younger on average.
-However, we note that such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars.
+However, such a trend is also expected in a sample of main-sequence stars as a natural consequence of the shorter main-sequence lifetimes of hotter, more massive stars.
 We also find that the dispersion in age is a sensitive function of \teff, with cooler stars on the ridge showing a broader range of ages.
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
-In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence.
+In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence \citep{vanSaders2019}.
 As a result, an age-gradient is expected along the pile-up as stars of different mass depart the pile-up at different ages.
@@
 
+{\color{blue} We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex.
+While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).
+Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pile-up stars.
+We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.}
+
 \begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{ages.pdf}
+    \includegraphics[width=\linewidth]{./figures/ages.pdf}
     \caption{Above, H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b).
@@
 
+\subsection{Where do the pile-ups end?} \label{subsec:extent}
+{\color{blue} It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff.
+While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
+We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
+In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window.
+We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, approximately halving in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
+The relative fraction of stars on the short-period edge declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
+Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/fraction.pdf}
+    \caption{The relative fraction of stars on the long- and short-period pile-ups in 10~K windows of \teff.}
+    \label{fig:fraction}
+\end{figure}
+
 \section{Discussion} \label{sec:discussion}
-\textbf{Some discussion of subgiants, some discussion of the other epoch of stalling noted in Curtis and how the two are not to be confused.}
+
+\subsection{Long-period behavior} \label{subsec:longperiod}
+{\color{blue}
+As discussed in \S\ref{subsec:models} and \S\ref{subsec:rossby}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
+The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021} and we suggest the two features are one and the same.
+Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
+
+Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
+Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
+By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
+{\color{blue} The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz.
+We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.}
+}
+
+\subsection{Short-period behavior} \label{subsec:shortperiod}
+{\color{blue}
+Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model.
+However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one.
+Works examining the rotation sequences of open clusters have found overlap between low-mass members of clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}.
+In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.
+
+\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused when the core transfers angular momentum to the envelope and offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
+Angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core-envelope coupling timescale.
+Theoretical predictions for this timescale are mass-dependent, and range from 30--110 Myr for a solar-mass star \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
+In the core-envelope coupling scenario, wind-driven spin-down stalls when the star's age becomes comparable to the core-envelope coupling timescale.
+Stars then resume spinning down via magnetic braking, and from open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
+
+Both the theoretically predicted core-envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
+Furthermore, the core-envelope coupling timescale is expected to increase towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
+
+\textbf{We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
+The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct.
+As mentioned above, the leading theory for the earlier stage of stalled braking is core-envelope coupling.
+Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core-envelope coupling.
+In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
+The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.}
+}
+
+\subsection{Implications for the period gap}
+{\color{blue}
+An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for stars of similar \teff.
+The effect was first noticed for M-dwarfs \citep{McQuillan2013b}, but was later shown to extend to $\sim$5000~K \citep{Reinhold2013, McQuillan2014, ReinholdHekker2020}, and even to $\sim$6500~K \citep{Davenport2017}.
+\citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}.
+However, \citet{Reinhold2019} found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability.
+Consequently, those authors proposed that the period bimodality may be the result of a transition between spot- and faculae-dominated photospheres.
+In this scenario, the period gap is due to bright faculae canceling out the effects of dark star spots.
+
+The short- and long-period pile-ups we examine here naturally produce a dearth of rotators at intermediate periods.
+This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \citet{McQuillan2014} sample.
+Moreover, as seen in Figure~\ref{fig:gap}, we recover the gradient in photometric variability across the gap pointed out by \citet{Reinhold2019}.
+The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum at the location of the gap.
+This supports the notion that the cause of the gap is due to changes in the stars themselves, rather than being the result of mixed stellar populations.
+However, the variability levels on both sides of the gap are not close to the detection limit, as evidenced by the fact that periods are securely detected for stars with similar properties at much lower \rper values.
+This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
+}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/gap.pdf}
+    \caption{The \teff--\prot distribution of the \lamostmcq (left) and \citet{McQuillan2014} samples color coded by the variability amplitude, \rper.
+Black contours show Gaussian kernel density estimation of the plotted distributions, and the dashed line shows a constant Rossby curve of Ro~=~0.5.
+    The long- and short-period pile-ups are separated by a relative dearth of stars with intermediate rotation periods.
+A strong gradient in \rper is apparent across this gap, such that variability amplitude reaches a local minimum near the gap's center.}
+    \label{fig:gap}
+\end{figure*}
+
+\subsection{Does the Sun reside on the long-period pileup?} \label{subsec:thesun}
+
+{\color{blue}
+It is unclear whether or not the Sun is a resident of the long-period pile-up.
+If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$~5--7 days above the long-period pile-up.
+This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
+However, a systematic offset in \teff could place the Sun on the long-period pile-up.
+Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
+Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required.
+To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K.
+However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+
+If the Sun is not in fact on the pile-up, one possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
+If this is the case, there is some moderate tension with the ages of pile-up stars in the CKS sample (Figure~\ref{fig:ages}).
+In the WMB model, hotter stars are expected to depart the pile-up on shorter timescales, yet some CKS stars in the pile-up appear to be both hotter and older than the Sun.
+It is of course possible that the isochrone ages are inaccurate, particularly for the cooler stars in the sample.
+If the Sun has indeed evolved through a period of weakened magnetic braking, this implies that the Sun's magnetic dynamo may be different from the majority of solar-type stars in the \citet{McQuillan2014} sample.
+
+It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
+As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}).
+However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
+Further work with a larger sample size may reveal if the pile-up does indeed extend to temperatures as cool as the Sun's.
+
+Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}).
+If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out.
+One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021}.
+Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question.
+}
 
@@
 
-    \item We find that the long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of $\mathrm{Ro_{crit}=1.467 \pm 0.007}$, in the \teff range of $\approx$~5000-6250~K.
-A pile-up of stars with a constant Rossby number and a variety of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
+    \item The long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
+The pile-up is also  populated by stars with a wide range of isochrone ages (\agerange).
+A pile-up of stars with a constant Rossby number and a broad range of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
+The precise value of \rocrit is highly sensitive to \teff scale shifts between observational data and the models used to compute \taucz.
 
-    \item The pile-up is populated by stars with a wide range of isochrone ages ($\sim$1--6~Gyr).
-Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile-up onto the ridge on a timescale of 1--3~Gyr, compatible with the predictions of \citet{vanSaders2019}.
-    While cooler stars appear to populate the pile-up until an age of $\sim$6~Gyr, the Sun, with an equatorial rotation period of $\approx$25~d and age of 4.567~Gyr, may have evolved away from the pile-up.
-\textbf{However, this statement depends on the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample.
-For example, stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}
+    \item Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$3~Gyr, compatible with the predictions of \citet{vanSaders2019}.
+Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pile-up until an age of $\sim$6~Gyr.
 
-    %Ruth: include some discussion of what this says about the Sun.
-Not just because it implies the Sun's dynamo might be different from McQuillan stars because it's been through WMB, but also how did the Sun get there?
-Its Rossby number is higher than the stars on the ridge...
-Can expansion alone explain the ~6 days of rotation period lengthening it has gone through since it passed through the ridge around 2.5 Gyr ago?
+    \item {\color{blue} It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it.
+Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up.
+If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.}
 
-    \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity, which we recover through kernel density estimation, is not as prominent as the long-period pile-up.
-\textbf{The pile-up on the short-period edge can also be fit with a constant Rossby model, though over a somewhat different range of \teff.
-}
+    %\textbf{Change this:} {\color{blue}However, this statement depends on (i) the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as (ii) any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample.
+Stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}
 
-    \item \textbf{Some statement about where the pile-up ends. At what Teff does density drop off by factor of X.}
+    \item {\color{blue} We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
+This overdensity is less prominent than the long-period pile-up, potentially due to the age distribution of the Kepler field or because this pile-up is shorter-lived relative to the long-period pile-up.
+The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core-envelope coupling, as proposed by \citet{Curtis2020}.
+The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.}
 
-    \item We observe a change in slope to the long-period edge of F-type stars at $\teff~=~6250~\pm~32$~K, coinciding with the Kraft break.
+    \item {\color{blue} The number density of stars on the long-period pile-up declines with \teff, though it remains unclear whether this is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
+The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.}
+
+    %...
+appears to vanish at \teff below \textbf{5800~K}, while the short-period pile-up vanishes below \textbf{NNNN~K}.
+{\color{blue}It is yet unclear whether the termination of the pile-ups is entirely astrophysical, an effect of the size and selection function of the samples considered here, or some combination of the aforementioned effects.}
+
+    %\item We observe a change in slope to the long-period edge of F-type stars at $\teff~\approx~6250$~K, coinciding with the Kraft break.
 \textbf{The statistical significance of the change in slope across this break is $\sim 3\sigma$.
-Note...this may just be due to a subgiant branch crossing through the main sequence Prot distribution}
+TJD: Remove this item from conclusions?
+Even if this is expected, is it worth noting?}
 
-    \item The existence of this pile-up, or ridge, limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the Rossby ridge.
-For example, a Sun-like star may spend several Gyr evolving across the ridge.
-
-    \item We find tentative evidence for an age-gradient along the ridge, such that hotter stars on the ridge are younger on average.
+    \item We find tentative evidence for an age-gradient along the long-period pile-up, such that hotter stars on the ridge are younger on average.
 Relatedly, the age dispersion along the ridge is non-uniform as a function of temperature, with hotter stars showing a smaller dispersion.
 This observation suggests that hotter stars reside on the ridge for a shorter period of time relative to cooler stars.
-However, a more careful analysis is required to conclusively show this effect is not due to the intrinsic age gradient expected among a sample of main sequence stars with different masses.
+However, a more careful analysis is required to conclusively show these observations are not due to the intrinsic age gradient expected among a sample of main sequence stars with different masses and the higher isochrone age uncertainties associated with cooler stars.
 
-    \item Authors using the method of gyro-tagging, which uses rotation periods to identify coeval stellar associations, should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.
+    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the ridge.
+For example, a Sun-like star may spend several Gyr evolving across the ridge.
+Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.
 
@@
 
-\newpage 
 \appendix
-\section{Additional figures}
-
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{rossby-rvar.pdf}
-%     \caption{The relationship between Rossby number and \rvar for the LAMOST--Kepler sample. \rvar values are sourced from \citet{Lu2021}.}
-%     \label{fig:rossby-rvar}
-% \end{figure*}
-
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{hrd.pdf}
-%     \caption{H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS and LAMOST--Kepler samples.
-\textbf{There seems to be a main cluster of orange points in the right panel, approximately tracing an isochrone or range of isochrones, and then a gap between that cluster and the orange points at upper right of the black box.
-Are these stars that have peeled off the MS, or are there multiple bursts of star formation?}}
-%     \label{fig:hrd}
-% \end{figure*}
-
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{abundance.pdf}
-%     \caption{Distribution of isochrone ages, [Y/Al], and [Y/Mg] ratios for CKS stars on the ridge and in an equivalent range of Teff/logg (control).
-The ridge stars are younger and tend to have moderately higher [Y/Al] and [Y/Mg], also indicating they are younger than average.}
-%     \label{fig:my_label}
-% \end{figure*}
+\section{Comparison of temperature-period distributions}\label{app:teffprot}
+In Figure~\ref{fig:comparison}, we show how the \teff--\prot distribution of the CKS sample changes when sourcing \teff and \prot from different, homogeneous catalogs in the literature.
+The sharpness of the long-period pile-up appears to be determined primarily by the source of \teff, rather than \prot.
+The CKS-Gaia catalog \citep{Fulton2018} appears to offer the highest internal precision.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{scatter-color.pdf}
-    \caption{LAMOST--Kepler sample in \teff--\prot plane color coded by \logg (left) and [Fe/H] (right).
-\textbf{This plot shows overlap between the ridge and lower gravity subgiants evolving through it (we think).}}
-    \label{fig:my_label}
+    \includegraphics[width=\linewidth]{./figures/comparison.pdf}
+    \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
+    \label{fig:comparison}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison.pdf}
-    \caption{Comparison of the \teff--\prot distribution for the CKS sample using rotation periods and \teff from the sources indicated by the axes labels.}
-    \label{fig:tp-comparison}
+    \includegraphics[width=\linewidth]{./figures/ridge.pdf}
+    \caption{The \teff-\prot plane for the CKS sample.
+Point colors are scaled to the CKS ages determined from isochrone fitting.
+The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements.
+The black trapezoid indicates the approximate area of the ridge.
+The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
+    \label{fig:ridge}
 \end{figure*}
 
+\section{Comparison of spectroscopic temperature scales}\label{app:teff}
+
+In Figure~\ref{fig:teffscales} we compare temperatures between the LAMOST--Kepler catalog and temperatures from other surveys.
+We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
+A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
+    \label{fig:teffscales}
+\end{figure}
+
 \begin{acknowledgments}
-We thank Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog and Rodrigo Luger for assistance with the \texttt{showyourwork} package.
-It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute and the Stars Group at the American Museum of Natural History for helpful discussions.
+We thank Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, and Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog.
+It is a pleasure to thank the Stars Group at the American Museum of Natural History, and the Astronomical Data Group at the Flatiron Institute for helpful discussions.
+
 This work made use of the gaia-kepler.fun crossmatch database created by Megan Bedell.
@@
 
-\bibliography{main}{}
+\facilities{Kepler; Gaia; LAMOST; Keck:I (HIRES); Sloan (APOGEE)}
+
+\software{\texttt{astropy} \citep{astropy13, astropy18},    
+          \texttt{emcee} \citep{emcee2013, emcee2019},
+          \texttt{jupyter} \citep{jupyter},
+          \texttt{matplotlib} \citep{matplotlib},
+          \texttt{numpy} \citep{numpy},
+          \texttt{pandas} \citep{pandas-soft, pandas-proc},
+          \texttt{scipy} \citep{scipy},
+          \texttt{seaborn} \citep{seaborn},
+          \texttt{showyourwork!} \citep{Luger2021}
+          }
+
+\bibliography{bib}{}
 \bibliographystyle{aasjournal}

### 2021-12-08 src/ms.tex
Updating tex file

@@
 \received{\today}
-%\revised{\today}
+\revised{\today}
 %\accepted{\today}
@@
 \graphicspath{{./}{figures/}}
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
 
 \begin{document}
-
-% \title{The Rossby Ridge: a Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages}
-%\title{A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking}
-%\title{Two Pile-ups in the Temperature--Period Distribution of Main-Sequence Stars:\\Possible Evidence for Weakened Magnetic Braking and Core-Envelope Coupling}
 \title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}
@@
 Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.
-{\color{blue}The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
-Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.}
+The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
 \end{abstract}
@@
 
-%Nevertheless, it is by now clear that the simple $t^{1/2}$ scaling of rotation rates with age breaks down in certain regimes of stellar mass and age.
-Efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}.
-More recently, low-mass stars were shown to experience a temporary epoch of stalled spin-down which is not accounted for by standard gyrochronology models \citep{Curtis2019a}.
-\citet{Spada2020} and \citet{Curtis2020} proposed that this epoch of apparent stalling may arise from rotational coupling of the core and envelope, in which redistribution of angular momentum from the interior to the surface temporarily offsets the effects of magnetic braking.
-Regardless of the true mechanism, rotation periods in this part of parameter space have diminished utility in predicting stellar ages.
-
 For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}.
@@
 We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, discuss some implications of these results in \S\ref{sec:discussion}, and present our conclusions in \S\ref{sec:conclusions}.
-
-%We argue that several aspects of the long-period pile-up feature are consistent with expectations from the WMB model.
-The short-period pile-up, by comparison, may be due a temporary epoch of stalled wind-driven losses due to core-envelope coupling.
 
@@
 Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}).
-{\color{blue} A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.}
+A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
+As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
 
@@
 A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
-{\color{blue}This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.}
+This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
 
@@
 While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
-{\color{blue} We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.}
+We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.
 
-{\color{blue} Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
+Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
 In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
@@
 Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
-A higher normalization factor would cause the low-mass stars to spin down more at fixed age.}
+A higher normalization factor would cause the low-mass stars to spin down more at fixed age.
 
-{\color{blue} Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models.
+Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models.
 At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
@@
 When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
-This hypothesis is discussed further in \S\ref{subsec:shortperiod}.}
+This hypothesis is discussed further in \S\ref{subsec:shortperiod}.
 
@@
 
-% \subsection{Piecewise polynomial model}
-% We fit a piecewise linear relation to the CKS stars on the ridge with $5800~K<$~\teff~$<6400~K$.
-We performed an initial Levenberg-Marquardt least-squares fit with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class.
-We then performed Markov chain Monte Carlo sampling (MCMC) with the \texttt{emcee} package \citep{emcee2013, emcee2019}.
-
 \subsection{Constant Rossby model}
@@
 Using a small sample of Kepler targets with rotation periods determined from brightness modulations, \citet{vanSaders2016} proposed this threshold happens at a critical Rossby number of $\mathrm{Ro_{crit}} \sim \mathrm{Ro_\odot}$.
-
-%{\color{blue} From a sample of Kepler targets with asteroseismic rotation rates, \citet{Hall2021} found a good match to observations when adopting a critical Rossby number of $\mathrm{Ro_{crit}} = 1.97$.}
 
@@
 
-%We computed constant Rossby models across the range of \teff bins using the equation for the convective turnover timescale presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}.
-This equation is valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K.
-%\textbf{Note: we can also try Landin 2010 and Barnes Kim 2010 formulas for convective turnover timescales.}
-
 We performed an initial Levenberg-Marquardt non-linear least-squares fit of a constant Rossby model to the long-period edge with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class to optimize the following likelihood:
@@
 
-{\color{blue}  We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set.
+We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set.
 For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d.
@@
 For the CKS sample, we additionally required $\logg~>~4$ and 5850~K~$<~\teff~<$~6500~K.
-In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.}
+In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.
 
@@
 
-{\color{blue} We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation.
+We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation.
 Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 may be compatible with all of the data sets if modest \teff systematics are present.
-Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff})}.
+Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
 
@@
 
-{\color{blue} We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex.
+We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex.
 While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).
 Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pile-up stars.
-We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.}
+We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.
 
@@
 \subsection{Where do the pile-ups end?} \label{subsec:extent}
-{\color{blue} It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff.
 While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
@@
 In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window.
-We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, approximately halving in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
+We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
 The relative fraction of stars on the short-period edge declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
-Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.}
+Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.
 
@@
 \subsection{Long-period behavior} \label{subsec:longperiod}
-{\color{blue}
+
 As discussed in \S\ref{subsec:models} and \S\ref{subsec:rossby}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
@@
 By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
-{\color{blue} The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz.
-We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.}
-}
+The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz.
+We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.
 
 \subsection{Short-period behavior} \label{subsec:shortperiod}
-{\color{blue}
+
 Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model.
@@
 
-\textbf{We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
+We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
 The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct.
@@
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
-The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.}
-}
+The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
 
-\subsection{Implications for the period gap}
-{\color{blue}
+\subsection{Implications for the period gap} \label{subsec:gap}
+
 An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for stars of similar \teff.
@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
-}
 
@@
 
-{\color{blue}
 It is unclear whether or not the Sun is a resident of the long-period pile-up.
@@
 Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question.
-}
 
@@
 
-    \item {\color{blue} It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it.
+    \item It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it.
 Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up.
-If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.}
+If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.
 
-    %\textbf{Change this:} {\color{blue}However, this statement depends on (i) the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as (ii) any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample.
-Stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}
-
-    \item {\color{blue} We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
+    \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
 This overdensity is less prominent than the long-period pile-up, potentially due to the age distribution of the Kepler field or because this pile-up is shorter-lived relative to the long-period pile-up.
 The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core-envelope coupling, as proposed by \citet{Curtis2020}.
-The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.}
+The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.
 
-    \item {\color{blue} The number density of stars on the long-period pile-up declines with \teff, though it remains unclear whether this is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
-The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.}
-
-    %...
-appears to vanish at \teff below \textbf{5800~K}, while the short-period pile-up vanishes below \textbf{NNNN~K}.
-{\color{blue}It is yet unclear whether the termination of the pile-ups is entirely astrophysical, an effect of the size and selection function of the samples considered here, or some combination of the aforementioned effects.}
-
-    %\item We observe a change in slope to the long-period edge of F-type stars at $\teff~\approx~6250$~K, coinciding with the Kraft break.
-\textbf{The statistical significance of the change in slope across this break is $\sim 3\sigma$.
-TJD: Remove this item from conclusions?
-Even if this is expected, is it worth noting?}
+    \item The number density of stars on the long-period pile-up declines with \teff, though it remains unclear whether this is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
+The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.
 
@@
 
-The code and data tables required to reproduce the figures and analysis presented here are publicly available on GitHub.\footnote{\url{https://github.com/trevordavid/rossby-ridge}}
+The code and data tables required to reproduce the figures and analysis presented here are publicly available through GitHub.\footnote{\url{https://github.com/trevordavid/rossby-ridge}}
 
@@
 
-\facilities{Kepler; Gaia; LAMOST; Keck:I (HIRES); Sloan (APOGEE)}
+\facilities{Gaia; Kepler; Keck:I (HIRES); LAMOST; Sloan (APOGEE)}
 
 \software{\texttt{astropy} \citep{astropy13, astropy18}, 
+          \texttt{corner} \citep{corner}, 
           \texttt{emcee} \citep{emcee2013, emcee2019},

### 2021-12-09 src/figures/ages.py
figure adjustments

@@ -162,7 +162,7 @@ ax.plot(lam["Teff_lam"][lam_ridge], lam["Prot"][lam_ridge], 'o', mfc="None", col
 # In[9]:
 
 
-sns.set(font_scale=1.2, context="paper", style="ticks")
+sns.set(font_scale=1.1, context="paper", style="ticks")
 
 mpl.rcParams["legend.markerscale"] = 2
 

### 2021-12-09 src/figures/fraction.py

@@ -120,7 +120,7 @@ lam_ridge = (frac_dist<0.05) & (lam["Teff_lam"]>5500) & (lam["Teff_lam"]<6500) &
 
 
 
-sns.set(font_scale=1.4, context="paper", style="ticks", palette="Blues")
+sns.set(font_scale=1.5, context="paper", style="ticks", palette="Blues")
 
 xeps = 5
 yeps = 1

### 2021-12-09 src/figures/percentiles.py

@@ -139,7 +139,7 @@ period_90th_pctl, e_period_90th_pctl = percentile_bootstrap(pctl=90.)
 period_10th_pctl, e_period_10th_pctl = percentile_bootstrap(pctl=10.)    
 
 
-sns.set(font_scale=1.4, context="paper", style="ticks", palette="Blues")
+sns.set(font_scale=1.5, context="paper", style="ticks", palette="Blues")
 
 teff_bin_centers = np.arange(4000,7020,20)
 roc_period_90th_pctl = np.zeros(len(teff_bin_centers))

### 2021-12-14 src/ms.tex
figure adjustments
paper edits
addressing co-author comments

@@
 Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.
-The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
 Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
@@
 
-\subsection{California--Kepler Survey} \label{sec:cks}
+\subsection{California--Kepler Survey} \label{subsec:cks}
 The CKS project gathered high-resolution spectroscopy for 1305 Kepler planet host stars \citep{Petigura2017}.
@@
 We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} rotation period catalog, which published rotation periods for $>$34000 Kepler targets.
-The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015}.
+The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
 
@@
 There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
+Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.
 
@@
 To place the Sun in the context of the long-period pile-up, we use up-to-date estimates of the Sun's effective temperature, rotation period, and age.
-Following the IAU 2015 Resolution B3 \citep{Mamajek2015}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
+Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
 For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25$~d with an upper error that encompasses the period of $\approx$~36~d measured near the poles from differential rotation studies \citep[][and references therein]{Thompson2003}.
@@
 The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
 As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
@@
 Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
+Additionally, a case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.
 By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
@@
 \citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused when the core transfers angular momentum to the envelope and offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
-Angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core-envelope coupling timescale.
+Angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core–envelope coupling timescale.
 Theoretical predictions for this timescale are mass-dependent, and range from 30--110 Myr for a solar-mass star \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
-In the core-envelope coupling scenario, wind-driven spin-down stalls when the star's age becomes comparable to the core-envelope coupling timescale.
+In the core–envelope coupling scenario, wind-driven spin-down stalls when the star's age becomes comparable to the core–envelope coupling timescale.
 Stars then resume spinning down via magnetic braking, and from open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
+This timescale is consistent with independent studies of open clusters, which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$300~Myr \citep{Fritzewski2020,Fritzewski2021}.
 
-Both the theoretically predicted core-envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
-Furthermore, the core-envelope coupling timescale is expected to increase towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
+Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
+Furthermore, the core–envelope coupling timescale is expected to increase towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
 
@@
 The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct.
-As mentioned above, the leading theory for the earlier stage of stalled braking is core-envelope coupling.
-Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core-envelope coupling.
+As mentioned above, the leading theory for the earlier stage of stalled braking is core–envelope coupling.
+Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling.
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
@@
 
+%The Sun's relationship to the long-period pile-up is also relevant to ongoing research into the question of whether the Sun is in the middle of a magnetic transition, as \citet{Metcalfe2016} suggested based on the Sun's unusual magnetic cycle period,
+
 \section{Conclusions} \label{sec:conclusions}
@@
 This overdensity is less prominent than the long-period pile-up, potentially due to the age distribution of the Kepler field or because this pile-up is shorter-lived relative to the long-period pile-up.
-The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core-envelope coupling, as proposed by \citet{Curtis2020}.
+The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.
 The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.
@@
 This paper includes data collected by the Kepler mission and obtained from the MAST data archive at the Space Telescope Science Institute (STScI).
+
 Funding for the Kepler mission is provided by the NASA Science Mission Directorate.
 STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555.
+
 Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences.
@@
 LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences.
+
 This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement.
+
 Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P.
@@
 SDSS-IV is managed by the Astrophysical Research Consortium for the Participating Institutions of the SDSS Collaboration including the Brazilian Participation Group, the Carnegie Institution for Science, Carnegie Mellon University, Center for Astrophysics | Harvard \& Smithsonian, the Chilean Participation Group, the French Participation Group, Instituto de Astrof\'isica de Canarias, The Johns Hopkins University, Kavli Institute for the Physics and Mathematics of the Universe (IPMU) / University of Tokyo, the Korean Participation Group, Lawrence Berkeley National Laboratory, Leibniz Institut f\"ur Astrophysik Potsdam (AIP),  Max-Planck-Institut f\"ur Astronomie (MPIA Heidelberg), Max-Planck-Institut f\"ur Astrophysik (MPA Garching), Max-Planck-Institut f\"ur Extraterrestrische Physik (MPE), National Astronomical Observatories of China, New Mexico State University, New York University, University of Notre Dame, Observat\'ario Nacional / MCTI, The Ohio State University, Pennsylvania State University, Shanghai Astronomical Observatory, United Kingdom Participation Group, Universidad Nacional Aut\'onoma de M\'exico, University of Arizona, University of Colorado Boulder, University of Oxford, University of Portsmouth, University of Utah, University of Virginia, University of Washington, University of Wisconsin, Vanderbilt University, and Yale University.
+
 This research has made use of NASA's Astrophysics Data System Bibliographic Services.

### 2021-12-14 src/ms.tex
addressing coauthor comments

@@
 Interestingly, a similar clustering of points is not observed in the LAMOST--Kepler sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
+Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pile-up.
+We conclude that the temperatures for these stars in the CKS catalog are too high by $\gtrsim$~100~K.
 
@@
 Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
-Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
+A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
+Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
@@
 
-\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused when the core transfers angular momentum to the envelope and offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
+\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused either by (a) core–envelope momentum transfer which offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}, or (b) a reduction in the magnetic braking torque.
 Angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core–envelope coupling timescale.
@@
 The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct.
-As mentioned above, the leading theory for the earlier stage of stalled braking is core–envelope coupling.
+As mentioned above, one theory for the earlier stage of stalled braking is core–envelope coupling.
 Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling.
@@
 \citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}.
-However, \citet{Reinhold2019} found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability.
+However, \citet{Curtis2020} demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation.
+Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, further invalidating the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories.
+An alternative explanation for the gap was proposed by \citet{Reinhold2019}, who found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability.
 Consequently, those authors proposed that the period bimodality may be the result of a transition between spot- and faculae-dominated photospheres.
@@
 Moreover, as seen in Figure~\ref{fig:gap}, we recover the gradient in photometric variability across the gap pointed out by \citet{Reinhold2019}.
-The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum at the location of the gap.
+The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum near the location of the gap.
 This supports the notion that the cause of the gap is due to changes in the stars themselves, rather than being the result of mixed stellar populations.

### 2021-12-14 src/figures/asteroseismic.py

@@ -80,8 +80,8 @@ sc_kws = {"marker":".", "color":"C0", "rasterized":True, "alpha":1}
 fig, ax = plt.subplots(nrows=2,ncols=1,figsize=(4.5,6.5))
 
 ax[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
-#ax[1].scatter(lam['Teff_lam'][ik], lam['Prot'][ik], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
-ax[1].scatter(lam['Teff_lam'], lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
+ax[1].scatter(lam['Teff_lam']+41, lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
+#Note: the 41 K offset is derived in the appendix
 
 for i in range(2):
     arg = hall["Type"] == "MS"

### 2021-12-15 src/ms.tex
addressing co-author comments

@@
 
-Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models.
+Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models.
 At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
@@
 Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model.
-However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one.
-Works examining the rotation sequences of open clusters have found overlap between low-mass members of clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}.
+However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one as cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
+Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages, notably Praesepe (0.67~Gyr), NGC 752 (1~Gyr), and NGC 6811 (1.4~Gyr) \citep{Agueros2018, Curtis2019a, Curtis2020}.
 In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.
+The short-period pile-up we observe may be the manifestation of the same type of slowed spin evolution, but for stars of higher masses and younger ages (since the pile-up more closely aligns with the Pleiades sequence) than the cluster members in the above-mentioned works.
 
-\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused either by (a) core–envelope momentum transfer which offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}, or (b) a reduction in the magnetic braking torque.
-Angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core–envelope coupling timescale.
-Theoretical predictions for this timescale are mass-dependent, and range from 30--110 Myr for a solar-mass star \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
+\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused either by (a)a reduction in the magnetic braking torque, or (b) core–envelope momentum transfer which offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
+In the core–envelope momentum transfer scenario angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core–envelope coupling timescale.
+Theoretical predictions for the core–envelope coupling are mass-dependent, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
 In the core–envelope coupling scenario, wind-driven spin-down stalls when the star's age becomes comparable to the core–envelope coupling timescale.
@@
 We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct.
+The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct, though it is interesting that both the long- and short-period pileups seem to be well-described by curves of constant Rossby number.
 As mentioned above, one theory for the earlier stage of stalled braking is core–envelope coupling.
@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
+The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling.
+The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
 
@@
 It is of course possible that the isochrone ages are inaccurate, particularly for the cooler stars in the sample.
-If the Sun has indeed evolved through a period of weakened magnetic braking, this implies that the Sun's magnetic dynamo may be different from the majority of solar-type stars in the \citet{McQuillan2014} sample.
+If the Sun has indeed evolved through a period of WMB, this implies that the Sun may have a fundamentally different dynamo mechanism to the stars that have not yet transitioned through WMB.
+This suggests that the Sun's dynamo and consequent manifestations of magnetic activity may not be directly comparable or analogous to the majority of stars with measured Kepler rotation periods (as the majority have not yet transitioned through WMB).
+This may have implications for the validity of using the Sun as a benchmark calibrator and exemplar in stellar rotation and magnetism models.
 
@@
 Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up.
-If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.
+If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up; some of the cooler long-period pile-up stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pile-up.
+However, this tension might be simply explained by inaccurate isochrone ages.
 
@@
 Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.
+
+    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence.
+(I think maybe Pisces Eridani is an example of this?
+Or perhaps Jason has a better example?).
+However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods.
+This is clearly demonstrated by the CKS sample in the right-hand panel of figure 1 -- taken out of context, these data look like a group of coeval stars with a slow-rotator sequence.
+The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
+When over-densities or ridges appear present in the rotation period distributions of stellar populations, care must be taken to ensure that these are not caused by WMB or core-envelope coupling before assuming that population is coeval and using the any over-densities to age-date it.
 

### 2021-12-15 src/ms.tex
addressing co-author comments

@@
 Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model.
-However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one as cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
+However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
 Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages, notably Praesepe (0.67~Gyr), NGC 752 (1~Gyr), and NGC 6811 (1.4~Gyr) \citep{Agueros2018, Curtis2019a, Curtis2020}.
 In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.
-The short-period pile-up we observe may be the manifestation of the same type of slowed spin evolution, but for stars of higher masses and younger ages (since the pile-up more closely aligns with the Pleiades sequence) than the cluster members in the above-mentioned works.
+The short-period pile-up we observe may be the manifestation of the same type of slowed spin evolution, but for stars of higher masses and younger ages than the cluster members in the above-mentioned works (since the pile-up is observed at shorter periods relative to the Praesepe sequence).
 
-\citet{Curtis2019a, Curtis2020} proposed that such an effect could be produced by a temporary epoch of stalled spin-down, caused either by (a)a reduction in the magnetic braking torque, or (b) core–envelope momentum transfer which offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
+\citet{Curtis2019a, Curtis2020} proposed that the overlapping cluster sequences could be produced by a temporary epoch of stalled spin-down, caused either by (i) a reduction in the magnetic braking torque, or (ii) core–envelope momentum transfer which offsets the effect of magnetic braking \citep[e.g.][]{MacGregor1991}.
 In the core–envelope momentum transfer scenario angular momentum is exchanged between the envelope and the core on a characteristic timescale known as the core–envelope coupling timescale.
+The angular momentum transfer spins up the envelope, temporarily offsetting the spin-down via magnetic braking.
+Thus, in the core–envelope coupling scenario, spin evolution is slowed when the star's age becomes comparable to the core–envelope coupling timescale.
 Theoretical predictions for the core–envelope coupling are mass-dependent, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
-In the core–envelope coupling scenario, wind-driven spin-down stalls when the star's age becomes comparable to the core–envelope coupling timescale.
-Stars then resume spinning down via magnetic braking, and from open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
-This timescale is consistent with independent studies of open clusters, which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$300~Myr \citep{Fritzewski2020,Fritzewski2021}.
+
+After a period of slowed spin-evolution, stars must resume spinning down via magnetic braking as evidenced from studies of older open clusters.
+From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
+Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience little spin evolution between $\sim$100~Myr and $\sim$200~Myr.
+This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.
 
@@
 For example, a Sun-like star may spend several Gyr evolving across the ridge.
-Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.
+\textbf{Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.}
 
     \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence.
-(I think maybe Pisces Eridani is an example of this?
-Or perhaps Jason has a better example?).
 However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods.
-This is clearly demonstrated by the CKS sample in the right-hand panel of figure 1 -- taken out of context, these data look like a group of coeval stars with a slow-rotator sequence.
+This is clearly demonstrated by the CKS sample in the right-hand panel of Figure~\ref{fig:surveys}.
+Taken out of context, this sample resembles a group of coeval stars with a slow-rotator sequence.
 The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
-When over-densities or ridges appear present in the rotation period distributions of stellar populations, care must be taken to ensure that these are not caused by WMB or core-envelope coupling before assuming that population is coeval and using the any over-densities to age-date it.
+When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
 

### 2021-12-15 src/ms.tex
addressing co-author comments

@@
 More recently, \citet{Hall2021} used the asteroseismic rotation rates of Kepler dwarfs, with a different selection bias from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
+\citet{Masuda2021} also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
 While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which varies with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
@@
     \includegraphics[width=\linewidth]{./figures/ages.pdf}
-    \caption{Above, H-R diagram placement of Rossby ridge stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b).
-Below, the \teff-age plane for CKS stars along the ridge using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
+    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b).
+Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
     \label{fig:ages}

### 2021-12-16 src/ms.tex
revising ages figure and discussion

@@
 
-\subsection{Ages of stars on the ridge}
+\subsection{Ages of stars on the long-period pile-up}
 \label{subsec:ages}
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
+However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
+In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun, and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
+
 In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence \citep{vanSaders2019}.

### 2021-12-16 src/ms.tex
revising ages discussion
co-author comments

@@
 In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun, and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
+The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and the concentration of stars at the upper age boundary at a given \teff (Figure~\ref{fig:ages}).
 
@@
 By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
-The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz.
+Furthermore, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$, which invalidates the claim of \citet{LorenzoOliveira2019}.
+We note that the \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
 We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.

### 2021-12-17 src/ms.tex
jvs comments

@@
 For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}.
-This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with masses $\gtrsim$~1\msun experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
+This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
+
 Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep{vanSaders2019}.
@@
     \includegraphics[width=\textwidth]{./figures/surveys.pdf}
-    \caption{The \teff-\prot plane using rotation periods from \citet{McQuillan2014} or, in the case of the CKS sample, \citet{McQuillan2013}, with \teff originating from the source denoted at top.
+    \caption{The \teff-\prot plane using rotation periods from \citet{McQuillan2014} or, in the case of the CKS sample, \citet{McQuillan2013} which applied an identical analysis to Kepler Objects of Interest (KOIs), with \teff originating from the source denoted at top.
 The \citet{McQuillan2014} \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars.
-The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the differentially rotating surface.}
+The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the differentially rotating surface.
+Many of the stars above the long-period pile-up are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \citet{vanSaders2019}.}
     \label{fig:surveys}
@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
+
 The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling.
 The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
+Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling.
+In this picture, spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability.
+After some time, spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
+\textbf{An increase in photometric variability accompanies this period of...}
 
@@
 
-    \item The number density of stars on the long-period pile-up declines with \teff, though it remains unclear whether this is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
+    \item The number density of stars on the long-period pile-up declines with \teff, in line with predictions from the WMB hypothesis, though it remains unclear whether this observation is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
 The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.

### 2021-12-17 src/ms.tex
jvs comments

@@
 \citet{Masuda2021} also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
-While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which varies with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
+While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which may vary with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
 

### 2021-12-17 src/ms.tex
jvs comments

@@
 
-The location of the pile-up approximately corresponds with an empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+The location of the short-period pile-up is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
@@
 the low-mass stars would be rotating more slowly than the model predictions).
-Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
+Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law.
+The models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
 A higher normalization factor would cause the low-mass stars to spin down more at fixed age.
@@
 The core and envelope at these times are thus assumed to be decoupled.
-However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015}.
+However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}.
 When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.

### 2021-12-17 src/ms.tex
jvs comments

@@
 
-We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set.
+We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
 For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d.
@@
 Furthermore, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$, which invalidates the claim of \citet{LorenzoOliveira2019}.
+Notably, the \citet{LorenzoOliveira2019} work did not include changes in the stellar moment of inertia in the spin-down calculations.
+Neglecting spin-down due to expansion will lead to more rapid rotation rates that allow standard spin-down models to match observations, leading to the erroneous conclusion that WMB is not required.
+
 We note that the \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
@@
 
-Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model.
+Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model or, more generally, any standard, solid-body braking model.
 However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.

### 2021-12-17 src/ms.tex
jvs comments

@@
 Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model or, more generally, any standard, solid-body braking model.
-However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
+However, this feature may also be due to an epoch of apparent stalling, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
 Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages, notably Praesepe (0.67~Gyr), NGC 752 (1~Gyr), and NGC 6811 (1.4~Gyr) \citep{Agueros2018, Curtis2019a, Curtis2020}.

### 2021-12-17 src/ms.tex
revised percentiles plot

@@
 Thus, in the core–envelope coupling scenario, spin evolution is slowed when the star's age becomes comparable to the core–envelope coupling timescale.
-Theoretical predictions for the core–envelope coupling are mass-dependent, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Spada2020}.
+Theoretical predictions for the core–envelope coupling are mass-dependent, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
 
-After a period of slowed spin-evolution, stars must resume spinning down via magnetic braking as evidenced from studies of older open clusters.
+After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
 From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
@@
 Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
-Furthermore, the core–envelope coupling timescale is expected to increase towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
+Furthermore, in order to match observations of rotation periods in young clusters, models require that the core–envelope coupling timescale increases towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
 
@@
     \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity is less prominent than the long-period pile-up, potentially due to the age distribution of the Kepler field or because this pile-up is shorter-lived relative to the long-period pile-up.
+This overdensity is less prominent than the long-period overdensity likely because the short-period pile-up is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pile-up ($>10^9$~yr).
 The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.
@@
 Relatedly, the age dispersion along the ridge is non-uniform as a function of temperature, with hotter stars showing a smaller dispersion.
-This observation suggests that hotter stars reside on the ridge for a shorter period of time relative to cooler stars.
+This observation suggests that hotter stars reside on the long-period pile-up for a shorter period of time relative to cooler stars.
+These observations are in accordance with predictions from the WMB model, which predicts that stars of different masses spend an approximately equal fraction of their respective main-sequence lifetimes on the long-period pile-up.
 However, a more careful analysis is required to conclusively show these observations are not due to the intrinsic age gradient expected among a sample of main sequence stars with different masses and the higher isochrone age uncertainties associated with cooler stars.

### 2021-12-17 src/ms.tex
jvs comments

@@
 However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}.
-When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
+When this happens, angular momentum can be transferred from the core to the envelope at a rate comparable to the rate at which angular momentum is lost via magnetized winds.
+Consequently, so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
 This hypothesis is discussed further in \S\ref{subsec:shortperiod}.
@@
     \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.
-The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.}
+The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.
+Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.}
     \label{fig:models}

### 2021-12-20 src/ms.tex
jvs comments

@@
 For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}.
-This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
+This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Rossby numbers of Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
 
@@
 The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct, though it is interesting that both the long- and short-period pileups seem to be well-described by curves of constant Rossby number.
-As mentioned above, one theory for the earlier stage of stalled braking is core–envelope coupling.
-Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling.
+As mentioned above, one theory for the earlier stage of stalled spin-down is core–envelope coupling.
+Crucially, in the stalled spin-down phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling.
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
@@
 After some time, spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
-\textbf{An increase in photometric variability accompanies this period of...}
+\textbf{An increase in photometric variability accompanies this period of post-stalling spin-down, }
 

### 2021-12-20 src/ms.tex
paper edits

@@
 
-We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
+We emphasize that the temporary epoch of slowed spin-down (also referred to as stalled magnetic braking) proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}.
 The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct, though it is interesting that both the long- and short-period pileups seem to be well-described by curves of constant Rossby number.

### 2021-12-20 src/ms.tex
paper edits

@@
 
-    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the ridge.
+    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up.
 For example, a Sun-like star may spend several Gyr evolving across the ridge.
-\textbf{Authors using rotation periods as means of identifying new members to coeval stellar associations should be cognizant of this pile-up for stars with \teff~$\gtrsim5800~K$ as it can mimic the the appearance of a coeval stellar association in the \teff--\prot plane.}
+Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
 
-    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence.
+    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane.
 However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods.

### 2021-12-20 src/ms.tex
paper edits

@@
 However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
-In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun, and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
 The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and the concentration of stars at the upper age boundary at a given \teff (Figure~\ref{fig:ages}).
+In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
 

### 2021-12-20 src/figures/percentiles.py

@@ -162,7 +162,7 @@ period_90th_pctl, e_period_90th_pctl = percentile_bootstrap(pctl=90.)
 period_10th_pctl, e_period_10th_pctl = percentile_bootstrap(pctl=10.)    
 
 
-sns.set(font_scale=1.6, context="paper", style="ticks", palette="Blues")
+sns.set(font_scale=1.7, context="paper", style="ticks", palette="Blues")
 
 teff_bin_centers = np.arange(4000,7020,20)
 roc_period_90th_pctl = np.zeros(len(teff_bin_centers))
@@ -191,7 +191,7 @@ plt.scatter(teff_bin_centers, period_10th_pctl, color='k', s=2)
 plt.xlim(7000,4500)
 plt.xlabel("Effective temperature [K]")
 plt.ylabel("Rotation period [d]")
-plt.legend()
+plt.legend(prop={'size':13})
 sns.despine()
 plt.savefig('../figures/percentiles.pdf')
 plt.show()

### 2021-12-21 src/ms.tex
paper edits

@@
 \begin{abstract}
-We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin.
+We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin.
 Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
@@
 These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin.
-Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution.
-The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
-Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
+Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the joint temperature–period distribution, which is also well-described by a curve of constant Rossby number.
+The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed surface spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
+These observations provide further support for the hypothesis that the period gap is due to stellar astrophysics, rather than a non-uniform star-formation history in the Kepler field.
 \end{abstract}
@@
 Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
+{\color{red} Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
+The samples considered here have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
+Thus, }
 
 Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
-Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
-Additionally, a case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.
-By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
+Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.\footnote{A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
 Furthermore, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$, which invalidates the claim of \citet{LorenzoOliveira2019}.
@@
 
-An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for stars of similar \teff.
+An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for dwarf stars of similar \teff.
 The effect was first noticed for M-dwarfs \citep{McQuillan2013b}, but was later shown to extend to $\sim$5000~K \citep{Reinhold2013, McQuillan2014, ReinholdHekker2020}, and even to $\sim$6500~K \citep{Davenport2017}.
+
 \citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}.
@@
 Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, further invalidating the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories.
+
 An alternative explanation for the gap was proposed by \citet{Reinhold2019}, who found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability.

### 2021-12-21 src/ms.tex
paper edits

@@
 The samples considered here have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
-Thus, }
+Exacerbating the problem of detection, slower rotation tends to be accompanied by smaller-scale and less periodic variability.
+The asteroseismic sample, by comparison, is less biased and more likely to contain pile-up stars that our sample may have missed.
+It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.
+This effect, in addition to the \teff offets mentioned above, can further bias our inference of \rocrit to lower values.}
 

### 2021-12-21 src/figures/models.py

@@ -41,6 +41,8 @@ cks = pd.read_parquet('../data/data.parquet')
 # where N is the number of KOIs detected around that star so we drop duplicates.
 cks = cks.drop_duplicates(subset=['kepid'], keep='first')
 cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
+# Filter stars that have discrepant Teff between CKS and SPOCS
+#cks_mask = abs(cks['cks_Teff']-cks['bf18_Teff'])/cks['cks_Teff'] < 0.02
 ######################################################################################
 
 

### 2021-12-22 src/ms.tex
paper edits

@@
 It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.
-This effect, in addition to the \teff offets mentioned above, can further bias our inference of \rocrit to lower values.}
+This effect, in addition to the \teff offets mentioned above, can further bias our inference of \rocrit to lower values.
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$>\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.}
 
@@
 As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}).
-However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
-Further work with a larger sample size may reveal if the pile-up does indeed extend to temperatures as cool as the Sun's.
+However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
+{\color{red} Furthermore...} Further work with a larger sample size may reveal if the pile-up does indeed extend to temperatures as cool as the Sun's.
 
@@
 We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates.
-Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim1$~\msun.
+Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
 
@@
 For example, a Sun-like star may spend several Gyr evolving across the ridge.
-Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
+Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim 5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
 

### 2021-12-22 src/figures/ages.py

@@ -265,8 +265,3 @@ plt.show()
 
 print('5th and 95th percentile range of CKS ages (Gyr)   :', np.nanpercentile(cks['cks_age'][ridge], [5,95]))
 print('5th and 95th percentile range of SPOCS ages (Gyr) :', np.nanpercentile(cks['bf18_Age'][ridge], [5,95]))
-# In[ ]:
-
-
-
-

### 2021-12-22 src/ms.tex
paper edits

@@
 Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the joint temperature–period distribution, which is also well-described by a curve of constant Rossby number.
-The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed surface spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
 The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
@@
 
-The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling.
+{\color{red} The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling.
 The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
 Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling.
-In this picture, spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability.
-After some time, spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
-\textbf{An increase in photometric variability accompanies this period of post-stalling spin-down, }
+In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability.
+After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
+A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.}
 
@@
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
+
 However, a systematic offset in \teff could place the Sun on the long-period pile-up.
@@
 If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out.
-One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021}.
+One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021a}.
 Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question.
@@
 
-    \item Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$3~Gyr, compatible with the predictions of \citet{vanSaders2019}.
+    \item Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \citet{vanSaders2019}.
 Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pile-up until an age of $\sim$6~Gyr.
@@
           \texttt{seaborn} \citep{seaborn},
-          \texttt{showyourwork!} \citep{Luger2021}
+          \texttt{showyourwork!} \citep{Luger2021c}
           }

### 2022-01-03 src/ms.tex
paper edits

@@
     \includegraphics[width=\linewidth]{./figures/fraction.pdf}
-    \caption{The relative fraction of stars on the long- and short-period pile-ups in 10~K windows of \teff.}
+    \caption{The relative fraction of stars on the long- and short-period pile-ups, in the \lamostmcq sample, in 10~K windows of \teff.}
     \label{fig:fraction}
@@
 {\color{red} Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
-The samples considered here have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
+The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
 Exacerbating the problem of detection, slower rotation tends to be accompanied by smaller-scale and less periodic variability.
@@
 This effect, in addition to the \teff offets mentioned above, can further bias our inference of \rocrit to lower values.
-Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$>\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.}
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.}
 
@@
 
-We note that the \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
+We note that the \rocrit values we found in \S\ref{subsec:rossby} are smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
 We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.
@@
 
-{\color{red} The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling.
+{\color{red} The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
 The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.

### 2022-01-03 src/ms.tex
paper edits

@@
 It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.
-This effect, in addition to the \teff offets mentioned above, can further bias our inference of \rocrit to lower values.
+This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values.
 Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.}
@@
 However, a systematic offset in \teff could place the Sun on the long-period pile-up.
-Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
+To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
+However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+
+%Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
 Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required.
 To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K.
-However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
 

### 2022-01-03 src/ms.tex
paper edits

@@
     \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
-    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}.
+    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
 The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.
-Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.}
+Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
+The black symbol in each panel indicates the position of the Sun.}
     \label{fig:models}

### 2022-01-03 src/figures/models.py

@@ -3,17 +3,30 @@
 
 import numpy as np
 import pandas as pd
-
+import astropy.constants as c
 from astropy.table import Table
 
 import matplotlib.pyplot as plt
 import matplotlib as mpl
+
 
 mpl.rcParams["figure.dpi"] = 100
 mpl.rcParams["savefig.bbox"] = "tight"
 mpl.rcParams["savefig.dpi"] = 300
 
 import seaborn as sns
+
+####################################################################
+# The Sun
+sun = {"teff": 5772,
+       "prot": 25.4,
+       "e_prot": 25.4-24.5,
+       "E_prot": 36-25.4
+      }
+
+sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
+####################################################################
+
 
 ######################################################################################
 #McQuillan et al. 2013
@@ -91,9 +104,12 @@ model.head()
 mpl.rcParams["legend.markerscale"] = 1
 sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":8, "rasterized":True}
+sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
 
 sns.displot(data=std, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
 plt.scatter(cks['cks_Teff'], cks['d21_prot'], label='California–Kepler Survey', **sc_kws)
+plt.plot(sun["teff"], sun["prot"], **sun_kws)
+plt.plot(sun["teff"], sun["prot"], 'k.')
 plt.gca().invert_xaxis()
 plt.xlim(6500,5000)
 plt.ylim(0,50)
@@ -108,6 +124,8 @@ plt.show()
 
 sns.displot(data=roc, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
 plt.scatter(cks['cks_Teff'], cks['d21_prot'], label='California–Kepler Survey', **sc_kws)
+plt.plot(sun["teff"], sun["prot"], **sun_kws)
[…]

### 2022-01-03 src/figures/teffscales.py

@@ -174,20 +174,17 @@ def residual(ax,
 
     ax.axhline(0, color='k', ls='--', lw=0.5)
     ax.scatter(x1, resid, **sc_kws)
+
     text1 = ax.text(0.05,0.17,'rms = '+str(int(rms))+' K', transform=ax.transAxes)
     text2 = ax.text(0.05,0.05,'median = '+str(int(np.nanmedian(resid)))+' K', transform=ax.transAxes)
     
-    text1.set_path_effects([path_effects.Stroke(linewidth=3, foreground='white'),
-                       path_effects.Normal()])
-    text2.set_path_effects([path_effects.Stroke(linewidth=3, foreground='white'),
-                       path_effects.Normal()])                          
-    
+    for text in [text1,text2]:
+        text.set_path_effects([path_effects.Stroke(linewidth=3, foreground='white'), path_effects.Normal()])
+       
     ax.set_xlabel(xlabel)
     ax.set_ylabel(ylabel)
     ax.set_xlim(xmin, xmax)
     ax.set_ylim(ymin, ymax)
-    
-
     
     return ax
 

### 2022-01-03 src/ms.tex
paper edits

@@
 \subsection{Does the Sun reside on the long-period pileup?} \label{subsec:thesun}
+{\color{red}
 
 It is unclear whether or not the Sun is a resident of the long-period pile-up.
-If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$~5--7 days above the long-period pile-up.
+If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pile-up.
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
 
-However, a systematic offset in \teff could place the Sun on the long-period pile-up.
+One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
+However, this scenario is considered implausible as the change in moment of inertia for the Sun over its lifetime is not enough to produce the 5--7 day gap between the Sun and the observed long-period pile-up.
+
+It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
+However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well.
+As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}).
+However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
+We thus consider the explanation that WMB only operates for $\gtrsim$~1~\msun stars unlikely, and suggest that further work with a larger sample size could reveal the true \teff range of the pile-up more clearly.
+
+We consider the most likely explanation for the observed gap between the Sun and the long-period pile-up to be systematic offsets in \teff.
 To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
-However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+Notably, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less subject to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
+As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
+We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases and small sample sizes.
+}
 
-%Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
-Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required.
-To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K.
-
-If the Sun is not in fact on the pile-up, one possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
-If this is the case, there is some moderate tension with the ages of pile-up stars in the CKS sample (Figure~\ref{fig:ages}).
+%If this is the case, there is some moderate tension with the ages of pile-up stars in the CKS sample (Figure~\ref{fig:ages}).
 In the WMB model, hotter stars are expected to depart the pile-up on shorter timescales, yet some CKS stars in the pile-up appear to be both hotter and older than the Sun.
@@
 
-It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
-As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}).
-However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
-{\color{red} Furthermore...} Further work with a larger sample size may reveal if the pile-up does indeed extend to temperatures as cool as the Sun's.
+%Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
+Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required.
+To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K.
 
-Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}).
+%Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}).
 If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out.
@@
 
-    \item The long-period pile-up is reasonably well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
+    \item The long-period pile-up is well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
 The pile-up is also populated by stars with a wide range of isochrone ages (\agerange).
 A pile-up of stars with a constant Rossby number and a broad range of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The precise value of \rocrit is highly sensitive to \teff scale shifts between observational data and the models used to compute \taucz.
+The precise value of \rocrit is sensitive to \teff scale shifts between observational data and the models used to compute \taucz.
 

### 2022-01-03 src/figures/models.py

@@ -143,7 +143,7 @@ sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":1, "rasterized":True, "alpha":0.75}
 
 sns.displot(data=std, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
-plt.scatter(lam['Teff_lam'], lam['Prot'], label='LAMOST–Kepler', **sc_kws)
+plt.scatter(lam['Teff_lam'], lam['Prot'], label='LAMOST–McQuillan', **sc_kws)
 plt.plot(sun["teff"], sun["prot"], **sun_kws)
 plt.plot(sun["teff"], sun["prot"], 'k.')
 plt.gca().invert_xaxis()
@@ -158,7 +158,7 @@ plt.savefig('../figures/std-model-lamost.pdf')
 plt.show()
 
 sns.displot(data=roc, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
-plt.scatter(lam['Teff_lam'], lam['Prot'], label='LAMOST–Kepler', **sc_kws)
+plt.scatter(lam['Teff_lam'], lam['Prot'], label='LAMOST–McQuillan', **sc_kws)
 plt.plot(sun["teff"], sun["prot"], **sun_kws)
 plt.plot(sun["teff"], sun["prot"], 'k.')
 plt.gca().invert_xaxis()

### 2022-01-04 src/ms.tex
minor edits

@@
 \begin{document}
-\title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}
+\title{Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}
 
@@
 After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
-A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.}
+A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.
+We note that this is a speculative scenario and the underlying physics are not known.}
 
@@
     \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up.
-For example, a Sun-like star may spend several Gyr evolving across the ridge.
+For example, a Sun-like star may spend several Gyr evolving through the long-period pile-up.
 Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim 5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.

### 2022-01-05 src/ms.tex
adding Santos periods

@@
 %% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{\today}
+%\received{\today}
 \revised{\today}
@@
 The LAMOST--Kepler project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
+The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
 We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} rotation period catalog, which published rotation periods for $>$34000 Kepler targets.
-The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
+{\color{red} We additionally cross-matched the LAMOST--Kepler catalog with the \citet{Santos2021} rotation period catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.}
 
-We matched 10844 LAMOST targets to 10650 unique Kepler IDs.
+\textbf{We matched 10844 LAMOST targets to 10650 unique Kepler IDs in the \citet{McQuillan2014} sample.
+In the \citet{Santos2021} sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
 For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
@@
 There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
-Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.
+Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.}
 

### 2022-01-10 src/ms.tex
Removing Barnes & Kim 2010 tau_c tables

@@
 \newcommand{\lamostmcq}{LAMOST--McQuillan\xspace}
-\newcommand{\teffmin}{5800~K\xspace}
+\newcommand{\teffmin}{5500~K\xspace}
 \newcommand{\agerange}{$\sim$2--6~Gyr\xspace}
@@
 
-\newcommand{\rhk}{$R'_\mathrm{HK}$\xspace}
-\newcommand{\shk}{$S_\mathrm{HK}$\xspace}
-\newcommand{\caiihk}{Ca II H\&K \xspace}
+% \newcommand{\rhk}{$R'_\mathrm{HK}$\xspace}
+% \newcommand{\shk}{$S_\mathrm{HK}$\xspace}
+% \newcommand{\caiihk}{Ca II H\&K \xspace}
 
@@
 
-It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
+\textbf{Needs revision} It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
 However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well.

### 2022-01-10 src/figures/ages.py

@@ -104,7 +104,7 @@ ridge &= mask
 ############################################
 
 ######################################################################################
-bk = pd.read_csv("../data/barnes_kim_2010/barnes-kim-2010.csv")
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff,
                                   ref='gunn1998'):
@@ -114,11 +114,11 @@ def convective_turnover_timescale(teff,
         #Gunn et al. 1998 relation, from Cranmer & Saar 2011
         return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
     
-    elif ref == 'barnes2010':
-        #Barnes & Kim 2010 relation for local tau_c
-        teff_pts = 10.**bk['logT']
-        tc_pts   = bk['Local_tau_c']
-        return np.interp(teff, teff_pts, tc_pts)
+    # elif ref == '2010':
+    #     # & Kim 2010 relation for local tau_c
+    #     teff_pts = 10.**bk['logT']
+    #     tc_pts   = bk['Local_tau_c']
+    #     return np.interp(teff, teff_pts, tc_pts)
 
 def constant_rossby(teff, ro):
     #Return locus of rotation periods corresponding to constant Rossby number

### 2022-01-10 src/figures/fraction.py

@@ -60,7 +60,7 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 ######################################################################################
 
 ######################################################################################
-bk = pd.read_csv("../data/barnes_kim_2010/barnes-kim-2010.csv")
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff,
                                   ref='gunn1998'):
@@ -70,11 +70,11 @@ def convective_turnover_timescale(teff,
         #Gunn et al. 1998 relation, from Cranmer & Saar 2011
         return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
     
-    elif ref == 'barnes2010':
-        #Barnes & Kim 2010 relation for local tau_c
-        teff_pts = 10.**bk['logT']
-        tc_pts   = bk['Local_tau_c']
-        return np.interp(teff, teff_pts, tc_pts)
+    # elif ref == '2010':
+    #     # & Kim 2010 relation for local tau_c
+    #     teff_pts = 10.**bk['logT']
+    #     tc_pts   = bk['Local_tau_c']
+    #     return np.interp(teff, teff_pts, tc_pts)
 
 def constant_rossby(teff, ro):
     #Return locus of rotation periods corresponding to constant Rossby number

### 2022-01-10 src/figures/gap.py

@@ -72,7 +72,7 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 
 
 ######################################################################################
-bk = pd.read_csv("../data/barnes_kim_2010/barnes-kim-2010.csv")
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff,
                                   ref='gunn1998'):
@@ -82,18 +82,18 @@ def convective_turnover_timescale(teff,
         #Gunn et al. 1998 relation, from Cranmer & Saar 2011
         return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
     
-    elif ref == 'barnes2010':
-        #Barnes & Kim 2010 relation for local tau_c
-        teff_pts = 10.**bk['logT']
-        tc_pts   = bk['Local_tau_c']
-        return np.interp(teff, teff_pts, tc_pts)
+    # elif ref == '2010':
+    #     # & Kim 2010 relation for local tau_c
+    #     teff_pts = 10.**bk['logT']
+    #     tc_pts   = bk['Local_tau_c']
+    #     return np.interp(teff, teff_pts, tc_pts)
 
 def constant_rossby(teff, ro):
     #Return locus of rotation periods corresponding to constant Rossby number
     return ro * convective_turnover_timescale(teff)
 
 lam["Ro"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='gunn1998')    
-lam["Ro_barnes"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='barnes2010')
+#lam["Ro_"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='2010')
 mcq["Ro"] = mcq["mcq_Prot"]/convective_turnover_timescale(mcq["mcq_Teff"], ref='gunn1998')    
 
 

### 2022-01-10 src/figures/kde.py

@@ -26,7 +26,7 @@ sun = {"teff": 5772,
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
 
-bk = pd.read_csv("../data/barnes_kim_2010/barnes-kim-2010.csv")
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff,
                                   ref='gunn1998'):
@@ -36,11 +36,11 @@ def convective_turnover_timescale(teff,
         #Gunn et al. 1998 relation, from Cranmer & Saar 2011
         return 314.24*np.exp(-(teff/1952.5) - (teff/6250.)**18.) + 0.002
     
-    elif ref == 'barnes2010':
-        #Barnes & Kim 2010 relation for local tau_c
-        teff_pts = 10.**bk['logT']
-        tc_pts   = bk['Local_tau_c']
-        return np.interp(teff, teff_pts, tc_pts)
+    # elif ref == '2010':
+    #     # & Kim 2010 relation for local tau_c
+    #     teff_pts = 10.**bk['logT']
+    #     tc_pts   = bk['Local_tau_c']
+    #     return np.interp(teff, teff_pts, tc_pts)
 
 
 def constant_rossby(teff, ro):

### 2022-01-10 src/figures/mcmc.py

@@ -77,7 +77,7 @@ lam = lam[lam_mask]
 
 
 ######################################################################################
-bk = pd.read_csv("../data/barnes_kim_2010/barnes-kim-2010.csv")
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff,
                                   ref='gunn1998'):
@@ -87,18 +87,18 @@ def convective_turnover_timescale(teff,
         #Gunn et al. 1998 relation, from Cranmer & Saar 2011
         return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
     
-    elif ref == 'barnes2010':
-        #Barnes & Kim 2010 relation for local tau_c
-        teff_pts = 10.**bk['logT']
-        tc_pts   = bk['Local_tau_c']
-        return np.interp(teff, teff_pts, tc_pts)
+    # elif ref == '2010':
+    #     # & Kim 2010 relation for local tau_c
+    #     teff_pts = 10.**bk['logT']
+    #     tc_pts   = bk['Local_tau_c']
+    #     return np.interp(teff, teff_pts, tc_pts)
 
 def constant_rossby(teff, ro):
     #Return locus of rotation periods corresponding to constant Rossby number
     return ro * convective_turnover_timescale(teff)
 
 lam["Ro"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='gunn1998')    
-lam["Ro_barnes"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='barnes2010')
+#lam["Ro_"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"], ref='2010')
 ######################################################################################
 
     

### 2022-01-10 src/figures/models.py
adding Santos 2021 data and updating JVS models

@@ -85,14 +85,14 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 
 ######################################################################################
 # van Saders et al. 2019 models
-std = pd.read_csv('../data/models/standard_model.csv')
-roc = pd.read_csv('../data/models/rocrit_model.csv')
+#std = pd.read_csv('../data/models/standard_model.csv')
+#roc = pd.read_csv('../data/models/rocrit_model.csv')
 
-# std = pd.read_hdf('../data/models/standard_population.h5', key='sample')
-# std = std[std['evo']==1]
+std = pd.read_hdf('../data/models/standard_population.h5', key='sample')
+std = std[std['evo']==1]
 
-# roc = pd.read_hdf('../data/models/rocrit_population.h5', key='sample')
-# roc = roc[roc['evo']==1]
+roc = pd.read_hdf('../data/models/rocrit_population.h5', key='sample')
+roc = roc[roc['evo']==1]
 
 std['flag'] = 'std'
 roc['flag'] = 'roc'

### 2022-01-10 src/ms.tex
adding Santos 2021 data and updating JVS models
paper updated

--- 2022-01-05 src/ms.tex
@@
 
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{./figures/lamost-mcquillan.pdf}
+    \includegraphics[width=\textwidth]{./figures/lamost-santos.pdf}
+    \caption{}
+    \label{fig:xmatch}
+\end{figure*}
+
 \section{Analysis}
@@
 The constant Rossby model provides a reasonably good description of the \lamostmcq long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
-Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, possibly because the cooler stars have not had enough time to evolve to that Rossby number \citep[see Figure 6 in][]{vanSaders2019}.
+Above and below this \teff range we see clear and significant divergence from the constant Rossby model, such that the model under-predicts periods of hotter stars and over-predicts periods of cooler stars, possibly because the cooler stars have not had enough time to evolve to the critical Rossby number associated with weakened magnetic braking \citep[see Figure 6 in][]{vanSaders2019}.
 

### 2022-01-11 src/ms.tex
paper edits

@@
 \newcommand{\lamostmcq}{LAMOST--McQuillan\xspace}
+\newcommand{\lamostsan}{LAMOST--Santos\xspace}
 \newcommand{\teffmin}{5500~K\xspace}
@@
 There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
-Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.}
+Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.
+Visualizations of the \lamostmcq and \lamostsan samples are shown in Figure~\ref{fig:xmatch}.}
 
@@
     \includegraphics[width=\textwidth]{./figures/lamost-santos.pdf}
-    \caption{}
+    \caption{Two-dimensional histograms indicating the number of Kepler target stars in the \teff–\prot plane for the \lamostmcq sample (top row) and the \lamostsan sample (bottom row).
+The effects of a simple cut in \logg to separate subgiants and dwarfs are shown in the middle and right columns.
+In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from its differentially rotating surface.
+The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature.
+The short-period pile-up is most clear in the \lamostmcq sample.}
     \label{fig:xmatch}
@@
 \begin{acknowledgments}
-We thank Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, and Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog.
+We thank Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, and Karl Jaehnig for plotting advice.
 It is a pleasure to thank the Stars Group at the American Museum of Natural History, and the Astronomical Data Group at the Flatiron Institute for helpful discussions.

### 2022-01-11 src/ms.tex
paper edits

@@
 The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature.
-The short-period pile-up is most clear in the \lamostmcq sample.}
+The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.}
     \label{fig:xmatch}
@@
 
-It it also possible that the long-period pile-up studied here only applies to stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
+Another possibility is that the long-period pile-up is only present for stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
 However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well.
-As seen in Figure~\ref{fig:asteroseismic}, it is not clear that a pile-up exists for \teff~$\lesssim$~5800~K, particularly when examining the CKS and \citet{Hall2021} samples (see also \S\ref{subsec:extent}).
-However, kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, suggests the pile-up may exist at \teff as low as 5500~K or even 5000~K.
-We thus consider the explanation that WMB only operates for $\gtrsim$~1~\msun stars unlikely, and suggest that further work with a larger sample size could reveal the true \teff range of the pile-up more clearly.
+As shown in Figure~\ref{fig:xmatch}, it appears that the long-period pile-up does extend to temperatures as low as the Sun's, though this is not apparent from smaller samples such as the CKS sample or the \citet{Hall2021} asteroseismic sample (see also \S\ref{subsec:extent}).
+Kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, also suggests the pile-up exists at \teff as low as 5500~K or even 5000~K.
+We thus reject the explanation that WMB only operates for $\gtrsim$~1~\msun stars.
 
@@
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
-We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases and small sample sizes.
+We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
 }

### 2022-01-12 src/ms.tex
paper edits

@@
 The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
-We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} rotation period catalog, which published rotation periods for $>$34000 Kepler targets.
-{\color{red} We additionally cross-matched the LAMOST--Kepler catalog with the \citet{Santos2021} rotation period catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.}
+We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, {\color{red} as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.}
 

### 2022-01-13 src/ms.tex
revising macros

@@
 As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
-Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim 5800~K$ have already piled up onto the edge by or before this timescale.
+Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim$~\teffmin have already piled up onto the edge by or before this timescale.
 Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge.
@@
 \begin{enumerate}
-    \item We observe an overdensity at the long-period edge of the \teff-\prot distribution of Kepler main-sequence stars with \teff~$\gtrsim$~5800~K.
+    \item We observe an overdensity at the long-period edge of the \teff-\prot distribution of Kepler main-sequence stars with \teff~$\gtrsim$~\teffmin.
 We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates.
@@
 For example, a Sun-like star may spend several Gyr evolving through the long-period pile-up.
-Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim 5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
+Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim$~\teffmin (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
 

### 2022-01-14 src/ms.tex
paper edits

@@
 However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
-The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and the concentration of stars at the upper age boundary at a given \teff (Figure~\ref{fig:ages}).
+The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
 In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.

### 2022-01-18 src/ms.tex
adjusting subgiant cuts

@@
     \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
-The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.
+Shown here are stars with \logg~$>$~4.1.
 Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
@@
 \end{figure*}
+
+%The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.
 
@@
 It is unclear whether or not the Sun is a resident of the long-period pile-up.
-If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pile-up.
+From Fig.~\ref{fig:xmatch} it is clear that the long-period pile-up extends to temperatures as cool as the Sun's.
+However, it is not clear whether there is a large systematic offset in the LAMOST temperature scale and the temperature scale of the Sun.
+If one assumes there are no systematic offsets between the two \teff scales, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pile-up.
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
@@
 
-Another possibility is that the long-period pile-up is only present for stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
+%Another possibility is that the long-period pile-up is only present for stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
 However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well.
-As shown in Figure~\ref{fig:xmatch}, it appears that the long-period pile-up does extend to temperatures as low as the Sun's, though this is not apparent from smaller samples such as the CKS sample or the \citet{Hall2021} asteroseismic sample (see also \S\ref{subsec:extent}).
+
+%As shown in Figure~\ref{fig:xmatch}, it appears that the long-period pile-up does extend to temperatures as low as the Sun's, though this is not apparent from smaller samples such as the CKS sample or the \citet{Hall2021} asteroseismic sample (see also \S\ref{subsec:extent}).
 Kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, also suggests the pile-up exists at \teff as low as 5500~K or even 5000~K.
-We thus reject the explanation that WMB only operates for $\gtrsim$~1~\msun stars.
+
+%We thus reject the explanation that WMB only operates for $\gtrsim$~1~\msun stars.
 
@@
 Notably, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
-Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less subject to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
+Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.

### 2022-01-18 src/figures/models.py

@@ -100,16 +100,20 @@ roc['flag'] = 'roc'
 model = pd.concat([std, roc], ignore_index=True, sort=True)
 model.head()
 ######################################################################################
-
-
 mpl.rcParams["legend.markerscale"] = 1
 sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":8, "rasterized":True}
 sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
 
 #sns.displot(data=std, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
+
+#We only want to plot dwarf stars
+logg_thresh = 4.1
+cks_ms = cks['p20_cks_slogg'] > logg_thresh
+lam_ms = lam['logg_lam'] > logg_thresh
+
 sns.displot(data=std, x="Teff", y="period", binwidth=(20, 0.5), cbar=True, vmin=0, vmax=100, cbar_kws={'label': r'N$_\mathregular{stars}$'})
-plt.scatter(cks['cks_Teff'], cks['d21_prot'], label='California–Kepler Survey', **sc_kws)
+plt.scatter(cks['cks_Teff'][cks_ms], cks['d21_prot'][cks_ms], label='California–Kepler Survey', **sc_kws)
 plt.plot(sun["teff"], sun["prot"], **sun_kws)
 plt.plot(sun["teff"], sun["prot"], 'k.')
 plt.gca().invert_xaxis()
@@ -121,12 +125,10 @@ plt.xlabel('Effective temperature [K]')
 plt.ylabel('Rotation period [d]')
 plt.text(1.15,1.05,"a",transform=plt.gca().transAxes,weight="bold",size=14)
 plt.savefig('../figures/std-model-cks.pdf')
-#plt.show()
-
 
 #sns.displot(data=roc, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
 sns.displot(data=roc, x="Teff", y="period", binwidth=(20, 0.5), cbar=True, vmin=0, vmax=100, cbar_kws={'label': r'N$_\mathregular{stars}$'})
[…]

### 2022-01-18 src/figures/surveys.py

@@ -120,4 +120,3 @@ sns.despine()
 plt.subplots_adjust(wspace=0)
 plt.savefig('../figures/surveys.pdf')
 plt.savefig('../static/surveys.png')
-#plt.show()

### 2022-01-18 src/figures/xmatch.py

@@ -85,13 +85,15 @@ sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
 x = xm["Teff"]
 y = xm["mcq_Prot"]
 
+logg_thresh = 4.1
+
 arg = (np.isfinite(x)) & (np.isfinite(y))
-ms  = arg & (xm["logg_x"]>4.2)
-sg  = arg & (xm["logg_x"]<4.2)
+ms  = arg & (xm["logg_x"]>logg_thresh)
+sg  = arg & (xm["logg_x"]<logg_thresh)
 
 titles = ["LAMOST-McQuillan\nall stars",
-          "LAMOST-McQuillan\nlog(g) < 4.2 (subgiants)",
-          "LAMOST-McQuillan\nlog(g) > 4.2 (dwarfs)"]
+          "LAMOST-McQuillan\nlog(g) < "+str(logg_thresh)+" (subgiants)",
+          "LAMOST-McQuillan\nlog(g) > "+str(logg_thresh)+" (dwarfs)"]
 
 h_kws = {"bins":150, "cmap": "Blues", "cmin": 1}
 
@@ -141,16 +143,19 @@ print(len(xm), 'unique stars in LAMOST-Santos cross-match')
 arg = (np.isfinite(xm["Teff"])&(np.isfinite(xm["san_Prot"])))
 print(len(xm[arg]), 'unique stars in LAMOST-Santos cross-match with Teff and Prot')
 
+xm = xm[arg]
+xm.to_parquet('../data/lamost-santos-xmatch.parquet')
+
 x = xm["Teff"]
 y = xm["san_Prot"]
 
 arg = (np.isfinite(x)) & (np.isfinite(y))
-ms  = arg & (xm["logg_x"]>4.2)
-sg  = arg & (xm["logg_x"]<4.2)
+ms  = arg & (xm["logg_x"]>logg_thresh)
+sg  = arg & (xm["logg_x"]<logg_thresh)
 
 titles = ["LAMOST-Santos\nall stars",
-          "LAMOST-Santos\nlog(g) < 4.2 (subgiants)",
-          "LAMOST-Santos\nlog(g) > 4.2 (dwarfs)"]
+          "LAMOST-Santos\nlog(g) < "+str(logg_thresh)+" (subgiants)",
+          "LAMOST-Santos\nlog(g) > "+str(logg_thresh)+" (dwarfs)"]
 
[…]

### 2022-01-18 src/ms.tex
revising table

@@
 The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature.
-The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.}
+The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.
+The secondary overdensity in the subgiant panel, most visible in the bottom center panel, appears to be at twice the period of the primary overdensity, potentially due to erroneously determined rotation periods.}
     \label{fig:xmatch}
@@
 We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
-For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d.
-To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
+{\color{red} For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.} To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
 We note that this selection is particular to the CKS \teff scale and is not general.
@@
 \hline
-\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.37 $\pm$ 0.02 & 0.005 $\pm$ 0.004 & 59 $\pm$ 15 \\
-\lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.41 $\pm$ 0.09 & 0.01 $\pm$ 0.01 & 42 $\pm$ 37 \\
-CKS (CKS \teff) & 5800~K--6250~K & 1.75 $\pm$ 0.07 & 0.05 $\pm$ 0.01 & -58 $\pm$ 16 \\ 
-CKS (SPOCS \teff) & 5800~K--6250~K & 1.9 $\pm$ 0.1 & 0.09 $\pm$ 0.01 & -123 $\pm$ 20 \\
-CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.1 & 0.17 $\pm$ 0.02 & 26 $\pm$ 38 \\
+\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.38 $\pm$ 0.04 & 0.011 $\pm$ 0.008 & 49 $\pm$ 19 \\
+\lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.4 $\pm$ 0.1 & 0.019 $\pm$ 0.015 & 44 $\pm$ 43 \\
+CKS (CKS \teff) & 5800~K--6250~K & 1.77 $\pm$ 0.07 & 0.02 $\pm$ 0.01 & -70 $\pm$ 15 \\ 
+CKS (SPOCS \teff) & 5800~K--6250~K & 1.9 $\pm$ 0.1 & 0.12 $\pm$ 0.01 & -136 $\pm$ 23 \\
+CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.2 & 0.18 $\pm$ 0.02 & 21 $\pm$ 42 \\
 \citet{Hall2021} main-sequence & 5000~K--6250~K & 1.7 $\pm$ 0.2 & 0.30 $\pm$ 0.04 & -5 $\pm$ 64 \\
 \citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & -29 $\pm$ 79 \\
-All samples & 5000~K--6250~K & 1.47 $\pm$ 0.05 & 0.132 $\pm$ 0.008 & 2 $\pm$ 23 (LAMOST) \\
- & & & & 10 $\pm$ 15 (CKS) \\
- & & & & 61 $\pm$ 21  (Hall) \\
-All samples & 5800~K--6250~K & 1.67 $\pm$ 0.08 & 0.116 $\pm$ 0.009 & -42 $\pm$ 24 (LAMOST) \\
-&  &  &  & -44 $\pm$ 20 (CKS) \\
-&  &  &  & 3 $\pm$ 23 (Hall) \\ 
+%All samples & 5000~K--6250~K & 1.47 $\pm$ 0.05 & 0.132 $\pm$ 0.008 & 2 $\pm$ 23 (LAMOST) \\
+% & & & & 10 $\pm$ 15 (CKS) \\
+% & & & & 61 $\pm$ 21  (Hall) \\
+%All samples & 5800~K--6250~K & 1.67 $\pm$ 0.08 & 0.116 $\pm$ 0.009 & -42 $\pm$ 24 (LAMOST) \\
+%&  &  &  & -44 $\pm$ 20 (CKS) \\
+%&  &  &  & 3 $\pm$ 23 (Hall) \\ 
 \enddata

### 2022-01-19 src/figures/fraction.py
paper edits

@@ -50,7 +50,8 @@ lam = lam.drop_duplicates(subset=['KIC'], keep='first')
 
 lam_mask = (lam["Teff_lam"]>3000)
 lam_mask = (lam["Teff_lam"]<8000)
-lam_mask &= (lam["logg_lam"]>3)
+#lam_mask &= (lam["logg_lam"]>3)
+lam_mask &= (lam["logg_lam"]>4.1)
 lam_mask &= (lam["logg_lam"]<5)
 lam_mask &= (abs(lam["feh_lam"])<2)
 lam = lam[lam_mask]

### 2022-01-19 src/figures/mcmc.py

@@ -375,7 +375,6 @@ def rossby_plot(teff_bin_centers, flat_samples_10, flat_samples_90):
     sns.despine()
     plt.tight_layout()
     plt.savefig('../figures/mcmc.pdf')
-    #plt.show()
     
     return
 

### 2022-01-20 src/ms.tex
paper edits

@@
 
-\textbf{We matched 10844 LAMOST targets to 10650 unique Kepler IDs in the \citet{McQuillan2014} sample.
+\textbf{We matched 10844 LAMOST targets to 10550 unique Kepler IDs in the \citet{McQuillan2014} sample.
 In the \citet{Santos2021} sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
@@
 
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
+    \caption{}
+    \label{fig:rocrit}
+\end{figure*}
+
 \begin{deluxetable*}{lllll}
@@
 \hline
-\lamostmcq 10th \prot pctl. & 4500~K--5800~K & 0.367 $\pm$ 0.009 & 0.008 $\pm$ 0.006 &  44 $\pm$ 36 \\
+\lamostmcq 10th \prot pctl. & 4500~K--5800~K & 0.37 $\pm$ 0.02 & 0.011 $\pm$ 0.009 &  46 $\pm$ 80 \\
 \hline 
@@
     \includegraphics[width=\linewidth]{./figures/fraction.pdf}
-    \caption{The relative fraction of stars on the long- and short-period pile-ups, in the \lamostmcq sample, in 10~K windows of \teff.}
+    \caption{The relative fraction of stars on the long- and short-period pile-ups, in the \lamostmcq sample, in 10~K windows of \teff (for stars with \logg~$>$4.1 dex).}
     \label{fig:fraction}
@@
 
-    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane.
+    \item {\color{red} An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane.
 However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods.
@@
 The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
-When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
+When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.}
 

### 2022-01-20 src/figures/rocrit.py

@@ -1,0 +1,272 @@
+import numpy as np
+import pandas as pd
+
+from astropy.table import Table
+
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+import seaborn as sns
+
+sns.set(
+        context="paper",
+        style="ticks",
+        font_scale=1.2,
+        palette="colorblind"
+        )
+
+
+######################################################################################
+#McQuillan et al. 2013
+mcq_koi = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/table1.dat",
+                readme="https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/ReadMe",
+                format="ascii.cds")
+mcq_koi = mcq_koi.to_pandas()
+mcq_koi = mcq_koi.add_prefix('mcq_')
+
+
+#McQuillan et al. 2014
+mcq = pd.read_parquet('../data/mcquillan2014_table1.parquet')
+######################################################################################
+
+
+######################################################################################
+# California-Kepler Survey (Fulton & Petigura 2018)
+# This data table has been augmented with data from other surveys (see David et al. 2021)
+cks = pd.read_parquet('../data/cks_merged.parquet')
+# The dataframe has a row entry for each KOI, meaning individual star are represented N times
+# where N is the number of KOIs detected around that star so we drop duplicates.
+cks = cks.drop_duplicates(subset=['kepid'], keep='first')
+cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
+######################################################################################
+
+
+######################################################################################
+# LAMOST-Kepler 
+lam = pd.read_csv('../data/kepler_lamost.csv')
+#print('LAMOST unique KIC targets:', len(np.unique(lam["KIC"])))
[…]

### 2022-01-20 src/figures/xmatch.py

@@ -123,8 +123,6 @@ for i,ax in enumerate(axes):
 sns.despine()
 plt.tight_layout()
 plt.savefig('../figures/lamost-mcquillan.pdf')
-#plt.show()
-
 
 
 #Santos et al. 2021

### 2022-01-20 src/ms.tex
paper edits

@@
 
-We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
+We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the \citet{Hall2021} asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
 {\color{red} For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.} To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
@@
 For the CKS sample, we additionally required $\logg~>~4$ and 5850~K~$<~\teff~<$~6500~K.
-In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.
+
+%, and all combined data sets of the long-period pile-up, 
+% In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.
 

### 2022-01-20 src/ms.tex
revising ages figure

@@
 
-We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex.
+We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
 While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).

### 2022-01-20 src/ms.tex
Adding Santos periods to comparison figure

@@
 Thus, in the core–envelope coupling scenario, spin evolution is slowed when the star's age becomes comparable to the core–envelope coupling timescale.
-Theoretical predictions for the core–envelope coupling are mass-dependent, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
+Theoretical predictions for the core–envelope coupling are mass-dependent {\color{red} (in order to match observed cluster sequences)}, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
 
-After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
+{\color{red} After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
 From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
 Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience little spin evolution between $\sim$100~Myr and $\sim$200~Myr.
-This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.
+This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.}
 

### 2022-01-20 src/ms.tex
cleaning up

@@
 The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.
-The secondary overdensity in the subgiant panel, most visible in the bottom center panel, appears to be at twice the period of the primary overdensity, potentially due to erroneously determined rotation periods.}
+The secondary overdensity observed in the subgiant samples, most visible in the bottom center panel, appears to be at twice the period of the primary overdensity, potentially due to erroneously determined rotation periods.}
     \label{fig:xmatch}

### 2022-01-20 src/figures/asteroseismic.py

@@ -24,7 +24,6 @@ import seaborn as sns
 hall = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/other/NatAs/5.707/table1.dat",
                   readme="https://cdsarc.cds.unistra.fr/ftp/J/other/NatAs/5.707/ReadMe",
                   format="ascii.cds")
-#hall.info()
 
 
 # ### CKS sample
@@ -34,25 +33,18 @@ mcq_koi = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/table1.dat
                 format="ascii.cds")
 mcq_koi = mcq_koi.to_pandas()
 mcq_koi = mcq_koi.add_prefix('mcq_')
-#mcq_koi.head()
 
 cks = pd.read_parquet('../data/cks_merged.parquet')
 # The dataframe has a row entry for each KOI, meaning individual star are represented N times
 # where N is the number of KOIs detected around that star so we drop duplicates.
-print(np.shape(cks))
 cks = cks.drop_duplicates(subset=['kepid'], keep='first')
-
 cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
-print(np.shape(cks))
-cks.head()
-
 
 # ### LAMOST-Kepler sample
 lam = pd.read_csv('../data/kepler_lamost.csv')
 
 #Drop duplicates
 lam = lam.drop_duplicates(subset=['KIC'], keep='first')
-lam.head()
 
 #ik = (lam['Teff_lam']>0)
 #ik &= (lam['Gmag']<15)
@@ -93,10 +85,8 @@ for i in range(2):
     ax[i].set_xlim(6500,5000)
     ax[i].set_xlabel("Effective temperature [K]")
     ax[i].set_ylabel("Rotation period [d]")
-    #ax[i].set_title(titles[i])
     ax[i].legend(prop={"size":9})
     
 sns.despine()    
 plt.tight_layout()
 plt.savefig('../figures/asteroseismic.pdf')
-#plt.show()

### 2022-01-20 src/figures/comparison.py

@@ -1,7 +1,6 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 import matplotlib as mpl
-
 import numpy as np
 
 mpl.rcParams["figure.dpi"] = 100
@@ -26,7 +25,6 @@ san = pd.read_csv('../data/santos2021/S21_rotators.csv')
 san = san.add_prefix('san_')
 san = san.sort_values(['san_KIC', 'san_Kp'], ascending = (True, True))
 san = san.drop_duplicates(subset=['san_KIC'], keep='first')
-# san.head()
 cks = cks.merge(san, left_on='kepid', right_on='san_KIC', how='left')
 
 
@@ -35,11 +33,9 @@ print(len(cks[np.isfinite(cks['ra'])]))
 
 cks = cks.dropna(subset=['gaia_ra'])
 print(np.shape(cks))
-cks.head()
 
 
 cks_pos = cks[['gaia_ra', 'gaia_dec']].copy()
-cks_pos.head()
 cks_pos.to_csv('../data/cks-gaia-coords.csv', index=False)
 
 

### 2022-01-20 src/figures/models.py

@@ -98,7 +98,6 @@ std['flag'] = 'std'
 roc['flag'] = 'roc'
 
 model = pd.concat([std, roc], ignore_index=True, sort=True)
-model.head()
 ######################################################################################
 mpl.rcParams["legend.markerscale"] = 1
 sns.set(font_scale=1.2, context="paper", style="ticks")

### 2022-01-20 src/figures/percentiles.py

@@ -108,7 +108,6 @@ std['flag'] = 'std'
 roc['flag'] = 'roc'
 
 model = pd.concat([std, roc], ignore_index=True, sort=True)
-model.head()
 ######################################################################################
 
 

### 2022-01-20 src/figures/ridge.py

@@ -97,7 +97,6 @@ mcq_koi = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/table1.dat
                 format="ascii.cds")
 mcq_koi = mcq_koi.to_pandas()
 mcq_koi = mcq_koi.add_prefix('mcq_')
-mcq_koi.head()
 
 
 #CKS catalog with auxiliary data
@@ -107,7 +106,6 @@ cks = pd.read_parquet('../data/cks_merged.parquet')
 # where N is the number of KOIs detected around that star so we drop duplicates.
 cks = cks.drop_duplicates(subset=['kepid'], keep='first')
 cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
-cks.head()
 
 sns.set(font_scale=2.1, context="paper", style="ticks")
 

### 2022-01-20 src/figures/surveys.py

@@ -82,6 +82,7 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 apo = Table.read('../data/kepler_apogee_dr16.fits')
 names = [name for name in apo.colnames if len(apo[name].shape) <= 1]
 apo  = apo[names].to_pandas()
+#apo = pd.read_parquet('../data/kepler_apogee_dr16.parquet')
 apok = apo.merge(mcq, how='inner', left_on='kepid', right_on='mcq_KIC')
 ######################################################################################
 

### 2022-01-21 src/ms.tex
paper edits
adding discussion of fixed Rocrit fits

@@
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+{\color{red} We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up.} As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
 
@@
 \end{figure*}
-
-%The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.
 
@@
 
+{\color{red} To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
+The first model assumes \rocrit = 2 and has one free parameter, a temperature shift added to the data.
+The second model assumes no temperature shift and allows \rocrit to vary (\rocrit being the only free parameter).
+The results of these fits are shown in Figure~\ref{fig:rocrit}.
+We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = 2 and variable \rocrit models was negligible.
+In other words, the \rocrit = 2 model fits the data well with a \teff shift of $\approx$~80--150~K depending on the \teff source used.
+For the \lamostmcq sample, there is weak support for the variable \rocrit model.
+However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
+Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure.
+Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs.
+\lamostmcq samples (see Fig.~\ref{fig:percentiles}).}
+
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
-    \caption{}
+    \caption{Non-linear least-squares fits to the long-period pile-up for a \rocrit~=~2 model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
+The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=2} - \chi^2_{\Delta T_{eff}=0}$.}
     \label{fig:rocrit}

### 2022-01-25 src/ms.tex
editing sample selection discussion

@@
 
-\textbf{We matched 10844 LAMOST targets to 10550 unique Kepler IDs in the \citet{McQuillan2014} sample.
+We matched 10844 LAMOST targets to 10550 unique Kepler IDs in the \citet{McQuillan2014}, resulting in a sample with well-determined \prot and spectroscopic \teff (having a median error of 24~K).
+For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
+(having a median error of 24~K).
 In the \citet{Santos2021} sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
-For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
-We additionally selected sources with 3000~K~$<~\teff~<$~8000~K, 3~$<$~\logg~$<$~5~dex, and -2~$<$~[Fe/H]~$<$~2~dex.
-The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K).
 The metallicity distribution of the LAMOST--McQuillan sample is centered near solar, with a mean and standard deviation of -0.1~dex and 0.26~dex, respectively.
@@
 Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.
-Visualizations of the \lamostmcq and \lamostsan samples are shown in Figure~\ref{fig:xmatch}.}
+Visualizations of the \lamostmcq and \lamostsan samples are shown in Figure~\ref{fig:xmatch}.
+
+%The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K).
+We additionally selected sources with 3000~K~$<~\teff~<$~8000~K, 3~$<$~\logg~$<$~5~dex, and -2~$<$~[Fe/H]~$<$~2~dex.
 

### 2022-01-25 src/ms.tex
edits

@@
 Thus, in the core–envelope coupling scenario, spin evolution is slowed when the star's age becomes comparable to the core–envelope coupling timescale.
-Theoretical predictions for the core–envelope coupling are mass-dependent {\color{red} (in order to match observed cluster sequences)}, and for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
+Theoretical predictions for the core–envelope coupling timescale for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
 

### 2022-02-01 src/ms.tex
addressing JC comments

@@
 For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
-(having a median error of 24~K).
 In the \citet{Santos2021} sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
-The metallicity distribution of the LAMOST--McQuillan sample is centered near solar, with a mean and standard deviation of -0.1~dex and 0.26~dex, respectively.
+The metallicity distribution of the LAMOST--McQuillan sample is centered near solar, with a mean and standard deviation of $-0.1$~dex and 0.26~dex, respectively.
 There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
@@
 \subsection{Initial observations}
-A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
+We first noted a pile-up of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
 Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}).
@@
 Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2020}, implying that it takes F-type stars $>1$~Gyr to converge onto the long-period edge.
+Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period edge.
 These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
@@
 
-We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
+We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}), 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
 While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).
@@
 However, this feature may also be due to an epoch of apparent stalling, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
-Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages, notably Praesepe (0.67~Gyr), NGC 752 (1~Gyr), and NGC 6811 (1.4~Gyr) \citep{Agueros2018, Curtis2019a, Curtis2020}.
+Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}, notably Praesepe (0.67~Gyr), NGC 6811 (1.4~Gyr), and NGC 752 (1.4~Gyr).
 In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.
@@
 {\color{red} After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
-From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after $\approx$~230~Myr.
+From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after an age of $\approx$~230~Myr.
 Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience little spin evolution between $\sim$100~Myr and $\sim$200~Myr.

### 2022-02-02 src/ms.tex
adding IC
JVS abstract edit
fixing KIC reference

@@
 
+\author[0000-0001-8196-516X]{Isabel L.~Colman}
+\affil{\amnh}
+
 \author[0000-0002-3011-4784]{Gabriella Contardo}
@@
 We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin.
+The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
 Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
 In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
-The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
 These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin.
@@
 Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep{vanSaders2019}.
-Those authors also predicted a pile-up of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the Kepler Input Catalog \citep[KIC,][]{Brown2011}.
+Those authors also predicted a pile-up of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
 While \citet{vanSaders2019} favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.

### 2022-02-02 src/ms.tex
asteroseismic sample footnote

@@
 
-More recently, \citet{Hall2021} used the asteroseismic rotation rates of Kepler dwarfs, with a different selection bias from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
+More recently, \citet{Hall2021} used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \citet{Hall2021} sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
 \citet{Masuda2021} also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.

### 2022-02-02 src/ms.tex
JVS comments

@@
 We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation.
-Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 may be compatible with all of the data sets if modest \teff systematics are present.
+Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 is compatible with all of the data sets if modest \teff systematics are present.
 Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).

### 2022-02-02 src/ms.tex
JVS comments

@@
 
+%Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff.
+For higher mass stars with MS lifetimes of 4 Gyr, it means that stars on the ridge should have a ~2 Gyr age spread.
+For stars with 10 Gyr MS lifetimes, it should be something close to a 5 Gyr age spread.
+
 We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}), 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
@@
 \subsection{Where do the pile-ups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
 While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
@@
 Exacerbating the problem of detection, slower rotation tends to be accompanied by smaller-scale and less periodic variability.
-The asteroseismic sample, by comparison, is less biased and more likely to contain pile-up stars that our sample may have missed.
+The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
 It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.

### 2022-02-02 src/ms.tex
JVS comments

@@
 Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.\footnote{A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
-Furthermore, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$, which invalidates the claim of \citet{LorenzoOliveira2019}.
-Notably, the \citet{LorenzoOliveira2019} work did not include changes in the stellar moment of inertia in the spin-down calculations.
-Neglecting spin-down due to expansion will lead to more rapid rotation rates that allow standard spin-down models to match observations, leading to the erroneous conclusion that WMB is not required.
+In contrast to the findings of those authors, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.
 

### 2022-02-02 src/ms.tex
JVS comments

@@
 Thus, in the core–envelope coupling scenario, spin evolution is slowed when the star's age becomes comparable to the core–envelope coupling timescale.
-Theoretical predictions for the core–envelope coupling timescale for a solar-mass star range from 30--110 Myr \citep{Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
+Theoretical predictions for the core–envelope coupling timescale for a solar-mass star range from 30--110 Myr \citep{MacGregor1991,Krishnamurthi1997, Bouvier2008, IrwinBouvier2009, Denissenkov2010, GalletBouvier2015, Lanzafame2015, Somers2016, Spada2020}.
 
@@
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
-The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
+
+%The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
 

### 2022-02-02 src/ms.tex
JVS comments

@@
 However, \citet{Curtis2020} demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation.
-Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, further invalidating the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories.
+Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, which is in tension with the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories (for sufficiently large volumes).
 
@@
 
-One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
+%One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
 However, this scenario is considered implausible as the change in moment of inertia for the Sun over its lifetime is not enough to produce the 5--7 day gap between the Sun and the observed long-period pile-up.

### 2022-02-02 src/ms.tex
abstract edit

@@
 Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the joint temperature–period distribution, which is also well-described by a curve of constant Rossby number.
-The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling.
+The same mechanism was previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
 The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.

### 2022-02-02 src/ms.tex
removing color text

@@
 The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
-We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, {\color{red} as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.}
+We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
 
@@
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-{\color{red} We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up.} As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up.
+As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
 
@@
 We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the \citet{Hall2021} asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
-{\color{red} For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.} To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
+For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
+To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
 We note that this selection is particular to the CKS \teff scale and is not general.
@@
 
-{\color{red} To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
+To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
 The first model assumes \rocrit = 2 and has one free parameter, a temperature shift added to the data.
@@
 Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs.
-\lamostmcq samples (see Fig.~\ref{fig:percentiles}).}
+\lamostmcq samples (see Fig.~\ref{fig:percentiles}).
 
@@
 Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
-{\color{red} Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
+Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
 The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
@@
 This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values.
-Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.}
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.
 
@@
 
-{\color{red} After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
+After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
 From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after an age of $\approx$~230~Myr.
 Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience little spin evolution between $\sim$100~Myr and $\sim$200~Myr.
-This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.}
+This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.
 
@@
 
-{\color{red} The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
+The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
 The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
@@
 A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.
-We note that this is a speculative scenario and the underlying physics are not known.}
+We note that this is a speculative scenario and the underlying physics are not known.
 
@@
 \subsection{Does the Sun reside on the long-period pileup?} \label{subsec:thesun}
-{\color{red}
 
@@
 We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
-}
 
@@
 
-    \item {\color{red} An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane.
+    \item An increasing number of open clusters are being discovered by searching for a clustering of rotation periods along a slow-rotator sequence in the \teff–\prot or color–\prot plane.
 However, the long-period pile up discovered here can mimic a slow-rotator sequence in a small sample of unassociated stars with different ages and precisely measured temperatures and rotation periods.
@@
 The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
-When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.}
+When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
 

### 2022-02-02 src/ms.tex
adding shifted models figure

@@
 
+\begin{figure*}
+    \centering 
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost-shifted.pdf}
+    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
+Shown here are stars with \logg~$>$~4.1.
+Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
+The black symbol in each panel indicates the position of the Sun.}
+    \label{fig:shifted}
+\end{figure*}
+
 \begin{figure}
@@
 
-%Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff.
+%This is a bit misleading.
+Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff.
 For higher mass stars with MS lifetimes of 4 Gyr, it means that stars on the ridge should have a ~2 Gyr age spread.
@@
 
+%JVS: I am not sure this is actually what you expect to see.
+If you go back and look at the Lanzafame and Spada paper from 2015, the spin down never stalls out entirely...
+there is always evolution in the period, even though it is slower than you would have predicted with a solid body model.
+In the solar mass stars, you never see something as stark as the Praesepe/NGC6811 overlap.
+You're also working in a regime where initial conditions, saturated spin down, and cluster age uncertainties can really get you.
+I'm not sure you can attribute the overlap in Fritzewski to core-envelope decoupling...
+it could also just be you are looking at clusters that are coeval within the age uncertainties (remember, you're dealing with isochrone ages that come from stars with convective cores at these ages, and they have *nasty* systematic shifts depending on what you assumed about the degree of convective core overshooting).
+I would tread carefully here.
+
 Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
@@
 
-The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
+\textbf{The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
 The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
@@
 A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.
-We note that this is a speculative scenario and the underlying physics are not known.
+We note that this is a speculative scenario and the underlying physics are not known.}
+
+%JVS: I am confused by this picture.
+Why would the slow in apparent spin down make the photometric variability drop?
+The variability doesn't care about the spin-down rate, it cares about the instantaneous rotation (and maybe differential rotation).
+In this picture you are still stuck without a reason for why a more slowly rotating star should be more active than a more rapidly rotating one.
+The information you add here doesn't allow you to support or reject the speculation in Gordon 2021, so I would vote to make this speculation more modest, saying that the gap is perhaps an epoch of rapid spin down and leave it at that.
 

### 2022-02-02 src/figures/shifted.py

@@ -1,0 +1,180 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+import numpy as np
+import pandas as pd
+import astropy.constants as c
+from astropy.table import Table
+
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+import seaborn as sns
+
+####################################################################
+# The Sun
+sun = {"teff": 5772,
+       "prot": 25.4,
+       "e_prot": 25.4-24.5,
+       "E_prot": 36-25.4
+      }
+
+sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
+####################################################################
+
+
+######################################################################################
+#McQuillan et al. 2013
+mcq_koi = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/table1.dat",
+                readme="https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/ReadMe",
+                format="ascii.cds")
+mcq_koi = mcq_koi.to_pandas()
+mcq_koi = mcq_koi.add_prefix('mcq_')
+
+
+#McQuillan et al. 2014
+# mcq = Table.read('../data/mcquillan2014/table1.dat',
+#                 readme='../data/mcquillan2014/ReadMe',
+#                 format='ascii.cds')
+# mcq = mcq.to_pandas()
+# mcq = mcq.add_prefix('mcq_')
+mcq = pd.read_parquet('../data/mcquillan2014_table1.parquet')
+######################################################################################
+
+
+######################################################################################
+# California-Kepler Survey (Fulton & Petigura 2018)
+# This data table has been augmented with data from other surveys (see David et al. 2021)
+cks = pd.read_parquet('../data/cks_merged.parquet')
+# The dataframe has a row entry for each KOI, meaning individual star are represented N times
[…]

### 2022-02-02 src/ms.tex
adding shifted models figure

@@
     \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost-shifted.pdf}
-    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
-Shown here are stars with \logg~$>$~4.1.
-Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
-The black symbol in each panel indicates the position of the Sun.}
+    \caption{Same as Figure~\ref{fig:models} with constant \teff offsets applied to the data.
+Shifts of +116~K and +140~K are applied to the CKS and LAMOST \teff, respectively.
+The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pile-up (as described in \S\ref{subsec:rossby}).
+The LAMOST shift is derived in a similar fashion, with an additional +34~K offset which represents the systematic offset between CKS and LAMOST (see Appendix~\ref{app:teff}).}
     \label{fig:shifted}

### 2022-02-02 src/ms.tex
changes to ages figure

@@
 The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations.
+Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
+The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
 
@@
     \includegraphics[width=\linewidth]{./figures/ages.pdf}
-    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the LAMOST--Kepler sample (b).
-Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.}
+    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the \lamostmcq sample (b).
+Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
+The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~<~2 (blue) and Ro~>~2 (orange).
+Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel b relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
     \label{fig:ages}

### 2022-02-02 src/figures/ages.py

@@ -127,6 +127,18 @@ def constant_rossby(teff, ro):
 ######################################################################################
 
 
+#Models
+std = pd.read_hdf('../data/models/standard_population.h5', key='sample')
+std['ro'] = std['period']/(std['taucz']/86400)
+std = std[std['evo']==1] # limit to main-sequence
+
+roc = pd.read_hdf('../data/models/rocrit_population.h5', key='sample')
+roc['ro'] = roc['period']/(roc['taucz']/86400)
+roc = roc[roc['evo']==1] # limit to main-sequence
+
+
+
+
 fig, ax = plt.subplots()
 sns.kdeplot(
     x=lam["Teff_lam"], 
@@ -218,6 +230,13 @@ spocs_teff_err = np.nanmedian(cks['bf18_e_Teff'])
 
 ax[0].plot(cks['p20_cks_steff'], cks['cks_age'], 
             '.', color='lightgrey', zorder=1)
+
+#Plot models
+for i in range(2):
+    ax[i].plot(roc['Teff'][roc['ro']>2], roc['age'][roc['ro']>2], 
+                ',', ms=0.2, color='orange', alpha=0.5, zorder=1, rasterized=True)
+    ax[i].plot(roc['Teff'][roc['ro']<2], roc['age'][roc['ro']<2], 
+                ',', ms=0.2, color='C0', alpha=0.5, zorder=1, rasterized=True)            
 
 ax[0].plot(cks['p20_cks_steff'][ridge], cks['cks_age'][ridge], 
             '.', color='k', zorder=2)

### 2022-02-03 src/ms.tex
JVS comments

@@
 In the WMB model, although wind-driven angular momentum losses cease, stars continue to evolve structurally which results in evolution in the moment of inertia and stellar spin, driven by expansion of the stars away from the main-sequence \citep{vanSaders2019}.
-As a result, an age-gradient is expected along the pile-up as stars of different mass depart the pile-up at different ages.
-Though we have not measured this gradient, the data suggest that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.
-
-%This is a bit misleading.
-Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff.
-For higher mass stars with MS lifetimes of 4 Gyr, it means that stars on the ridge should have a ~2 Gyr age spread.
-For stars with 10 Gyr MS lifetimes, it should be something close to a 5 Gyr age spread.
+Stars reach \rocrit~$\approx$~2 approximately halfway through their main-sequence lifetimes and remain there until the main-sequence turnoff.
+Thus, higher mass stars with shorter main-sequence lifetimes should show a smaller age spread on the long-period pile-up relative to lower mass stars with longer main-sequence lifetimes.
+Though we have not quantified such an age-gradient, the data suggest that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.
 

### 2022-02-03 src/ms.tex
JVS comments

@@
 From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after an age of $\approx$~230~Myr.
-Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience little spin evolution between $\sim$100~Myr and $\sim$200~Myr.
+Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience slowed spin evolution between $\sim$100~Myr and $\sim$200~Myr.
 This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.
-
-%JVS: I am not sure this is actually what you expect to see.
-If you go back and look at the Lanzafame and Spada paper from 2015, the spin down never stalls out entirely...
-there is always evolution in the period, even though it is slower than you would have predicted with a solid body model.
-In the solar mass stars, you never see something as stark as the Praesepe/NGC6811 overlap.
-You're also working in a regime where initial conditions, saturated spin down, and cluster age uncertainties can really get you.
-I'm not sure you can attribute the overlap in Fritzewski to core-envelope decoupling...
-it could also just be you are looking at clusters that are coeval within the age uncertainties (remember, you're dealing with isochrone ages that come from stars with convective cores at these ages, and they have *nasty* systematic shifts depending on what you assumed about the degree of convective core overshooting).
-I would tread carefully here.
+However, an alternative explanation for the overlapping cluster sequences observed by those authors could be that the clusters are in fact coeval.
 

### 2022-02-04 src/ms.tex
cleaning up
JVS comments

@@
 
-%The final sample contains 10516 unique targets with well-determined \prot and \teff (having a median error of 24~K).
-We additionally selected sources with 3000~K~$<~\teff~<$~8000~K, 3~$<$~\logg~$<$~5~dex, and -2~$<$~[Fe/H]~$<$~2~dex.
-
 \subsection{APOGEE--Kepler}
@@
 
-%, and all combined data sets of the long-period pile-up, 
-% In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.
-
 The constant Rossby model provides a reasonably good description of the \lamostmcq long-period edge in the \teff range of $\approx$~5000--6250~K, with fractional residuals $\lesssim5\%$ over this range.
@@
 The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~<~2 (blue) and Ro~>~2 (orange).
-Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel b relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
+Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel (b) relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
     \label{fig:ages}
@@
 Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience slowed spin evolution between $\sim$100~Myr and $\sim$200~Myr.
-This picture is compatible with open cluster studies which show multiple cluster sequences of Sun-like stars overlapping at ages of $\sim$100--200~Myr, with spin-down resuming by $\sim$200--300~Myr \citep{Fritzewski2020,Fritzewski2021}.
-However, an alternative explanation for the overlapping cluster sequences observed by those authors could be that the clusters are in fact coeval.
 
@@
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
-
-%The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
 
@@
 
-%One possibility is that the Sun evolved through a phase of weakened magnetic braking in the past and its subsequent period-lengthening was driven by its expansion away from the main sequence \citep[e.g.][]{vanSaders2019}.
-However, this scenario is considered implausible as the change in moment of inertia for the Sun over its lifetime is not enough to produce the 5--7 day gap between the Sun and the observed long-period pile-up.
-
-%Another possibility is that the long-period pile-up is only present for stars with masses $\gtrsim$~1~\msun, in which case the Sun's period and age require no explanation.
-However, in this scenario we would have no explanation for why hotter stars pile-up while cooler stars do not, especially given that the pile-up appears to occur at low enough Ro that the cooler stars should appear to pile up as well.
-
-%As shown in Figure~\ref{fig:xmatch}, it appears that the long-period pile-up does extend to temperatures as low as the Sun's, though this is not apparent from smaller samples such as the CKS sample or the \citet{Hall2021} asteroseismic sample (see also \S\ref{subsec:extent}).
-Kernel density estimation of the joint \teff--\prot distribution in the \lamostmcq sample, as well as a cross-match between LAMOST and \citet{Santos2021}, also suggests the pile-up exists at \teff as low as 5500~K or even 5000~K.
-
-%We thus reject the explanation that WMB only operates for $\gtrsim$~1~\msun stars.
-
 We consider the most likely explanation for the observed gap between the Sun and the long-period pile-up to be systematic offsets in \teff.
@@
 We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
-
-%If this is the case, there is some moderate tension with the ages of pile-up stars in the CKS sample (Figure~\ref{fig:ages}).
-In the WMB model, hotter stars are expected to depart the pile-up on shorter timescales, yet some CKS stars in the pile-up appear to be both hotter and older than the Sun.
-It is of course possible that the isochrone ages are inaccurate, particularly for the cooler stars in the sample.
-If the Sun has indeed evolved through a period of WMB, this implies that the Sun may have a fundamentally different dynamo mechanism to the stars that have not yet transitioned through WMB.
-This suggests that the Sun's dynamo and consequent manifestations of magnetic activity may not be directly comparable or analogous to the majority of stars with measured Kepler rotation periods (as the majority have not yet transitioned through WMB).
-This may have implications for the validity of using the Sun as a benchmark calibrator and exemplar in stellar rotation and magnetism models.
-
-%Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up.
-Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required.
-To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K.
-
-%Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}).
-If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out.
-One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021a}.
-Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question.
-
-%The Sun's relationship to the long-period pile-up is also relevant to ongoing research into the question of whether the Sun is in the middle of a magnetic transition, as \citet{Metcalfe2016} suggested based on the Sun's unusual magnetic cycle period,
 

### 2022-02-04 src/ms.tex
JVS comments

@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
-
-\textbf{The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling.
-The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling.
-Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling.
-In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability.
-After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there.
-A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.
-We note that this is a speculative scenario and the underlying physics are not known.}
-
-%JVS: I am confused by this picture.
-Why would the slow in apparent spin down make the photometric variability drop?
-The variability doesn't care about the spin-down rate, it cares about the instantaneous rotation (and maybe differential rotation).
-In this picture you are still stuck without a reason for why a more slowly rotating star should be more active than a more rapidly rotating one.
-The information you add here doesn't allow you to support or reject the speculation in Gordon 2021, so I would vote to make this speculation more modest, saying that the gap is perhaps an epoch of rapid spin down and leave it at that.
+\citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
+We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here.
+Notably, if the short-period pile-up is indeed due to core-envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
 

### 2022-02-04 src/ms.tex
abstract revision

@@
 %% Reintroduced the \received and \accepted commands from AASTeX v5.2
-%\received{\today}
+\received{February 4, 2022}
 \revised{\today}
@@
 \begin{abstract}
-We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin.
-The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
-Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age.
-In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
-These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin.
-Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the joint temperature–period distribution, which is also well-described by a curve of constant Rossby number.
-The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling.
-The same mechanism was previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
+We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic temperatures to demonstrate the existence of pile-ups at the long-period and short-period edges of the temperature--period distribution for main-sequence stars with temperatures exceeding $\sim$\teffmin.
+The long-period pile-up is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
+The long-period pile-up was predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
+Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but weakly predictive of its age.
+The short-period pile-up, which is also well-described by a curve of constant Rossby number, is not a prediction of the weakened magnetic braking hypothesis but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling.
+The same mechanism was proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
 The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.

### 2022-02-10 src/ms.tex
caption typo fix

@@
     \includegraphics[width=\linewidth]{./figures/mcmc.pdf}
-    \caption{\textit{Top:} Gaussian kernel density estimation of the LAMOST--McQuillan sample and the 10th and 90th \prot percentiles (which and black points, respectively), computed as described in \S\ref{subsec:rossby}.
+    \caption{\textit{Top:} Gaussian kernel density estimation of the LAMOST--McQuillan sample and the 10th and 90th \prot percentiles (white and black points, respectively), computed as described in \S\ref{subsec:rossby}.
 \textit{Middle:} The same 10th and 90th percentile curves shown in relation to constant Rossby curves.

### 2022-03-17 src/ms.tex
referee comments

@@
 %% using aastex version 6.3
-\documentclass[linenumbers,tighten,trackchanges,twocolumn]{aastex631}
+\documentclass[trackchanges,twocolumn]{aastex631}
 
@@
 CKS spectra were acquired with the Keck/HIRES spectrograph \citep{Vogt1994} and spectroscopic parameters were determined by averaging parameters from the SpecMatch pipeline \citep{Petigura2015} and SME@XSEDE, a Python implementation of the Spectroscopy Made Easy pipeline \citep{Valenti1996}.
-The internal errors on \teff from the CKS catalog are 60~K.
+\edit1{The internal (relative) errors on \teff from the CKS catalog are estimated at $\pm$60~K, with systematic errors of $\pm$100~K estimated from comparison to other catalogs \citep[see Table 7 of][]{Petigura2017}}.
 The metallicity distribution of the CKS sample is centered near solar, with a mean and standard deviation of +0.03~dex and 0.18~dex, respectively.
@@
 \subsection{Where do the pile-ups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
-While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \textbf{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
+\textbf{The number density of stars on the short-period pile-up similarly declines towards cooler \teff.
+If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
+However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and LAMOST (or whatever spectroscopic survey \teff is gathered from).} While it not clear whether or not these declines are astrophysical in nature, the result of \textbf{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
 We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
@@
 We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
-The relative fraction of stars on the short-period edge declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
+The relative fraction of stars on the short-period \textbf{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
 Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.
@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
-\citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
+\citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core–envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
 We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here.
-Notably, if the short-period pile-up is indeed due to core-envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
+Notably, if the short-period pile-up is indeed due to core–envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
 
@@
 The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
-When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core-envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
+When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core–envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
 
@@
 This research has made use of NASA's Astrophysics Data System Bibliographic Services.
-
 \end{acknowledgments}
@@
           \texttt{seaborn} \citep{seaborn},
-          \texttt{showyourwork!} \citep{Luger2021c}
-          }
+          \texttt{showyourwork!} \citep{Luger2021c}}
 

### 2022-03-17 src/ms.tex
referee comments

@@
 If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
-However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and LAMOST (or whatever spectroscopic survey \teff is gathered from).} While it not clear whether or not these declines are astrophysical in nature, the result of \textbf{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
+However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
+LAMOST).} While it not clear whether or not these declines are astrophysical in nature, the result of \textbf{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
 We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.

### 2022-03-17 src/ms.tex
adding CMD figure
paper edits

@@
 The \citet{Brewer2018} study, referred to here as SPOCS, also published elemental abundances and ages from isochrone fitting for the CKS sample.
-We additionally cross-matched the CKS catalog with the LAMOST--Kepler catalog \citep{Xiang2019} which is described further in \S\ref{subsec:lamost}.
+We additionally cross-matched the CKS catalog with the \edit1{LAMOST DR5} catalog \citep{Xiang2019} which is described further in \S\ref{subsec:lamost}.
 We compare the \teff--\prot distributions of the CKS sample using \teff and rotation periods from a variety of sources in Appendix~\ref{app:teffprot}.
 
-\subsection{LAMOST--Kepler} \label{subsec:lamost}
-The LAMOST--Kepler project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
+\subsection{\edit1{LAMOST DR5}} \label{subsec:lamost}
+The LAMOST project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
 The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
-We cross-matched the LAMOST--Kepler stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
+We cross-matched the \edit1{LAMOST DR5} stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
 
@@
 The periods of the stars in the cluster above the ridge appear to be just as reliable as those on the ridge.
-Interestingly, a similar clustering of points is not observed in the LAMOST--Kepler sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
+Interestingly, a similar clustering of points is not observed in the \edit1{\lamostmcq} sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
 Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pile-up.
@@
 
-We compared the \teff--\prot distribution of the CKS and LAMOST--Kepler samples with the theoretical predictions of \citet{vanSaders2019}.
+We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}.
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
@@
 While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
-We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.
+We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \edit1{\lamostmcq} may be slightly different.
 
@@
 \citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the LAMOST--Kepler sample while such an offset is either absent or not as apparent relative to the CKS sample.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
 Such an offset could be due to differing spectroscopic temperature scales between the two studies.
@@
     \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left) and LAMOST--Kepler samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
+    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
 Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
@@
 
-We tested the hypothesis that the long-period pile-up observed in the LAMOST--Kepler sample is due to WMB by fitting constant Rossby models to the \prot boundary.
+We tested the hypothesis that the long-period pile-up observed in the \edit1{\lamostmcq} sample is due to WMB by fitting constant Rossby models to the \prot boundary.
 For a given \teff, this model predicts \prot as:
@@
 We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
-We show these curves in relation to the full LAMOST--Kepler sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
+We show these curves in relation to the full \edit1{\lamostmcq} sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
 
@@
 We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.
+
+\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \citet{Hall2021} asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
+As expected, the stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass stars on the main-sequence.
+The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
+CKS stars on the long-period pile-up clearly occupy a well-defined region of the main-sequence, which overlaps well with the \citet{Hall2021} asteroseismic sample.}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{./figures/cmd.pdf}
+    \caption{Color-magnitude diagram.}
+    \label{fig:cmd}
+\end{figure}
 
@@
 
-In Figure~\ref{fig:teffscales} we compare temperatures between the LAMOST--Kepler catalog and temperatures from other surveys.
+In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
 We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
@@
 \begin{acknowledgments}
-We thank Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, and Karl Jaehnig for plotting advice.
+We thank the anonymous referee for a thoughtful review, Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, and Karl Jaehnig for plotting advice.
 It is a pleasure to thank the Stars Group at the American Museum of Natural History, and the Astronomical Data Group at the Flatiron Institute for helpful discussions.

### 2022-03-17 src/figures/cmd.py

@@ -110,7 +110,7 @@ for i in range(2):
     #change the marker size manually for both lines
     #lgnd.legendHandles[0]._legmarker.set_markersize(6)
     #lgnd.legendHandles[1]._legmarker.set_markersize(6)
-    lgnd.legendHandles[0]._sizes = [30]
-    lgnd.legendHandles[1]._sizes = [30]
+    #lgnd.legendHandles[0]._sizes = [30]
+    #lgnd.legendHandles[1]._sizes = [30]
     
 plt.savefig('../figures/cmd.pdf')

### 2022-03-17 src/ms.tex
paper edits

@@
 All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data.
-For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys.
+For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys \edit1{(Figure~\ref{fig:surveys})}.
 
@@
 
-\subsection{\edit1{LAMOST DR5}} \label{subsec:lamost}
+\subsection{\edit1{LAMOST} \label{subsec:lamost}}
 The LAMOST project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
@@
 
-\subsection{APOGEE--Kepler}
+\subsection{\edit1{APOGEE}}
 The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV;][]{Blanton2017}.
@@
 
-The location of the short-period pile-up is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+The location of the \edit1{long}-period pile-up is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
 As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
-Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\lesssim$~\teffmin have already piled up onto the edge by or before this timescale.
+Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
 Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period edge.
@@
 
-\begin{figure}
+\begin{figure*}
     \includegraphics[width=\linewidth]{./figures/cmd.pdf}
@@
     \label{fig:cmd}
-\end{figure}
+\end{figure*}
 
@@
 \subsection{Where do the pile-ups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \textbf{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
-\textbf{The number density of stars on the short-period pile-up similarly declines towards cooler \teff.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
+\edit1{The number density of stars on the short-period pile-up similarly declines towards cooler \teff.
 If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
 However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
-LAMOST).} While it not clear whether or not these declines are astrophysical in nature, the result of \textbf{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
+LAMOST).} While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
 We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
@@
 We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
-The relative fraction of stars on the short-period \textbf{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
+The relative fraction of stars on the short-period \edit1{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
 Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.
@@
 
-    \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity is less prominent than the long-period overdensity likely because the short-period pile-up is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pile-up ($>10^9$~yr).
+    \item We \edit1{tentatively detect} a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
+This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly} because the short-period pile-up is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pile-up ($>10^9$~yr).
 The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.

### 2022-03-17 src/ms.tex
CMD edits

@@
     \includegraphics[width=\linewidth]{./figures/cmd.pdf}
-    \caption{Color-magnitude diagram.}
+    \caption{Color-magnitude diagram.
+The two panels are identical with exception of the orange points.
+In the background, a 2d histogram of the Kepler field, where darker values represent a higher density of stars, up to a threshold beyond which the highest density is saturated (shown in white).
+Shown by orange points are stars with photometric rotation period detections from either the \citet{McQuillan2014} or \citet{Santos2021} catalogs, on the left and right, respectively.
+The CKS long-period pile-up stars identified in this work are shown by black points.
+A Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by black contours.}
     \label{fig:cmd}

### 2022-03-17 src/figures/cmd.py

@@ -4,10 +4,12 @@ import matplotlib as mpl
 from astropy.table import Table
 import pandas as pd
 import seaborn as sns
+import scipy.stats as st
 
 mpl.rcParams["figure.dpi"] = 150
 mpl.rcParams["savefig.bbox"] = "tight"
 mpl.rcParams["savefig.dpi"] = 300
+mpl.rcParams["legend.markerscale"] = 10
 
 sns.set(
         context="paper",
@@ -81,7 +83,7 @@ bprp = dr2['dr2_bp_rp']
 MG   = dr2['MG']
 arg  = np.isfinite(bprp) & np.isfinite(MG)
 
-h_kws = {"bins":400, "cmap": "Blues", "cmin": 0, "cmax":5, "density":False}
+h_kws = {"bins":200, "cmap": "Blues_r", "cmin": 1, "cmax":100, "density":False}
 
 phot_kws = {"ms": 0.1,
             "color": "orange",
@@ -98,19 +100,33 @@ for i in range(2):
     axes[i].set_xlabel(r'G$_\mathregular{BP}$-G$_\mathregular{RP}$ [mag]')
     axes[i].set_ylabel(r'M$_\mathregular{G}$ [mag]')
 
-axes[0].plot(bprp[mcq_rot], MG[mcq_rot], '.', label='photometric periods\n(McQuillan et al. 2014)', **phot_kws)
-axes[1].plot(bprp[san_rot], MG[san_rot], '.', label='photometric periods\n(Santos et al. 2021)', **phot_kws)
+axes[0].plot(bprp[mcq_rot], MG[mcq_rot], ',', **phot_kws)
+axes[1].plot(bprp[san_rot], MG[san_rot], ',', **phot_kws)
+
+#Hack for legend troubles
+axes[0].plot(bprp[mcq_rot].iloc[0], MG[mcq_rot].iloc[0], '.', ms=1, color='orange', alpha=1, label='photometric periods\n(McQuillan et al. 2014)')
+axes[1].plot(bprp[san_rot].iloc[0], MG[san_rot].iloc[0], '.', ms=1, color='orange', alpha=1, label='photometric periods\n(Santos et al. 2021)')
 
 for i in range(2):        
     
     axes[i].plot(cks['gaia_bp_rp'][ridge], cks['MG'][ridge], 'k.', ms=1, label='long-period pile-up (this work)', zorder=999)
[…]

### 2022-03-17 src/ms.tex
CMD edits

@@
 The two panels are identical with exception of the orange points.
-In the background, a 2d histogram of the Kepler field, where darker values represent a higher density of stars, up to a threshold beyond which the highest density is saturated (shown in white).
+In the background, a 2-d histogram of the Kepler field, where lighter shades represent a higher number of stars.
 Shown by orange points are stars with photometric rotation period detections from either the \citet{McQuillan2014} or \citet{Santos2021} catalogs, on the left and right, respectively.

### 2022-03-17 src/figures/cmd.py

@@ -126,6 +126,10 @@ for i in range(2):
 
     cset = axes[i].contour(xx, yy, f, colors='k', levels=4, linewidths=0.5, zorder=998)
     #cset.collections[0].set_label('asteroseismic periods\n(Hall et al. 2021)')
+
+    #Hack for asteroseismic legend label
+    axes[i].plot([-100,-99],[12,12],'k',label='asteroseismic periods\n(Hall et al. 2021)')
+
     lgnd = axes[i].legend(loc='lower left', prop={'size':8})
     
 sns.despine()

### 2022-03-18 src/ms.tex
minor edits

@@
 We first noted a pile-up of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}).
+Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (\edit1{see Appendix~\ref{app:teffprot}}).
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
@@
 
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/ages.pdf}
+    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the \lamostmcq sample (b).
+Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
+The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~$<$~2 (blue) and Ro~$>$~2 (orange).
+Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel (b) relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
+    \label{fig:ages}
+\end{figure*}
+
 \edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \citet{Hall2021} asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
@@
     \label{fig:cmd}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/ages.pdf}
-    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the \lamostmcq sample (b).
-Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
-The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~<~2 (blue) and Ro~>~2 (orange).
-Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel (b) relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
-    \label{fig:ages}
 \end{figure*}

### 2022-03-18 src/ms.tex
minor edits

@@
 %% using aastex version 6.3
-\documentclass[trackchanges,twocolumn]{aastex631}
+\documentclass[linenumbers,trackchanges,twocolumn]{aastex631}
 
@@
 \subsection{California--Kepler Survey} \label{subsec:cks}
-The CKS project gathered high-resolution spectroscopy for 1305 Kepler planet host stars \citep{Petigura2017}.
+The CKS project acquired high-resolution spectroscopy for 1305 Kepler planet host stars \citep{Petigura2017}.
 CKS spectra were acquired with the Keck/HIRES spectrograph \citep{Vogt1994} and spectroscopic parameters were determined by averaging parameters from the SpecMatch pipeline \citep{Petigura2015} and SME@XSEDE, a Python implementation of the Spectroscopy Made Easy pipeline \citep{Valenti1996}.
@@
 Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs.
-\lamostmcq samples (see Fig.~\ref{fig:percentiles}).
+\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pile-up and the \lamostmcq 90th percentile curve, we observe a systematic trend in which most CKS pile-up stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.}}
 
@@
 \edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \citet{Hall2021} asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
-As expected, the stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass stars on the main-sequence.
+As expected, stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs.
 The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.

### 2022-03-18 src/ms.tex
adding harmonic figure

@@
 
+\subsection{Detection biases}
+\edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
+As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
+This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
+\citet{vanSaders2019} addressed this issue in detail...}
+
 \subsection{Comparison with theory}
@@
 \citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & -29 $\pm$ 79 \\
-%All samples & 5000~K--6250~K & 1.47 $\pm$ 0.05 & 0.132 $\pm$ 0.008 & 2 $\pm$ 23 (LAMOST) \\
-% & & & & 10 $\pm$ 15 (CKS) \\
-% & & & & 61 $\pm$ 21  (Hall) \\
-%All samples & 5800~K--6250~K & 1.67 $\pm$ 0.08 & 0.116 $\pm$ 0.009 & -42 $\pm$ 24 (LAMOST) \\
-%&  &  &  & -44 $\pm$ 20 (CKS) \\
-%&  &  &  & 3 $\pm$ 23 (Hall) \\ 
 \enddata
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{./figures/harmonic.pdf}
+    \caption{}
+    \label{fig:harmonic}    
+\end{figure}
+
 \begin{acknowledgments}

### 2022-03-18 src/figures/harmonic.py

@@ -1,0 +1,151 @@
+import scipy.stats as st
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+import numpy as np
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+import seaborn as sns
+sns.set_context("paper")
+
+from astropy.table import Table
+import astropy.units as u
+import astropy.constants as c
+from astroquery.xmatch import XMatch
+import pandas as pd
+
+#McQuillan et al. 2014
+mcq = pd.read_parquet('../data/mcquillan2014_table1.parquet')
+
+#Gaia-Kepler cross-match from Megan Bedell
+gk = pd.read_parquet('../data/kepler_dr2_1arcsec.parquet')
+
+mcq = mcq.merge(gk, how="left", left_on="mcq_KIC", right_on="kepid")
+#np.shape(mcq)
+mcq = mcq.sort_values(by=["kepid", "kepler_gaia_ang_dist"])
+mcq = mcq.drop_duplicates(subset=['kepid'], keep='first')
+
+def lamost_xmatch(df):
+    
+    df_pos = df[['ra', 'dec']].copy()
+    #df_pos.head()
+    df_pos.to_csv('../data/coords.csv', index=False)
+    
+    table = XMatch.query(cat1=open('../data/coords.csv'),
+                         cat2='vizier:J/ApJS/245/34/catalog',
+                         max_distance=1 * u.arcsec,
+                         colRA1='ra',
+                         colDec1='dec',
+                         colRA2='RAJ2000',
+                         colDec2='DEJ2000')
+
+    #type(table)
+    #print(table)
+    table = table.to_pandas()
+    
+    unq = np.unique(table['ra'], return_index=True, return_counts=True)
+
+    table['xmatch_count'] = np.zeros(len(table))
+
+    for i in range(len(table)):
+        arg = unq[0] == table['ra'].iloc[i]
+        table['xmatch_count'].iloc[i] = unq[2][arg]
+
+    #table = table.sort_values(by=["gaia_ra"])
+    table = table.merge(df, how='right', left_on='ra', right_on='ra')
[…]

### 2022-03-18 src/ms.tex
adding appendix

@@
 
+\section{Confounding scenarios}\label{app:harmonics}
+
+\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pile-up, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
+Figure~\ref{fig:harmonic} shows a Gaussian kernel density estimation of the \lamostmcq \teff–\prot distribution for stars with \logg~$>$~4.1.
+The short-period pile-up is observed in this sample, particularly for \teff~$>$~6000~K.
+We found through inspection a constant Rossby curve that traces the long-period pile-up.
+Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily threads the short-period pile-up, though they both bracket the feature seen through density estimation.
+We interpret this as evidence that the short-period pile-up is not simply due to measurement error, although we have not definitively ruled out that scenario.}
+
 \begin{figure}
@@
     \includegraphics[width=0.5\linewidth]{./figures/harmonic.pdf}
-    \caption{}
+    \caption{Gaussian kernel density estimation of the \teff–\prot distribution in the \lamostmcq sample for stars with \logg~$>$~4.1.
+A curve of constant Rossby number is shown by the solid line.
+The \prot/2 and \prot/3 harmonics of this curve are shown by the dashed and dotted lines, respectively.
+The short-period pile-up is seen near the short-period edge, particularly for \teff~$>$~6000~K.
+The short-period pile-up does not appear to be simply a harmonic of the long-period pile-up, as one might expect if the periods for the stars in the short-period pile-up were erroneously determined.}
     \label{fig:harmonic}    

### 2022-03-18 src/ms.tex
adding LAMOST reference

@@
 The LAMOST project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
-The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
+\edit1{A description of the LAMOST Kepler field observations is provided in \citet{deCat2015}.} The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
 We cross-matched the \edit1{LAMOST DR5} stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.

### 2022-03-21 src/ms.tex
adding to appendix

@@
 We found through inspection a constant Rossby curve that traces the long-period pile-up.
-Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily threads the short-period pile-up, though they both bracket the feature seen through density estimation.
+Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pile-up, though they both bracket the feature seen through density estimation.
 We interpret this as evidence that the short-period pile-up is not simply due to measurement error, although we have not definitively ruled out that scenario.}
+
+\edit1{We also note that the short-period pile-up is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
+The differences between the \citet{McQuillan2014} and \citet{Santos2021} catalogs may provide an explanation for this observation: the latter catalog contains a much larger number of detections at longer rotation periods.
+The higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pile-up appear weaker relative to the long-period pile-up, as expected since open cluster obsbervations suggest that the short-period pile-up can not be a long-lived feature.}
 

### 2022-03-21 src/ms.tex
appendix edits

@@
 \edit1{We also note that the short-period pile-up is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
-The differences between the \citet{McQuillan2014} and \citet{Santos2021} catalogs may provide an explanation for this observation: the latter catalog contains a much larger number of detections at longer rotation periods.
-The higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pile-up appear weaker relative to the long-period pile-up, as expected since open cluster obsbervations suggest that the short-period pile-up can not be a long-lived feature.}
+The differences between the \citet{McQuillan2014} and \citet{Santos2021} catalogs may provide an explanation for this observation.
+The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pile-up appear weaker relative to the long-period pile-up.
+This is as expected since open cluster obsbervations suggest that the short-period pile-up can not be a long-lived feature, while the long-period pile-up appears to be much longer lived.
+Consequently, the \citet{Santos2021} catalog may more accurately reflect the relative strengths of these two features.}
 

### 2022-03-21 src/ms.tex
detection bias discussion

@@
 As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
-This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
-\citet{vanSaders2019} addressed this issue in detail...}
+This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.}
+
+\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution.
+However, those authors presented two arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
+First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge.
+For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods).
+Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}
+
+\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
 

### 2022-03-21 src/ms.tex
detection bias discussion

@@
 As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
-This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.}
+This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
+This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here may depend sensitively on how detectability depends on Ro.
+For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space, then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
 
@@
 \edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
-To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
+We also note that the asteroseismic method of measuring rotation periods has separate observational biases, yet still seemingly yields a pile-up of stars at the long-period edge (as discussed in \S\ref{sec:asteroseismic}).}
 

### 2022-03-21 src/ms.tex
detection bias discussion

@@
 \edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution.
-However, those authors presented two arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
+In this scenario, the detection threshold sets the upper edge of the observed \teff–\prot distribution at high \teff, while the age of the galaxy sets the upper edge at low \teff (since low-mass stars have not yet had enough time to spin down to Ro$_\mathrm{thresh}$).
+However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
 First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge.
@@
 To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
-We also note that the asteroseismic method of measuring rotation periods has separate observational biases, yet still seemingly yields a pile-up of stars at the long-period edge (as discussed in \S\ref{sec:asteroseismic}).}
+Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge.
+These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
 

### 2022-03-21 src/ms.tex
referee comments

@@
 We first noted a pile-up of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (\edit1{see Appendix~\ref{app:teffprot}}).
+\edit1{The pile-up in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2014}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.

### 2022-03-21 src/ms.tex
referee comments

--- 2022-03-21 src/ms.tex
@@
 We first noted a pile-up of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (\edit1{see Appendix~\ref{app:teffprot}}).
+\edit1{The pile-up in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \citet{Santos2021} revealed that this pile-up is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up.
-As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up \edit1{(see Appendix~\ref{app:harmonics})}.
+As shown in \edit1{the middle panel of} Figure~\ref{fig:kde}, this secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+\edit1{The short-period pile-up is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
+the second panel of Figure~\ref{fig:surveys}, and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
 
@@
 
-\section{Comparison of spectroscopic temperature scales}\label{app:teff}
-
-In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
-We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
-A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
-    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
-    \label{fig:teffscales}
-\end{figure}
-
 \section{Confounding scenarios}\label{app:harmonics}
@@
     \label{fig:harmonic}    
+\end{figure}
+
+\section{Comparison of spectroscopic temperature scales}\label{app:teff}
+
+In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
+We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
+A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
+    \label{fig:teffscales}
 \end{figure}

### 2022-03-21 src/ms.tex
referee comments

@@
 \edit1{The short-period pile-up is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
-the second panel of Figure~\ref{fig:surveys}, and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
+the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
 
@@
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
-As we show in \S\ref{subsec:ages}, stars along the pile-up have a range of ages.
-Therefore, the observation that the edge approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the edge is well-separated from the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period edge.
+The observation that the \edit1{long-period pile-up} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
+Similarly, the \edit1{long-period pile-up clearly lies above} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period edge.
 These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
+\edit1{We show in \S\ref{subsec:ages} that stars along the pile-up have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
 

### 2022-03-21 src/ms.tex
re-arranging sections

@@
 We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up \edit1{(see Appendix~\ref{app:harmonics})}.
-As shown in \edit1{the middle panel of} Figure~\ref{fig:kde}, this secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
 \edit1{The short-period pile-up is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
@@
 The observation that the \edit1{long-period pile-up} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the \edit1{long-period pile-up clearly lies above} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period edge.
+Similarly, the \edit1{long-period pile-up clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pile-up}.
 These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
@@
 \subsection{Detection biases}
+\label{subsec:detectionbias}
 \edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
@@
 These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
-
-\subsection{Comparison with theory}
-\label{subsec:models}
-
-We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}.
-Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
-Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations.
-Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
-
-To quantify the degree of agreement between the theoretical models and observations we computed the 10th and 90th percentile \prot ranges of the standard and WMB models in overlapping \teff bins, analogous to how the upper and lower boundaries of the observed \prot distribution were found in \S\ref{subsec:rossby}.
-We computed the $\chi^2$ values between the observed upper edge and the 90th percentile ranges of the standard and WMB models, finding the WMB model is preferred with a $\Delta \chi^2 = 154$.
-Moreover, the WMB model better reproduces the slope of the observed long-period edge between 5300--6000~K (Figure~\ref{fig:percentiles}).
-
-While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
-We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \edit1{\lamostmcq} may be slightly different.
-
-Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
-In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
-the low-mass stars would be rotating more slowly than the model predictions).
-Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law.
-The models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
-A higher normalization factor would cause the low-mass stars to spin down more at fixed age.
-
-Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models.
-At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
-The core and envelope at these times are thus assumed to be decoupled.
-However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}.
-When this happens, angular momentum can be transferred from the core to the envelope at a rate comparable to the rate at which angular momentum is lost via magnetized winds.
-Consequently, so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
-This hypothesis is discussed further in \S\ref{subsec:shortperiod}.
-
-\begin{figure*}
-    \centering 
-    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
-    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
-Shown here are stars with \logg~$>$~4.1.
-Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
-The black symbol in each panel indicates the position of the Sun.}
-    \label{fig:models}
-\end{figure*}
-
-\begin{figure*}
-    \centering 
-    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks-shifted.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks-shifted.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost-shifted.pdf}
-    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost-shifted.pdf}
-    \caption{Same as Figure~\ref{fig:models} with constant \teff offsets applied to the data.
-Shifts of +116~K and +140~K are applied to the CKS and LAMOST \teff, respectively.
-The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pile-up (as described in \S\ref{subsec:rossby}).
-The LAMOST shift is derived in a similar fashion, with an additional +34~K offset which represents the systematic offset between CKS and LAMOST (see Appendix~\ref{app:teff}).}
-    \label{fig:shifted}
-\end{figure*}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/percentiles.pdf}
-    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
-Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
-    \label{fig:percentiles}
-\end{figure}
-
-\subsection{Comparison with asteroseismic rotation rates}
-\label{subsec:asteroseismic}
-
-\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
-Such an offset could be due to differing spectroscopic temperature scales between the two studies.
-To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019}.
-We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average.
-Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
-Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
-A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
-Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
-    \label{fig:asteroseismic}
-\end{figure}
 
@@
 \end{deluxetable*}
+
+\subsection{Comparison with theory}
+\label{subsec:models}
+
+We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}.
+Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
+The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
+Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
+The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
+Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
+The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
+
+To quantify the degree of agreement between the theoretical models and the \lamostmcq observations we computed the 10th and 90th percentile \prot ranges of the standard and WMB models in overlapping \teff bins, analogous to how the upper and lower boundaries of the observed \prot distribution were found in \S\ref{subsec:rossby}.
+We computed the $\chi^2$ values between the observed upper edge and the 90th percentile ranges of the standard and WMB models, finding the WMB model is preferred with a $\Delta \chi^2 = 154$.
+Moreover, the WMB model better reproduces the slope of the observed long-period edge between 5300--6000~K (Figure~\ref{fig:percentiles}).
+
+While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
+We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \edit1{\lamostmcq} may be slightly different.
+
+Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
+In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
+the low-mass stars would be rotating more slowly than the model predictions).
+Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law.
+The models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$.
+A higher normalization factor would cause the low-mass stars to spin down more at fixed age.
+
+Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models.
+At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
+The core and envelope at these times are thus assumed to be decoupled.
+However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}.
+When this happens, angular momentum can be transferred from the core to the envelope at a rate comparable to the rate at which angular momentum is lost via magnetized winds.
+Consequently, so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age.
+This hypothesis is discussed further in \S\ref{subsec:shortperiod}.
+
+\begin{figure*}
+    \centering 
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
+    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
+Shown here are stars with \logg~$>$~4.1.
+Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.
+The black symbol in each panel indicates the position of the Sun.}
+    \label{fig:models}
+\end{figure*}
+
+\begin{figure*}
+    \centering 
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost-shifted.pdf}
+    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost-shifted.pdf}
+    \caption{Same as Figure~\ref{fig:models} with constant \teff offsets applied to the data.
+Shifts of +116~K and +140~K are applied to the CKS and LAMOST \teff, respectively.
+The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pile-up (as described in \S\ref{subsec:rossby}).
+The LAMOST shift is derived in a similar fashion, with an additional +34~K offset which represents the systematic offset between CKS and LAMOST (see Appendix~\ref{app:teff}).}
+    \label{fig:shifted}
+\end{figure*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/percentiles.pdf}
+    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
+Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
+    \label{fig:percentiles}
+\end{figure}
+
+\subsection{Comparison with asteroseismic rotation rates}
+\label{subsec:asteroseismic}
+
+\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
+Such an offset could be due to differing spectroscopic temperature scales between the two studies.
+To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019}.
+We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average.
+Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
+    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
+Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
+A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
+Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
+    \label{fig:asteroseismic}
+\end{figure}
 

### 2022-03-21 src/ms.tex
referee comments

@@
 Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs.
-\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pile-up and the \lamostmcq 90th percentile curve, we observe a systematic trend in which most CKS pile-up stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.}}
+\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pile-up and the \lamostmcq 90th percentile curve, we observe in Figure~\ref{fig:percentiles} a systematic trend in which most CKS pile-up stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.
+This is consistent with there being a \teff-dependent systematic trend between the \teff scales of CKS and LAMOST, as discussed in Appendix~\ref{app:teff}.}}
 
@@
 Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
+The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
 
@@
 \edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \citet{Hall2021} asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
-As expected, stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs.
+Stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
 The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
-CKS stars on the long-period pile-up clearly occupy a well-defined region of the main-sequence, which overlaps well with the \citet{Hall2021} asteroseismic sample.}
+CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \citet{Hall2021} asteroseismic sample.}
 

### 2022-03-21 src/ms.tex
adding Gaia

@@
 
-\subsection{Detection biases}
-\label{subsec:detectionbias}
-\edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
-As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
-This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
-This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here may depend sensitively on how detectability depends on Ro.
-For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space, then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
-
-\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution.
-In this scenario, the detection threshold sets the upper edge of the observed \teff–\prot distribution at high \teff, while the age of the galaxy sets the upper edge at low \teff (since low-mass stars have not yet had enough time to spin down to Ro$_\mathrm{thresh}$).
-However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
-First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge.
-For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods).
-Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}
-
-\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
-To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
-Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge.
-These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
-
 \subsection{Constant Rossby model}
@@
 
+\subsection{Detection biases}
+\label{subsec:detectionbias}
+\edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
+As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
+This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
+This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here may depend sensitively on how detectability depends on Ro.
+For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space, then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
+
+\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution.
+In this scenario, the detection threshold sets the upper edge of the observed \teff–\prot distribution at high \teff, while the age of the galaxy sets the upper edge at low \teff (since low-mass stars have not yet had enough time to spin down to Ro$_\mathrm{thresh}$).
+However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
+First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge.
+For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods).
+Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}
+
+\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
+Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge.
+These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
+
 \subsection{Long-period behavior} \label{subsec:longperiod}
 
-As discussed in \S\ref{subsec:models} and \S\ref{subsec:rossby}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
+As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
 The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021} and we suggest the two features are one and the same.
 Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
+
 Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
@@
 The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
+
 It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.
@@
 We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.
+\edit1{As shown in \citet{See2021}, the \taucz relation of \citet{Amard2019} also lies above the \citet{Gunn1998} relation, so adopting the newer \citet{Amard2019} relation would also not alleviate the \rocrit discrepancy.}
 
@@
 
+\section{Gaia}\label{app:gaia}
+\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
+We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3}.
+We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves.
+For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators.}
+
 \begin{acknowledgments}
-We thank the anonymous referee for a thoughtful review, Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, and Karl Jaehnig for plotting advice.
+We thank the anonymous referee for a thoughtful review, Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, Karl Jaehnig for plotting advice, and Matteo Cantiello for helpful discussions.
 It is a pleasure to thank the Stars Group at the American Museum of Natural History, and the Astronomical Data Group at the Flatiron Institute for helpful discussions.

### 2022-03-22 src/ms.tex
referee report

@@
 Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
-For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25$~d with an upper error that encompasses the period of $\approx$~36~d measured near the poles from differential rotation studies \citep[][and references therein]{Thompson2003}.
-The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
+For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
+\edit1{While differential rotation studies have inferred polar rotation periods of $\approx$~36~d from helioseismology \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
+We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
 
@@
 \edit1{The pile-up in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \citet{Santos2021} revealed that this pile-up is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
+
 A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
@@
 This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.
+
 \citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core–envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/ridge.pdf}
-    \caption{The \teff-\prot plane for the CKS sample.
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\linewidth]{./figures/ridge.pdf}
+%     \caption{The \teff-\prot plane for the CKS sample.
 Point colors are scaled to the CKS ages determined from isochrone fitting.
@@
 The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
-    \label{fig:ridge}
-\end{figure*}
+%     \label{fig:ridge}
+% \end{figure*}
 
@@
 \begin{acknowledgments}
-We thank the anonymous referee for a thoughtful review, Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, Karl Jaehnig for plotting advice, and Matteo Cantiello for helpful discussions.
+We thank the anonymous referee for a thoughtful review, as well as Travis Metcalfe and Matteo Cantiello for helpful discussions.
+We also thank Rodrigo Luger for helpful discussions and assistance with the \texttt{showyourwork!} package, Adrian Price-Whelan for providing the APOGEE--Kepler cross-match catalog, and Karl Jaehnig for plotting advice.
 It is a pleasure to thank the Stars Group at the American Museum of Natural History, and the Astronomical Data Group at the Flatiron Institute for helpful discussions.

### 2022-03-22 src/figures/gap.py

@@ -21,7 +21,7 @@ import seaborn as sns
 sun = {"teff": 5772,
        "prot":25.4,
        "e_prot":0,
-       "E_prot":36-25.4
+       "E_prot":27-25.4
       }
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)

### 2022-03-22 src/figures/kde.py

@@ -21,7 +21,7 @@ from labellines import labelLine, labelLines
 sun = {"teff": 5772,
        "prot": 25.4,
        "e_prot": 25.4-24.5,
-       "E_prot": 36-25.4
+       "E_prot": 27-25.4
       }
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)

### 2022-03-22 src/figures/percentiles.py

@@ -19,7 +19,7 @@ import seaborn as sns
 sun = {"teff": 5772,
        "prot":26.09,
        "e_prot":26.09-24.5,
-       "E_prot":36-26.09
+       "E_prot":27-26.09
       }
 
 

### 2022-03-22 src/figures/shifted.py

@@ -21,7 +21,7 @@ import seaborn as sns
 sun = {"teff": 5772,
        "prot": 25.4,
        "e_prot": 25.4-24.5,
-       "E_prot": 36-25.4
+       "E_prot": 27-25.4
       }
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)

### 2022-03-22 src/figures/surveys.py

@@ -8,8 +8,9 @@ import astropy.constants as c
 from astropy.table import Table
 from astropy.table import join
 
-import matplotlib.pyplot as plt
 import matplotlib as mpl
+import matplotlib.pyplot as plt
+
 
 mpl.rcParams["figure.dpi"] = 100
 mpl.rcParams["savefig.bbox"] = "tight"
@@ -20,7 +21,7 @@ import seaborn as sns
 sun = {"teff": 5772,
        "prot": 25.4,
        "e_prot": 25.4-24.5,
-       "E_prot": 36-25.4
+       "E_prot": 27-25.4
       }
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
@@ -103,6 +104,7 @@ ax[2].scatter(apok["TEFF"], apok["mcq_Prot"], **sc_kws, s=0.5)
 ax[3].scatter(cks["cks_Teff"], cks["mcq_Prot"], **sc_kws, s=2)
 
 for i in range(4):
+
     ax[i].set_title(titles[i])
     ax[i].set_xlabel("Effective temperature [K]")        
     ax[i].set_xlim(7250,3250)

### 2022-03-22 src/figures/xmatch.py

@@ -75,7 +75,7 @@ sns.set(font_scale=1.3, context="paper", style="ticks")
 sun = {"teff": 5772,
        "prot": 25.4,
        "e_prot": 25.4-24.5,
-       "E_prot": 36-25.4
+       "E_prot": 27-25.4
       }
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)

### 2022-03-22 src/ms.tex
revising cmd

@@
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
+\edit1{For the Gaia color of the Sun in we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.82$ \citep{Casagrande2018}.}
 
@@
     \caption{Color-magnitude diagram.
-The two panels are identical with exception of the orange points.
-In the background, a 2-d histogram of the Kepler field, where lighter shades represent a higher number of stars.
-Shown by orange points are stars with photometric rotation period detections from either the \citet{McQuillan2014} or \citet{Santos2021} catalogs, on the left and right, respectively.
-The CKS long-period pile-up stars identified in this work are shown by black points.
-A Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by black contours.}
+In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell.
+In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
+In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
     \label{fig:cmd}
@@
 This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here may depend sensitively on how detectability depends on Ro.
-For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space, then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
+For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
 

### 2022-03-22 src/figures/cmd.py

@@ -6,10 +6,21 @@ import pandas as pd
 import seaborn as sns
 import scipy.stats as st
 
+import matplotlib.colors as colors
+
 mpl.rcParams["figure.dpi"] = 150
 mpl.rcParams["savefig.bbox"] = "tight"
 mpl.rcParams["savefig.dpi"] = 300
 mpl.rcParams["legend.markerscale"] = 10
+
+def truncate_colormap(cmap, minval=0.0, maxval=1.0, n=100):
+    new_cmap = colors.LinearSegmentedColormap.from_list(
+        'trunc({n},{a:.2f},{b:.2f})'.format(n=cmap.name, a=minval, b=maxval),
+        cmap(np.linspace(minval, maxval, n)))
+    return new_cmap
+
+cmap = plt.get_cmap('Greys_r')
+new_cmap = truncate_colormap(cmap, 0.2, 0.8)
 
 sns.set(
         context="paper",
@@ -83,7 +94,7 @@ bprp = dr2['dr2_bp_rp']
 MG   = dr2['MG']
 arg  = np.isfinite(bprp) & np.isfinite(MG)
 
-h_kws = {"bins":200, "cmap": "Blues_r", "cmin": 1, "cmax":100, "density":False}
+h_kws = {"bins":200, "cmap": new_cmap, "cmin": 1, "cmax":100, "density":False}
 
 phot_kws = {"ms": 0.1,
             "color": "orange",
@@ -113,7 +124,7 @@ axes[1].legend(loc='lower left', prop={'size':8})
 
 for i in range(2,3):        
     
-    axes[i].plot(cks['gaia_bp_rp'][ridge], cks['MG'][ridge], 'k.', ms=1, label='long-period pile-up (this work)', zorder=999)
+    axes[i].plot(cks['gaia_bp_rp'][ridge], cks['MG'][ridge], '.', color='C0', ms=1, label='long-period pile-up (this work)', zorder=999)
    
     #Represent the asteroseismic sample with Guassian KDE contours
     x, y = bprp[hall_ast], MG[hall_ast]
@@ -128,7 +139,7 @@ for i in range(2,3):
     kernel = st.gaussian_kde(values)
     f = np.reshape(kernel(positions).T, xx.shape)
 
-    cset = axes[i].contour(xx, yy, f, colors='k', levels=4, linewidths=0.5, zorder=998)
+    cset = axes[i].contour(xx, yy, f, colors='C0', levels=4, linewidths=0.5, zorder=998)
[…]

### 2022-03-22 src/ms.tex
revising cmd

@@
     \includegraphics[width=\linewidth]{./figures/cmd.pdf}
-    \caption{Color-magnitude diagram.
+    \caption{Gaia DR2 color-magnitude diagram.
 In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell.
@@
 However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
-LAMOST).} While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
+LAMOST).}
+
+While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
 We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.

### 2022-03-22 src/figures/cmd.py

@@ -25,7 +25,7 @@ new_cmap = truncate_colormap(cmap, 0.2, 0.8)
 sns.set(
         context="paper",
         style="ticks",
-        font_scale=1.2,
+        font_scale=1.4,
         palette="colorblind"
         )
 
@@ -124,7 +124,7 @@ axes[1].legend(loc='lower left', prop={'size':8})
 
 for i in range(2,3):        
     
-    axes[i].plot(cks['gaia_bp_rp'][ridge], cks['MG'][ridge], '.', color='C0', ms=1, label='long-period pile-up (this work)', zorder=999)
+    axes[i].plot(cks['gaia_bp_rp'][ridge], cks['MG'][ridge], '.', color='C0', ms=1, label='long-period pile-up\n(David et al. 2022)', zorder=999)
    
     #Represent the asteroseismic sample with Guassian KDE contours
     x, y = bprp[hall_ast], MG[hall_ast]
@@ -143,7 +143,7 @@ for i in range(2,3):
     #cset.collections[0].set_label('asteroseismic periods\n(Hall et al. 2021)')
 
     #Hack for asteroseismic legend label
-    axes[i].plot([-100,-99],[12,12],'k',label='asteroseismic periods\n(Hall et al. 2021)')
+    axes[i].plot([-100,-99],[12,12],'C0',label='asteroseismic periods\n(Hall et al. 2021)')
 
     lgnd = axes[i].legend(loc='lower left', prop={'size':8})
     

### 2022-03-23 src/ms.tex
referee comments
Gaia appendix

@@
 \citet{Skumanich1972} presented the first attempt to calibrate this age-rotation relationship using the rotation periods of Sun-like stars in open clusters with independently determined ages, finding a $P_\mathrm{rot} \propto t^{1/2}$ scaling, where $t$ is stellar age.
-In the intervening decades, observational determinations of stellar rotation periods among open cluster members revealed how stellar spin rates evolve in more detail, leading to the calibration of the so-called gyrochronology method \citep{Barnes2003, Barnes2007, Barnes2010, MamajekHillenbrand2008, Meibom2009}.
+In the intervening decades, observational determinations of stellar rotation periods among open cluster members revealed how stellar spin rates evolve in more detail, leading to the calibration of the so-called gyrochronology method \citep{Barnes2003, Barnes2007, Barnes2010, MamajekHillenbrand2008, Meibom2009, Angus2019}.
 
@@
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
-\edit1{For the Gaia color of the Sun in we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.82$ \citep{Casagrande2018}.}
+\edit1{For the color of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.82$ \citep{Casagrande2018}.}
 
@@
 The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
+\edit1{We note that in our comparisons we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
+The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
 
@@
 
-\section{Gaia}\label{app:gaia}
+\section{The Gaia Color–Period Plane}\label{app:gaia}
 \edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
@@
 We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves.
-For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators.}
+For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators.
+We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
+In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=0.5$ which is approximately the solar value on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{./figures/gaia.pdf}
+    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
+The top rows show 2–d histograms of the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.
+Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above.
+The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
+    \label{fig:gaia}
+\end{figure}
 
@@
           \texttt{corner} \citep{corner},
+          \texttt{dustmaps} \citep{dustmaps}, 
           \texttt{emcee} \citep{emcee2013, emcee2019},

### 2022-03-23 src/ms.tex
referee comments

@@
 
-\subsection{Comparison with asteroseismic rotation rates}
+\subsection{Comparison with \edit1{literature surveys}}
 \label{subsec:asteroseismic}
@@
 
-\begin{figure}
+\edit1{\citet{Masuda2021} inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from their precise stellar radii and spectroscopically determined \vsini.
+As with the asteroseismic sample, we compare the CKS and \lamostmcq samples with the results of those authors.
+We find that the long-period pile-up we observe appears to overlap with the \citet{Masuda2021} data.
+We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the \citet{Hall2021} and \citet{Masuda2021} studies.}
+
+\begin{figure*}
     \centering
@@
     \label{fig:asteroseismic}
-\end{figure}
+\end{figure*}
 
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
-However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
+\edit1{These ranges are consistent with the range of ages observed in the \citet{Hall2021} and \citet{Masuda2021} samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
 The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
@@
 Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge.
-These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
+These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.
+We do note, however, that we can not rule out the possibility that the long-period pile-up is wider in period space than we observe, with stars just above the pile-up being undetected through rotational brightness modulations.}
 
@@
 
+\section{Comparison of spectroscopic temperature scales}\label{app:teff}
+
+In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
+We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
+A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
+    \label{fig:teffscales}
+\end{figure}
+
 \section{Confounding scenarios}\label{app:harmonics}
@@
 
-\section{Comparison of spectroscopic temperature scales}\label{app:teff}
-
-In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
-We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
-A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
-    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
-    \label{fig:teffscales}
-\end{figure}
-
 \section{The Gaia Color–Period Plane}\label{app:gaia}
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{./figures/gaia.pdf}
+    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
     \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
-The top rows show 2–d histograms of the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.
+The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.
 Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above.
 The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
-    \label{fig:gaia}
+    \label{fig:gaia-mcquillan}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{./figures/gaia-santos.pdf}
+    \caption{The same as Figure~\ref{fig:gaia-mcquillan} but for the \citet{Santos2021} sample.}
+    \label{fig:gaia-santos}
 \end{figure}

### 2022-03-23 src/figures/asteroseismic.py

@@ -65,27 +65,48 @@ sns.set(font_scale=1.2, context="paper", style="ticks")
 
 mpl.rcParams["legend.markerscale"] = 2
 titles = ["California–Kepler Survey","LAMOST–McQuillan"]
-ebar_kws = {"fmt":"o", "lw":1, "ms": 4, "ecolor":"orange", "color":"orange", "alpha":1, "mec":"grey", "mew":0.5}
+ebar_kws = {"fmt":"o", "lw":1, "ecolor":"orange", "color":"orange", "alpha":1, "ms":4, "mec":"grey", "mew":0.5, "zorder":999}
 
 sc_kws = {"marker":".", "color":"C0", "rasterized":True, "alpha":1}
 
-fig, ax = plt.subplots(nrows=2,ncols=1,figsize=(4.5,6.5))
+fig, (ax1,ax2) = plt.subplots(nrows=2,ncols=2,figsize=(2*4.5,6.5))
 
-ax[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
-ax[1].scatter(lam['Teff_lam']+41, lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
+ax1[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
+ax1[1].scatter(lam['Teff_lam']+41, lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
 #Note: the 41 K offset is derived in the appendix
+
+ax2[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
+ax2[1].scatter(lam['Teff_lam'], lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
+
+
+#Masuda et al. 2021 data
+specteff = np.array([5996.3610916725,6163.47095871239,6351.57452764171])
+specrot= np.array([15.1398954901294,10.9922199999722,5.40119754702563])
+spectefferr = np.array([[81.4555633310,58.78236529041,103.289013296011],[103.289013296011,79.77606717985,282.15535339398]])
[…]

### 2022-03-23 src/ms.tex
referee comments

@@
     \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane.
+    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the \citet{Hall2021} main-sequence asteroseismic sample (top row) \edit1{and the \citet{Masuda2021} sample (bottom row)} in the \teff--\prot plane.
 Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
-A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
+A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper left panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
 Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}

### 2022-03-23 src/ms.tex
paper edits

@@
 \newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
-\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
+%\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
+\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~\mathrm{Ro_\odot}$\xspace}
 \newcommand{\lamostkep}{LAMOST--Kepler\xspace}
@@
 
-where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}:
+where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{CranmerSaar2011} \edit1{from a fit to the zero-age main-sequence stellar models of \citet{Gunn1998}}:
 
@@
 \end{multline}
+
+\edit1{For reference, this relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d}.
 
@@
 \begin{multline} \label{eq:3}
-    \ln{p} (y | T_\mathrm{eff}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
+    \ln{p} (y | T_\mathrm{eff}, T_\mathrm{sys}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
 \end{multline}
@@
 
-We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation.
+We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, \edit1{including} the \teff scale of the models used to derive the \taucz relation.
 Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 is compatible with all of the data sets if modest \teff systematics are present.
@@
 We note that the \rocrit values we found in \S\ref{subsec:rossby} are smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
-We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.
-\edit1{As shown in \citet{See2021}, the \taucz relation of \citet{Amard2019} also lies above the \citet{Gunn1998} relation, so adopting the newer \citet{Amard2019} relation would also not alleviate the \rocrit discrepancy.}
+We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which exacerbates the \rocrit mismatch mentioned above.
+\edit1{As shown in \citet{See2021}, the \taucz relation of \citet{Amard2019} also lies above the \citet{CranmerSaar2011} relation, so adopting the newer \citet{Amard2019} relation would also not alleviate the \rocrit discrepancy.}
 
@@
 We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
-In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=0.5$ which is approximately the solar value on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
+In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=0.5$ which is approximately the solar value ($\mathrm{Ro_\odot}=0.496$) on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
 

### 2022-03-23 src/ms.tex
paper edits

@@
 We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
-In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=0.5$ which is approximately the solar value ($\mathrm{Ro_\odot}=0.496$) on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
+In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
+
+\edit1{We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$.
+This break is not apparent in the Gaia–McQuillan sample.
+Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}.
+At present, we do not have a satisfactory explanation for this behavior.}
 
@@
     \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
-    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
+    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
 The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.

### 2022-03-24 src/ms.tex
adding Gaia scripts

@@
 The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge.
-In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up.
+In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up.
 Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting.
@@
 In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell.
+The estimated position of the Sun is indicated by the black point, based on the calibration of \citet{Casagrande2018}.
 In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.

### 2022-03-24 src/figures/cmd.py

@@ -13,6 +13,15 @@ mpl.rcParams["savefig.bbox"] = "tight"
 mpl.rcParams["savefig.dpi"] = 300
 mpl.rcParams["legend.markerscale"] = 10
 
+#Solar data
+import astropy.units as u
+import astropy.constants as c
+dpc_sun = c.au.to(u.pc).value
+mG_sun = -26.895
+MG_sun = mG_sun - 5.*(np.log10(dpc_sun) - 1.)
+bprp_sun = 0.818
+
+
 def truncate_colormap(cmap, minval=0.0, maxval=1.0, n=100):
     new_cmap = colors.LinearSegmentedColormap.from_list(
         'trunc({n},{a:.2f},{b:.2f})'.format(n=cmap.name, a=minval, b=maxval),
@@ -94,6 +103,7 @@ bprp = dr2['dr2_bp_rp']
 MG   = dr2['MG']
 arg  = np.isfinite(bprp) & np.isfinite(MG)
 
+sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
 h_kws = {"bins":200, "cmap": new_cmap, "cmin": 1, "cmax":100, "density":False}
 
 phot_kws = {"ms": 0.1,
@@ -106,6 +116,8 @@ fig,axes = plt.subplots(nrows=1,ncols=3,
 
 for i in range(3):
     cb1 = axes[i].hist2d(bprp[arg], MG[arg], **h_kws)
+    axes[i].plot(bprp_sun, MG_sun,  zorder=999, **sun_kws)
+    axes[i].plot(bprp_sun, MG_sun, 'k.', zorder=999)
     axes[i].set_xlim(-0.5,4)
     axes[i].set_ylim(13,-3)
     axes[i].set_xlabel(r'G$_\mathregular{BP}$-G$_\mathregular{RP}$ [mag]')

### 2022-03-24 src/figures/gaia-mcquillan.py

@@ -1,0 +1,229 @@
+from dustmaps.bayestar import BayestarWebQuery
+from dustmaps.config import config
+config.reset()
+
+from dustmaps.bayestar import BayestarWebQuery
+dustquery = BayestarWebQuery(version='bayestar2019')
+
+import numpy as np
+import astropy.units as units
+from astropy.coordinates import SkyCoord
+from astroquery.vizier import Vizier
+from astroquery.xmatch import XMatch
+
+def curtis_gyrochrone(bprp, kind):
+    bprp = np.array(bprp)
+    
+    if kind=='kepler': #Kepler lower envelope
+        bprp_min, bprp_max = 0.6, 2.1
+        coeff = [36.4756, -202.718, 414.752, -395.161, 197.800, -50.0287, 5.05738]
+        
+    elif kind=='pleiades-ro':
+        bprp_min, bprp_max = 0.6, 1.3
+        coeff = [37.068, -188.02, 332.32, -235.78, 60.395]
+
+    elif kind=='pleiades-quad':
+        bprp_min, bprp_max = 0.6, 1.3
+        coeff = [-8.467, 19.64, -5.438]
+        
+    elif kind=='praesepe':
+        bprp_min, bprp_max = 0.6, 2.4
+        coeff = [-330.810, 1462.48, -2569.35, 2347.13, -1171.90, 303.620, -31.9227]
+        
+    elif kind=='ngc6811':
+        bprp_min, bprp_max = 0.65, 1.95 
+        coeff = [-594.019, 2671.90, -4791.80, 4462.64, -2276.40, 603.772, -65.0830]
+        
+    elif kind=='ngc752':
+        bprp_min, bprp_max = 1.32, 2.24
+        coeff = [6.80, 5.63] 
+    
+    elif kind=='ngc6819+ruprecht147':
+        bprp_min, bprp_max = 0.62, 2.07
+        coeff = [-271.783, 932.879, -1148.51, 695.539, -210.562, 25.8119]
+        
+    prot  = np.array([np.sum([co*_bprp**i for i,co in enumerate(coeff)]) for _bprp in bprp])
+    mask  = (bprp>=bprp_min) & (bprp<=bprp_max)
+    prot[~mask] = np.nan
+    
+    return prot
+
+gyro_sequences = ['pleiades-ro', 'praesepe', 'ngc6811', 'ngc6819+ruprecht147']
[…]

### 2022-03-24 src/figures/gaia-santos.py

@@ -1,0 +1,226 @@
+from dustmaps.bayestar import BayestarWebQuery
+from dustmaps.config import config
+config.reset()
+
+from dustmaps.bayestar import BayestarWebQuery
+dustquery = BayestarWebQuery(version='bayestar2019')
+
+import numpy as np
+import astropy.units as units
+from astropy.coordinates import SkyCoord
+from astroquery.vizier import Vizier
+from astroquery.xmatch import XMatch
+
+def curtis_gyrochrone(bprp, kind):
+    bprp = np.array(bprp)
+    
+    if kind=='kepler': #Kepler lower envelope
+        bprp_min, bprp_max = 0.6, 2.1
+        coeff = [36.4756, -202.718, 414.752, -395.161, 197.800, -50.0287, 5.05738]
+        
+    elif kind=='pleiades-ro':
+        bprp_min, bprp_max = 0.6, 1.3
+        coeff = [37.068, -188.02, 332.32, -235.78, 60.395]
+
+    elif kind=='pleiades-quad':
+        bprp_min, bprp_max = 0.6, 1.3
+        coeff = [-8.467, 19.64, -5.438]
+        
+    elif kind=='praesepe':
+        bprp_min, bprp_max = 0.6, 2.4
+        coeff = [-330.810, 1462.48, -2569.35, 2347.13, -1171.90, 303.620, -31.9227]
+        
+    elif kind=='ngc6811':
+        bprp_min, bprp_max = 0.65, 1.95 
+        coeff = [-594.019, 2671.90, -4791.80, 4462.64, -2276.40, 603.772, -65.0830]
+        
+    elif kind=='ngc752':
+        bprp_min, bprp_max = 1.32, 2.24
+        coeff = [6.80, 5.63] 
+    
+    elif kind=='ngc6819+ruprecht147':
+        bprp_min, bprp_max = 0.62, 2.07
+        coeff = [-271.783, 932.879, -1148.51, 695.539, -210.562, 25.8119]
+        
+    prot  = np.array([np.sum([co*_bprp**i for i,co in enumerate(coeff)]) for _bprp in bprp])
+    mask  = (bprp>=bprp_min) & (bprp<=bprp_max)
+    prot[~mask] = np.nan
+    
+    return prot
+
+gyro_sequences = ['pleiades-ro', 'praesepe', 'ngc6811', 'ngc6819+ruprecht147']
+gyro_ages = ['0.12 Gyr', '0.67 Gyr', '1 Gyr', '2.5 Gyr']
+
[…]

### 2022-03-24 src/figures/gap.py

@@ -147,10 +147,12 @@ cb2 = ax2.scatter(mcq["mcq_Teff"],
 for ax in [ax1,ax2]:
     _teff = np.linspace(3000,7000,1000)
     ax.plot(_teff, constant_rossby(_teff, 0.5), '--', color='k', lw=2, alpha=0.5)
-    ax.set_xlim(7000,3000)
     ax.set_ylim(0,50)  
     ax.set_xlabel("Effective temperature [K]")
     ax.set_ylabel("Rotation period [d]")    
+
+ax1.set_xlim(7000,4000)
+ax2.set_xlim(7000,3000)
     
 ax1.set_title('LAMOST–McQuillan')
 ax2.set_title('McQuillan et al. 2014')

### 2022-03-24 src/ms.tex
referee comments

@@
 \newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
+\newcommand{\rothresh}{$\mathrm{Ro_{thresh}}$\xspace}
 %\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
@@
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
-\edit1{For the color of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.82$ \citep{Casagrande2018}.}
+\edit1{For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.}
 
@@
 Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.
+\edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
 
@@
 Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above.
-The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
+The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.
+The grey lines indicate empirical cluster sequences derived in \citet{Curtis2020}.
+From top to bottom, the cluster sequences are Pleiades (0.12~Gyr), Praesepe (0.67~Gyr), NGC 6811 (1~Gyr), and NGC 6819 + Ruprecht 147 (2.5~Gyr).}
     \label{fig:gaia-mcquillan}

### 2022-03-24 src/figures/gaia-mcquillan.py

@@ -165,6 +165,7 @@ sun = {"teff": 5772,
 ####################################################################
 
 _bprp = np.linspace(0.55,0.97,1000)
+_bprp_cluster = np.linspace(0.6,2,1000)
 
 h_kws = {"bins":100, "cmap": "Blues", "density":True}
 kde_kws = {"fill": True, "bw_adjust": 0.25, "cmap": "Blues"}
@@ -218,7 +219,7 @@ for ax in zip(axes,axs):
         ax[i].legend()
 
         for j,seq in enumerate(gyro_sequences):
-            ax[i].plot(_bprp, curtis_gyrochrone(_bprp, kind=seq), label=gyro_ages[j], color='k', lw=1, alpha=0.25)
+            ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
 
 
 axes[0].set_title("All stars\n(McQuillan et al. 2014 + Gaia EDR3)")

### 2022-03-24 src/figures/gaia-santos.py

@@ -161,6 +161,7 @@ sun = {"teff": 5772,
 ####################################################################
 
 _bprp = np.linspace(0.55,0.97,1000)
+_bprp_cluster = np.linspace(0.6,2,1000)
 
 h_kws = {"bins":100, "cmap": "Blues", "density":True}
 kde_kws = {"fill": True, "bw_adjust": 0.25, "cmap": "Blues"}
@@ -215,7 +216,7 @@ for ax in zip(axes,axs):
         ax[i].legend()
         
         for j,seq in enumerate(gyro_sequences):
-            ax[i].plot(_bprp, curtis_gyrochrone(_bprp, kind=seq), label=gyro_ages[j], color='k', lw=1, alpha=0.25)
+            ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
 
 
 axes[0].set_title("All stars\n(Santos et al. 2021 + Gaia EDR3)")

### 2022-03-24 src/ms.tex
updating solar Gaia data

@@
 For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
-\edit1{While differential rotation studies have inferred polar rotation periods of $\approx$~36~d from helioseismology \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
+\edit1{While helioseismic differential rotation studies have inferred rotation periods of $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.

### 2022-03-24 src/figures/gaia-santos.py

@@ -156,7 +156,7 @@ sun = {"teff": 5772,
     "prot": 25.4,
     "e_prot": 25.4-24.5,
     "E_prot": 27-25.4,
-    "bprp": 0.82,
+    "bprp": 0.818,
     }
 ####################################################################
 

### 2022-03-25 src/ms.tex
minor edits

@@
 
-We tested the hypothesis that the long-period pile-up observed in the \edit1{\lamostmcq} sample is due to WMB by fitting constant Rossby models to the \prot boundary.
+We tested the hypothesis that the long-period pile-up observed in the \edit1{\lamostmcq} sample is \edit1{compatible with the} WMB \edit1{model} by fitting constant Rossby models to the \prot boundary.
 For a given \teff, this model predicts \prot as:
@@
     \caption{Gaia DR2 color-magnitude diagram.
-In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with lighter shades representing a higher number of targets in each cell.
+In each panel, a 2-d histogram of all Kepler targets is shown in grayscale, with darker shades representing a higher number of targets in each cell.
 The estimated position of the Sun is indicated by the black point, based on the calibration of \citet{Casagrande2018}.

### 2022-03-25 src/figures/cmd.py

@@ -28,7 +28,7 @@ def truncate_colormap(cmap, minval=0.0, maxval=1.0, n=100):
         cmap(np.linspace(minval, maxval, n)))
     return new_cmap
 
-cmap = plt.get_cmap('Greys_r')
+cmap = plt.get_cmap('Greys')
 new_cmap = truncate_colormap(cmap, 0.2, 0.8)
 
 sns.set(
@@ -104,7 +104,8 @@ MG   = dr2['MG']
 arg  = np.isfinite(bprp) & np.isfinite(MG)
 
 sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
-h_kws = {"bins":200, "cmap": new_cmap, "cmin": 1, "cmax":100, "density":False}
+#h_kws = {"bins":200, "cmap": new_cmap, "cmin": 1, "cmax":100, "density":False}
+h_kws = {"bins":200, "cmap": new_cmap, "cmin": 5, "density":False}
 
 phot_kws = {"ms": 0.1,
             "color": "orange",
@@ -118,7 +119,7 @@ for i in range(3):
     cb1 = axes[i].hist2d(bprp[arg], MG[arg], **h_kws)
     axes[i].plot(bprp_sun, MG_sun,  zorder=999, **sun_kws)
     axes[i].plot(bprp_sun, MG_sun, 'k.', zorder=999)
-    axes[i].set_xlim(-0.5,4)
+    axes[i].set_xlim(-0.8,3.5)
     axes[i].set_ylim(13,-3)
     axes[i].set_xlabel(r'G$_\mathregular{BP}$-G$_\mathregular{RP}$ [mag]')
     axes[i].set_ylabel(r'M$_\mathregular{G}$ [mag]')

### 2022-03-28 src/ms.tex
adding weights figure

@@
 To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
-The first model assumes \rocrit = 2 and has one free parameter, a temperature shift added to the data.
-The second model assumes no temperature shift and allows \rocrit to vary (\rocrit being the only free parameter).
+The first model assumes \rocrit = 2 and has one free parameter, a temperature offset added to the data.
+The second model assumes no temperature offset and allows \rocrit to vary (\rocrit being the only free parameter).
 The results of these fits are shown in Figure~\ref{fig:rocrit}.
@@
 \edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
-As rotation periods become longer, it becomes more difficult to measure periods directly from rotational brightness modulations in Kepler time series.
+As rotation periods become longer, they become more difficult to measure directly from rotational brightness modulations in Kepler time series.
 This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
@@
 
+\edit1{Finally, \citet{McQuillan2014} published weight parameters for each star, a metric which was designed to serve as a proxy for reliability, based on a star's autocorrelation function (ACF) properties and its position in \teff–\prot–LPH space, where LPH designates the local peak height of the ACF.
+Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
+In Figure~\ref{fig:weights}, we show that the long-period pile-up is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pile-up are likely to be genuine detections and not an artifact of some detection threshold.}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/weights.pdf}
+    \caption{The \teff–\prot distribution of the \lamostmcq sample for different weight thresholds (indicated above each panel) computed by \citet{McQuillan2014}.}
+    \label{fig:weights}    
+\end{figure*}
+
 \subsection{Long-period behavior} \label{subsec:longperiod}
@@
 As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021} and we suggest the two features are one and the same.
-Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
+The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021}, \edit1{as well as the peak of the period distribution inferred by \citet{Masuda2021} from combining \vsini and stellar radii.
+We suggest that these features are one and the same.} Offsets in both observational and theoretical \teff scales between different studies \edit1{may help} explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as \edit1{differences} in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
 
@@
 The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
+\edit1{The same argument is true for the \citet{Masuda2021} sample.}
 
@@
 This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values.
-Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.
-\edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that \edit1{a pile-up exists at the long-period edge} for stars with \teff~$\gtrsim$~\teffmin.
+\edit1{By construction, this pile-up occurs at a Rossby number lower than that associated with a detection edge.} \edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
 
@@
 We note that the \rocrit values we found in \S\ref{subsec:rossby} are smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
-We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which exacerbates the \rocrit mismatch mentioned above.
-\edit1{As shown in \citet{See2021}, the \taucz relation of \citet{Amard2019} also lies above the \citet{CranmerSaar2011} relation, so adopting the newer \citet{Amard2019} relation would also not alleviate the \rocrit discrepancy.}
+We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, \edit1{and \citet{Amard2019}}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which exacerbates the \rocrit mismatch mentioned above.
+\edit1{We also explored using a polynomial fit to the \taucz computed in \citet{vanSaders2019} for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
+However, even adopting this different \taucz prescription changes the inferred \rocrit upwards by only 15\%, not enough to obtain \rocrit~=~2.}
 

### 2022-03-28 src/figures/gaia-santos.py

@@ -218,7 +218,6 @@ for ax in zip(axes,axs):
         for j,seq in enumerate(gyro_sequences):
             ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
 
-
 axes[0].set_title("All stars\n(Santos et al. 2021 + Gaia EDR3)")
 axes[1].set_title("Low reddening sample\n(A$_\mathregular{V}$<0.2 mag)")
 axes[2].set_title("Low reddening sample\nwith -0.1 mag shift")

### 2022-03-28 src/figures/teffscales.py

@@ -218,6 +218,7 @@ ax8 = fig.add_subplot(spec[1,3])
 ax9 = fig.add_subplot(spec[0,4])
 ax10 = fig.add_subplot(spec[1,4])
 
+
 one_to_one(ax1, table["Teff"], table["cks_Teff"], 'LAMOST'+tstr, 'CKS'+tstr, xmin=xmin, xmax=xmax, sc_kws=sc_kws)
 residual(ax2, table["Teff"], table["cks_Teff"], 'LAMOST'+tstr, 'LAMOST-\nCKS [K]', xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, sc_kws=sc_kws)    
 

### 2022-03-28 src/figures/weights.py

@@ -1,0 +1,88 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+import numpy as np
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+import seaborn as sns
+sns.set_context("paper")
+
+from astropy.table import Table
+import astropy.units as u
+import astropy.constants as c
+from astroquery.xmatch import XMatch
+import pandas as pd
+
+
+#McQuillan et al. 2014
+mcq = pd.read_parquet('../data/mcquillan2014_table1.parquet')
+
+#Gaia-Kepler cross-match from Megan Bedell
+gk = pd.read_parquet('../data/kepler_dr2_1arcsec.parquet')
+
+mcq = mcq.merge(gk, how="left", left_on="mcq_KIC", right_on="kepid")
+#np.shape(mcq)
+mcq = mcq.sort_values(by=["kepid", "kepler_gaia_ang_dist"])
+mcq = mcq.drop_duplicates(subset=['kepid'], keep='first')
+
+def lamost_xmatch(df):
+    
+    df_pos = df[['ra', 'dec']].copy()
+    #df_pos.head()
+    df_pos.to_csv('../data/coords.csv', index=False)
+    
+    table = XMatch.query(cat1=open('../data/coords.csv'),
+                         cat2='vizier:J/ApJS/245/34/catalog',
+                         max_distance=1 * u.arcsec,
+                         colRA1='ra',
+                         colDec1='dec',
+                         colRA2='RAJ2000',
+                         colDec2='DEJ2000')
+
+    #type(table)
+    #print(table)
+    table = table.to_pandas()
+    
+    unq = np.unique(table['ra'], return_index=True, return_counts=True)
+
+    table['xmatch_count'] = np.zeros(len(table))
+
+    for i in range(len(table)):
+        arg = unq[0] == table['ra'].iloc[i]
+        table['xmatch_count'].iloc[i] = unq[2][arg]
+
+    #table = table.sort_values(by=["gaia_ra"])
+    table = table.merge(df, how='right', left_on='ra', right_on='ra')
+    table = table.sort_values(["ra","angDist"], ascending = (True, True))
[…]

### 2022-03-29 src/ms.tex
adding inflection figure

@@
 
-In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
+In both the CKS and LAMOST samples, there is a change of slope along the ridge (Figure~\ref{fig:inflection}), with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
 A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
@@
 
-In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in the top panels of Figure~\ref{fig:models}).
+In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in Figure~\ref{fig:inflection}).
 This cluster of points has a similar slope in the \teff--\prot plane as the long-period pile-up, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
@@
 We conclude that the temperatures for these stars in the CKS catalog are too high by $\gtrsim$~100~K.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/inflection.pdf}
+    \caption{The \teff–\prot plane for the \lamostmcq sample (grey points) and the CKS sample (blue points) with periods sourced from \citet{David2021}.
+The selection used to identify long-period pile-up stars in the CKS sample is shown by the region enclosed by the black lines.
+Stars with \teff differences $>100$~K between the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs are shown by open red circles.
+The orange point indicates the Sun.
+Random draws from the MCMC chain of a piecewise linear fit to the CKS long-period pile-up are shown by the blue lines.
+An inflection point in the long-period edge is apparent near \teff=~6250~K.}
+    \label{fig:inflection}
+\end{figure}
 
@@
 In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
-In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.}
+In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.
+No reddening corrections were performed to the photometry in this figure.}
     \label{fig:cmd}
@@
 
-\edit1{Finally, \citet{McQuillan2014} published weight parameters for each star, a metric which was designed to serve as a proxy for reliability, based on a star's autocorrelation function (ACF) properties and its position in \teff–\prot–LPH space, where LPH designates the local peak height of the ACF.
+\edit1{Finally, \citet{McQuillan2014} published weight parameters for each star, a metric which was designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in \teff–\prot–LPH space, where LPH designates the local peak height of the ACF.
 Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.

### 2022-03-29 src/figures/inflection.py

@@ -1,0 +1,131 @@
+import numpy as np
+import pandas as pd
+import emcee
+from scipy.optimize import curve_fit
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
+
+import seaborn as sns 
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+#CKS (Fulton & Petigura 2018)
+cks = pd.read_parquet('../data/cks_merged.parquet')
+
+#LAMOST-McQuillan
+lam = pd.read_csv('../data/kepler_lamost.csv')
+lam = lam.drop_duplicates(subset=['KIC'], keep='first')
+
+def ridge_hi(teff):
+    m = (2-24)/(6500-5800)
+    b = (2 - m*6500) 
+    return m*teff + b
+
+def ridge_lo(teff):
+    m = (2-24)/(6500-5800)
+    b = (-5 - m*6500) 
+    return m*teff + b
+
+mask = (cks['p20_cks_slogg']>4) #main sequence
+ridge = (cks['p20_cks_steff']>5850)
+ridge &= (cks['p20_cks_steff']<6500)
+ridge &= (cks['d21_prot']<ridge_hi(cks['p20_cks_steff']))
+ridge &= (cks['d21_prot']>ridge_lo(cks['p20_cks_steff']))
+ridge &= mask
+
+# Parallelogram
+xp = [6500,6500,5850,5850]
+yp = [ridge_lo(6500),ridge_hi(6500),ridge_hi(5850),ridge_lo(5850)]
+
+# Select stars on the pile-up
+x, y = np.array(cks['cks_Teff'][ridge]), np.array(cks['d21_prot'][ridge])
+yerr = np.ones(len(y))
+
+
+def piecewise_linear(x, x0, y0, k1, k2):
+    y = np.piecewise(x, [x < x0, x >= x0],
+                     [lambda x:k1*x + y0-k1*x0, lambda x:k2*x + y0-k2*x0])
+    return y
+
+def log_likelihood(theta, x, y, yerr):
+    x0, y0, k1, k2, f = theta
+    model = piecewise_linear(x, x0, y0, k1, k2)
+    sigma2 = yerr ** 2 + model ** 2 * f ** 2
+    return -0.5 * np.sum((y - model) ** 2 / sigma2 + np.log(sigma2))
+
[…]

### 2022-03-29 src/ms.tex
minor edits

@@
 \newcommand{\rothresh}{$\mathrm{Ro_{thresh}}$\xspace}
+\newcommand{\rosun}{$\mathrm{Ro_{\odot}}$\xspace}
 %\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
@@
 The results of these fits are shown in Figure~\ref{fig:rocrit}.
-We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = 2 and variable \rocrit models was negligible.
-In other words, the \rocrit = 2 model fits the data well with a \teff shift of $\approx$~80--150~K depending on the \teff source used.
+We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = \rosun and variable \rocrit models was negligible.
+In other words, the \rocrit = \rosun model fits the data well with a \teff shift of $\approx$~80--150~K depending on the \teff source used.
 For the \lamostmcq sample, there is weak support for the variable \rocrit model.
@@
 The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021}, \edit1{as well as the peak of the period distribution inferred by \citet{Masuda2021} from combining \vsini and stellar radii.
-We suggest that these features are one and the same.} Offsets in both observational and theoretical \teff scales between different studies \edit1{may help} explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as \edit1{differences} in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
+We suggest that these features are one and the same.
+The modest differences between the long-period pile-up we observe and the features observed by \citet{Hall2021} and \citet{Masuda2021} may be partially explained by differing \teff scales between the samples.}
+
+%Additionally, both observational and theoretical \teff scales between different studies \edit1{may help} explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as \edit1{differences} in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
 
@@
 \edit1{By construction, this pile-up occurs at a Rossby number lower than that associated with a detection edge.} \edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
+
+\edit1{We also note that the WMB model makes the simplistic assumption that magnetic braking ceases when $Ro$ reaches \rocrit.
+However, it is possible that magnetic braking becomes gradually weaker, a scenario which would also lead to a pile-up but may also cause such a pile-up to have some intrinsic width which is not due to measurement uncertainties.}
 

### 2022-03-29 src/figures/rocrit.py

@@ -19,6 +19,17 @@ sns.set(
         palette="colorblind"
         )
 
+####################################################################
+# The Sun
+sun = {"teff": 5772,
+       "prot": 25.4,
+       "e_prot": 25.4-24.5,
+       "E_prot": 27-25.4
+      }
+
+sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}      
+####################################################################
+
 
 ######################################################################################
 #McQuillan et al. 2013
@@ -77,7 +88,7 @@ def constant_rossby(teff, ro):
 
 def rocrit_teff_shift(teff, teff_shift):
     #Return locus of rotation periods corresponding to constant Rossby number
-    return 2 * convective_turnover_timescale(teff+teff_shift)
+    return 1.97 * convective_turnover_timescale(teff+teff_shift)
 
 lam["Ro"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"])    
 ######################################################################################
@@ -223,9 +234,10 @@ for i,teff in enumerate([cks['cks_Teff'],
     axes[i].errorbar(_xdata[_arg], _ydata[_arg], yerr = 0.1*_ydata[_arg], 
                      **ebar_kws)
     axes[i].plot(model_teff, _model, color='orange', lw=3, alpha=0.5,
-                 label=r'Ro$_\mathregular{crit}$ = 2 ($\mathregular{\Delta}$T$_\mathregular{eff}$ = '+str(int(_popt[0]))+')')
+                 label=r'Ro$_\mathregular{crit}$ = Ro$_\mathregular{\odot}$ ($\mathregular{\Delta}$T$_\mathregular{eff}$ = '+str(int(_popt[0]))+')')
     axes[i].plot(model_teff, __model, color='C2', lw=3, alpha=0.5,
                  label='Ro$_\mathregular{crit}$ = '+str(np.round(__popt[0],2))+' ($\mathregular{\Delta}$T$_\mathregular{eff}$ = 0)')
+
     axes[i].set_xlim(6400,5600)
     axes[i].set_ylim(-2,30)
     axes[i].text(0.05,0.05, 
@@ -253,7 +265,7 @@ axes[3].errorbar(x_lam_det, y_lam_det, yerr = 0.1*y_lam_det,
                  **ebar_kws)
 
[…]

### 2022-03-29 src/ms.tex
changes to Gaia appendix

@@
 \edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
-We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3}.
+We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
+We computed reddening estimates for each target using the \texttt{dustmaps} package \citep{dustmaps}.
 We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves.
@@
 \software{\texttt{astropy} \citep{astropy13, astropy18},
+          \texttt{astroquery} \citep{astroquery},
           \texttt{corner} \citep{corner},

### 2022-03-30 src/ms.tex
adding CKS data

@@
 We conclude that the temperatures for these stars in the CKS catalog are too high by $\gtrsim$~100~K.
+
+\edit1{Given that the long-period pile-up becomes apparent when adopting high-precision \teff from spectroscopic catalogs, it is also natural to consider the high-precision photometry of the Gaia mission \citep{Gaia2016}.
+While the Gaia photometry are susceptible to reddening, we show in Appendix~\ref{app:gaia} that the long-period pile-up can indeed be recovered in the color–period plane when restricting to stars with low interstellar reddening.}
 
@@
 We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
-We computed reddening estimates for each target using the \texttt{dustmaps} package \citep{dustmaps}.
+We computed reddening estimates for each target by querying the \texttt{Bayestar19} 3D dust map using the \texttt{dustmaps} package \citep{dustmaps, Green2019}.
 We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves.

### 2022-03-31 src/ms.tex
adding references

@@
 Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
-Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.\footnote{A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
+Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
+A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.
+\edit1{Additionally, \citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
 In contrast to the findings of those authors, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.

### 2022-03-31 src/ms.tex
revising Gaia appendix

@@
 We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
-Such an offset could be due to differing spectroscopic temperature scales between the two studies.
-To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019}.
-We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average.
-Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.
+Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of the two effects}.
+
+To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019} \edit1{and the CKS sample}.
+We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
+\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \citet{Hall2021} samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
+
+{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.}
 
@@
 This break is not apparent in the Gaia–McQuillan sample.
-Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}.
-At present, we do not have a satisfactory explanation for this behavior.}
+Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}, suggesting the origin of the discontinuity is in the rotation periods rather than the temperatures or colors.
+At present, we do not have a satisfactory explanation for this behavior, though we note that the \citet{Santos2021} catalog employed various time series filters which might introduce systematic artifacts.}
 

### 2022-03-31 src/ms.tex
paper edits

@@
 \newcommand{\rosun}{$\mathrm{Ro_{\odot}}$\xspace}
-%\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
 \newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~\mathrm{Ro_\odot}$\xspace}
@@
 \newcommand{\agemin}{$\sim$2--3~Gyr\xspace}
+
+% \newcommand{\hall}{H21\xspace}
+% \newcommand{\masuda}{MPH21\xspace}
+% \newcommand{\vs19}{vS19\xspace}
 
@@
 %% Adds "Submitted to " the argument.
-%\submitjournal{PSJ}
+\submitjournal{ApJ}
 
@@
 For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}.
-This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Rossby numbers of Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
+This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Rossby numbers of Ro~$\gtrsim$~\rosun experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.
 
@@
 
-More recently, \citet{Hall2021} used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \citet{Hall2021} sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
-\citet{Masuda2021} also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
+More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the H21 sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
+\citet{Masuda2021}, \edit1{hereafter MPH21}, also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
 While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which may vary with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
@@
 The \citet{McQuillan2014} \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars.
-The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the differentially rotating surface.
+The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the \edit1{activity belts}.
 Many of the stars above the long-period pile-up are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \citet{vanSaders2019}.}
@@
 For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
-\edit1{While helioseismic differential rotation studies have inferred rotation periods of $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
+\edit1{While helioseismic differential rotation studies have inferred rotation periods of up to $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
@@
 The effects of a simple cut in \logg to separate subgiants and dwarfs are shown in the middle and right columns.
-In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from its differentially rotating surface.
+In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from \edit1{the activity belts}.
 The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature.
@@
     \includegraphics[width=\linewidth]{./figures/kde.pdf}
-    \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic \citet{Hall2021} samples, from left to right.
+    \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic H21 samples, from left to right.
 Empirical cluster sequences from \citet{Curtis2020} are shown by the dark grey lines.
@@
 The short-period pile-up can be observed in the LAMOST--McQuillan sample for \teff~$\gtrsim$~5500~K.
-The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from its differentially rotating surface.}
+The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from \edit1{the activity belts}.}
     \label{fig:kde}
@@
 A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
-This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
+This change in slope \edit1{may be} due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
 
@@
 
-\edit1{Given that the long-period pile-up becomes apparent when adopting high-precision \teff from spectroscopic catalogs, it is also natural to consider the high-precision photometry of the Gaia mission \citep{Gaia2016}.
-While the Gaia photometry are susceptible to reddening, we show in Appendix~\ref{app:gaia} that the long-period pile-up can indeed be recovered in the color–period plane when restricting to stars with low interstellar reddening.}
+\edit1{In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pile-up in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
 
@@
 Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K.
-To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling.
-The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values.
-We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
+{\color{red} To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling.
+The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values.} We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
 We show these curves in relation to the full \edit1{\lamostmcq} sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
@@
 
-We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the \citet{Hall2021} asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
+We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the H21 asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
 For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
@@
 We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, \edit1{including} the \teff scale of the models used to derive the \taucz relation.
-Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 is compatible with all of the data sets if modest \teff systematics are present.
-Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
+Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
 
@@
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
+\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
+}.}}
+
 The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
@@
 The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
-Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
+{\color{red}Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
+The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff})}.
 \edit1{We note that in our comparisons we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
@@
 
-\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
+H21 determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the H21 sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
 Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of the two effects}.
 
-To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019} \edit1{and the CKS sample}.
-We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
-\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \citet{Hall2021} samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
+To investigate the origin of this offset, we cross-matched the H21 sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
+We found a root mean square (RMS) of the residuals between the H21 and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
+\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and H21 samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
 
@@
 
-\edit1{\citet{Masuda2021} inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from their precise stellar radii and spectroscopically determined \vsini.
+\edit1{MPH21 inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from their precise stellar radii and spectroscopically determined \vsini.
 As with the asteroseismic sample, we compare the CKS and \lamostmcq samples with the results of those authors.
-We find that the long-period pile-up we observe appears to overlap with the \citet{Masuda2021} data.
-We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the \citet{Hall2021} and \citet{Masuda2021} studies.}
+We find that the long-period pile-up we observe appears to overlap with the MPH21 data.
+We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the H21 and MPH21 studies.}
 
@@
     \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the \citet{Hall2021} main-sequence asteroseismic sample (top row) \edit1{and the \citet{Masuda2021} sample (bottom row)} in the \teff--\prot plane.
-Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines.
+    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the H21 main-sequence asteroseismic sample (top row) \edit1{and the MPH21 sample (bottom row)} in the \teff--\prot plane.
+Note, the CKS, LAMOST, and H21 samples derive \teff from distinct pipelines.
 A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper left panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
-Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
+Using LAMOST \teff, where it exists, for the H21 sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
-\edit1{These ranges are consistent with the range of ages observed in the \citet{Hall2021} and \citet{Masuda2021} samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
+\edit1{These ranges are consistent with the range of ages observed in the H21 and MPH21 samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
 The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
@@
 
-\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \citet{Hall2021} asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
+\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the H21 asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
 Stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
 The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
-CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \citet{Hall2021} asteroseismic sample.}
+CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the H21 asteroseismic sample.}
 
@@
 In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
-In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \citet{Hall2021} asteroseismic sample is shown by blue contours.
+In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the H21 asteroseismic sample is shown by blue contours.
 No reddening corrections were performed to the photometry in this figure.}
@@
 This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
-This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here may depend sensitively on how detectability depends on Ro.
+This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here depends sensitively on how detectability depends on Ro.
 For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
@@
 \edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
-To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
-Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge.
-These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.
-We do note, however, that we can not rule out the possibility that the long-period pile-up is wider in period space than we observe, with stars just above the pile-up being undetected through rotational brightness modulations.}
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
+
+\edit1{Furthermore, two independent samples with different observational biases than the surveys considered here also appear to yield a pile-up of stars at the long-period edge.
+These samples are the H21 Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the MPH21 sample of Kepler stars with Keck/HIRES spectroscopy, from which MPH21 inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.}
+
+\edit1{We do note, however, that we can not rule out the possibility that the long-period pile-up is wider in period space than we observe, with stars just above the pile-up being undetected through rotational brightness modulations.
+Such a scenario might result if magnetic braking is gradually weakened as stars approach \rocrit, as opposed to ceasing entirely at \rocrit.}
 
@@
 As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021}, \edit1{as well as the peak of the period distribution inferred by \citet{Masuda2021} from combining \vsini and stellar radii.
+The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by H21, \edit1{as well as the peak of the period distribution inferred by MPH21 from combining \vsini and stellar radii.
 We suggest that these features are one and the same.
-The modest differences between the long-period pile-up we observe and the features observed by \citet{Hall2021} and \citet{Masuda2021} may be partially explained by differing \teff scales between the samples.}
+The modest differences between the long-period pile-up we observe and the features observed by H21 and MPH21 may be partially explained by differing \teff scales between the samples.}
 
@@
 
-Furthermore, we note that there are important differences between the samples considered here and the \citet{Hall2021} asteroseismic sample.
+Furthermore, we note that there are important differences between the samples considered here and the H21 asteroseismic sample.
 The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
@@
 The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
-\edit1{The same argument is true for the \citet{Masuda2021} sample.}
+\edit1{The same argument is true for the MPH21 sample.}
 
@@
 To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
-Notably, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
-Furthermore, as discussed in \S\ref{subsec:longperiod}, the \citet{Hall2021} sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
+Notably, comparing the Sun to the H21 asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+Furthermore, as discussed in \S\ref{subsec:longperiod}, the H21 sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
@@
     \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
-    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \citet{Hall2021}.}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and H21.}
     \label{fig:teffscales}
@@
 \section{The Gaia Color–Period Plane}\label{app:gaia}
-\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
-We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
-We computed reddening estimates for each target by querying the \texttt{Bayestar19} 3D dust map using the \texttt{dustmaps} package \citep{dustmaps, Green2019}.
+\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
+We cross-matched the \citet{McQuillan2014} and \citet{Santos2021} samples with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
+We retrieved reddening estimates for each target by querying the \texttt{Bayestar19} 3D dust map using the \texttt{dustmaps} package \citep{dustmaps, Green2019}.
 We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves.
@@
 We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
-In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}
+In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required (or, equivalently, a +0.1 mag shift applied to the constant Rossby curve).
+While there is no justification for such a large shift, it may indicate the presence of a systematic offset in the \taucz relation.}
 

### 2022-04-01 src/ms.tex
minor edits

@@
 
-To obtain a reasonable description of the long-period edge we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
+\edit1{We approximated} the long-period \edit1{pile-up by computing} the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
 We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
@@
     \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
-    \caption{Non-linear least-squares fits to the long-period pile-up for a \rocrit~=~2 model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
-The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=2} - \chi^2_{\Delta T_{eff}=0}$.}
+    \caption{Non-linear least-squares fits to the long-period pile-up for a \rocrit~=~\edit1{\rosun} model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
+The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=\text{Ro}_\odot} - \chi^2_{\Delta T_{eff}=0}$.}
     \label{fig:rocrit}

### 2022-04-01 src/ms.tex
new macros

@@
 
-% \newcommand{\hall}{H21\xspace}
-% \newcommand{\masuda}{MPH21\xspace}
-% \newcommand{\vs19}{vS19\xspace}
+% Commonly cited references
+\newcommand{\jvs}{vS19\xspace}
+\newcommand{\mma}{MMA14\xspace}
+\newcommand{\hall}{H21\xspace}
+\newcommand{\masuda}{MPH21\xspace}
 
@@
 
-Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep{vanSaders2019}.
+Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep[][hereafter vS19]{vanSaders2019}.
 Those authors also predicted a pile-up of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
-While \citet{vanSaders2019} favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.
-
-More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the H21 sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
+While \jvs favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.
+
+More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \hall sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
 \citet{Masuda2021}, \edit1{hereafter MPH21}, also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
@@
 
-Here we examine the rotation period distribution of main-sequence stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pile-ups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
+Here we examine the rotation period distribution of {\color{red} main-sequence} stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pile-ups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
 We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, discuss some implications of these results in \S\ref{sec:discussion}, and present our conclusions in \S\ref{sec:conclusions}.
@@
     \includegraphics[width=\linewidth]{./figures/kde.pdf}
-    \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic H21 samples, from left to right.
+    \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic \hall samples, from left to right.
 Empirical cluster sequences from \citet{Curtis2020} are shown by the dark grey lines.
@@
 
-We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the H21 asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
+We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the \hall asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
 For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
@@
 
-H21 determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the H21 sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
+\hall determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \hall sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
 Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of the two effects}.
 
-To investigate the origin of this offset, we cross-matched the H21 sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
-We found a root mean square (RMS) of the residuals between the H21 and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
-\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and H21 samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
+To investigate the origin of this offset, we cross-matched the \hall sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
+We found a root mean square (RMS) of the residuals between the \hall and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
+\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \hall samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
 
@@
 
-\edit1{MPH21 inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from their precise stellar radii and spectroscopically determined \vsini.
-As with the asteroseismic sample, we compare the CKS and \lamostmcq samples with the results of those authors.
-We find that the long-period pile-up we observe appears to overlap with the MPH21 data.
-We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the H21 and MPH21 studies.}
+\edit1{\masuda inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from precise stellar radii and spectroscopically determined \vsini.
+As with the asteroseismic sample, we compared the CKS and \lamostmcq samples with the results of those authors.
+We find that the long-period pile-up we observe appears to overlap with the \masuda data, within the uncertainties of those authors.
+We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the \hall and \masuda studies.}
 
@@
     \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the H21 main-sequence asteroseismic sample (top row) \edit1{and the MPH21 sample (bottom row)} in the \teff--\prot plane.
-Note, the CKS, LAMOST, and H21 samples derive \teff from distinct pipelines.
+    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the \hall main-sequence asteroseismic sample (top row) \edit1{and the \masuda sample (bottom row)} in the \teff--\prot plane.
+Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
 A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper left panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
-Using LAMOST \teff, where it exists, for the H21 sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
+Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
-\edit1{These ranges are consistent with the range of ages observed in the H21 and MPH21 samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
+\edit1{These ranges are consistent with the range of ages observed in the \hall and \masuda samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
 The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
@@
 
-\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the H21 asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
+\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
 Stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
 The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
-CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the H21 asteroseismic sample.}
+CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.}
 
@@
 In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
-In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the H21 asteroseismic sample is shown by blue contours.
+In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \hall asteroseismic sample is shown by blue contours.
 No reddening corrections were performed to the photometry in this figure.}
@@
 \edit1{Furthermore, two independent samples with different observational biases than the surveys considered here also appear to yield a pile-up of stars at the long-period edge.
-These samples are the H21 Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the MPH21 sample of Kepler stars with Keck/HIRES spectroscopy, from which MPH21 inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.}
+These samples are the \hall Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the \masuda sample of Kepler stars with Keck/HIRES spectroscopy, from which \masuda inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.}
 
@@
 As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by H21, \edit1{as well as the peak of the period distribution inferred by MPH21 from combining \vsini and stellar radii.
+The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, \edit1{as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
 We suggest that these features are one and the same.
-The modest differences between the long-period pile-up we observe and the features observed by H21 and MPH21 may be partially explained by differing \teff scales between the samples.}
+The modest differences between the long-period pile-up we observe and the features observed by \hall and \masuda may be partially explained by differing \teff scales between the samples.}
 
@@
 
-Furthermore, we note that there are important differences between the samples considered here and the H21 asteroseismic sample.
+Furthermore, we note that there are important differences between the samples considered here and the \hall asteroseismic sample.
 The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
@@
 The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
-\edit1{The same argument is true for the MPH21 sample.}
+\edit1{The same argument is true for the \masuda sample.}
 
@@
 To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
-Notably, comparing the Sun to the H21 asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
-Furthermore, as discussed in \S\ref{subsec:longperiod}, the H21 sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
+Notably, comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+Furthermore, as discussed in \S\ref{subsec:longperiod}, the \hall sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
@@
     \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
-    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and H21.}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \hall.}
     \label{fig:teffscales}

### 2022-04-05 src/ms.tex
minor edits

@@
 \begin{document}
-\title{Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}
+\title{Further Evidence of Modified Spin-down in Sun-like Stars:\\Pileups in the Temperature--Period Distribution}
 
@@
 \begin{abstract}
-We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic temperatures to demonstrate the existence of pile-ups at the long-period and short-period edges of the temperature--period distribution for main-sequence stars with temperatures exceeding $\sim$\teffmin.
-The long-period pile-up is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
-The long-period pile-up was predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
-Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but weakly predictive of its age.
-The short-period pile-up, which is also well-described by a curve of constant Rossby number, is not a prediction of the weakened magnetic braking hypothesis but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling.
+We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic temperatures to demonstrate the existence of pileups at the long-period and short-period edges of the temperature--period distribution for main-sequence stars with temperatures exceeding $\sim$\teffmin.
+The long-period pileup is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal.
+The long-period pileup was predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number.
+Stars in the long-period pileup are found to have a wide range of ages (\agerange), meaning that, along the pileup, rotation period is strongly predictive of a star's surface temperature but weakly predictive of its age.
+The short-period pileup, which is also well-described by a curve of constant Rossby number, is not a prediction of the weakened magnetic braking hypothesis but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling.
 The same mechanism was proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters.
-The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
+The relative dearth of stars with intermediate rotation periods between the short- and long-period pileups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
 These observations provide further support for the hypothesis that the period gap is due to stellar astrophysics, rather than a non-uniform star-formation history in the Kepler field.
@@
 Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep[][hereafter vS19]{vanSaders2019}.
-Those authors also predicted a pile-up of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
+Those authors also predicted a pileup of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
 While \jvs favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.
@@
 
-Here we examine the rotation period distribution of {\color{red} main-sequence} stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pile-ups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
+Here we examine the rotation period distribution of {\color{red} main-sequence} stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pileups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
 We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, discuss some implications of these results in \S\ref{sec:discussion}, and present our conclusions in \S\ref{sec:conclusions}.
@@
 The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the \edit1{activity belts}.
-Many of the stars above the long-period pile-up are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \citet{vanSaders2019}.}
+Many of the stars above the long-period pileup are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \citet{vanSaders2019}.}
     \label{fig:surveys}
@@
 \subsection{The Sun}
-To place the Sun in the context of the long-period pile-up, we use up-to-date estimates of the Sun's effective temperature, rotation period, and age.
+To place the Sun in the context of the long-period pileup, we use up-to-date estimates of the Sun's effective temperature, \edit1{color}, rotation period, and age.
 Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
@@
 In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from \edit1{the activity belts}.
-The long-period pile-up for dwarf stars is clearly seen to extend to the solar temperature.
-The short-period pile-up is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.
+The long-period pileup for dwarf stars is clearly seen to extend to the solar temperature.
+The short-period pileup is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.
 The secondary overdensity observed in the subgiant samples, most visible in the bottom center panel, appears to be at twice the period of the primary overdensity, potentially due to erroneously determined rotation periods.}
@@
 \subsection{Initial observations}
-We first noted a pile-up of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-\edit1{The pile-up in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \citet{Santos2021} revealed that this pile-up is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
-
-A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-We verified through inspection that the secondary pile-up does not lie along the \prot/2 harmonic line of the long-period pile-up \edit1{(see Appendix~\ref{app:harmonics})}.
-As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
-\edit1{The short-period pile-up is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
+We first noted a pileup of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
+\edit1{The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \citet{Santos2021} revealed that this pileup is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
+
+A secondary pileup at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
+We verified through inspection that the secondary pileup does not lie along the \prot/2 harmonic line of the long-period pileup \edit1{(see Appendix~\ref{app:harmonics})}.
+As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pileup is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+\edit1{The short-period pileup is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
 the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
@@
 The orange dashed lines show constant Rossby curves of fiducial values (see \S\ref{subsec:rossby}).
-The short-period pile-up can be observed in the LAMOST--McQuillan sample for \teff~$\gtrsim$~5500~K.
+The short-period pileup can be observed in the LAMOST--McQuillan sample for \teff~$\gtrsim$~5500~K.
 The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from \edit1{the activity belts}.}
@@
 
-The location of the \edit1{long}-period pile-up is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+The location of the \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
-The observation that the \edit1{long-period pile-up} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the \edit1{long-period pile-up clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pile-up}.
-These observations are in accordance with predictions from the WMB model which suggest the pile-up forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
-\edit1{We show in \S\ref{subsec:ages} that stars along the pile-up have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
+The observation that the \edit1{long-period pileup} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
+Similarly, the \edit1{long-period pileup clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pileup}.
+These observations are in accordance with predictions from the WMB model which suggest the pileup forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
+\edit1{We show in \S\ref{subsec:ages} that stars along the pileup have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
 
@@
 In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in Figure~\ref{fig:inflection}).
-This cluster of points has a similar slope in the \teff--\prot plane as the long-period pile-up, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
+This cluster of points has a similar slope in the \teff--\prot plane as the long-period pileup, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
 After inspecting the light curves of stars on the ridge and above the ridge in the temperature range 6100~K~$<$~ \teff~$<6300$~K we did not notice any obvious trends or differences between the two samples.
@@
 Interestingly, a similar clustering of points is not observed in the \edit1{\lamostmcq} sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
-Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pile-up.
+Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pileup.
 We conclude that the temperatures for these stars in the CKS catalog are too high by $\gtrsim$~100~K.
 
-\edit1{In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pile-up in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
+\edit1{In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pileup in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
 
@@
     \caption{The \teff–\prot plane for the \lamostmcq sample (grey points) and the CKS sample (blue points) with periods sourced from \citet{David2021}.
-The selection used to identify long-period pile-up stars in the CKS sample is shown by the region enclosed by the black lines.
+The selection used to identify long-period pileup stars in the CKS sample is shown by the region enclosed by the black lines.
 Stars with \teff differences $>100$~K between the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs are shown by open red circles.
 The orange point indicates the Sun.
-Random draws from the MCMC chain of a piecewise linear fit to the CKS long-period pile-up are shown by the blue lines.
+Random draws from the MCMC chain of a piecewise linear fit to the CKS long-period pileup are shown by the blue lines.
 An inflection point in the long-period edge is apparent near \teff=~6250~K.}
@@
 In the WMB model of \citet{vanSaders2016, vanSaders2019}, a star spins down until it reaches a critical Rossby number, at which point magnetic braking ceases.
-Since Rossby number is highly sensitive to temperature through its dependence on the convective turnover timescale ($\mathrm{Ro} = P_\mathrm{rot}/\tau_\mathrm{cz}$, where $\tau_\mathrm{cz}$ is the convective turnover timescale), this critical threshold corresponds to different rotation periods for stars of different \teff, leading to a pile-up in the \teff-\prot plane.
+Since Rossby number is highly sensitive to temperature through its dependence on the convective turnover timescale ($\mathrm{Ro} = P_\mathrm{rot}/\tau_\mathrm{cz}$, where $\tau_\mathrm{cz}$ is the convective turnover timescale), this critical threshold corresponds to different rotation periods for stars of different \teff, leading to a pileup in the \teff-\prot plane.
 Using a small sample of Kepler targets with rotation periods determined from brightness modulations, \citet{vanSaders2016} proposed this threshold happens at a critical Rossby number of $\mathrm{Ro_{crit}} \sim \mathrm{Ro_\odot}$.
 
-We tested the hypothesis that the long-period pile-up observed in the \edit1{\lamostmcq} sample is \edit1{compatible with the} WMB \edit1{model} by fitting constant Rossby models to the \prot boundary.
+We tested the hypothesis that the long-period pileup observed in the \edit1{\lamostmcq} sample is \edit1{compatible with the} WMB \edit1{model} by fitting constant Rossby models to the \prot boundary.
 For a given \teff, this model predicts \prot as:
@@
 
-\edit1{We approximated} the long-period \edit1{pile-up by computing} the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
+\edit1{We approximated} the long-period \edit1{pileup by computing} the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
 We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
@@
 
-We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources) and the \hall asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
-For the CKS pile-up stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
-To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
+We additionally fit the CKS long-period pileup (using three different homogeneous \teff sources) and the \hall asteroseismic main-sequence sample, allowing for \teff offsets in each data set (Table~\ref{tab:mcmc}).
+For the CKS pileup stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
+To isolate the long-period pileup stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
 We note that this selection is particular to the CKS \teff scale and is not general.
@@
 We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, \edit1{including} the \teff scale of the models used to derive the \taucz relation.
-Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
-
-To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pile-up and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
+Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pileups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
+
+To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
 The first model assumes \rocrit = 2 and has one free parameter, a temperature offset added to the data.
@@
 Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure.
-Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pile-up in the CKS vs.
-\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pile-up and the \lamostmcq 90th percentile curve, we observe in Figure~\ref{fig:percentiles} a systematic trend in which most CKS pile-up stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.
+Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pileup in the CKS vs.
+\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pileup and the \lamostmcq 90th percentile curve, we observe in Figure~\ref{fig:percentiles} a systematic trend in which most CKS pileup stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.
 This is consistent with there being a \teff-dependent systematic trend between the \teff scales of CKS and LAMOST, as discussed in Appendix~\ref{app:teff}.}}
@@
     \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
-    \caption{Non-linear least-squares fits to the long-period pile-up for a \rocrit~=~\edit1{\rosun} model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
+    \caption{Non-linear least-squares fits to the long-period pileup for a \rocrit~=~\edit1{\rosun} model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
 The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=\text{Ro}_\odot} - \chi^2_{\Delta T_{eff}=0}$.}
@@
 \startdata
-\textit{Short-period pile-up} \\
+\textit{Short-period pileup} \\
 \hline
@@
 \hline 
-\textit{Long-period pile-up} \\
+\textit{Long-period pileup} \\
 \hline
@@
 Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pile-up that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
+The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
 {\color{red}Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pile-up in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff})}.
+The \teff offsets lead to better agreement between the data and models, although the long-period pileup in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff})}.
 \edit1{We note that in our comparisons we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
@@
 
-Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models.
+Notably, both models fail to reproduce the observed short-period pileup, possibly due to the assumption of solid-body rotation in both models.
 At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence.
@@
 Shifts of +116~K and +140~K are applied to the CKS and LAMOST \teff, respectively.
-The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pile-up (as described in \S\ref{subsec:rossby}).
+The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pileup (as described in \S\ref{subsec:rossby}).
 The LAMOST shift is derived in a similar fashion, with an additional +34~K offset which represents the systematic offset between CKS and LAMOST (see Appendix~\ref{app:teff}).}
@@
 \hall determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \hall sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
-Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of the two effects}.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pileup we observe, although the \hall sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
+Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of effects}.
 
@@
 
-{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.}
+{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pileup as seen in the two samples.}
 
@@
 As with the asteroseismic sample, we compared the CKS and \lamostmcq samples with the results of those authors.
-We find that the long-period pile-up we observe appears to overlap with the \masuda data, within the uncertainties of those authors.
-We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the \hall and \masuda studies.}
+We find that the long-period pileup we observe appears to overlap with the \masuda data, within the uncertainties of those authors.
+We therefore consider it likely that the long-period pileup is the same feature, or an extension of the same feature, that was detected in the \hall and \masuda studies.}
 
@@
 A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper left panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
-Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
+Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
@@
 
-\subsection{Ages of stars on the long-period pile-up}
+\subsection{Ages of stars on the long-period pileup}
 \label{subsec:ages}
 The WMB model predicts that hotter stars pile up on the long-period edge at younger ages, producing an age gradient across the edge.
-In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pile-up.
+In Figure~\ref{fig:ages} we show the \teff-age distributions of stars on the ridge using ages from the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs, where we use the trapezoidal selection described in \S\ref{subsec:rossby} to select stars on the long-period pileup.
 Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting.
@@
 The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
-In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
+In fact, there are only a handful of long-period pileup stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pileup stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
 
@@
 Stars reach \rocrit~$\approx$~2 approximately halfway through their main-sequence lifetimes and remain there until the main-sequence turnoff.
-Thus, higher mass stars with shorter main-sequence lifetimes should show a smaller age spread on the long-period pile-up relative to lower mass stars with longer main-sequence lifetimes.
+Thus, higher mass stars with shorter main-sequence lifetimes should show a smaller age spread on the long-period pileup relative to lower mass stars with longer main-sequence lifetimes.
 Though we have not quantified such an age-gradient, the data suggest that some stars spend several Gyr occupying the ridge with only modest evolution of their spin rates.
 
-We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}), 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
-While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).
-Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pile-up stars.
-We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.
+We additionally selected stars on the long-period pileup from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}), 5500~K~$<$~\teff~$<$~6500~K, and 4.1~dex~$<$~\logg~$<$~4.75~dex.
+While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pileup is populated by stars with a broad range of ages (Figure~\ref{fig:ages}).
+Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pileup stars.
+We discuss the Sun in context of the long-period pileup further in \S\ref{subsec:thesun}.
 
@@
     \includegraphics[width=\linewidth]{./figures/ages.pdf}
-    \caption{Above, H-R diagram placement of long-period pile-up stars relative to the Sun and the CKS sample (a) and similarly for the \lamostmcq sample (b).
-Below, the \teff-age plane for CKS stars along the long-period pile-up using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
+    \caption{Above, H-R diagram placement of long-period pileup stars relative to the Sun and the CKS sample (a) and similarly for the \lamostmcq sample (b).
+Below, the \teff-age plane for CKS stars along the long-period pileup using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
 The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~$<$~2 (blue) and Ro~$>$~2 (orange).
@@
 
-\edit1{To further assess the evolutionary state of stars on the long-period pile-up, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pile-up stars from this work (Figure~\ref{fig:cmd}).
+\edit1{To further assess the evolutionary state of stars on the long-period pileup, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pileup stars from this work (Figure~\ref{fig:cmd}).
 Stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
 The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
-CKS stars on the long-period pile-up clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.}
+CKS stars on the long-period pileup clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.}
 
@@
 In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
-In the right panel, the CKS long-period pile-up stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \hall asteroseismic sample is shown by blue contours.
+In the right panel, the CKS long-period pileup stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \hall asteroseismic sample is shown by blue contours.
 No reddening corrections were performed to the photometry in this figure.}
@@
 
-\subsection{Where do the pile-ups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pile-up} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
-\edit1{The number density of stars on the short-period pile-up similarly declines towards cooler \teff.
-If the short-period pile-up is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
+\subsection{Where do the pileups end?} \label{subsec:extent}
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pileup} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
+\edit1{The number density of stars on the short-period pileup similarly declines towards cooler \teff.
+If the short-period pileup is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
 However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
@@
 
-While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach.
-We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
+While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pileups through the following approach.
+We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pileups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
 In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window.
-We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
-The relative fraction of stars on the short-period \edit1{pile-up} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal.
-Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.
+We found that the relative fraction of stars on the long-period pileup declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
+The relative fraction of stars on the short-period \edit1{pileup} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pileups are nearly equal.
+Thus, it is not yet clear if either pileup extends to temperatures cooler than \tsun.
 
@@
     \includegraphics[width=\linewidth]{./figures/fraction.pdf}
-    \caption{The relative fraction of stars on the long- and short-period pile-ups, in the \lamostmcq sample, in 10~K windows of \teff (for stars with \logg~$>$4.1 dex).}
+    \caption{The relative fraction of stars on the long- and short-period pileups, in the \lamostmcq sample, in 10~K windows of \teff (for stars with \logg~$>$4.1 dex).}
     \label{fig:fraction}
@@
 \edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
-As rotation periods become longer, they become more difficult to measure directly from rotational brightness modulations in Kepler time series.
-This is true for at least three reasons: (1) the Kepler baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
-This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pile-up in the catalogs considered here depends sensitively on how detectability depends on Ro.
-For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pile-up has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pile-up.}
-
-\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution.
-In this scenario, the detection threshold sets the upper edge of the observed \teff–\prot distribution at high \teff, while the age of the galaxy sets the upper edge at low \teff (since low-mass stars have not yet had enough time to spin down to Ro$_\mathrm{thresh}$).
+As periods become longer, they become more difficult to measure directly from Kepler time series.
+This is true for at least three reasons: (1) the observational baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
+The last two points are due to astrophysical reasons.
+This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pileup in the catalogs considered here depends sensitively on how detectability depends on Ro.
+For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pileup has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pileup.}
+
+\edit1{\jvs addressed detection biases in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \mma \teff–\prot distribution.
+In this scenario, if \rothresh $\lesssim$ \rocrit, it becomes difficult to distinguish between the standard and WMB models in the \mma sample.
 However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
@@
 
-\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge.
+\edit1{\mma published weight parameters for each star, a metric which was designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in $T_\mathrm{eff}$–$P_\mathrm{rot}$–LPH space, where LPH designates the local peak height of the ACF.
+Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
+In Figure~\ref{fig:weights}, we show that the long-period pileup is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pileup are likely to be genuine detections and not an artifact of some detection threshold.
+Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pileup or overdensity at this edge.
 To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
 
-\edit1{Furthermore, two independent samples with different observational biases than the surveys considered here also appear to yield a pile-up of stars at the long-period edge.
+\edit1{Most importantly, two independent samples with different observational biases than the surveys considered here also appear to yield a pileup of stars at the long-period edge.
 These samples are the \hall Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the \masuda sample of Kepler stars with Keck/HIRES spectroscopy, from which \masuda inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.}
 
-\edit1{We do note, however, that we can not rule out the possibility that the long-period pile-up is wider in period space than we observe, with stars just above the pile-up being undetected through rotational brightness modulations.
-Such a scenario might result if magnetic braking is gradually weakened as stars approach \rocrit, as opposed to ceasing entirely at \rocrit.}
-
-\edit1{Finally, \citet{McQuillan2014} published weight parameters for each star, a metric which was designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in \teff–\prot–LPH space, where LPH designates the local peak height of the ACF.
-Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
-In Figure~\ref{fig:weights}, we show that the long-period pile-up is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pile-up are likely to be genuine detections and not an artifact of some detection threshold.}
+\edit1{We do note, however, that we can not rule out the possibility that the long-period pileup is wider in period or Rossby space than we observe, with stars just above the pileup being undetected through rotational brightness modulations.
+Such a scenario might result if magnetic braking is gradually weakened as stars approach \rocrit, as opposed to ceasing entirely at \rocrit.
+However, if the pileup does have a broader width in Rossby number, the close agreement between our observations and the \hall and \masuda samples suggests the true pileup width cannot be much larger than we observe.}
 
@@
 
-As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
-The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, \edit1{as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
+As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pileup of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
+The location of the long-period pileup coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, \edit1{as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
 We suggest that these features are one and the same.
-The modest differences between the long-period pile-up we observe and the features observed by \hall and \masuda may be partially explained by differing \teff scales between the samples.}
+The modest differences between the long-period pileup we observe and the features observed by \hall and \masuda may be partially explained by differing \teff scales between the samples.}
 
@@
 
-Furthermore, we note that there are important differences between the samples considered here and the \hall asteroseismic sample.
+Furthermore, there are important differences between the samples considered here and the \hall asteroseismic sample.
 The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
 Exacerbating the problem of detection, slower rotation tends to be accompanied by smaller-scale and less periodic variability.
-The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pile-up stars that our sample may have missed.
+The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pileup stars that our sample may have missed.
 \edit1{The same argument is true for the \masuda sample.}
 
-It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability.
+It is evident from Figure~\ref{fig:gap} that the long-period pileup also coincides with a steep gradient in variability amplitude, such that a hidden population of pileup stars may lie just beyond the edge of detectability.
 This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values.
-Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that \edit1{a pile-up exists at the long-period edge} for stars with \teff~$\gtrsim$~\teffmin.
-\edit1{By construction, this pile-up occurs at a Rossby number lower than that associated with a detection edge.} \edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
-
-\edit1{We also note that the WMB model makes the simplistic assumption that magnetic braking ceases when $Ro$ reaches \rocrit.
-However, it is possible that magnetic braking becomes gradually weaker, a scenario which would also lead to a pile-up but may also cause such a pile-up to have some intrinsic width which is not due to measurement uncertainties.}
-
-Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that \edit1{a pileup exists at the long-period edge} for stars with \teff~$\gtrsim$~\teffmin.
+\edit1{By construction, this pileup occurs at a Rossby number lower than that associated with a detection edge.} \edit1{In other words, the fact that we observe a pileup indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
+
+\edit1{We also note that the WMB model makes the simplistic assumption that magnetic braking ceases when Ro reaches \rocrit.
+However, it is possible that magnetic braking becomes gradually weaker, a scenario which would also lead to a pileup but may also cause such a pileup to have some intrinsic width which is not due to measurement uncertainties.}
+
+\edit1{Our observation of a long-period pileup stands in contrast with recent rotation studies of solar analogs.} \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
 Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
 A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.
-\edit1{Additionally, \citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr.
-In contrast to the findings of those authors, we find that the long-period pile-up lies clearly below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.
+\edit1{Additionally, \citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~\rosun and at ages in the range of $\sim$2--6~Gyr.
+In contrast to the findings of those authors, we find that the long-period pileup lies far below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.
+\edit1{This implies the braking index must drop to a much lower value at some time after 2.5~Gyr for Sun-like stars.}
 
@@
 
+%Sentence about 95% vs 90% percentiles.
+
 \subsection{Short-period behavior} \label{subsec:shortperiod}
 
-Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model or, more generally, any standard, solid-body braking model.
-However, this feature may also be due to an epoch of apparent stalling, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pile-up.
+Unlike the long-period pileup, the short-period pileup is not predicted by the WMB model or, more generally, any standard, solid-body braking model.
+However, this feature may also be due to an epoch of apparent stalling, albeit a temporary one since cluster studies demonstrate that stars continue to spin down beyond the short-period pileup.
 Works examining the \teff–\prot sequences of open clusters have found overlap between low-mass members in clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}, notably Praesepe (0.67~Gyr), NGC 6811 (1.4~Gyr), and NGC 752 (1.4~Gyr).
 In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.
-The short-period pile-up we observe may be the manifestation of the same type of slowed spin evolution, but for stars of higher masses and younger ages than the cluster members in the above-mentioned works (since the pile-up is observed at shorter periods relative to the Praesepe sequence).
+The short-period pileup we observe may be the manifestation of the same type of slowed spin evolution, but for stars of higher masses and younger ages than the cluster members in the above-mentioned works (since the pileup is observed at shorter periods relative to the Praesepe sequence).
 
@@
 
-Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pile-up is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
+Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pileup is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
 Furthermore, in order to match observations of rotation periods in young clusters, models require that the core–envelope coupling timescale increases towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
@@
 
-The short- and long-period pile-ups we examine here naturally produce a dearth of rotators at intermediate periods.
+The short- and long-period pileups we examine here naturally produce a dearth of rotators at intermediate periods.
 This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \citet{McQuillan2014} sample.
@@
 We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here.
-Notably, if the short-period pile-up is indeed due to core–envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
+Notably, if the short-period pileup is indeed due to core–envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
 
@@
 Black contours show Gaussian kernel density estimation of the plotted distributions, and the dashed line shows a constant Rossby curve of Ro~=~0.5.
-    The long- and short-period pile-ups are separated by a relative dearth of stars with intermediate rotation periods.
+    The long- and short-period pileups are separated by a relative dearth of stars with intermediate rotation periods.
 A strong gradient in \rper is apparent across this gap, such that variability amplitude reaches a local minimum near the gap's center.}
@@
 
-It is unclear whether or not the Sun is a resident of the long-period pile-up.
-From Fig.~\ref{fig:xmatch} it is clear that the long-period pile-up extends to temperatures as cool as the Sun's.
+It is unclear whether or not the Sun is a resident of the long-period pileup.
+From Fig.~\ref{fig:xmatch} it is clear that the long-period pileup extends to temperatures as cool as the Sun's.
 However, it is not clear whether there is a large systematic offset in the LAMOST temperature scale and the temperature scale of the Sun.
-If one assumes there are no systematic offsets between the two \teff scales, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pile-up.
+If one assumes there are no systematic offsets between the two \teff scales, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pileup.
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
 
-We consider the most likely explanation for the observed gap between the Sun and the long-period pile-up to be systematic offsets in \teff.
-To place the Sun on the long-period pile-up would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
-Notably, comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.
+We consider the most likely explanation for the observed gap between the Sun and the long-period pileup to be systematic offsets in \teff.
+To place the Sun on the long-period pileup would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
+Notably, comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pileup, or very close to it, without the need for a \teff shift.
 Furthermore, as discussed in \S\ref{subsec:longperiod}, the \hall sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
-We conclude that the Sun likely does reside on the long-period pile-up, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
+We conclude that the Sun likely does reside on the long-period pileup, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
 
@@
     \item We observe an overdensity at the long-period edge of the \teff-\prot distribution of Kepler main-sequence stars with \teff~$\gtrsim$~\teffmin.
-We hypothesize that this pile-up was previously obfuscated by imprecise \teff estimates.
-Both the existence of the pile-up and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
-
-    \item The long-period pile-up is well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
-The pile-up is also populated by stars with a wide range of isochrone ages (\agerange).
-A pile-up of stars with a constant Rossby number and a broad range of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
+We hypothesize that this pileup was previously obfuscated by imprecise \teff estimates.
+Both the existence of the pileup and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
+
+    \item The long-period pileup is well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
+The pileup is also populated by stars with a wide range of isochrone ages (\agerange).
+A pileup of stars with a constant Rossby number and a broad range of ages is a prediction of the WMB model of \citet{vanSaders2016, vanSaders2019}.
 The precise value of \rocrit is sensitive to \teff scale shifts between observational data and the models used to compute \taucz.
@@
     \item Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \citet{vanSaders2019}.
-Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pile-up until an age of $\sim$6~Gyr.
-
-    \item It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it.
-Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up.
-If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up; some of the cooler long-period pile-up stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pile-up.
+Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pileup until an age of $\sim$6~Gyr.
+
+    \item It is yet unclear whether the Sun resides on the long-period pileup or has already evolved through it.
+Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pileup.
+If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
 However, this tension might be simply explained by inaccurate isochrone ages.
@@
     \item We \edit1{tentatively detect} a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly} because the short-period pile-up is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pile-up ($>10^9$~yr).
-The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.
-The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.
-
-    \item The number density of stars on the long-period pile-up declines with \teff, in line with predictions from the WMB hypothesis, though it remains unclear whether this observation is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
+This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly} because the short-period pileup is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pileup ($>10^9$~yr).
+The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.
+The short-period pileup can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pileup.
+
+    \item The number density of stars on the long-period pileup declines with \teff, in line with predictions from the WMB hypothesis, though it remains unclear whether this observation is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects.
 The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.
 
-    \item We find tentative evidence for an age-gradient along the long-period pile-up, such that hotter stars on the ridge are younger on average.
+    \item We find tentative evidence for an age-gradient along the long-period pileup, such that hotter stars on the ridge are younger on average.
 Relatedly, the age dispersion along the ridge is non-uniform as a function of temperature, with hotter stars showing a smaller dispersion.
-This observation suggests that hotter stars reside on the long-period pile-up for a shorter period of time relative to cooler stars.
-These observations are in accordance with predictions from the WMB model, which predicts that stars of different masses spend an approximately equal fraction of their respective main-sequence lifetimes on the long-period pile-up.
+This observation suggests that hotter stars reside on the long-period pileup for a shorter period of time relative to cooler stars.
+These observations are in accordance with predictions from the WMB model, which predicts that stars of different masses spend an approximately equal fraction of their respective main-sequence lifetimes on the long-period pileup.
 However, a more careful analysis is required to conclusively show these observations are not due to the intrinsic age gradient expected among a sample of main sequence stars with different masses and the higher isochrone age uncertainties associated with cooler stars.
 
-    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up.
-For example, a Sun-like star may spend several Gyr evolving through the long-period pile-up.
+    \item The existence of the long-period pileup limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pileup.
+For example, a Sun-like star may spend several Gyr evolving through the long-period pileup.
 Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim$~\teffmin (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
@@
 Taken out of context, this sample resembles a group of coeval stars with a slow-rotator sequence.
-The discovery of pile-ups in stellar rotation periods therefore has consequences for open cluster studies.
+The discovery of pileups in stellar rotation periods therefore has consequences for open cluster studies.
 When an overdensity or ridge is present in the rotation period distribution of a stellar population, care must be taken to ensure that it is not caused by WMB or core–envelope coupling before assuming that population is coeval and using the overdensity to age-date it.
@@
 In Figure~\ref{fig:comparison}, we show how the \teff--\prot distribution of the CKS sample changes when sourcing \teff and \prot from different, homogeneous catalogs in the literature.
-The sharpness of the long-period pile-up appears to be determined primarily by the source of \teff, rather than \prot.
+The sharpness of the long-period pileup appears to be determined primarily by the source of \teff, rather than \prot.
 The CKS-Gaia catalog \citep{Fulton2018} appears to offer the highest internal precision.
@@
 
-\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pile-up, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
+\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pileup, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
 Figure~\ref{fig:harmonic} shows a Gaussian kernel density estimation of the \lamostmcq \teff–\prot distribution for stars with \logg~$>$~4.1.
-The short-period pile-up is observed in this sample, particularly for \teff~$>$~6000~K.
-We found through inspection a constant Rossby curve that traces the long-period pile-up.
-Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pile-up, though they both bracket the feature seen through density estimation.
-We interpret this as evidence that the short-period pile-up is not simply due to measurement error, although we have not definitively ruled out that scenario.}
-
-\edit1{We also note that the short-period pile-up is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
+The short-period pileup is observed in this sample, particularly for \teff~$>$~6000~K.
+We found through inspection a constant Rossby curve that traces the long-period pileup.
+Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pileup, though they both bracket the feature seen through density estimation.
+We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.}
+
+\edit1{We also note that the short-period pileup is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
 The differences between the \citet{McQuillan2014} and \citet{Santos2021} catalogs may provide an explanation for this observation.
-The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pile-up appear weaker relative to the long-period pile-up.
-This is as expected since open cluster obsbervations suggest that the short-period pile-up can not be a long-lived feature, while the long-period pile-up appears to be much longer lived.
+The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pileup appear weaker relative to the long-period pileup.
+This is as expected since open cluster obsbervations suggest that the short-period pileup can not be a long-lived feature, while the long-period pileup appears to be much longer lived.
 Consequently, the \citet{Santos2021} catalog may more accurately reflect the relative strengths of these two features.}
@@
 The \prot/2 and \prot/3 harmonics of this curve are shown by the dashed and dotted lines, respectively.
-The short-period pile-up is seen near the short-period edge, particularly for \teff~$>$~6000~K.
-The short-period pile-up does not appear to be simply a harmonic of the long-period pile-up, as one might expect if the periods for the stars in the short-period pile-up were erroneously determined.}
+The short-period pileup is seen near the short-period edge, particularly for \teff~$>$~6000~K.
+The short-period pileup does not appear to be simply a harmonic of the long-period pileup, as one might expect if the periods for the stars in the short-period pileup were erroneously determined.}
     \label{fig:harmonic}    
@@
 \section{The Gaia Color–Period Plane}\label{app:gaia}
-\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
+\edit1{Given that the long-period pileup was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
 We cross-matched the \citet{McQuillan2014} and \citet{Santos2021} samples with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
@@
 For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators.
-We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
+We found that the long-period pileup is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
 In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required (or, equivalently, a +0.1 mag shift applied to the constant Rossby curve).

### 2022-04-05 src/figures/ages.py

@@ -272,13 +272,13 @@ for i in range(2):
     ax[i].set_xlim(6600,5600)
     ax[i].set_ylim(0,12)
     ax[i].scatter(sun["teff"], 4.567, color='orange', zorder=3)
-    ax[i].text(6500, 11.5, 'typical error', size=10)
+    ax[i].text(6450, 11, 'typical error', size=10)
     
 for i,let in enumerate("cd"):
     ax[i].text(1.05,1.05,let,transform=ax[i].transAxes,weight='bold')
     
 plt.tight_layout()
-sns.despine()
+#sns.despine()
 plt.savefig('../figures/ages.pdf')
 
 print('5th and 95th percentile range of CKS ages (Gyr)   :', np.nanpercentile(cks['cks_age'][ridge], [5,95]))

### 2022-04-05 src/figures/asteroseismic.py

@@ -65,7 +65,6 @@ sns.set(font_scale=1.2, context="paper", style="ticks")
 
 mpl.rcParams["legend.markerscale"] = 2
 titles = ["California–Kepler Survey","LAMOST–McQuillan"]
-ebar_kws = {"fmt":"o", "lw":1, "ecolor":"orange", "color":"orange", "alpha":1, "ms":4, "mec":"grey", "mew":0.5, "zorder":999}
 
 sc_kws = {"marker":".", "color":"C0", "rasterized":True, "alpha":1}
 
@@ -79,6 +78,8 @@ ax2[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"Californ
 ax2[1].scatter(lam['Teff_lam'], lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
 
 
+ebar_kws = {"fmt":"o", "ecolor":"orange", "color":"orange", "alpha":1, "mfc":"orange", "mec":"orange", "mew":1, "zorder":999}
+
 #Masuda et al. 2021 data
 specteff = np.array([5996.3610916725,6163.47095871239,6351.57452764171])
 specrot= np.array([15.1398954901294,10.9922199999722,5.40119754702563])
@@ -90,14 +91,15 @@ for i in range(2):
     ax1[i].errorbar(hall["Teff"][arg], hall["P"][arg], 
                    xerr=hall["e_Teff"][arg], 
                    yerr=[hall["e_P"][arg], hall["E_P"][arg]], 
-                   label="Hall et al. 2021\nasteroseismic sample",
+                   label="Hall et al. 2021\nasteroseismic sample", ms=2, lw=0.5, 
                    **ebar_kws)
     ax2[i].errorbar(specteff, specrot, 
                     xerr=spectefferr,
                     yerr=specroterr,
-                    label="Masuda et al. 2021",
+                    label="Masuda et al. 2021", ms=4, lw=1, 
                     **ebar_kws)
     ax1[i].set_xlim(6500,5000)
+    #ax1[i].set_ylim(0,60)
     ax1[i].set_xlabel("Effective temperature [K]")
     ax1[i].set_ylabel("Rotation period [d]")
     ax1[i].legend(loc="upper left", prop={"size":9})

### 2022-04-07 src/ms.tex
restructuring paper

@@
 %% using aastex version 6.3
-\documentclass[linenumbers,trackchanges,twocolumn]{aastex631}
+%\documentclass[linenumbers,trackchanges,twocolumn]{aastex631}
+\documentclass[trackchanges,twocolumn]{aastex631}
 
@@
 
-Here we examine the rotation period distribution of {\color{red} main-sequence} stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pileups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
+Here we examine the rotation period distribution of stars observed by Kepler, leveraging the recent release of precise spectroscopic parameters from large-scale surveys, to demonstrate the existence of pileups at the long- and short-period edges of the \teff--\prot distribution of solar-type stars.
 We discuss our sample in \S\ref{sec:sample}, describe the steps of our analysis in \S\ref{sec:analysis}, discuss some implications of these results in \S\ref{sec:discussion}, and present our conclusions in \S\ref{sec:conclusions}.
@@
 Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
-For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
+\edit1{For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.} For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
 \edit1{While helioseismic differential rotation studies have inferred rotation periods of up to $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
 We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
-\edit1{For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.}
 
@@
 
-A secondary pileup at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-We verified through inspection that the secondary pileup does not lie along the \prot/2 harmonic line of the long-period pileup \edit1{(see Appendix~\ref{app:harmonics})}.
-As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pileup is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
-\edit1{The short-period pileup is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
-the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
-
 \begin{figure*}
@@
 
-The location of the \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+\edit1{Comparing the long-period pileup to rotation period sequences from open clusters can provide insight into the ages of stars on the pileup.} The \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
@@
 
-In both the CKS and LAMOST samples, there is a change of slope along the ridge (Figure~\ref{fig:inflection}), with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
-A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
+\edit1{A change in slope along the long-period edge is observed} in both the CKS and LAMOST samples (Figure~\ref{fig:inflection}), with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
+A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6224~$\pm$~24~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
 This change in slope \edit1{may be} due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
 
-In the CKS sample, there appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in Figure~\ref{fig:inflection}).
+In the CKS sample, there also appears to be a clustering of stars above the ridge with \teff~$>6100$~K (seen most clearly in Figure~\ref{fig:inflection}).
 This cluster of points has a similar slope in the \teff--\prot plane as the long-period pileup, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
-After inspecting the light curves of stars on the ridge and above the ridge in the temperature range 6100~K~$<$~ \teff~$<6300$~K we did not notice any obvious trends or differences between the two samples.
-The periods of the stars in the cluster above the ridge appear to be just as reliable as those on the ridge.
-Interestingly, a similar clustering of points is not observed in the \edit1{\lamostmcq} sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
+A similar clustering of points is not observed in the \edit1{\lamostmcq} sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
 Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pileup.
 We conclude that the temperatures for these stars in the CKS catalog are too high by $\gtrsim$~100~K.
+
+A secondary pileup at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
+We verified through inspection that the secondary pileup does not lie along the \prot/2 harmonic line of the long-period pileup \edit1{(see Appendix~\ref{app:harmonics})}.
+As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pileup is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+\edit1{The short-period pileup is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
+the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
 
@@
 The selection used to identify long-period pileup stars in the CKS sample is shown by the region enclosed by the black lines.
-Stars with \teff differences $>100$~K between the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs are shown by open red circles.
+Stars with \teff differences $>100$~K between the CKS \citep{Fulton2018} and SPOCS \citep{Brewer2018} catalogs are shown by open black circles.
 The orange point indicates the Sun.
@@
 
-\edit1{For reference, this relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d}.
-
-\edit1{We approximated} the long-period \edit1{pileup by computing} the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
-We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
+\edit1{The above relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d}.
+
+\edit1{We approximated the long-period edge in the following manner.
+For 100 bootstrapped resamplings of the \teff–\prot distribution, leaving out 50\% of the data for each bootstrapped sample, we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
+The final long-period edge curve was then computed as the mean of the bootstrapped 90th percentile values.
+We assumed constant 10\% fractional errors on the \prot values computed in this way.} We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
 Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K.
-{\color{red} To compute approximate uncertainties on the 90th percentile curve we performed 100 bootstrapped resamplings in each bin, leaving out 50\% of the data for each resampling.
-The 90th percentile curve and its uncertainty was then computed from the mean and standard deviation of the boostrapped values.} We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
+We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
 We show these curves in relation to the full \edit1{\lamostmcq} sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
@@
 We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
-We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro $f$, and $T_\mathrm{sys}$ with ranges of (0.1,10), (0,10), and (-1000~K, 1000~K), respectively.
+We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro, $f$, and $T_\mathrm{sys}$ with ranges of (0.1,10), (0,10), and (-1000~K, 1000~K), respectively.
 Convergence was assessed by ensuring the chain length was at least 100 times longer than the chain autocorrelation lengths for each parameter.
@@
 For the CKS pileup stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
-To isolate the long-period pileup stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$.
+To isolate the long-period pileup stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$ \edit1{(shown in Figure~\ref{fig:inflection})}.
 We note that this selection is particular to the CKS \teff scale and is not general.
@@
 
-We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, \edit1{including} the \teff scale of the models used to derive the \taucz relation.
-Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pileups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
-
-To further explore whether our \rocrit inference procedure is biased by systematics in the data, we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
-The first model assumes \rocrit = 2 and has one free parameter, a temperature offset added to the data.
+\edit1{In some cases, our constant Rossby fits to the long-period pileup prefer Rossby numbers lower than the Sun's, in contrast with \jvs.
+We caution against overinterpreting the specific \rocrit values inferred here, and discuss our interpretations further in \S\ref{subsec:longperiod}.
+To further quantify the preference for \rocrit~$<$~\rosun} we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
+The first model assumes \rocrit = \rosun and has one free parameter, a temperature offset added to the data.
 The second model assumes no temperature offset and allows \rocrit to vary (\rocrit being the only free parameter).
@@
 For the \lamostmcq sample, there is weak support for the variable \rocrit model.
-However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
+However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend \edit1{(i.e.
+a \teff-dependent \teff offset)} in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
 Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure.
@@
 \hline
-\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.38 $\pm$ 0.04 & 0.011 $\pm$ 0.008 & 49 $\pm$ 19 \\
+%\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.38 $\pm$ 0.04 & 0.011 $\pm$ 0.008 & 49 $\pm$ 19 \\
 \lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.4 $\pm$ 0.1 & 0.019 $\pm$ 0.015 & 44 $\pm$ 43 \\
@@
 CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.2 & 0.18 $\pm$ 0.02 & 21 $\pm$ 42 \\
-\citet{Hall2021} main-sequence & 5000~K--6250~K & 1.7 $\pm$ 0.2 & 0.30 $\pm$ 0.04 & -5 $\pm$ 64 \\
+%\citet{Hall2021} main-sequence & 5000~K--6250~K & 1.7 $\pm$ 0.2 & 0.30 $\pm$ 0.04 & -5 $\pm$ 64 \\
 \citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & -29 $\pm$ 79 \\
@@
 \edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
-}.}}
+}.}} \edit1{We note that we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
+The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
 
@@
 Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
-{\color{red}Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pileup in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff})}.
-\edit1{We note that in our comparisons we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
-The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
+The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08 \edit1{(using a different \taucz prescription than the one used here)}, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
+Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
+The \teff offsets lead to better agreement between the data and models, although the long-period pileup in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
 
@@
     \caption{Same as Figure~\ref{fig:models} with constant \teff offsets applied to the data.
-Shifts of +116~K and +140~K are applied to the CKS and LAMOST \teff, respectively.
-The CKS \teff shift originates from a least-squares fit of a \rocrit~=~2 curve (with a \teff offset) to the long-period pileup (as described in \S\ref{subsec:rossby}).
-The LAMOST shift is derived in a similar fashion, with an additional +34~K offset which represents the systematic offset between CKS and LAMOST (see Appendix~\ref{app:teff}).}
+\edit1{Shifts of +111~K and +116~K are applied to the CKS and LAMOST \teff, respectively.
+The \teff shifts originate from a least-squares fit of a \rocrit~=~\rosun curve (with a \teff offset) to the long-period pileup (as described in \S\ref{subsec:rossby})}.}
     \label{fig:shifted}
@@
 Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
-A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper left panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
+A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper right panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
 Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
@@
 
-\edit1{\mma published weight parameters for each star, a metric which was designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in $T_\mathrm{eff}$–$P_\mathrm{rot}$–LPH space, where LPH designates the local peak height of the ACF.
+\edit1{To understand whether the period detections on the long-period pileup are reliable, we examined the weight parameters published by \mma.
+The weight, $w$, is a metric \mma designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in $T_\mathrm{eff}$–$P_\mathrm{rot}$–LPH space, where LPH designates the local peak height of the ACF.
 Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
@@
 
-As discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}, the existence of a pileup of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}.
+The existence of a pileup of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}, as discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}.
 The location of the long-period pileup coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, \edit1{as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
-We suggest that these features are one and the same.
-The modest differences between the long-period pileup we observe and the features observed by \hall and \masuda may be partially explained by differing \teff scales between the samples.}
-
-%Additionally, both observational and theoretical \teff scales between different studies \edit1{may help} explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as \edit1{differences} in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.
-
-Furthermore, there are important differences between the samples considered here and the \hall asteroseismic sample.
-The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns.
-Exacerbating the problem of detection, slower rotation tends to be accompanied by smaller-scale and less periodic variability.
-The asteroseismic sample, by comparison, is not biased against quiet, unspotted stars and is more likely to contain pileup stars that our sample may have missed.
-\edit1{The same argument is true for the \masuda sample.}
+We suggest that these features are one and the same.}
+
+\edit1{The modest differences between the long-period pileup we observe and the features observed by \hall and \masuda might be explained by differing \teff scales and differing observational biases of the samples.
+The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns (see \S\ref{subsec:detectionbias}).
+The \hall asteroseismic sample and \masuda sample, by comparison, are not biased against quiet, unspotted stars and are more likely to contain pileup stars that our sample may have missed.}
+
+\edit1{In some cases, the \rocrit values we inferred in \S\ref{subsec:rossby} were smaller than the value of \rocrit~$\approx$~\rosun~$\approx$~2 found by \citet{vanSaders2016,vanSaders2019}, sometimes with high statistical significance.
+We caution against overinterpreting the specific \rocrit values inferred here, and enumerate below the ways in which our \rocrit inference may be biased.}
+
+\edit1{(1) The specific \taucz relation adopted has a significant impact on the inferred \rocrit.
+We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, and \citet{Amard2019}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which leads to even smaller inferred \rocrit values.
+We also explored using a fit to the \taucz computed in \citet{vanSaders2019} for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
+This relation revises our \rocrit estimates upwards by 15\%.}
+
+\edit1{(2) While our \rocrit inference procedure includes a constant \teff offset between the data and the \taucz relation, it does not include any \teff-dependent offsets.
+In Appendix~\ref{app:teff}, we show that the LAMOST \teff appear to have a non-linear mapping to the other \teff scales considered here.
+Perhaps not coincidentally, the \lamostmcq sample favors a \rocrit that is the least compatible with other samples.}
+
+\edit1{(3) Defining the long-period pileup or edge is a choice that impacts \rocrit.
+For example, though we chose the 90th \prot percentile in the \lamostmcq sample, \jvs chose the 95th percentile.
+Similarly, our boundary for the CKS long-period pileup was chosen subjectively.}
+
+%We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between \edit1{observations and the models}, \edit1{including} the \taucz relation.
+Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pileups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
 
 It is evident from Figure~\ref{fig:gap} that the long-period pileup also coincides with a steep gradient in variability amplitude, such that a hidden population of pileup stars may lie just beyond the edge of detectability.
-This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values.
+This effect, in addition to the effects mentioned above, can further bias our inference of \rocrit to lower values.
 Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that \edit1{a pileup exists at the long-period edge} for stars with \teff~$\gtrsim$~\teffmin.
@@
 
-\edit1{Our observation of a long-period pileup stands in contrast with recent rotation studies of solar analogs.} \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
+\edit1{Our observation of a long-period pileup stands in contrast with recent rotation studies of solar analogs.
+\citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} Additionally, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
 Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
 A case study of an $\sim$8~Gyr solar twin further reinforced these conclusions \citep{LorenzoOliveira2020}.
-\edit1{Additionally, \citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~\rosun and at ages in the range of $\sim$2--6~Gyr.
+By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~\rosun and at ages in the range of $\sim$2--6~Gyr.
 In contrast to the findings of those authors, we find that the long-period pileup lies far below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.
 \edit1{This implies the braking index must drop to a much lower value at some time after 2.5~Gyr for Sun-like stars.}
-
-We note that the \rocrit values we found in \S\ref{subsec:rossby} are smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however, these results are sensitive to different \teff scales and different prescriptions for \taucz.
-We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, \edit1{and \citet{Amard2019}}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which exacerbates the \rocrit mismatch mentioned above.
-\edit1{We also explored using a polynomial fit to the \taucz computed in \citet{vanSaders2019} for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
-However, even adopting this different \taucz prescription changes the inferred \rocrit upwards by only 15\%, not enough to obtain \rocrit~=~2.}
-
-%Sentence about 95% vs 90% percentiles.
 
@@
 In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia.
+
+\edit1{An alternative explanation for the short-period overdensity is contamination by subgiants.
+Stars which have \teff on the ZAMS higher than $\sim$6250 will never undergo magnetic braking, spinning rapidly for the entirety of their main-sequence lifetimes.
+As these stars evolve away from the main-sequence, their temperatures drop while \logg changes minimally, while the spin rate slows due to radial expansion.
+Removing subgiants using spectroscopic parameters alone is difficult...}
 
@@
 
-    \item Comparison of the feature with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \citet{vanSaders2019}.
+    \item Comparison of the \edit1{long-period pileup} with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \citet{vanSaders2019}.
 Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pileup until an age of $\sim$6~Gyr.
@@
     \item It is yet unclear whether the Sun resides on the long-period pileup or has already evolved through it.
-Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pileup.
-If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
+{\color{red} Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pileup.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
 However, this tension might be simply explained by inaccurate isochrone ages.

### 2022-04-07 src/figures/inflection.py

@@ -87,6 +87,9 @@ print(tau)
 flat_samples = sampler.get_chain(discard=100, thin=15, flat=True)
 print(len(flat_samples)/tau)
 
+for j in range(np.shape(flat_samples)[1]):
+    print(j, np.median(flat_samples[:,j]), np.std(flat_samples[:,j]))
+
 #Stars with large temperature differences between CKS (Fulton & Petigura 2018) & SPOCS (Brewer & Fischer 2018)
 arg = (abs(cks['bf18_Teff']-cks['cks_Teff'])>100)
 
@@ -104,7 +107,7 @@ fig,ax = plt.subplots(nrows=1,ncols=1,figsize=(5,5))
 
 ax.plot(lam['Teff_lam'], lam['Prot'], 'o', ms=4, alpha=0.4, color='lightgrey', label='LAMOST–McQuillan', mew=0, rasterized=True)
 ax.plot(cks['cks_Teff'], cks['d21_prot'], 'o', ms=2, label='CKS–David', rasterized=True)
-ax.plot(cks['cks_Teff'][arg], cks['d21_prot'][arg], 'o', ms=4, mfc='None', mew=0.5, color='r')
+ax.plot(cks['cks_Teff'][arg], cks['d21_prot'][arg], 'o', ms=5, mfc='None', mew=0.5, color='k')
 
 ax.errorbar(5200,1,xerr=np.nanmedian(cks['p20_cks_steff_err1']),yerr=0,fmt='.',color='k', zorder=999)
 ax.errorbar(5200,3,xerr=np.nanmedian(lam['e_Teff_lam'][lam['Teff_lam']>5800]),yerr=0,fmt='.',color='k', zorder=999)

### 2022-04-07 src/figures/mcmc.py

@@ -372,7 +372,7 @@ def rossby_plot(teff_bin_centers, flat_samples_10, flat_samples_90):
     ax3.set_xlabel("Effective temperature [K]")
     ax3.set_ylabel("Residuals [%]")
 
-    sns.despine()
+    #sns.despine()
     plt.tight_layout()
     plt.savefig('../figures/mcmc.pdf')
     

### 2022-04-07 src/figures/rocrit.py

@@ -77,10 +77,24 @@ lam = lam[lam_mask]
 
 
 ######################################################################################
-def convective_turnover_timescale(teff):
+def convective_turnover_timescale(teff, author='CS11'):
     #Returns convective turnover timescale in days
-    #Gunn et al. 1998 relation, from Cranmer & Saar 2011
-    return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
+    if author=='CS11':
+        #Gunn et al. 1998 relation, from Cranmer & Saar 2011
+        return 314.24*np.exp( -(teff/1952.5) - (teff/6250.)**18. ) + 0.002
+    elif author=='VS19':
+        # coeff = [1.61117853e-14, 
+        #      -4.48118958e-10,  
+        #      4.97516474e-06, 
+        #      -2.75645645e-02,
+        #      7.62100664e+01, 
+        #      -8.40866926e+04]
+        # return sum([co*teff**i for i,co in enumerate(coeff[::-1])])
+        a = 1.50290736e+02
+        b = 2.71200127e+03
+        c = 6.35361887e+03
+        d = -3.42283628e+00
+        return a * np.exp(- (teff/b) - (teff/c)**18) + d
 
 def constant_rossby(teff, ro):
     #Return locus of rotation periods corresponding to constant Rossby number
@@ -88,7 +102,8 @@ def constant_rossby(teff, ro):
 
 def rocrit_teff_shift(teff, teff_shift):
     #Return locus of rotation periods corresponding to constant Rossby number
-    return 1.97 * convective_turnover_timescale(teff+teff_shift)
+    rosun = sun["prot"]/convective_turnover_timescale(sun["teff"])
+    return rosun * convective_turnover_timescale(teff+teff_shift)
 
 lam["Ro"] = lam["Prot"]/convective_turnover_timescale(lam["Teff_lam"])    
 ######################################################################################
@@ -155,7 +170,7 @@ def lamost_teff_detrend(teff):
     dteff = 2.55513439e-13*teff**5 - 7.18129973e-09*teff**4 + 8.04175914e-05*teff**3 - 4.48417848e-01*teff**2 + 1.24490338e+03*teff - 1.37649898e+06
     return teff-dteff
 
-lam["Teff_lam"] = lamost_teff_detrend(lam["Teff_lam"])
[…]

### 2022-04-07 src/figures/shifted.py

@@ -100,8 +100,15 @@ roc['flag'] = 'roc'
 model = pd.concat([std, roc], ignore_index=True, sort=True)
 ######################################################################################
 
-cks_shift = 116
-lamost_shift = 106 + 34
+cks_shift = 111
+lamost_shift = 116
+
+#Detrending LAMOST Teff
+# def lamost_teff_detrend(teff):   
+#     dteff = 2.55513439e-13*teff**5 - 7.18129973e-09*teff**4 + 8.04175914e-05*teff**3 - 4.48417848e-01*teff**2 + 1.24490338e+03*teff - 1.37649898e+06
+#     return teff-dteff
+
+# lam["Teff_lam"] = lamost_teff_detrend(lam["Teff_lam"])
 
 
 mpl.rcParams["legend.markerscale"] = 1

### 2022-04-07 src/ms.tex
paper edits

@@
 \newcommand{\masuda}{MPH21\xspace}
+\newcommand{\curtis}{C20\xspace}
+\newcommand{\santos}{S21\xspace}
 
@@
 Forward modeling simulations of the observed Kepler rotation period distribution also provided support for the WMB hypothesis over standard spin-down models, in that WMB models are better able to match the observed long-period edge \citep[][hereafter vS19]{vanSaders2019}.
-Those authors also predicted a pileup of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
+Those authors also predicted a pileup of stars along the long-period edge, which they hypothesized could not be seen in the \citet{McQuillan2014} sample (hereafter \mma) due to large errors on \teff in the revised Kepler Input Catalog \citep[KIC,][]{Huber2014}.
 While \jvs favored the WMB hypothesis to explain observations, those authors were also careful to point out that a long-period edge can be caused by detection biases, as stars with longer rotation periods (and larger Rossby numbers) have smaller amplitude variations which pose more difficulty to period-detection algorithms.
 
-More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \hall sample had rotation periods that were also detected from rotational brightness modulations.} from the \citet{vanSaders2019} study and the present work, to argue support for the WMB model.
+More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \hall sample had rotation periods that were also detected from rotational brightness modulations.} from the \jvs study and the present work, to argue support for the WMB model.
 \citet{Masuda2021}, \edit1{hereafter MPH21}, also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
@@
     \includegraphics[width=\textwidth]{./figures/surveys.pdf}
-    \caption{The \teff-\prot plane using rotation periods from \citet{McQuillan2014} or, in the case of the CKS sample, \citet{McQuillan2013} which applied an identical analysis to Kepler Objects of Interest (KOIs), with \teff originating from the source denoted at top.
-The \citet{McQuillan2014} \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars.
+    \caption{The \teff-\prot plane using rotation periods from \mma or, in the case of the CKS sample, \citet{McQuillan2013} which applied an identical analysis to Kepler Objects of Interest (KOIs), with \teff originating from the source denoted at top.
+The \mma \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars.
 The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the \edit1{activity belts}.
-Many of the stars above the long-period pileup are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \citet{vanSaders2019}.}
+Many of the stars above the long-period pileup are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \jvs.}
     \label{fig:surveys}
@@
 \edit1{A description of the LAMOST Kepler field observations is provided in \citet{deCat2015}.} The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
-We cross-matched the \edit1{LAMOST DR5} stellar parameter catalog of \citet{Xiang2019} with the \citet{McQuillan2014} catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog, which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
-
-We matched 10844 LAMOST targets to 10550 unique Kepler IDs in the \citet{McQuillan2014}, resulting in a sample with well-determined \prot and spectroscopic \teff (having a median error of 24~K).
+We cross-matched the \edit1{LAMOST DR5} stellar parameter catalog of \citet{Xiang2019} with the \mma catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog (hereafter S21), which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
+
+We matched 10844 LAMOST targets to 10550 unique Kepler IDs in \mma, resulting in a sample with well-determined \prot and spectroscopic \teff (having a median error of 24~K).
 For the Kepler sources with duplicate cross-matched LAMOST sources we kept the source with a brighter Gaia DR2 $G$ magnitude.
-In the \citet{Santos2021} sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
+In the \santos sample we found 54982 unique cross-matched sources in LAMOST, of which 18990 have published temperatures and rotation periods.
 The metallicity distribution of the LAMOST--McQuillan sample is centered near solar, with a mean and standard deviation of $-0.1$~dex and 0.26~dex, respectively.
-There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since the \citet{McQuillan2014} did not publish rotation periods for KOIs, which were the targets of the CKS project.
+There is negligible overlap (3 stars) between our LAMOST--McQuillan sample and the CKS sample since \mma did not publish rotation periods for KOIs, which were the targets of the CKS project.
 Rotation periods for KOIs are instead published in \citet{McQuillan2013}, as discussed in \S\ref{subsec:cks}.
@@
 
+\newpage
 \subsection{\edit1{APOGEE}}
@@
 We used Gaia DR2 source IDs \citep{Gaia2016, Gaia2018} to cross-match Megan Bedell's Gaia--Kepler catalog\footnote{\url{https://gaia-kepler.fun/}} with the APOGEE DR16 catalog \citep{Ahumada2020}.
-Kepler IDs were then used to cross-match this table with the \citet{McQuillan2014} catalog.
+Kepler IDs were then used to cross-match this table with the \mma catalog.
 While the current overlap between Kepler targets and APOGEE is small compared to the LAMOST catalog, APOGEE DR17 will contain more dwarf stars and provide a better resource for studies such as ours.
@@
 The long-period pileup for dwarf stars is clearly seen to extend to the solar temperature.
-The short-period pileup is clearer in the smaller \lamostmcq dwarf sample, potentially because the \citet{Santos2021} catalog detected more stars at longer periods.
+The short-period pileup is clearer in the smaller \lamostmcq dwarf sample, potentially because the \santos catalog detected more stars at longer periods.
 The secondary overdensity observed in the subgiant samples, most visible in the bottom center panel, appears to be at twice the period of the primary overdensity, potentially due to erroneously determined rotation periods.}
@@
 We first noted a pileup of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-\edit1{The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \citet{Santos2021} revealed that this pileup is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
+\edit1{The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \santos revealed that this pileup is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
+
+\edit1{Comparing the long-period pileup to rotation period sequences from open clusters can provide insight into the ages of stars on the pileup.} The \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020}, hereafter \curtis, from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
+The observation that the \edit1{long-period pileup} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
+Similarly, the \edit1{long-period pileup clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pileup}.
+These observations are in accordance with predictions from the WMB model which suggest the pileup forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
+\edit1{We show in \S\ref{subsec:ages} that stars along the pileup have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
 
@@
     \caption{Gaussian kernel density estimation (blue contours) of the \teff--\prot distributions of the CKS--McQuillan, \lamostmcq, and asteroseismic \hall samples, from left to right.
-Empirical cluster sequences from \citet{Curtis2020} are shown by the dark grey lines.
+Empirical cluster sequences from \curtis are shown by the dark grey lines.
 The orange dashed lines show constant Rossby curves of fiducial values (see \S\ref{subsec:rossby}).
@@
 \end{figure*}
-
-\edit1{Comparing the long-period pileup to rotation period sequences from open clusters can provide insight into the ages of stars on the pileup.} The \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020} from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
-Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
-The observation that the \edit1{long-period pileup} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the \edit1{long-period pileup clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pileup}.
-These observations are in accordance with predictions from the WMB model which suggest the pileup forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
-\edit1{We show in \S\ref{subsec:ages} that stars along the pileup have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
 
@@
 For 100 bootstrapped resamplings of the \teff–\prot distribution, leaving out 50\% of the data for each bootstrapped sample, we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
-The final long-period edge curve was then computed as the mean of the bootstrapped 90th percentile values.
-We assumed constant 10\% fractional errors on the \prot values computed in this way.} We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
-Since we observed a break in the 90th percentile \prot curve near \teff~=~4250~K, due to smaller sample sizes at low \teff, we restricted the analysis below to stars with \teff$>~4250$~K.
+The final long-period edge curve was then computed as the mean of the bootstrapped 90th percentile values.} We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
 We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
@@
 \begin{multline} \label{eq:3}
-    \ln{p} (y | T_\mathrm{eff}, T_\mathrm{sys}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
+    \ln{p} (y | T_\mathrm{eff}, T_\mathrm{sys}, \sigma, \mathrm{Ro}, f) =\\ -\frac{1}{2}\sum_n \left [ \frac{(y_n - P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} + T_\mathrm{sys}))^2}{s_n^2}  + \ln{(2\pi s_n^2)} \right ],
 \end{multline}
@@
 \begin{equation} \label{eq:4}
-    s_n^2 = \sigma^2 + f^2 P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} - T_\mathrm{sys})^2,
+    s_n^2 = \sigma^2 + f^2 P_\mathrm{rot}(\mathrm{Ro}, T_\mathrm{eff} + T_\mathrm{sys})^2,
 \end{equation}
@@
 \edit1{In some cases, our constant Rossby fits to the long-period pileup prefer Rossby numbers lower than the Sun's, in contrast with \jvs.
-We caution against overinterpreting the specific \rocrit values inferred here, and discuss our interpretations further in \S\ref{subsec:longperiod}.
-To further quantify the preference for \rocrit~$<$~\rosun} we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K) using the \texttt{curve\_fit} function in \texttt{scipy.optimize}.
+We caution against overinterpreting the specific \rocrit values inferred here, and discuss our interpretations further in \S\ref{subsec:longperiod}.}
+
+\edit1{To further quantify the preference for \rocrit~$<$~\rosun} we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K).
 The first model assumes \rocrit = \rosun and has one free parameter, a temperature offset added to the data.
@@
 We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = \rosun and variable \rocrit models was negligible.
-In other words, the \rocrit = \rosun model fits the data well with a \teff shift of $\approx$~80--150~K depending on the \teff source used.
+In other words, the \rocrit = \rosun model fits the data well with a \teff shift of $\approx$~\edit1{70--140}~K depending on the \teff source used.
 For the \lamostmcq sample, there is weak support for the variable \rocrit model.
@@
 \hline
-\lamostmcq 10th \prot pctl. & 4500~K--5800~K & 0.37 $\pm$ 0.02 & 0.011 $\pm$ 0.009 &  46 $\pm$ 80 \\
+\lamostmcq 10th \prot pctl. & 4500~K--5800~K & 0.37 $\pm$ 0.02 & 0.011 $\pm$ 0.009 &  -46 $\pm$ 80 \\
 \hline 
@@
 %\lamostmcq 90th \prot pctl. & 5000~K--6250~K & 1.38 $\pm$ 0.04 & 0.011 $\pm$ 0.008 & 49 $\pm$ 19 \\
-\lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.4 $\pm$ 0.1 & 0.019 $\pm$ 0.015 & 44 $\pm$ 43 \\
-CKS (CKS \teff) & 5800~K--6250~K & 1.77 $\pm$ 0.07 & 0.02 $\pm$ 0.01 & -70 $\pm$ 15 \\ 
-CKS (SPOCS \teff) & 5800~K--6250~K & 1.9 $\pm$ 0.1 & 0.12 $\pm$ 0.01 & -136 $\pm$ 23 \\
-CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.2 & 0.18 $\pm$ 0.02 & 21 $\pm$ 42 \\
+\lamostmcq 90th \prot pctl. & 5800~K--6250~K & 1.4 $\pm$ 0.1 & 0.019 $\pm$ 0.015 & -44 $\pm$ 43 \\
+CKS (CKS \teff) & 5800~K--6250~K & 1.77 $\pm$ 0.07 & 0.02 $\pm$ 0.01 & 70 $\pm$ 15 \\ 
+CKS (SPOCS \teff) & 5800~K--6250~K & 1.9 $\pm$ 0.1 & 0.12 $\pm$ 0.01 & 136 $\pm$ 23 \\
+CKS (M19 \teff) & 5800~K--6250~K & 1.6 $\pm$ 0.2 & 0.18 $\pm$ 0.02 & -21 $\pm$ 42 \\
 %\citet{Hall2021} main-sequence & 5000~K--6250~K & 1.7 $\pm$ 0.2 & 0.30 $\pm$ 0.04 & -5 $\pm$ 64 \\
-\citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & -29 $\pm$ 79 \\
+\citet{Hall2021} main-sequence & 5800~K--6250~K & 1.9 $\pm$ 0.4 & 0.27 $\pm$ 0.05 & 29 $\pm$ 79 \\
 \enddata
@@
 
-We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}.
+We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \jvs.
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
 \edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
-}.}} \edit1{We note that we did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
+}.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
 The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
 
-The theoretically predicted \teff--\prot distributions of \citet{vanSaders2019} are shown in relation to the observations in Figure~\ref{fig:models}.
+The theoretically predicted \teff--\prot distributions of \jvs are shown in relation to the observations in Figure~\ref{fig:models}.
 Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-The specific WMB prescription of \citet{vanSaders2019} adopted a critical Rossby number of \rocrit~=~2.08 \edit1{(using a different \taucz prescription than the one used here)}, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations \edit1{(and assuming no \teff offset between the models and the data)}.
+The specific WMB prescription of \jvs adopted a critical Rossby number of \rocrit~=~2.08 \edit1{(using a different \taucz prescription than the one used here)}, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations, \edit1{assuming no \teff offset between the models and the data}.
 Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
@@
 
-While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.
+While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \jvs, as well as differences in the computation of \taucz.
 We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \edit1{\lamostmcq} may be slightly different.
 
-Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models.
+Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models, \edit1{and in Appendix~\ref{app:teff} we show the LAMOST \teff are $\sim$50--100~K cooler than the other surveys considered here}.
 In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
@@
     \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
-    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}.
+    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \jvs.
 Shown here are stars with \logg~$>$~4.1.
@@
     \includegraphics[width=\linewidth]{./figures/percentiles.pdf}
-    \caption{Discrepancies between the observed and theoretically predicted \teff--\prot distribution of Kepler stars.
-Comparison of the 10th and 90th \prot percentiles predicted from the standard (light blue) and WMB (blue) models of \citet{vanSaders2019} with the same percentile ranges from the LAMOST--McQuillan sample (black points).}
+    \caption{\edit1{Comparison of the theoretically predicted \teff--\prot distribution of Kepler stars from \jvs (given by the 10th and 90th \prot percentiles) with the same percentile ranges from the \lamostmcq sample (black points) and the CKS long-period pileup (open circles).}}
     \label{fig:percentiles}
@@
 
-{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pileup as seen in the two samples.}
+%{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pileup as seen in the two samples.}
 
@@
 Both catalogs derive spectroscopic parameters from the CKS spectra, but use different pipelines for both the spectroscopic parameters and the isochrone fitting.
+
 In both cases, there appears to be an age gradient such that hotter stars are younger on average.
@@
 \edit1{These ranges are consistent with the range of ages observed in the \hall and \masuda samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
-The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
+\edit1{For example, it is curious that, when using SPOCS ages} only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
 In fact, there are only a handful of long-period pileup stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pileup stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
@@
 Below, the \teff-age plane for CKS stars along the long-period pileup using isochrone ages from the CKS (c) and SPOCS (d) catalogs.
-The light shaded points represent predictions from the WMB model of \citet{vanSaders2019} for stars with Ro~$<$~2 (blue) and Ro~$>$~2 (orange).
+The light shaded points represent predictions from the WMB model of \jvs for stars with Ro~$<$~2 (blue) and Ro~$>$~2 (orange).
 Note, panels (a), (c), and (d) all pertain to the CKS sample, whereas panel (b) relates to the \lamostmcq sample, for which homogeneous ages have not been published.}
@@
 
-\edit1{To further assess the evolutionary state of stars on the long-period pileup, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \citet{McQuillan2014} and \citet{Santos2021} rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pileup stars from this work (Figure~\ref{fig:cmd}).
-Stars with detected periods from rotational brightness modulations are almost exclusively solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
-The \citet{Santos2021} catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \citet{McQuillan2014}.
+\edit1{To further assess the evolutionary state of stars on the long-period pileup, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \mma and \santos rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pileup stars from this work (Figure~\ref{fig:cmd}).
+Stars with detected periods from rotational brightness modulations are primarily solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
+The \santos catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \mma.
 CKS stars on the long-period pileup clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.}
@@
 The estimated position of the Sun is indicated by the black point, based on the calibration of \citet{Casagrande2018}.
-In the middle and right panels, the \citet{Santos2021} targets with photometric rotation period detections are overplotted in orange.
+In the middle and right panels, the \santos targets with photometric rotation period detections are overplotted in orange.
 In the right panel, the CKS long-period pileup stars identified in this work are shown by blue points and a Gaussian kernel density estimation of the \hall asteroseismic sample is shown by blue contours.
@@
 \subsection{Where do the pileups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pileup} declines towards cooler \teff, as predicted by the WMB model \citep[see Figure 13 of][]{vanSaders2019}.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pileup} declines towards cooler \teff, as predicted by the WMB model (see Figure~13 of \jvs).
 \edit1{The number density of stars on the short-period pileup similarly declines towards cooler \teff.
@@
     \includegraphics[width=\linewidth]{./figures/weights.pdf}
-    \caption{The \teff–\prot distribution of the \lamostmcq sample for different weight thresholds (indicated above each panel) computed by \citet{McQuillan2014}.}
+    \caption{The \teff–\prot distribution of the \lamostmcq sample for different weight thresholds (indicated above each panel) computed by \mma.}
     \label{fig:weights}    
@@
 \edit1{In some cases, the \rocrit values we inferred in \S\ref{subsec:rossby} were smaller than the value of \rocrit~$\approx$~\rosun~$\approx$~2 found by \citet{vanSaders2016,vanSaders2019}, sometimes with high statistical significance.
-We caution against overinterpreting the specific \rocrit values inferred here, and enumerate below the ways in which our \rocrit inference may be biased.}
-
-\edit1{(1) The specific \taucz relation adopted has a significant impact on the inferred \rocrit.
-We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, and \citet{Amard2019}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, which leads to even smaller inferred \rocrit values.
-We also explored using a fit to the \taucz computed in \citet{vanSaders2019} for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
+We caution against overinterpreting the \rocrit values inferred here, and enumerate below the ways in which the \rocrit inference may be biased.}
+
+\begin{enumerate}
+    \item \edit1{The \taucz relation adopted impacts the inferred \rocrit.
+We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, and \citet{Amard2019}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, leading to even smaller inferred \rocrit values.
+We also explored using a fit to the \taucz computed in \jvs for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
 This relation revises our \rocrit estimates upwards by 15\%.}
-
-\edit1{(2) While our \rocrit inference procedure includes a constant \teff offset between the data and the \taucz relation, it does not include any \teff-dependent offsets.
+    \item \edit1{While our \rocrit inference procedure includes a constant \teff offset between the data and the \taucz relation, it does not include any \teff-dependent offsets.
 In Appendix~\ref{app:teff}, we show that the LAMOST \teff appear to have a non-linear mapping to the other \teff scales considered here.
-Perhaps not coincidentally, the \lamostmcq sample favors a \rocrit that is the least compatible with other samples.}
-
-\edit1{(3) Defining the long-period pileup or edge is a choice that impacts \rocrit.
+Perhaps not coincidentally, the \lamostmcq sample favors a \rocrit that is the least compatible with other samples.
+We also note that the MCMC results for the \lamostmcq sample preferred the LAMOST \teff to be shifted lower (Table~\ref{tab:mcmc}), which is in inconsistent with our finding that the LAMOST \teff are already lower than those from other spectroscopic surveys (Appendix~\ref{app:teff}).}
+    \item \edit1{Defining the long-period pileup or edge is a choice that impacts \rocrit.
 For example, though we chose the 90th \prot percentile in the \lamostmcq sample, \jvs chose the 95th percentile.
 Similarly, our boundary for the CKS long-period pileup was chosen subjectively.}
+\end{enumerate}
 
@@
 By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~\rosun and at ages in the range of $\sim$2--6~Gyr.
-In contrast to the findings of those authors, we find that the long-period pileup lies far below an empirical 2.5~Gyr gyrochrone of \citet{Curtis2020} that is evolved forward to 5.3~Gyr for braking indices of $n=0.5$ or $n=0.65$.
+In contrast to the findings of those authors, we find that the long-period pileup lies far below an empirical 2.5~Gyr gyrochrone of \curtis that is evolved forward to 5~Gyr for braking indices of $n=0.5$ or $n=0.65$ (Figure~\ref{fig:skumanich}).
 \edit1{This implies the braking index must drop to a much lower value at some time after 2.5~Gyr for Sun-like stars.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{./figures/skumanich.pdf}
+    \caption{A 2-d histogram of the \lamostmcq \teff–\prot distribution compared to an empirical $\approx$~2.5~Gyr hybrid cluster sequence from \curtis (solid line).
+The dashed and dotted lines show the same sequence evolved forward to 5~Gyr assuming braking indices of $n=0.5$ (Skumanich-like spin-down) and $n=0.65$, which is favored by \curtis.
+For comparison, the white points indicate the peak of the \prot distribution from \masuda and the black point indicates the Sun.}
+    \label{fig:skumanich}    
+\end{figure}
 
@@
 After a period of slowed spin-evolution, stars must resume spinning down as evidenced from studies of older open clusters.
-From open cluster data \citet{Curtis2020} estimated that solar-mass stars resume spin-down after an age of $\approx$~230~Myr.
+From open cluster data \curtis estimated that solar-mass stars resume spin-down after an age of $\approx$~230~Myr.
 Thus, if core–envelope coupling is responsible for delaying the spin-down of stars, and if the theoretical core–envelope coupling timescales are accurate, then we may expect Sun-like stars to experience slowed spin evolution between $\sim$100~Myr and $\sim$200~Myr.
@@
 Both the theoretically predicted core–envelope coupling timescale and the observationally inferred timescale for the resumption of spin-down are consistent with our observation that the short-period pileup is intermediate to the Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) cluster sequences.
-Furthermore, in order to match observations of rotation periods in young clusters, models require that the core–envelope coupling timescale increases towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars \citep{Curtis2020}.
+Furthermore, in order to match observations of rotation periods in young clusters, models require that the core–envelope coupling timescale increases towards lower stellar masses \citep[e.g.][]{Irwin2007, Denissenkov2010, GalletBouvier2015}, which provides a natural explanation for why this stalling occurs at older ages for lower-mass stars in \curtis.
 
@@
 \citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}.
-However, \citet{Curtis2020} demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation.
+However, \curtis demonstrated that cluster sequences cross the gap, invalidating the claim that the feature is caused at a specific age, as one might expect from a period of decreased star formation.
 Additionally, \citet{Gordon2021} found that the gap is observed across the many fields observed by the K2 mission, which is in tension with the star formation history hypothesis as different Galactic sight lines are expected to have different different star formation histories (for sufficiently large volumes).
@@
 The short- and long-period pileups we examine here naturally produce a dearth of rotators at intermediate periods.
-This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \citet{McQuillan2014} sample.
+This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \mma sample.
 Moreover, as seen in Figure~\ref{fig:gap}, we recover the gradient in photometric variability across the gap pointed out by \citet{Reinhold2019}.
-The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum near the location of the gap.
+The photometric variability, as measured through the \rper metric published by \mma, reaches a local minimum near the location of the gap.
 This supports the notion that the cause of the gap is due to changes in the stars themselves, rather than being the result of mixed stellar populations.
@@
 \citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core–envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
-We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here.
+We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here \edit1{(see also Appendix~\ref{app:gaia})}.
 Notably, if the short-period pileup is indeed due to core–envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
@@
     \includegraphics[width=\linewidth]{./figures/gap.pdf}
-    \caption{The \teff--\prot distribution of the \lamostmcq (left) and \citet{McQuillan2014} samples color coded by the variability amplitude, \rper.
+    \caption{The \teff--\prot distribution of the \lamostmcq (left) and \mma samples color coded by the variability amplitude, \rper.
 Black contours show Gaussian kernel density estimation of the plotted distributions, and the dashed line shows a constant Rossby curve of Ro~=~0.5.
@@
 From Fig.~\ref{fig:xmatch} it is clear that the long-period pileup extends to temperatures as cool as the Sun's.
-However, it is not clear whether there is a large systematic offset in the LAMOST temperature scale and the temperature scale of the Sun.
-If one assumes there are no systematic offsets between the two \teff scales, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pileup.
+If one assumes there are no systematic offsets between \edit1{the spectroscopic \teff scales considered here and that of the Sun}, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pileup.
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
 
-We consider the most likely explanation for the observed gap between the Sun and the long-period pileup to be systematic offsets in \teff.
-To place the Sun on the long-period pileup would require shifting the spectroscopic temperatures higher by $\sim$50--100~K for the CKS sample, $\sim$100--200~K for the SPOCS sample, or $\sim$200--300~K for the LAMOST sample.
-Notably, comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pileup, or very close to it, without the need for a \teff shift.
+\edit1{If the Sun is indeed on the long-period pileup, it likely the presence of \teff systematics and/or observational biases that obfuscate that fact.} To place the Sun on the long-period pileup would require shifting the spectroscopic temperatures higher by \edit1{111}~K for the CKS sample, \edit1{143}~K for the SPOCS sample, or \edit1{116}~K for the LAMOST sample.
+\edit1{We also note that there are \teff-dependent systematics in the LAMOST \teff that, if corrected for, would bring the \lamostmcq long-period pileup into closer agreement with a solar Rossby curve.} Comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pileup, or very close to it, without the need for a \edit1{large} \teff shift.
+\edit1{This apears to be at odds with the notion that a \teff shift alone can place the Sun on the long-period pileup, since the \hall and CKS temperature scales agree to within $\sim$30~K.
+However, it is still possible that there are \teff-dependend systematics in the temperature scales.}
+
 Furthermore, as discussed in \S\ref{subsec:longperiod}, the \hall sample is less prone to detection bias which acts to censor stars with higher Ro and lower variability amplitudes in the samples considered here.
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
-We conclude that the Sun likely does reside on the long-period pileup, but that this is obfuscated in the present samples by temperature offsets as well as potentially detection biases.
+\edit1{Regardless, it is clear from the comparisons in \S\ref{subsec:asteroseismic} that the long-period pileup can not be much broader (in \prot or Rossby space) than we observe.}
+
+\edit1{If the Sun is not on the long-period pileup, this would seem to imply any number of unusual explanations.
+For example, the Sun's angular momentum evolution may have been different from those stars on the pileup.
+Alternatively, spin-down may proceed on the main-sequence after the pileup either through magnetic braking or some other mechanism.
+We consider these explanations unlikely, and suggest it is likely some combination of \teff systematics and observational biases that complicate contextualizing the Sun.}
 
@@
 We hypothesize that this pileup was previously obfuscated by imprecise \teff estimates.
-Both the existence of the pileup and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
+Both the existence of the pileup and its obscuration by large \teff errors were predicted by \jvs as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
 
@@
 
-    \item Comparison of the \edit1{long-period pileup} with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \citet{vanSaders2019}.
+    \item Comparison of the \edit1{long-period pileup} with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \jvs.
 Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pileup until an age of $\sim$6~Gyr.
@@
 This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly} because the short-period pileup is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pileup ($>10^9$~yr).
-The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \citet{Curtis2020}.
+The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \curtis.
 The short-period pileup can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pileup.
@@
 The black trapezoid indicates the approximate area of the ridge.
-The grey curves indicate empirical cluster sequences from \citet{Curtis2020}, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
+The grey curves indicate empirical cluster sequences from \curtis, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
 %     \label{fig:ridge}
@@
 In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
-We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \citet{McQuillan2014} study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
+We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \mma study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
 A LAMOST \teff scale which is systematically cooler provides support to the notion that \rocrit determined from LAMOST temperatures will be systematically underestimated.
@@
     \includegraphics[width=\linewidth]{./figures/teffscales.pdf}
-    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \citet{McQuillan2014}, and \hall.}
+    \caption{Comparison of \teff estimates from different catalogs: LAMOST \citep{Xiang2019}, CKS \citep{Fulton2018}, SPOCS \citep{Brewer2018}, M19 \citep{Martinez2019}, \mma, and \hall.}
     \label{fig:teffscales}
@@
 \edit1{We also note that the short-period pileup is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
-The differences between the \citet{McQuillan2014} and \citet{Santos2021} catalogs may provide an explanation for this observation.
-The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \citet{Santos2021} may then make the short-period pileup appear weaker relative to the long-period pileup.
+The differences between the \mma and \santos catalogs may provide an explanation for this observation.
+The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \santos may then make the short-period pileup appear weaker relative to the long-period pileup.
 This is as expected since open cluster obsbervations suggest that the short-period pileup can not be a long-lived feature, while the long-period pileup appears to be much longer lived.
-Consequently, the \citet{Santos2021} catalog may more accurately reflect the relative strengths of these two features.}
+Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.}
 
@@
 \edit1{Given that the long-period pileup was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
-We cross-matched the \citet{McQuillan2014} and \citet{Santos2021} samples with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
+We cross-matched the \mma and \santos samples with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
 We retrieved reddening estimates for each target by querying the \texttt{Bayestar19} 3D dust map using the \texttt{dustmaps} package \citep{dustmaps, Green2019}.
@@
 Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}, suggesting the origin of the discontinuity is in the rotation periods rather than the temperatures or colors.
-At present, we do not have a satisfactory explanation for this behavior, though we note that the \citet{Santos2021} catalog employed various time series filters which might introduce systematic artifacts.}
+At present, we do not have a satisfactory explanation for this behavior, though we note that the \santos catalog employed various time series filters which might introduce systematic artifacts.}
 
@@
     \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
-    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
+    \caption{The color–\prot plane for the \mma sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
 The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.
@@
 The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.
-The grey lines indicate empirical cluster sequences derived in \citet{Curtis2020}.
+The grey lines indicate empirical cluster sequences derived in \curtis.
 From top to bottom, the cluster sequences are Pleiades (0.12~Gyr), Praesepe (0.67~Gyr), NGC 6811 (1~Gyr), and NGC 6819 + Ruprecht 147 (2.5~Gyr).}
@@
     \includegraphics[width=\linewidth]{./figures/gaia-santos.pdf}
-    \caption{The same as Figure~\ref{fig:gaia-mcquillan} but for the \citet{Santos2021} sample.}
+    \caption{The same as Figure~\ref{fig:gaia-mcquillan} but for the \santos sample.}
     \label{fig:gaia-santos}

### 2022-04-07 src/figures/inflection.py

@@ -103,7 +103,7 @@ sun = {"teff": 5772,
 sns.set(font_scale=1.3, context="paper", style="ticks")
 
 
-fig,ax = plt.subplots(nrows=1,ncols=1,figsize=(5,5))
+fig,ax = plt.subplots(nrows=1,ncols=1,figsize=(5,3))
 
 ax.plot(lam['Teff_lam'], lam['Prot'], 'o', ms=4, alpha=0.4, color='lightgrey', label='LAMOST–McQuillan', mew=0, rasterized=True)
 ax.plot(cks['cks_Teff'], cks['d21_prot'], 'o', ms=2, label='CKS–David', rasterized=True)

### 2022-04-07 src/figures/kde.py

@@ -192,10 +192,6 @@ hall.info()
 
 sns.set(style='ticks', font_scale=1.4, context='paper')
 
-
-# In[4]:
-
-
 sns.set(style='ticks', font_scale=1.6, context='paper')
 
 fig,(ax1,ax2,ax3) = plt.subplots(nrows=1, ncols=3, 

### 2022-04-07 src/ms.tex
adding Skumanich figure

@@
 For the \lamostmcq sample, there is weak support for the variable \rocrit model.
-However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend \edit1{(i.e.
+However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend not adequately captured by a constant offset \edit1{(i.e.
 a \teff-dependent \teff offset)} in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
 Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure.
-Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pileup in the CKS vs.
+
+%Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pileup in the CKS vs.
 \lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pileup and the \lamostmcq 90th percentile curve, we observe in Figure~\ref{fig:percentiles} a systematic trend in which most CKS pileup stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.
@@
 We hypothesize that this pileup was previously obfuscated by imprecise \teff estimates.
-Both the existence of the pileup and its obscuration by large \teff errors were predicted by \jvs as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
+Both the existence of the pileup and its obscuration by large \teff errors were predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking for stars with $M\gtrsim$1~\msun.
 
@@
     \item It is yet unclear whether the Sun resides on the long-period pileup or has already evolved through it.
-{\color{red} Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pileup.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
+\edit1{Offsets of $\approx$110--140~K between the observational and theoretical \teff scales allow for the long-period pileup to be fit with a solar value Rossby curve, thereby placing the Sun on the pileup.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
 However, this tension might be simply explained by inaccurate isochrone ages.
@@
     \item We \edit1{tentatively detect} a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly} because the short-period pileup is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pileup ($>10^9$~yr).
+This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly indicating that} the short-period pileup is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pileup ($>10^9$~yr).
 The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \curtis.

### 2022-04-07 src/figures/inflection.py

@@ -103,7 +103,7 @@ sun = {"teff": 5772,
 sns.set(font_scale=1.3, context="paper", style="ticks")
 
 
-fig,ax = plt.subplots(nrows=1,ncols=1,figsize=(5,3))
+fig,ax = plt.subplots(nrows=1,ncols=1,figsize=(5,5))
 
 ax.plot(lam['Teff_lam'], lam['Prot'], 'o', ms=4, alpha=0.4, color='lightgrey', label='LAMOST–McQuillan', mew=0, rasterized=True)
 ax.plot(cks['cks_Teff'], cks['d21_prot'], 'o', ms=2, label='CKS–David', rasterized=True)

### 2022-04-07 src/figures/skumanich.py

@@ -1,0 +1,186 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+import numpy as np
+import pandas as pd
+
+from astropy.table import Table
+
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+mpl.rcParams["figure.dpi"] = 100
+mpl.rcParams["savefig.bbox"] = "tight"
+mpl.rcParams["savefig.dpi"] = 300
+
+import seaborn as sns
+import astropy.constants as c
+from scipy import interpolate
+from labellines import labelLine, labelLines
+
+sun = {"teff": 5772,
+       "prot": 25.4,
+       "e_prot": 25.4-24.5,
+       "E_prot": 27-25.4
+      }
+
+sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
+
+#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
+
+def convective_turnover_timescale(teff):
+    #Returns convective turnover timescale in days
+    #Gunn et al. 1998 relation, from Cranmer & Saar 2011
+    return 314.24*np.exp(-(teff/1952.5) - (teff/6250.)**18.) + 0.002
+
+
+def constant_rossby(teff, ro):
+    #Return locus of rotation periods corresponding to constant Rossby number
+    return ro * convective_turnover_timescale(teff)
+
+# Curtis et al. 2020 table
+def curtis_bprp_teff(bprp):
+    #Estimating effective temperature from the dereddened Gaia DR2 (Bp-Rp) color
+    bprp  = np.array(bprp)
+    coeff = [-416.585, 39780.0, -84190.5, 85203.9, -48225.9, 15598.5, -2694.76, 192.865]    
+    teff  = np.array([np.sum([co*_bprp**i for i,co in enumerate(coeff)]) for _bprp in bprp])
+    mask  = (bprp>=0.55) & (bprp<=3.25)
+    teff[~mask] = np.nan
+    return teff
+
+def curtis_gyrochrone(bprp, kind):
+    bprp = np.array(bprp)
+    
+    if kind=='kepler': #Kepler lower envelope
+        bprp_min, bprp_max = 0.6, 2.1
+        coeff = [36.4756, -202.718, 414.752, -395.161, 197.800, -50.0287, 5.05738]
+        
+    elif kind=='pleiades-ro':
+        bprp_min, bprp_max = 0.6, 1.3
+        coeff = [37.068, -188.02, 332.32, -235.78, 60.395]
+
[…]

### 2022-04-08 src/ms.tex
more real estate for Teff residuals

@@
 
-\edit1{An alternative explanation for the short-period overdensity is contamination by subgiants.
+\edit1{Finally, we note that we regard the short-period pileup as less secure than the long-period pileup.
+In Appendix~\ref{app:harmonics} we show that the short-period pileup does not appear to be a harmonic of the long-period pileup, as one might expect if the periods on the short-period pileup were erroneously determined.
+However, the short-period pileup is primarily observed in the \lamostmcq sample, while it is unclear whether it is present in the \lamostsan sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
+The differences between the \mma and \santos catalogs may provide an explanation for this observation.
+The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \santos may then make the short-period pileup appear weaker relative to the long-period pileup.
+This is as expected since open cluster observations suggest that the short-period pileup can not be a long-lived feature, while the long-period pileup appears to be much longer lived.
+Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.}
+
+%An alternative explanation for the short-period overdensity is contamination by subgiants.
 Stars which have \teff on the ZAMS higher than $\sim$6250 will never undergo magnetic braking, spinning rapidly for the entirety of their main-sequence lifetimes.
 As these stars evolve away from the main-sequence, their temperatures drop while \logg changes minimally, while the spin rate slows due to radial expansion.
-Removing subgiants using spectroscopic parameters alone is difficult...}
+While cursory attempts to isolate subgiants have been taken, it is
 
@@
 
-    \item The long-period pileup is well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5000-6250~K.
+    \item The long-period pileup is well-described by a constant Rossby number, with a critical value of \rocritfinal, in the \teff range of $\approx$~5500-6250~K.
 The pileup is also populated by stars with a wide range of isochrone ages (\agerange).
@@
     \item We \edit1{tentatively detect} a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly indicating that} the short-period pileup is shorter-lived ($\mathcal{O}(10^8 \text{yr})$) relative to the long-period pileup ($>10^9$~yr).
+This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly indicating that} the short-period pileup is shorter-lived, $\mathcal{O}(10^8 \text{yr})$, relative to the long-period pileup ($>10^9$~yr).
 The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \curtis.
@@
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{./figures/ridge.pdf}
-%     \caption{The \teff-\prot plane for the CKS sample.
-Point colors are scaled to the CKS ages determined from isochrone fitting.
-The source of \prot is denoted above each panel, where \citet{David2021} is a compilation of vetted periods, rather than a source of original measurements.
-The black trapezoid indicates the approximate area of the ridge.
-The grey curves indicate empirical cluster sequences from \curtis, corresponding to ages of $\sim$2.7, 1, 0.67, and 0.12~Gyr from top to bottom.}
-%     \label{fig:ridge}
-% \end{figure*}
-
 \section{Comparison of spectroscopic temperature scales}\label{app:teff}
@@
 We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.}
-
-\edit1{We also note that the short-period pileup is not as prominent in the \lamostsan sample, compared to the \lamostmcq sample (see the rightmost panels of Figure~\ref{fig:xmatch}).
-The differences between the \mma and \santos catalogs may provide an explanation for this observation.
-The latter catalog contains a much larger number of detections at longer rotation periods, and the higher sensitivity to more slowly rotating stars within the \santos may then make the short-period pileup appear weaker relative to the long-period pileup.
-This is as expected since open cluster obsbervations suggest that the short-period pileup can not be a long-lived feature, while the long-period pileup appears to be much longer lived.
-Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.}
 
@@
 We found that the long-period pileup is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded.
-In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required (or, equivalently, a +0.1 mag shift applied to the constant Rossby curve).
-While there is no justification for such a large shift, it may indicate the presence of a systematic offset in the \taucz relation.}
+In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a $\approx$-0.1 mag shift to the Gaia colors of the data was required (or, equivalently, a +0.1 mag shift applied to the constant Rossby curve).
+While there is no justification for such a large shift, it may indicate the presence of a systematic offset in the \taucz relation.
+We also note that we do not have an explanation for why the solar value Rossby curve does not pass through the Sun.}
 

### 2022-04-08 src/figures/teffscales.py

@@ -178,6 +178,7 @@ def residual(ax,
     ax.set_ylabel(ylabel)
     ax.set_xlim(xmin, xmax)
     ax.set_ylim(ymin, ymax)
+    ax.set_yticks(np.arange(-500,750,250))
     
     return ax
 
@@ -191,9 +192,9 @@ sns.set(context="paper",
 
 tstr = " Teff [K]"
 
-fig = plt.figure(constrained_layout=False, figsize=(20,5))
+fig = plt.figure(constrained_layout=False, figsize=(18,5))
 widths = [1,1,1,1,1]
-heights = [2,1]
+heights = [1,1]
 spec = fig.add_gridspec(ncols=5, nrows=2, 
                         width_ratios=widths,
                         height_ratios=heights)

### 2022-04-08 src/ms.tex
fixing Gaia plots

@@
     \caption{The color–\prot plane for the \mma sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$.
-The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data.
+The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data and cluster sequences.
 Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above.

### 2022-04-08 src/figures/gaia-mcquillan.py

@@ -216,11 +216,24 @@ for ax in zip(axes,axs):
         ax[i].set_ylim(0,50)
         ax[i].set_xlabel(r"$G_\mathregular{BP}-G_\mathregular{RP}$ [mag]")
         ax[i].set_ylabel("Rotation period [d]")
-        ax[i].plot(_bprp, constant_rossby(_bprp, 0.496), color='orange', ls='--', label=r'$\mathregular{Ro} = \mathregular{R}_\odot$')
-        ax[i].legend()
+        ax[i].plot(_bprp, constant_rossby(_bprp, 1.2*0.496), color='orange', ls='-', label=r'$\mathregular{Ro} = 1.2 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].plot(_bprp, constant_rossby(_bprp, 0.496), color='orange', ls='--', label=r'$\mathregular{Ro} = 1.0 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].plot(_bprp, constant_rossby(_bprp, 0.8*0.496), color='orange', ls=':', label=r'$\mathregular{Ro} = 0.8 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].legend(prop={"size":8})
 
-        for j,seq in enumerate(gyro_sequences):
-            ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+        # for j,seq in enumerate(gyro_sequences):
+        #     ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+
+
+for i in range(3):
+    for j,seq in enumerate(gyro_sequences):
+        if i==2:
+            axes[i].plot(_bprp_cluster-0.1, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+            axs[i].plot(_bprp_cluster-0.1, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+        else:
[…]

### 2022-04-08 src/figures/gaia-santos.py

@@ -212,11 +212,22 @@ for ax in zip(axes,axs):
         ax[i].set_ylim(0,50)
         ax[i].set_xlabel(r"$G_\mathregular{BP}-G_\mathregular{RP}$ [mag]")
         ax[i].set_ylabel("Rotation period [d]")
-        ax[i].plot(_bprp, constant_rossby(_bprp, 0.496), color='orange', ls='--', label=r'$\mathregular{Ro} = \mathregular{R}_\odot$')
-        ax[i].legend()
+        ax[i].plot(_bprp, constant_rossby(_bprp, 1.2*0.496), color='orange', ls='-', label=r'$\mathregular{Ro} = 1.2 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].plot(_bprp, constant_rossby(_bprp, 0.496), color='orange', ls='--', label=r'$\mathregular{Ro} = 1.0 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].plot(_bprp, constant_rossby(_bprp, 0.8*0.496), color='orange', ls=':', label=r'$\mathregular{Ro} = 0.8 \mathregular{R}_\odot$', zorder=np.inf)
+        ax[i].legend(prop={"size":8})
         
-        for j,seq in enumerate(gyro_sequences):
-            ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+        #for j,seq in enumerate(gyro_sequences):
+        #    ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+
+for i in range(3):
+    for j,seq in enumerate(gyro_sequences):
+        if i==2:
+            axes[i].plot(_bprp_cluster-0.1, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+            axs[i].plot(_bprp_cluster-0.1, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
+        else:
+            axes[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
[…]

### 2022-04-11 src/ms.tex
revising asteroseismic figure

@@
 \edit1{The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \santos revealed that this pileup is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
+\edit1{It is thus the \teff precision afforded by spectroscopic catalogs that reveal the long-period pileup.
+In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pileup in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
 
@@
 the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
-
-\edit1{In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pileup in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
 
@@
 
-To investigate the origin of this offset, we cross-matched the \hall sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
+To \edit1{assess the consistency of the \teff scales between catalogs}, we cross-matched the \hall sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
 We found a root mean square (RMS) of the residuals between the \hall and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
-\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \hall samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}
-
-%{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pileup as seen in the two samples.}
+\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \hall samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.
+These offsets are modest, and do not account for non-linear systematics which may exist (particularly in the LAMOST \teff).}
 
@@
     \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the \hall main-sequence asteroseismic sample (top row) \edit1{and the \masuda sample (bottom row)} in the \teff--\prot plane.
-Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
-A constant offset of +41~K was applied to the LAMOST \teff \edit1{in the upper right panel} to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}).
-Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
+\edit1{The black point indicates the Sun.} Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
+\edit1{In the top panels, constant offsets of -29~K and +41~K were added to the CKS and LAMOST \teff, respectively, based on comparisons to those stars with overlap in the H21 sample (see \S~\ref{subsec:asteroseismic} and Appendix~\ref{app:teff}).} Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
@@
 
-%An alternative explanation for the short-period overdensity is contamination by subgiants.
+\edit1{An alternative explanation for the short-period overdensity is contamination by subgiants.
 Stars which have \teff on the ZAMS higher than $\sim$6250 will never undergo magnetic braking, spinning rapidly for the entirety of their main-sequence lifetimes.
 As these stars evolve away from the main-sequence, their temperatures drop while \logg changes minimally, while the spin rate slows due to radial expansion.
-While cursory attempts to isolate subgiants have been taken, it is
+While the feature persists even after crude subgiant cuts, it is possible a more careful selection would weaken the feature.
+We leave this to future investigators.}
 
@@
 
-\section{Confounding scenarios}\label{app:harmonics}
-
-\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pileup, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
-Figure~\ref{fig:harmonic} shows a Gaussian kernel density estimation of the \lamostmcq \teff–\prot distribution for stars with \logg~$>$~4.1.
-The short-period pileup is observed in this sample, particularly for \teff~$>$~6000~K.
-We found through inspection a constant Rossby curve that traces the long-period pileup.
-Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pileup, though they both bracket the feature seen through density estimation.
-We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\linewidth]{./figures/harmonic.pdf}
-    \caption{Gaussian kernel density estimation of the \teff–\prot distribution in the \lamostmcq sample for stars with \logg~$>$~4.1.
-A curve of constant Rossby number is shown by the solid line.
-The \prot/2 and \prot/3 harmonics of this curve are shown by the dashed and dotted lines, respectively.
-The short-period pileup is seen near the short-period edge, particularly for \teff~$>$~6000~K.
-The short-period pileup does not appear to be simply a harmonic of the long-period pileup, as one might expect if the periods for the stars in the short-period pileup were erroneously determined.}
-    \label{fig:harmonic}    
-\end{figure}
-
-\section{The Gaia Color–Period Plane}\label{app:gaia}
+\section{The Gaia color–period plane}\label{app:gaia}
 \edit1{Given that the long-period pileup was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
@@
     \label{fig:gaia-santos}
+\end{figure}
+
+\section{Confounding scenarios for the short-period pileup}\label{app:harmonics}
+
+\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pileup, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
+Figure~\ref{fig:harmonic} shows a Gaussian kernel density estimation of the \lamostmcq \teff–\prot distribution for stars with \logg~$>$~4.1.
+The short-period pileup is observed in this sample, particularly for \teff~$>$~6000~K.
+We found through inspection a constant Rossby curve that traces the long-period pileup.
+Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pileup, though they both bracket the feature seen through density estimation.
+We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{./figures/harmonic.pdf}
+    \caption{Gaussian kernel density estimation of the \teff–\prot distribution in the \lamostmcq sample for stars with \logg~$>$~4.1.
+A curve of constant Rossby number is shown by the solid line.
+The \prot/2 and \prot/3 harmonics of this curve are shown by the dashed and dotted lines, respectively.
+The short-period pileup is seen near the short-period edge, particularly for \teff~$>$~6000~K.
+The short-period pileup does not appear to be simply a harmonic of the long-period pileup, as one might expect if the periods for the stars in the short-period pileup were erroneously determined.}
+    \label{fig:harmonic}    
 \end{figure}

### 2022-04-11 src/figures/asteroseismic.py

@@ -19,6 +19,16 @@ mpl.rcParams["savefig.dpi"] = 300
 
 import seaborn as sns
 
+####################################################################
+# The Sun
+sun = {"teff": 5772,
+       "prot": 25.4,
+       "e_prot": 25.4-24.5,
+       "E_prot": 27-25.4
+      }
+      
+sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
+####################################################################
 
 # ### Hall et al. 2021 asteroseismic sample
 hall = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/other/NatAs/5.707/table1.dat",
@@ -70,8 +80,13 @@ sc_kws = {"marker":".", "color":"C0", "rasterized":True, "alpha":1}
 
 fig, (ax1,ax2) = plt.subplots(nrows=2,ncols=2,figsize=(2*4.5,6.5))
 
-ax1[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
-ax1[1].scatter(lam['Teff_lam']+41, lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
+
+cks_hall_offset = -29
+lamost_hall_offset = 41
+
+
+ax1[0].scatter(cks["p20_cks_steff"]+cks_hall_offset, cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
+ax1[1].scatter(lam['Teff_lam']+lamost_hall_offset, lam['Prot'], **sc_kws, label=r"LAMOST–McQuillan", s=0.5)
 #Note: the 41 K offset is derived in the appendix
 
 ax2[0].scatter(cks["p20_cks_steff"], cks["d21_prot"], **sc_kws, label=r"California–Kepler Survey", zorder=999, s=4)
@@ -100,6 +115,13 @@ for i in range(2):
                     **ebar_kws)
     ax1[i].set_xlim(6500,5000)
     #ax1[i].set_ylim(0,60)
+
+    ax1[i].plot(sun["teff"], sun["prot"], **sun_kws)
+    ax1[i].plot(sun["teff"], sun["prot"], 'k.')
+
+    ax2[i].plot(sun["teff"], sun["prot"], **sun_kws)
+    ax2[i].plot(sun["teff"], sun["prot"], 'k.')    
+
[…]

### 2022-04-11 src/ms.tex
referee comments

@@
 Here $T_\mathrm{sys}$ is a constant to allow for a systematic offset between the data and the models used to calibrate the \taucz relation.
-We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
+\edit1{This offset can equivalently be thought of as a correction to the data, or a correction to the model \teff scale.} We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
 We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro, $f$, and $T_\mathrm{sys}$ with ranges of (0.1,10), (0,10), and (-1000~K, 1000~K), respectively.

### 2022-04-11 src/figures/asteroseismic.py

@@ -26,7 +26,7 @@ sun = {"teff": 5772,
        "e_prot": 25.4-24.5,
        "E_prot": 27-25.4
       }
-      
+
 sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
 ####################################################################
 
@@ -116,11 +116,11 @@ for i in range(2):
     ax1[i].set_xlim(6500,5000)
     #ax1[i].set_ylim(0,60)
 
-    ax1[i].plot(sun["teff"], sun["prot"], **sun_kws)
-    ax1[i].plot(sun["teff"], sun["prot"], 'k.')
+    ax1[i].plot(sun["teff"], sun["prot"], zorder=np.inf, **sun_kws)
+    ax1[i].plot(sun["teff"], sun["prot"], 'k.', zorder=np.inf)
 
-    ax2[i].plot(sun["teff"], sun["prot"], **sun_kws)
-    ax2[i].plot(sun["teff"], sun["prot"], 'k.')    
+    ax2[i].plot(sun["teff"], sun["prot"], zorder=np.inf, **sun_kws)
+    ax2[i].plot(sun["teff"], sun["prot"], 'k.', zorder=np.inf)    
 
     ax1[i].set_xlabel("Effective temperature [K]")
     ax1[i].set_ylabel("Rotation period [d]")

### 2022-04-15 src/ms.tex
coauthor comments

@@
 \edit1{While helioseismic differential rotation studies have inferred rotation periods of up to $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
-We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
+We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.\footnote{We note that the \jvs models used in \S\ref{subsec:models} are calibrated to the Sun's equatorial rotation period of 25.4~d.}} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
 

### 2022-04-15 src/ms.tex
coauthor comments

@@
 
-\edit1{The above relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d}.
+\edit1{The above relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d.
+For comparison, in the model grids used to create the population models in \S\ref{subsec:models} the solar Rossby number is $\approx$2.16.}
 

### 2022-04-18 src/ms.tex
coauthor comments

@@
 
-We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \jvs.
+We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}, \edit1{which were updated for \citet{Hall2021} and utilized in that work}.
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
+\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
 }.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.

### 2022-04-18 src/ms.tex
coauthor comments

@@
 If the short-period pileup is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
-However, the observed \teff--\prot distribution will also depend sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
-LAMOST).}
+However, the observed \teff--\prot distribution depends sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
+LAMOST), which likely explains the observed number density trend.}
 
@@
 As periods become longer, they become more difficult to measure directly from Kepler time series.
-This is true for at least three reasons: (1) the observational baseline is finite, (2) photometric variation amplitude declines with increasing rotation period, and (3) photometric variations tend to become less periodic with increasing rotation period.
-The last two points are due to astrophysical reasons.
+This is true for at least three reasons: (1) photometric variation amplitude declines with increasing Rossby number, (2) the existence of significant spot groups appears to be intermittent for high Rossby numbers, and (3) the observational baseline is finite.
+We note that the last point is not so important for Kepler's long baseline ($\sim$1400~d), while the first two points are due to astrophysical reasons.
 This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pileup in the catalogs considered here depends sensitively on how detectability depends on Ro.
@@
 However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution.
-First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge.
-For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods).
-Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}
+We direct readers to \S4.2 of that work for a summary of those arguments.
+Critically, we note here that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pileup or overdensity at this edge.
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
 
@@
 Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
-In Figure~\ref{fig:weights}, we show that the long-period pileup is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pileup are likely to be genuine detections and not an artifact of some detection threshold.
-Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pileup or overdensity at this edge.
-To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
+In Figure~\ref{fig:weights}, we show that the long-period pileup is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pileup are likely to be genuine detections and not an artifact of some detection threshold.}
 
@@
 \edit1{We do note, however, that we can not rule out the possibility that the long-period pileup is wider in period or Rossby space than we observe, with stars just above the pileup being undetected through rotational brightness modulations.
-Such a scenario might result if magnetic braking is gradually weakened as stars approach \rocrit, as opposed to ceasing entirely at \rocrit.
-However, if the pileup does have a broader width in Rossby number, the close agreement between our observations and the \hall and \masuda samples suggests the true pileup width cannot be much larger than we observe.}
-
-\begin{figure*}
+Such a scenario might result if magnetic braking is gradually weakened once stars reach \rocrit, as opposed to ceasing entirely at \rocrit.
+However, even in this scenario there are observational constraints on the rate at which the braking index changes.
+If the braking index starts to decline as Ro approaches \rocrit then there may be tension with the open cluster data.
+Additionally, if the pileup does have a broader width in Rossby number, the close agreement between our observations and the \hall and \masuda samples suggests the true pileup width cannot be much larger than we observe.}
+
+\begin{figure}
     \centering
@@
     \label{fig:weights}    
-\end{figure*}
+\end{figure}
 
@@
 We also note that the MCMC results for the \lamostmcq sample preferred the LAMOST \teff to be shifted lower (Table~\ref{tab:mcmc}), which is in inconsistent with our finding that the LAMOST \teff are already lower than those from other spectroscopic surveys (Appendix~\ref{app:teff}).}
-    \item \edit1{Defining the long-period pileup or edge is a choice that impacts \rocrit.
+    \item \edit1{Defining the long-period pileup or edge is a choice that impacts the inferred value of \rocrit.
 For example, though we chose the 90th \prot percentile in the \lamostmcq sample, \jvs chose the 95th percentile.
@@
 \edit1{Finally, we note that we regard the short-period pileup as less secure than the long-period pileup.
+Preliminary tests with other density estimation methods \citep{Contardo2022}\footnote{\url{https://github.com/contardog/FindTheGap}} also reveal the short-period pileup, but further work is needed to better understand this feature.
 In Appendix~\ref{app:harmonics} we show that the short-period pileup does not appear to be a harmonic of the long-period pileup, as one might expect if the periods on the short-period pileup were erroneously determined.
@@
 Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.}
-
-\edit1{An alternative explanation for the short-period overdensity is contamination by subgiants.
-Stars which have \teff on the ZAMS higher than $\sim$6250 will never undergo magnetic braking, spinning rapidly for the entirety of their main-sequence lifetimes.
-As these stars evolve away from the main-sequence, their temperatures drop while \logg changes minimally, while the spin rate slows due to radial expansion.
-While the feature persists even after crude subgiant cuts, it is possible a more careful selection would weaken the feature.
-We leave this to future investigators.}
 
@@
 
-\edit1{If the Sun is not on the long-period pileup, this would seem to imply any number of unusual explanations.
-For example, the Sun's angular momentum evolution may have been different from those stars on the pileup.
-Alternatively, spin-down may proceed on the main-sequence after the pileup either through magnetic braking or some other mechanism.
-We consider these explanations unlikely, and suggest it is likely some combination of \teff systematics and observational biases that complicate contextualizing the Sun.}
-
 \section{Conclusions} \label{sec:conclusions}
@@
     \item It is yet unclear whether the Sun resides on the long-period pileup or has already evolved through it.
-\edit1{Offsets of $\approx$110--140~K between the observational and theoretical \teff scales allow for the long-period pileup to be fit with a solar value Rossby curve, thereby placing the Sun on the pileup.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
+\edit1{Offsets of $\approx$110--140~K between the observational and theoretical \teff scales would place the long-period pileup at Rossby numbers consistent with other literature values of the location of the WMB transition.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
 However, this tension might be simply explained by inaccurate isochrone ages.

### 2022-04-18 src/figures/skumanich.py

@@ -26,7 +26,6 @@ sun = {"teff": 5772,
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
 
-#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
 def convective_turnover_timescale(teff):
     #Returns convective turnover timescale in days
@@ -118,12 +117,29 @@ def curtis_teff_bprp(teff):
     return f(teff)
 
 
+# from scipy.interpolate import interp1d
+
+# std = pd.read_hdf('../data/models/standard_population.h5', key='sample')
+# #roc = pd.read_hdf('../models/rocrit_population.h5', key='sample')
+
+# idx = (abs(std['age']-5)<0.01) & (std['evo']==1) & (abs(std['[Fe/H]'])<0.05)
+# model_teff = np.array(std['Teff'][idx])
+# model_prot = np.array(std['period'][idx])
+
+# order = np.argsort(model_teff)
+# model_teff = model_teff[order]
+# model_prot = model_prot[order]
+
+# df = pd.DataFrame({"x": model_teff,
+#                    "y": model_prot})
+# df = df.drop_duplicates()
+
+# x = np.array(df.x)
+# y = np.array(df.y)
+# f = interp1d(x, y, kind='linear', fill_value='extrapolate')
+
+
 #McQuillan et al. 2014
-# mcq = Table.read('../data/mcquillan2014/table1.dat',
-#                 readme='../data/mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet('../data/mcquillan2014_table1.parquet')
 
 
@@ -156,7 +172,7 @@ lam = lam[lam_mask]
 
 sns.set(style='ticks', font_scale=1.5, context='paper')
 
-_teff = np.linspace(4500,6250,1000)
+_teff = np.linspace(5000,6250,1000)
 
 seq = 'ngc6819+ruprecht147'
 
@@ -176,6 +192,7 @@ plt.plot(sun["teff"], sun["prot"], **sun_kws)
 plt.plot(sun["teff"], sun["prot"], 'k.')
 
[…]

### 2022-04-18 src/figures/weights.py

@@ -67,9 +67,9 @@ xm = lamost_xmatch(mcq)
 
 sns.set(font_scale=1.3, context="paper", style="ticks")
 
-fig, axes = plt.subplots(nrows=1,ncols=5,figsize=(14,2))
+fig, axes = plt.subplots(nrows=1,ncols=3,figsize=(8,1.5))
 
-for i,j in enumerate(np.arange(0.1,0.6,0.1)):
+for i,j in enumerate(np.arange(0.1,0.31,0.1)):
     
     arg = xm['mcq_w']>j
     axes[i].scatter(xm['Teff'][arg], xm['mcq_Prot'][arg], s=0.05, alpha=0.2, rasterized=True)

### 2022-04-18 src/ms.tex
minor edits

@@
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6391526
+\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{New URL TBD
 }.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.

### 2022-04-20 src/ms.tex
updating Zenodo repos

@@
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{New URL TBD
-}.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
+\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6471539}.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
 The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
@@
 
-The code and data tables required to reproduce the figures and analysis presented here are publicly available through GitHub.\footnote{\url{https://github.com/trevordavid/rossby-ridge}}
+The code and data tables required to reproduce the figures and analysis presented here are publicly available through GitHub.\footnote{\url{https://github.com/trevordavid/rossby-ridge}} The data tables are also available through Zenodo.\footnote{\url{https://doi.org/10.5281/zenodo.6391526}}
 

### 2022-05-05 src/scripts/cmd.py
upgrade to new showyourwork [skip ci]

@@ -1,3 +1,4 @@
+import paths
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib as mpl

### 2022-05-05 src/scripts/comparison.py

@@ -1,3 +1,4 @@
+import paths
 import pandas as pd
 import matplotlib.pyplot as plt
 import matplotlib as mpl

### 2022-05-05 src/scripts/gaia-mcquillan.py

@@ -1,3 +1,4 @@
+import paths
 from dustmaps.bayestar import BayestarWebQuery
 from dustmaps.config import config
 config.reset()

### 2022-05-05 src/scripts/harmonic.py

@@ -1,3 +1,4 @@
+import paths
 import scipy.stats as st
 import pandas as pd
 import matplotlib.pyplot as plt

### 2022-05-05 src/scripts/inflection.py

@@ -1,3 +1,4 @@
+import paths
 import numpy as np
 import pandas as pd
 import emcee

### 2022-05-05 src/scripts/mcmc.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # coding: utf-8
 
+import paths
 import numpy as np
 import pandas as pd
 

### 2022-05-05 src/scripts/models.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # coding: utf-8
 
+import paths
 import numpy as np
 import pandas as pd
 import astropy.constants as c
@@ -85,9 +86,6 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 
 ######################################################################################
 # van Saders et al. 2019 models
-# std = pd.read_csv('../data/models/standard_model.csv')
-# roc = pd.read_csv('../data/models/rocrit_model.csv')
-
 std = pd.read_hdf('../data/standard_population.h5', key='sample')
 std = std[std['evo']==1]
 
@@ -99,18 +97,6 @@ roc['flag'] = 'roc'
 
 model = pd.concat([std, roc], ignore_index=True, sort=True)
 ######################################################################################
-
-cks_shift = 111
-lamost_shift = 116
-
-#Detrending LAMOST Teff
-# def lamost_teff_detrend(teff):   
-#     dteff = 2.55513439e-13*teff**5 - 7.18129973e-09*teff**4 + 8.04175914e-05*teff**3 - 4.48417848e-01*teff**2 + 1.24490338e+03*teff - 1.37649898e+06
-#     return teff-dteff
-
-# lam["Teff_lam"] = lamost_teff_detrend(lam["Teff_lam"])
-
-
 mpl.rcParams["legend.markerscale"] = 1
 sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":8, "rasterized":True}
@@ -124,7 +110,7 @@ cks_ms = cks['p20_cks_slogg'] > logg_thresh
 lam_ms = lam['logg_lam'] > logg_thresh
 
 sns.displot(data=std, x="Teff", y="period", binwidth=(20, 0.5), cbar=True, vmin=0, vmax=100, cbar_kws={'label': r'N$_\mathregular{stars}$'})
-plt.scatter(cks['cks_Teff'][cks_ms]+cks_shift, cks['d21_prot'][cks_ms], label='California–Kepler Survey', **sc_kws)
+plt.scatter(cks['cks_Teff'][cks_ms], cks['d21_prot'][cks_ms], label='California–Kepler Survey', **sc_kws)
[…]

### 2022-05-05 src/scripts/rocrit.py

@@ -1,3 +1,4 @@
+import paths
 import numpy as np
 import pandas as pd
 

### 2022-05-05 src/scripts/shifted.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # coding: utf-8
 
+import paths
 import numpy as np
 import pandas as pd
 import astropy.constants as c
@@ -85,6 +86,9 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 
 ######################################################################################
 # van Saders et al. 2019 models
+# std = pd.read_csv('../data/models/standard_model.csv')
+# roc = pd.read_csv('../data/models/rocrit_model.csv')
+
 std = pd.read_hdf('../data/standard_population.h5', key='sample')
 std = std[std['evo']==1]
 
@@ -96,6 +100,18 @@ roc['flag'] = 'roc'
 
 model = pd.concat([std, roc], ignore_index=True, sort=True)
 ######################################################################################
+
+cks_shift = 111
+lamost_shift = 116
+
+#Detrending LAMOST Teff
+# def lamost_teff_detrend(teff):   
+#     dteff = 2.55513439e-13*teff**5 - 7.18129973e-09*teff**4 + 8.04175914e-05*teff**3 - 4.48417848e-01*teff**2 + 1.24490338e+03*teff - 1.37649898e+06
+#     return teff-dteff
+
+# lam["Teff_lam"] = lamost_teff_detrend(lam["Teff_lam"])
+
+
 mpl.rcParams["legend.markerscale"] = 1
 sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":8, "rasterized":True}
@@ -109,7 +125,7 @@ cks_ms = cks['p20_cks_slogg'] > logg_thresh
 lam_ms = lam['logg_lam'] > logg_thresh
 
 sns.displot(data=std, x="Teff", y="period", binwidth=(20, 0.5), cbar=True, vmin=0, vmax=100, cbar_kws={'label': r'N$_\mathregular{stars}$'})
-plt.scatter(cks['cks_Teff'][cks_ms], cks['d21_prot'][cks_ms], label='California–Kepler Survey', **sc_kws)
+plt.scatter(cks['cks_Teff'][cks_ms]+cks_shift, cks['d21_prot'][cks_ms], label='California–Kepler Survey', **sc_kws)
[…]

### 2022-05-05 src/scripts/skumanich.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # coding: utf-8
 
+import paths
 import numpy as np
 import pandas as pd
 
@@ -26,21 +27,11 @@ sun = {"teff": 5772,
 
 sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
 
-#bk = pd.read_csv("../data/_kim_2010/-kim-2010.csv")
 
-def convective_turnover_timescale(teff,
-                                  ref='gunn1998'):
+def convective_turnover_timescale(teff):
     #Returns convective turnover timescale in days
-    
-    if ref == 'gunn1998':
-        #Gunn et al. 1998 relation, from Cranmer & Saar 2011
-        return 314.24*np.exp(-(teff/1952.5) - (teff/6250.)**18.) + 0.002
-    
-    # elif ref == '2010':
-    #     # & Kim 2010 relation for local tau_c
-    #     teff_pts = 10.**bk['logT']
-    #     tc_pts   = bk['Local_tau_c']
-    #     return np.interp(teff, teff_pts, tc_pts)
+    #Gunn et al. 1998 relation, from Cranmer & Saar 2011
+    return 314.24*np.exp(-(teff/1952.5) - (teff/6250.)**18.) + 0.002
 
 
 def constant_rossby(teff, ro):
@@ -127,36 +118,30 @@ def curtis_teff_bprp(teff):
     return f(teff)
 
 
+# from scipy.interpolate import interp1d
 
+# std = pd.read_hdf('../data/standard_population.h5', key='sample')
+# #roc = pd.read_hdf('../models/rocrit_population.h5', key='sample')
 
-######################################################################################
-#McQuillan et al. 2013
-mcq_koi = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/table1.dat",
-                readme="https://cdsarc.cds.unistra.fr/ftp/J/ApJ/775/L11/ReadMe",
-                format="ascii.cds")
-mcq_koi = mcq_koi.to_pandas()
-mcq_koi = mcq_koi.add_prefix('mcq_')
+# idx = (abs(std['age']-5)<0.01) & (std['evo']==1) & (abs(std['[Fe/H]'])<0.05)
+# model_teff = np.array(std['Teff'][idx])
+# model_prot = np.array(std['period'][idx])
+
+# order = np.argsort(model_teff)
[…]

### 2022-05-05 src/scripts/teffscales.py

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 # coding: utf-8
 
+import paths
 import pandas as pd
 import matplotlib.pyplot as plt
 import matplotlib as mpl

### 2022-05-05 src/scripts/surveys.py
woot

@@ -123,4 +123,5 @@ ax[0].set_ylabel("Rotation period [d]")
 sns.despine()
 plt.subplots_adjust(wspace=0)
 plt.savefig(paths.figures / 'surveys.pdf')
-plt.savefig('../static/surveys.png')
+if False:
+    plt.savefig(paths.static / 'surveys.png')

### 2022-05-09 src/tex/ms.tex
upgrade to new showyourwork [skip ci]
woot
fixing yml file

--- 2022-04-20 src/ms.tex
@@
 \usepackage{longtable}
+\usepackage{showyourwork}
 
@@
 \received{February 4, 2022}
-\revised{\today}
-%\accepted{\today}
+\revised{April 20, 2022}
+\accepted{May 5, 2022}
 
@@
 
-More recently, \citet{Hall2021}, \edit1{hereafter H21}, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \hall sample had rotation periods that were also detected from rotational brightness modulations.} from the \jvs study and the present work, to argue support for the WMB model.
-\citet{Masuda2021}, \edit1{hereafter MPH21}, also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
+More recently, \citet{Hall2021}, hereafter H21, used the asteroseismic rotation rates of Kepler dwarfs, with different selection and detection biases\footnote{Only 48/91 stars in the \hall sample had rotation periods that were also detected from rotational brightness modulations.} from the \jvs study and the present work, to argue support for the WMB model.
+\citet{Masuda2021}, hereafter MPH21, also found support for the WMB hypothesis from inference of the rotation period distribution of Sun-like stars using stellar radii and projected rotational velocities.
 While the physics responsible for the weakened magnetic braking of solar-type stars is unknown, one hypothesis is that the declining efficiency of wind-driven angular momentum loss is connected to the magnetic field complexity, which may vary with Rossby number \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.
@@
 All stars characterized here were targets of NASA's Kepler mission \citep{Borucki2010} and have published rotation periods derived from the Kepler data.
-For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys \edit1{(Figure~\ref{fig:surveys})}.
+For each subsample of the Kepler field described below, we combined published rotation periods from a variety of literature sources with spectroscopic parameters provided by large-scale surveys (Figure~\ref{fig:surveys}).
 
@@
 The \mma \teff values originate from the Kepler Input Catalog \citep[KIC,][]{Brown2011} or \citet{Dressing2013} for low-mass stars.
-The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the \edit1{activity belts}.
+The orange point in each panel indicates the Sun's equatorial rotation period, with the errorbar capturing the range of periods measured from the activity belts.
 Many of the stars above the long-period pileup are subgiants which have experienced spin-down due to expansion off the main-sequence, as pointed out in \jvs.}
     \label{fig:surveys}
+    \script{surveys.py}
 \end{figure*}
@@
 CKS spectra were acquired with the Keck/HIRES spectrograph \citep{Vogt1994} and spectroscopic parameters were determined by averaging parameters from the SpecMatch pipeline \citep{Petigura2015} and SME@XSEDE, a Python implementation of the Spectroscopy Made Easy pipeline \citep{Valenti1996}.
-\edit1{The internal (relative) errors on \teff from the CKS catalog are estimated at $\pm$60~K, with systematic errors of $\pm$100~K estimated from comparison to other catalogs \citep[see Table 7 of][]{Petigura2017}}.
+The internal (relative) errors on \teff from the CKS catalog are estimated at $\pm$60~K, with systematic errors of $\pm$100~K estimated from comparison to other catalogs \citep[see Table 7 of][]{Petigura2017}.
 The metallicity distribution of the CKS sample is centered near solar, with a mean and standard deviation of +0.03~dex and 0.18~dex, respectively.
@@
 The \citet{Brewer2018} study, referred to here as SPOCS, also published elemental abundances and ages from isochrone fitting for the CKS sample.
-We additionally cross-matched the CKS catalog with the \edit1{LAMOST DR5} catalog \citep{Xiang2019} which is described further in \S\ref{subsec:lamost}.
+We additionally cross-matched the CKS catalog with the LAMOST DR5 catalog \citep{Xiang2019} which is described further in \S\ref{subsec:lamost}.
 We compare the \teff--\prot distributions of the CKS sample using \teff and rotation periods from a variety of sources in Appendix~\ref{app:teffprot}.
 
-\subsection{\edit1{LAMOST} \label{subsec:lamost}}
+\subsection{LAMOST \label{subsec:lamost}}
 The LAMOST project derived homogeneous spectroscopic parameters from low-resolution ($R\sim$~1800) LAMOST DR5 spectra for approximately 40\% of the Kepler field \citep{Zong2018, Xiang2019}.
-\edit1{A description of the LAMOST Kepler field observations is provided in \citet{deCat2015}.} The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
-We cross-matched the \edit1{LAMOST DR5} stellar parameter catalog of \citet{Xiang2019} with the \mma catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog (hereafter S21), which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
+A description of the LAMOST Kepler field observations is provided in \citet{deCat2015}.
+The \citet{Xiang2019} catalog derived stellar parameters using the DD-Payne pipeline, which builds on the method of \citet{Ting2017b} by incorporating elements of the Cannon \citep{Ness2015} and uses the overlap with GALAH DR2 and APOGEE DR14 as training data.
+We cross-matched the LAMOST DR5 stellar parameter catalog of \citet{Xiang2019} with the \mma catalog, which published rotation periods for $>$34000 Kepler targets, as well as the \citet{Santos2021} catalog (hereafter S21), which provides rotation periods for $>$55000 FGKM stars observed by Kepler.
 
@@
 \newpage
-\subsection{\edit1{APOGEE}}
+\subsection{APOGEE}
 The Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017} is a large-scale, high-resolution ($R \sim 22500$) stellar spectroscopic survey conducted at $H$-band as part of the Sloan Digital Sky Survey \citep[SDSS-IV;][]{Blanton2017}.
@@
 \subsection{The Sun}
-To place the Sun in the context of the long-period pileup, we use up-to-date estimates of the Sun's effective temperature, \edit1{color}, rotation period, and age.
+To place the Sun in the context of the long-period pileup, we use up-to-date estimates of the Sun's effective temperature, color, rotation period, and age.
 Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K.
-\edit1{For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.} For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
-\edit1{While helioseismic differential rotation studies have inferred rotation periods of up to $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
-We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.\footnote{We note that the \jvs models used in \S\ref{subsec:models} are calibrated to the Sun's equatorial rotation period of 25.4~d.}} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
+For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.
+For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d.
+While helioseismic differential rotation studies have inferred rotation periods of up to $\approx$~36~d near the poles \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}.
+We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.\footnote{We note that the \jvs models used in \S\ref{subsec:models} are calibrated to the Sun's equatorial rotation period of 25.4~d.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}.
 
@@
 The effects of a simple cut in \logg to separate subgiants and dwarfs are shown in the middle and right columns.
-In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from \edit1{the activity belts}.
+In each panel the Sun is indicated by the orange point, with an errorbar reflecting the range of periods measured from the activity belts.
 The long-period pileup for dwarf stars is clearly seen to extend to the solar temperature.
@@
     \label{fig:xmatch}
+    \script{xmatch.py}
 \end{figure*}
@@
 We first noted a pileup of stars along the long-period edge for stars with \teff~$>5800$~K when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}.
-\edit1{The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.} Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \santos revealed that this pileup is still apparent when adopting periods uniformly from \edit1{other catalogs as well} (\edit1{see Appendix~\ref{app:teffprot}}).
-\edit1{It is thus the \teff precision afforded by spectroscopic catalogs that reveal the long-period pileup.
-In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pileup in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.}
-
-\edit1{Comparing the long-period pileup to rotation period sequences from open clusters can provide insight into the ages of stars on the pileup.} The \edit1{long}-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020}, hereafter \curtis, from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
+The pileup in that sample is visible even when sourcing \prot uniformly from \citet{McQuillan2013}, as shown in the rightmost panel of Figure~\ref{fig:surveys}.
+Sourcing rotation periods from \citet{Mazeh2015}, \citet{Angus2018}, and \santos revealed that this pileup is still apparent when adopting periods uniformly from other catalogs as well (see Appendix~\ref{app:teffprot}).
+It is thus the \teff precision afforded by spectroscopic catalogs that reveal the long-period pileup.
+In Appendix~\ref{app:gaia}, we demonstrate that it is possible to recover the long-period pileup in the color–period plane using high-precision photometry from the Gaia mission \citep{Gaia2016} and selecting stars with minimal interstellar reddening.
+
+Comparing the long-period pileup to rotation period sequences from open clusters can provide insight into the ages of stars on the pileup.
+The long-period pileup is close to the empirical hybrid cluster sequence derived by \citet{Curtis2020}, hereafter \curtis, from members of the NGC~6819 (age~$\sim$~2.5~Gyr) and Ruprecht~147 (age~$\sim$~2.7~Gyr) open clusters (Figure~\ref{fig:kde}).
 Notably, we use the color--\teff relation presented by those authors to recast the cluster sequences in terms of \teff.
-The observation that the \edit1{long-period pileup} approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
-Similarly, the \edit1{long-period pileup clearly lies at longer periods than} the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period \edit1{pileup}.
+The observation that the long-period pileup approximately corresponds with the $\sim$2.5--2.7~Gyr cluster sequence implies that stars with \teff$\gtrsim$~\teffmin have already piled up onto the edge by or before this timescale.
+Similarly, the long-period pileup clearly lies at longer periods than the empirical $\sim$1~Gyr cluster sequence based on rotation rates in the NGC~6811 cluster \citep{Curtis2019a, Curtis2020}, implying that it takes F-type stars $>1$~Gyr to reach the long-period pileup.
 These observations are in accordance with predictions from the WMB model which suggest the pileup forms on a timescale of 2--3~Gyr \citep{vanSaders2019}.
-\edit1{We show in \S\ref{subsec:ages} that stars along the pileup have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.}
+We show in \S\ref{subsec:ages} that stars along the pileup have a range of ages, with a lower bound that is consistent with $\sim$2~Gyr.
 
@@
 The short-period pileup can be observed in the LAMOST--McQuillan sample for \teff~$\gtrsim$~5500~K.
-The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from \edit1{the activity belts}.}
+The orange point indicates the Sun's temperature and equatorial rotation period, with the errorbar capturing the range of periods measured from the activity belts.}
     \label{fig:kde}
+    \script{kde.py}
 \end{figure*}
 
-\edit1{A change in slope along the long-period edge is observed} in both the CKS and LAMOST samples (Figure~\ref{fig:inflection}), with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
+A change in slope along the long-period edge is observed in both the CKS and LAMOST samples (Figure~\ref{fig:inflection}), with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}.
 A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6224~$\pm$~24~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling.
-This change in slope \edit1{may be} due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
+This change in slope may be due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.
 
@@
 This cluster of points has a similar slope in the \teff--\prot plane as the long-period pileup, and does not reside on the harmonic of the ridge line as one might expect if the periods were erroneously determined.
-A similar clustering of points is not observed in the \edit1{\lamostmcq} sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
+A similar clustering of points is not observed in the \lamostmcq sample and is less pronounced or absent when substituting the CKS temperatures with \teff from either \citet{Brewer2018} or \citet{Martinez2019}, two studies that independently derived spectroscopic parameters from the CKS spectra.
 Further inspection of the stars in this cluster revealed that they have anomalously large \teff discrepancies between the CKS and SPOCS catalogs, such that the SPOCS temperatures shift the stars onto the long-period pileup.
@@
 A secondary pileup at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample.
-We verified through inspection that the secondary pileup does not lie along the \prot/2 harmonic line of the long-period pileup \edit1{(see Appendix~\ref{app:harmonics})}.
-As shown in \edit1{the middle panel of} Figure~\ref{fig:kde} \edit1{and the top panel of Figure~\ref{fig:mcmc}}, this secondary pileup is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
-\edit1{The short-period pileup is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
-the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.}
+We verified through inspection that the secondary pileup does not lie along the \prot/2 harmonic line of the long-period pileup (see Appendix~\ref{app:harmonics}).
+As shown in the middle panel of Figure~\ref{fig:kde} and the top panel of Figure~\ref{fig:mcmc}, this secondary pileup is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.
+The short-period pileup is also subtly apparent in the upper right panel of Figure~\ref{fig:xmatch}, as well as scatter plots of the \lamostmcq sample, e.g.
+the second panel of Figure~\ref{fig:surveys} and the bottom panels of Figures~\ref{fig:models} and \ref{fig:shifted}.
 
@@
     \label{fig:inflection}
+    \script{inflection.py}
 \end{figure}
@@
 
-We tested the hypothesis that the long-period pileup observed in the \edit1{\lamostmcq} sample is \edit1{compatible with the} WMB \edit1{model} by fitting constant Rossby models to the \prot boundary.
+We tested the hypothesis that the long-period pileup observed in the \lamostmcq sample is compatible with the WMB model by fitting constant Rossby models to the \prot boundary.
 For a given \teff, this model predicts \prot as:
@@
 
-where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{CranmerSaar2011} \edit1{from a fit to the zero-age main-sequence stellar models of \citet{Gunn1998}}:
+where we used the equation for the convective turnover timescale (valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K) presented in \citet{CranmerSaar2011} from a fit to the zero-age main-sequence stellar models of \citet{Gunn1998}:
 
@@
 
-\edit1{The above relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d.
-For comparison, in the model grids used to create the population models in \S\ref{subsec:models} the solar Rossby number is $\approx$2.16.}
-
-\edit1{We approximated the long-period edge in the following manner.
+The above relation yields a convective turnover timescale for the Sun of $\tau_\mathrm{cz,\odot} = 12.88$~d, leading to a Rossby number of $\mathrm{Ro_\odot}=1.97$ given the mean equatorial rotation period of $P_\mathrm{rot,\odot} = 25.4$~d.
+For comparison, in the model grids used to create the population models in \S\ref{subsec:models} the solar Rossby number is $\approx$2.16.
+
+We approximated the long-period edge in the following manner.
 For 100 bootstrapped resamplings of the \teff–\prot distribution, leaving out 50\% of the data for each bootstrapped sample, we computed the 90th percentile of \prot values in overlapping \teff bins with centers located every 20~K between 4000~K and 7000~K and half-widths of 100~K.
-The final long-period edge curve was then computed as the mean of the bootstrapped 90th percentile values.} We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
+The final long-period edge curve was then computed as the mean of the bootstrapped 90th percentile values.
+We found it was also necessary to omit stars with Ro~$>$~5/3 in this computation to match the long-period edge.
 We computed the 10th percentile curve and its uncertainty similarly, as an approximation to the lower boundary of the \teff-\prot plane.
-We show these curves in relation to the full \edit1{\lamostmcq} sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
+We show these curves in relation to the full \lamostmcq sample and to constant Rossby models in Figure~\ref{fig:mcmc}.
 
@@
 Here $T_\mathrm{sys}$ is a constant to allow for a systematic offset between the data and the models used to calibrate the \taucz relation.
-\edit1{This offset can equivalently be thought of as a correction to the data, or a correction to the model \teff scale.} We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
+This offset can equivalently be thought of as a correction to the data, or a correction to the model \teff scale.
+We performed Markov chain Monte Carlo sampling (MCMC) of this likelihood with the \texttt{emcee} package \citep{emcee2013, emcee2019} to estimate the mean and uncertainty of the critical Rossby number that best matches the long-period edge in the range of 5000~K$\lesssim \teff \lesssim$~6250~K.
 We instantiated 32 walkers around the least-squares solution and sampled for $10^5$ steps, adopting uniform priors on Ro, $f$, and $T_\mathrm{sys}$ with ranges of (0.1,10), (0,10), and (-1000~K, 1000~K), respectively.
@@
 For the CKS pileup stars and the \lamostmcq percentile curves, we assumed constant fractional \prot uncertainties of 10\%.
-To isolate the long-period pileup stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$ \edit1{(shown in Figure~\ref{fig:inflection})}.
+To isolate the long-period pileup stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$ (shown in Figure~\ref{fig:inflection}).
 We note that this selection is particular to the CKS \teff scale and is not general.
@@
 
-\edit1{In some cases, our constant Rossby fits to the long-period pileup prefer Rossby numbers lower than the Sun's, in contrast with \jvs.
-We caution against overinterpreting the specific \rocrit values inferred here, and discuss our interpretations further in \S\ref{subsec:longperiod}.}
-
-\edit1{To further quantify the preference for \rocrit~$<$~\rosun} we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K).
+In some cases, our constant Rossby fits to the long-period pileup prefer Rossby numbers lower than the Sun's, in contrast with \jvs.
+We caution against overinterpreting the specific \rocrit values inferred here, and discuss our interpretations further in \S\ref{subsec:longperiod}.
+
+To further quantify the preference for \rocrit~$<$~\rosun we performed non-linear least-squares fits of two models to the CKS long-period pileup and the \lamostmcq 90th percentile curve (both in the \teff range of 5800--6250~K).
 The first model assumes \rocrit = \rosun and has one free parameter, a temperature offset added to the data.
@@
 We found that for the CKS sample, particularly when using the more precise temperatures from \citet{Fulton2018}, that the $\Delta \chi^2$ between the \rocrit = \rosun and variable \rocrit models was negligible.
-In other words, the \rocrit = \rosun model fits the data well with a \teff shift of $\approx$~\edit1{70--140}~K depending on the \teff source used.
+In other words, the \rocrit = \rosun model fits the data well with a \teff shift of $\approx$~70--140~K depending on the \teff source used.
 For the \lamostmcq sample, there is weak support for the variable \rocrit model.
-However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend not adequately captured by a constant offset \edit1{(i.e.
-a \teff-dependent \teff offset)} in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
+However, as shown in Appendix~\ref{app:teff}, there is a strong systematic trend not adequately captured by a constant offset (i.e.
+a \teff-dependent \teff offset) in the LAMOST temperatures when compared to other spectroscopic catalogs (particularly SPOCS).
 Correcting for these systematic trends weakens the support for the variable \rocrit model, and we speculate that temperature systematics bias the \rocrit inference procedure.
-
-%Lending further support to this hypothesis is the observation of visibly different slopes for the long-period pileup in the CKS vs.
-\lamostmcq samples (see Fig.~\ref{fig:percentiles}).\footnote{\edit1{While there is reasonably good agreement between the CKS long-period pileup and the \lamostmcq 90th percentile curve, we observe in Figure~\ref{fig:percentiles} a systematic trend in which most CKS pileup stars lie above the \lamostmcq curve for \teff~$\lesssim$~6000~K, and below it for \teff~$\gtrsim$~6000~K.
-This is consistent with there being a \teff-dependent systematic trend between the \teff scales of CKS and LAMOST, as discussed in Appendix~\ref{app:teff}.}}
 
@@
     \label{fig:mcmc}
+    \script{mcmc.py}
 \end{figure}
@@
     \includegraphics[width=\linewidth]{./figures/rocrit.pdf}
-    \caption{Non-linear least-squares fits to the long-period pileup for a \rocrit~=~\edit1{\rosun} model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
+    \caption{Non-linear least-squares fits to the long-period pileup for a \rocrit~=~\rosun model with constant \teff offset (orange) and a variable \rocrit model with no \teff offset (blue).
 The $\Delta \chi^2$ values printed in each panel are computed as $\chi^2_{\text{Ro}_\text{crit}=\text{Ro}_\odot} - \chi^2_{\Delta T_{eff}=0}$.}
     \label{fig:rocrit}
+    \script{rocrit.py}
 \end{figure*}
@@
 
-We compared the \teff--\prot distribution of the CKS and \edit1{\lamostmcq} samples with the theoretical predictions of \citet{vanSaders2019}, \edit1{which were updated for \citet{Hall2021} and utilized in that work}.
+We compared the \teff--\prot distribution of the CKS and \lamostmcq samples with the theoretical predictions of \citet{vanSaders2019}, which were updated for \citet{Hall2021} and utilized in that work.
 Those authors presented forward modeling simulations of the observed Kepler \prot distribution including theoretical models of stellar angular momentum evolution (for both the standard spin-down and WMB scenarios), a galactic population model, and a prescribed observational selection function.
-\edit1{For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6471539}.}} \edit1{We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
-The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.}
+For this exercise we selected main-sequence stars from the simulations using an evolutionary state flag, defining main-sequence stars as those with a He core mass fraction $<$0.0002.\footnote{These models are packaged in a Zenodo repository at \url{https://doi.org/10.5281/zenodo.6471539}.} We did not apply a cut on Rossby number to the models to mimic a detection threshold, as one might expect if detectability is Ro-dependent.
+The issue of detection bias is discussed further in \S\ref{subsec:detectionbias}.
 
@@
 Neither model satisfactorily matches the observations, though the WMB model more closely matches the long-period edge of F-type and early G-type stars.
-The specific WMB prescription of \jvs adopted a critical Rossby number of \rocrit~=~2.08 \edit1{(using a different \taucz prescription than the one used here)}, leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations, \edit1{assuming no \teff offset between the models and the data}.
+The specific WMB prescription of \jvs adopted a critical Rossby number of \rocrit~=~2.08 (using a different \taucz prescription than the one used here), leading to a pileup that is located at larger \prot values (at fixed \teff) when compared to the observations, assuming no \teff offset between the models and the data.
 Figure~\ref{fig:shifted} shows the same models in relation to the data with constant \teff offsets applied to the data, which are derived in \S\ref{subsec:rossby}.
-The \teff offsets lead to better agreement between the data and models, although the long-period pileup in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \edit1{\teff-dependent} systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
+The \teff offsets lead to better agreement between the data and models, although the long-period pileup in the \lamostmcq sample appears to overlay the models only for \teff~$\gtrsim 6000$~K, possibly due to strong \teff-dependent systematics in the LAMOST \teff scale (see Appendix~\ref{app:teff}).
 
@@
 While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \jvs, as well as differences in the computation of \taucz.
-We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \edit1{\lamostmcq} may be slightly different.
-
-Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models, \edit1{and in Appendix~\ref{app:teff} we show the LAMOST \teff are $\sim$50--100~K cooler than the other surveys considered here}.
+We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the \lamostmcq may be slightly different.
+
+Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models, and in Appendix~\ref{app:teff} we show the LAMOST \teff are $\sim$50--100~K cooler than the other surveys considered here.
 In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e.
@@
     \label{fig:models}
+    \script{models.py}
 \end{figure*}
@@
     \caption{Same as Figure~\ref{fig:models} with constant \teff offsets applied to the data.
-\edit1{Shifts of +111~K and +116~K are applied to the CKS and LAMOST \teff, respectively.
-The \teff shifts originate from a least-squares fit of a \rocrit~=~\rosun curve (with a \teff offset) to the long-period pileup (as described in \S\ref{subsec:rossby})}.}
+Shifts of +111~K and +116~K are applied to the CKS and LAMOST \teff, respectively.
+The \teff shifts originate from a least-squares fit of a \rocrit~=~\rosun curve (with a \teff offset) to the long-period pileup (as described in \S\ref{subsec:rossby}).}
     \label{fig:shifted}
+    \script{shifted.py}
 \end{figure*}
@@
     \includegraphics[width=\linewidth]{./figures/percentiles.pdf}
-    \caption{\edit1{Comparison of the theoretically predicted \teff--\prot distribution of Kepler stars from \jvs (given by the 10th and 90th \prot percentiles) with the same percentile ranges from the \lamostmcq sample (black points) and the CKS long-period pileup (open circles).}}
+    \caption{Comparison of the theoretically predicted \teff--\prot distribution of Kepler stars from \jvs (given by the 10th and 90th \prot percentiles) with the same percentile ranges from the \lamostmcq sample (black points) and the CKS long-period pileup (open circles).}
     \label{fig:percentiles}
+    \script{percentiles.py}
 \end{figure}
 
-\subsection{Comparison with \edit1{literature surveys}}
+\subsection{Comparison with literature surveys}
 \label{subsec:asteroseismic}
@@
 \hall determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies.
-We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pileup we observe, although the \hall sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample.
-Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of effects}.
-
-To \edit1{assess the consistency of the \teff scales between catalogs}, we cross-matched the \hall sample with the LAMOST DR5 catalog \edit1{and the CKS sample}.
+We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pileup we observe, although the \hall sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \lamostmcq sample while such an offset is either absent or not as apparent relative to the CKS sample.
+Such an offset could be due to differing spectroscopic temperature scales between the two studies, differing observational biases, or some combination of effects.
+
+To assess the consistency of the \teff scales between catalogs, we cross-matched the \hall sample with the LAMOST DR5 catalog and the CKS sample.
 We found a root mean square (RMS) of the residuals between the \hall and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}).
-\edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \hall samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.
-These offsets are modest, and do not account for non-linear systematics which may exist (particularly in the LAMOST \teff).}
-
-\edit1{\masuda inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from precise stellar radii and spectroscopically determined \vsini.
+Similarly, we found excellent agreement between the \teff scales of the CKS and \hall samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.
+These offsets are modest, and do not account for non-linear systematics which may exist (particularly in the LAMOST \teff).
+
+\masuda inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from precise stellar radii and spectroscopically determined \vsini.
 As with the asteroseismic sample, we compared the CKS and \lamostmcq samples with the results of those authors.
 We find that the long-period pileup we observe appears to overlap with the \masuda data, within the uncertainties of those authors.
-We therefore consider it likely that the long-period pileup is the same feature, or an extension of the same feature, that was detected in the \hall and \masuda studies.}
+We therefore consider it likely that the long-period pileup is the same feature, or an extension of the same feature, that was detected in the \hall and \masuda studies.
 
@@
     \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
-    \caption{Comparison of CKS (left column) and \edit1{\lamostmcq} samples (right column) with the \hall main-sequence asteroseismic sample (top row) \edit1{and the \masuda sample (bottom row)} in the \teff--\prot plane.
-\edit1{The black point indicates the Sun.} Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
-\edit1{In the top panels, constant offsets of -29~K and +41~K were added to the CKS and LAMOST \teff, respectively, based on comparisons to those stars with overlap in the H21 sample (see \S~\ref{subsec:asteroseismic} and Appendix~\ref{app:teff}).} Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
+    \caption{Comparison of CKS (left column) and \lamostmcq samples (right column) with the \hall main-sequence asteroseismic sample (top row) and the \masuda sample (bottom row) in the \teff--\prot plane.
+The black point indicates the Sun.
+Note, the CKS, LAMOST, and \hall samples derive \teff from distinct pipelines.
+In the top panels, constant offsets of -29~K and +41~K were added to the CKS and LAMOST \teff, respectively, based on comparisons to those stars with overlap in the H21 sample (see \S~\ref{subsec:asteroseismic} and Appendix~\ref{app:teff}).
+Using LAMOST \teff, where it exists, for the \hall sample brings that sample into even closer agreement with the long-period pileup in the LAMOST-McQuillan sample.}
     \label{fig:asteroseismic}
+    \script{asteroseismic.py}
 \end{figure*}
@@
 We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages).
-\edit1{These ranges are consistent with the range of ages observed in the \hall and \masuda samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
-\edit1{For example, it is curious that, when using SPOCS ages} only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
+These ranges are consistent with the range of ages observed in the \hall and \masuda samples, as shown in Figure~11 of the latter reference.
+However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties.
+For example, it is curious that, when using SPOCS ages only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}).
 In fact, there are only a handful of long-period pileup stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pileup stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.
@@
     \label{fig:ages}
+    \script{ages.py}
 \end{figure*}
 
-\edit1{To further assess the evolutionary state of stars on the long-period pileup, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \mma and \santos rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pileup stars from this work (Figure~\ref{fig:cmd}).
+To further assess the evolutionary state of stars on the long-period pileup, we constructed a color-magnitude diagram (CMD) from the Gaia DR2 photometry and parallaxes for the Kepler field, the \mma and \santos rotation period catalogs, the \hall asteroseismic sample, and the CKS long-period pileup stars from this work (Figure~\ref{fig:cmd}).
 Stars with detected periods from rotational brightness modulations are primarily solar-type and lower-mass dwarfs, with a comparatively small number of subgiants.
 The \santos catalog contains more stars, in part due to the higher sensitivity to more slowly rotating, evolved stars, relative to \mma.
-CKS stars on the long-period pileup clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.}
+CKS stars on the long-period pileup clearly occupy a well-defined region of the upper main-sequence which overlaps well with the \hall asteroseismic sample.
 
@@
     \label{fig:cmd}
+    \script{cmd.py}
 \end{figure*}
@@
 \subsection{Where do the pileups end?} \label{subsec:extent}
-It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the \edit1{long-period pileup} declines towards cooler \teff, as predicted by the WMB model (see Figure~13 of \jvs).
-\edit1{The number density of stars on the short-period pileup similarly declines towards cooler \teff.
+It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long-period pileup declines towards cooler \teff, as predicted by the WMB model (see Figure~13 of \jvs).
+The number density of stars on the short-period pileup similarly declines towards cooler \teff.
 If the short-period pileup is due to core–envelope coupling, one might expect an opposite trend of increasing number density toward cooler \teff, since the core–envelope coupling timescales and hence the ``stalled'' braking phases are longer for lower-mass stars \citep{Curtis2020}.
 However, the observed \teff--\prot distribution depends sensitively on the selection functions and observational biases inherent to both Kepler and the source of \teff (e.g.
-LAMOST), which likely explains the observed number density trend.}
-
-While it not clear whether or not these declines are astrophysical in nature, the result of \edit1{the selection functions or observational biases inherent to Kepler or LAMOST}, or some combination of effects, we attempted to characterize the extent of the pileups through the following approach.
+LAMOST), which likely explains the observed number density trend.
+
+While it not clear whether or not these declines are astrophysical in nature, the result of the selection functions or observational biases inherent to Kepler or LAMOST, or some combination of effects, we attempted to characterize the extent of the pileups through the following approach.
 We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pileups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample.
@@
 We found that the relative fraction of stars on the long-period pileup declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}).
-The relative fraction of stars on the short-period \edit1{pileup} declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pileups are nearly equal.
+The relative fraction of stars on the short-period pileup declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pileups are nearly equal.
 Thus, it is not yet clear if either pileup extends to temperatures cooler than \tsun.
@@
     \label{fig:fraction}
+    \script{fraction.py}
 \end{figure}
@@
 \label{subsec:detectionbias}
-\edit1{To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
+To this point, we have not addressed the detection biases inherent to Kepler rotation period catalogs.
 As periods become longer, they become more difficult to measure directly from Kepler time series.
@@
 This is important because, if a period detection threshold is Ro-dependent and the detection threshold Ro$_\mathrm{thresh}$ is close to the value \rocrit at which WMB becomes important, then the location of the pileup in the catalogs considered here depends sensitively on how detectability depends on Ro.
-For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pileup has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pileup.}
-
-\edit1{\jvs addressed detection biases in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \mma \teff–\prot distribution.
+For example, if Ro$_\mathrm{thresh}$ is only slightly larger than \rocrit and the long-period pileup has some non-negligible width in period space (due either to astrophysics or measurement error), then it's possible the current data set only reveals the lower edge of the long-period pileup.
+
+\jvs addressed detection biases in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \mma \teff–\prot distribution.
 In this scenario, if \rothresh $\lesssim$ \rocrit, it becomes difficult to distinguish between the standard and WMB models in the \mma sample.
@@
 Critically, we note here that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pileup or overdensity at this edge.
-To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.}
-
-\edit1{To understand whether the period detections on the long-period pileup are reliable, we examined the weight parameters published by \mma.
+To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold.
+
+To understand whether the period detections on the long-period pileup are reliable, we examined the weight parameters published by \mma.
 The weight, $w$, is a metric \mma designed to serve as a proxy for the reliability of a rotation period detection, based on a star's autocorrelation function (ACF) properties and its position in $T_\mathrm{eff}$–$P_\mathrm{rot}$–LPH space, where LPH designates the local peak height of the ACF.
 Those authors selected $w>0.25$ for an acceptable compromise between real detections and false positives.
-In Figure~\ref{fig:weights}, we show that the long-period pileup is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pileup are likely to be genuine detections and not an artifact of some detection threshold.}
-
-\edit1{Most importantly, two independent samples with different observational biases than the surveys considered here also appear to yield a pileup of stars at the long-period edge.
-These samples are the \hall Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the \masuda sample of Kepler stars with Keck/HIRES spectroscopy, from which \masuda inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.}
-
-\edit1{We do note, however, that we can not rule out the possibility that the long-period pileup is wider in period or Rossby space than we observe, with stars just above the pileup being undetected through rotational brightness modulations.
+In Figure~\ref{fig:weights}, we show that the long-period pileup is apparent even for $w>0.3$, and we conclude that the majority of stars on the long-period pileup are likely to be genuine detections and not an artifact of some detection threshold.
+
+Most importantly, two independent samples with different observational biases than the surveys considered here also appear to yield a pileup of stars at the long-period edge.
+These samples are the \hall Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the \masuda sample of Kepler stars with Keck/HIRES spectroscopy, from which \masuda inferred the peak of the \teff–\prot distribution using spectroscopically-determined \vsini and constraints on the stellar radii.
+
+We do note, however, that we can not rule out the possibility that the long-period pileup is wider in period or Rossby space than we observe, with stars just above the pileup being undetected through rotational brightness modulations.
 Such a scenario might result if magnetic braking is gradually weakened once stars reach \rocrit, as opposed to ceasing entirely at \rocrit.
@@
 If the braking index starts to decline as Ro approaches \rocrit then there may be tension with the open cluster data.
-Additionally, if the pileup does have a broader width in Rossby number, the close agreement between our observations and the \hall and \masuda samples suggests the true pileup width cannot be much larger than we observe.}
+Additionally, if the pileup does have a broader width in Rossby number, the close agreement between our observations and the \hall and \masuda samples suggests the true pileup width cannot be much larger than we observe.
 
@@
     \label{fig:weights}
+    \script{weights.py}    
 \end{figure}
@@
 The existence of a pileup of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}, as discussed in \S\ref{subsec:rossby} and \S\ref{subsec:models}.
-The location of the long-period pileup coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, \edit1{as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
-We suggest that these features are one and the same.}
-
-\edit1{The modest differences between the long-period pileup we observe and the features observed by \hall and \masuda might be explained by differing \teff scales and differing observational biases of the samples.
+The location of the long-period pileup coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \hall, as well as the peak of the period distribution inferred by \masuda from combining \vsini and stellar radii.
+We suggest that these features are one and the same.
+
+The modest differences between the long-period pileup we observe and the features observed by \hall and \masuda might be explained by differing \teff scales and differing observational biases of the samples.
 The present samples have rotation periods detected from photometric variations and, for a survey of finite baseline and sensitivity like Kepler, photometric rotation periods are harder to detect for slower rotators, smaller amplitude variations, and more stochastic variability patterns (see \S\ref{subsec:detectionbias}).
-The \hall asteroseismic sample and \masuda sample, by comparison, are not biased against quiet, unspotted stars and are more likely to contain pileup stars that our sample may have missed.}
-
-\edit1{In some cases, the \rocrit values we inferred in \S\ref{subsec:rossby} were smaller than the value of \rocrit~$\approx$~\rosun~$\approx$~2 found by \citet{vanSaders2016,vanSaders2019}, sometimes with high statistical significance.
-We caution against overinterpreting the \rocrit values inferred here, and enumerate below the ways in which the \rocrit inference may be biased.}
+The \hall asteroseismic sample and \masuda sample, by comparison, are not biased against quiet, unspotted stars and are more likely to contain pileup stars that our sample may have missed.
+
+In some cases, the \rocrit values we inferred in \S\ref{subsec:rossby} were smaller than the value of \rocrit~$\approx$~\rosun~$\approx$~2 found by \citet{vanSaders2016,vanSaders2019}, sometimes with high statistical significance.
+We caution against overinterpreting the \rocrit values inferred here, and enumerate below the ways in which the \rocrit inference may be biased.
 
 \begin{enumerate}
-    \item \edit1{The \taucz relation adopted impacts the inferred \rocrit.
+    \item The \taucz relation adopted impacts the inferred \rocrit.
 We explored the \taucz formulae of \citet{BarnesKim2010}, \citet{Landin2010}, and \citet{Amard2019}, but these produce larger \taucz at fixed \teff than the \citet{CranmerSaar2011} relation, leading to even smaller inferred \rocrit values.
 We also explored using a fit to the \taucz computed in \jvs for solar-metallicity stars, which produces a smaller \taucz at fixed \teff than the \citet{CranmerSaar2011} relation.
-This relation revises our \rocrit estimates upwards by 15\%.}
-    \item \edit1{While our \rocrit inference procedure includes a constant \teff offset between the data and the \taucz relation, it does not include any \teff-dependent offsets.
+This relation revises our \rocrit estimates upwards by 15\%.
+    \item While our \rocrit inference procedure includes a constant \teff offset between the data and the \taucz relation, it does not include any \teff-dependent offsets.
 In Appendix~\ref{app:teff}, we show that the LAMOST \teff appear to have a non-linear mapping to the other \teff scales considered here.
 Perhaps not coincidentally, the \lamostmcq sample favors a \rocrit that is the least compatible with other samples.
-We also note that the MCMC results for the \lamostmcq sample preferred the LAMOST \teff to be shifted lower (Table~\ref{tab:mcmc}), which is in inconsistent with our finding that the LAMOST \teff are already lower than those from other spectroscopic surveys (Appendix~\ref{app:teff}).}
-    \item \edit1{Defining the long-period pileup or edge is a choice that impacts the inferred value of \rocrit.
+We also note that the MCMC results for the \lamostmcq sample preferred the LAMOST \teff to be shifted lower (Table~\ref{tab:mcmc}), which is in inconsistent with our finding that the LAMOST \teff are already lower than those from other spectroscopic surveys (Appendix~\ref{app:teff}).
+    \item Defining the long-period pileup or edge is a choice that impacts the inferred value of \rocrit.
 For example, though we chose the 90th \prot percentile in the \lamostmcq sample, \jvs chose the 95th percentile.
-Similarly, our boundary for the CKS long-period pileup was chosen subjectively.}
+Similarly, our boundary for the CKS long-period pileup was chosen subjectively.
 \end{enumerate}
-
-%We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between \edit1{observations and the models}, \edit1{including} the \taucz relation.
-Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 {\color{red} is compatible with all of the data sets if modest \teff systematics are present.} Comparison of the long-period pileups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).
 
@@
 This effect, in addition to the effects mentioned above, can further bias our inference of \rocrit to lower values.
-Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that \edit1{a pileup exists at the long-period edge} for stars with \teff~$\gtrsim$~\teffmin.
-\edit1{By construction, this pileup occurs at a Rossby number lower than that associated with a detection edge.} \edit1{In other words, the fact that we observe a pileup indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}
-
-\edit1{We also note that the WMB model makes the simplistic assumption that magnetic braking ceases when Ro reaches \rocrit.
-However, it is possible that magnetic braking becomes gradually weaker, a scenario which would also lead to a pileup but may also cause such a pileup to have some intrinsic width which is not due to measurement uncertainties.}
-
-\edit1{Our observation of a long-period pileup stands in contrast with recent rotation studies of solar analogs.
-\citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.} Additionally, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
+Regardless of the biases mentioned above and the true value of \rocrit, the sample considered here clearly indicates that a pileup exists at the long-period edge for stars with \teff~$\gtrsim$~\teffmin.
+By construction, this pileup occurs at a Rossby number lower than that associated with a detection edge.
+In other words, the fact that we observe a pileup indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.
+
+We also note that the WMB model makes the simplistic assumption that magnetic braking ceases when Ro reaches \rocrit.
+However, it is possible that magnetic braking becomes gradually weaker, a scenario which would also lead to a pileup but may also cause such a pileup to have some intrinsic width which is not due to measurement uncertainties.
+
+Our observation of a long-period pileup stands in contrast with recent rotation studies of solar analogs.
+\citet{doNascimento2020} studied the rotation period distribution of 193 solar analogs, concluding that some solar-mass main-sequence stars appear to rotate significantly more slowly than the Sun, seemingly at odds with the WMB model.
+Additionally, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model.
 Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr.
@@
 In contrast to the findings of those authors, we find that the long-period pileup lies far below an empirical 2.5~Gyr gyrochrone of \curtis that is evolved forward to 5~Gyr for braking indices of $n=0.5$ or $n=0.65$ (Figure~\ref{fig:skumanich}).
-\edit1{This implies the braking index must drop to a much lower value at some time after 2.5~Gyr for Sun-like stars.}
+This implies the braking index must drop to a much lower value at some time after 2.5~Gyr for Sun-like stars.
 
@@
     \label{fig:skumanich}
+    \script{skumanich.py}    
 \end{figure}
@@
 
-\edit1{Finally, we note that we regard the short-period pileup as less secure than the long-period pileup.
+Finally, we note that we regard the short-period pileup as less secure than the long-period pileup.
 Preliminary tests with other density estimation methods \citep{Contardo2022}\footnote{\url{https://github.com/contardog/FindTheGap}} also reveal the short-period pileup, but further work is needed to better understand this feature.
@@
 This is as expected since open cluster observations suggest that the short-period pileup can not be a long-lived feature, while the long-period pileup appears to be much longer lived.
-Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.}
+Consequently, the \santos catalog may more accurately reflect the relative strengths of these two features.
 
@@
 \citet{Gordon2021} proposed that the gap could instead be due to a period of accelerated spin-down immediately proceeding the stalling due to core–envelope coupling (such that stars evolve quickly through the gap and are rarely observed there).
-We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here \edit1{(see also Appendix~\ref{app:gaia})}.
+We can neither confirm nor reject this scenario, and we note that the gap is significantly emptier when using Gaia colors \citep{DavenportCovey2018, Gordon2021} compared to when using spectroscopic \teff as we do here (see also Appendix~\ref{app:gaia}).
 Notably, if the short-period pileup is indeed due to core–envelope coupling, and the gap due to a period of accelerated spin-down after such a coupling episode, the observed gradient in photometric variability still requires a physical explanation.
@@
     \label{fig:gap}
+    \script{gap.py}
 \end{figure*}
@@
 From Fig.~\ref{fig:xmatch} it is clear that the long-period pileup extends to temperatures as cool as the Sun's.
-If one assumes there are no systematic offsets between \edit1{the spectroscopic \teff scales considered here and that of the Sun}, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pileup.
+If one assumes there are no systematic offsets between the spectroscopic \teff scales considered here and that of the Sun, the Sun's equatorial rotation period places it $\sim$5--7 days above the long-period pileup.
 This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase.
 
-\edit1{If the Sun is indeed on the long-period pileup, it likely the presence of \teff systematics and/or observational biases that obfuscate that fact.} To place the Sun on the long-period pileup would require shifting the spectroscopic temperatures higher by \edit1{111}~K for the CKS sample, \edit1{143}~K for the SPOCS sample, or \edit1{116}~K for the LAMOST sample.
-\edit1{We also note that there are \teff-dependent systematics in the LAMOST \teff that, if corrected for, would bring the \lamostmcq long-period pileup into closer agreement with a solar Rossby curve.} Comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pileup, or very close to it, without the need for a \edit1{large} \teff shift.
-\edit1{This apears to be at odds with the notion that a \teff shift alone can place the Sun on the long-period pileup, since the \hall and CKS temperature scales agree to within $\sim$30~K.
-However, it is still possible that there are \teff-dependend systematics in the temperature scales.}
+If the Sun is indeed on the long-period pileup, it likely the presence of \teff systematics and/or observational biases that obfuscate that fact.
+To place the Sun on the long-period pileup would require shifting the spectroscopic temperatures higher by 111~K for the CKS sample, 143~K for the SPOCS sample, or 116~K for the LAMOST sample.
+We also note that there are \teff-dependent systematics in the LAMOST \teff that, if corrected for, would bring the \lamostmcq long-period pileup into closer agreement with a solar Rossby curve.
+Comparing the Sun to the \hall asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pileup, or very close to it, without the need for a large \teff shift.
+This apears to be at odds with the notion that a \teff shift alone can place the Sun on the long-period pileup, since the \hall and CKS temperature scales agree to within $\sim$30~K.
+However, it is still possible that there are \teff-dependend systematics in the temperature scales.
 
@@
 As shown by \citet{Aigrain2015}, the pipelines used to measure rotation periods from Kepler data would not be guaranteed to detect the solar rotation period.
-\edit1{Regardless, it is clear from the comparisons in \S\ref{subsec:asteroseismic} that the long-period pileup can not be much broader (in \prot or Rossby space) than we observe.}
+Regardless, it is clear from the comparisons in \S\ref{subsec:asteroseismic} that the long-period pileup can not be much broader (in \prot or Rossby space) than we observe.
 
@@
 
-    \item Comparison of the \edit1{long-period pileup} with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \jvs.
+    \item Comparison of the long-period pileup with empirical rotation sequences from open clusters implies that stars with $M\gtrsim1$~\msun pile up onto the ridge on a timescale $>$~1~Gyr but $\lesssim$2.5~Gyr, compatible with the predictions of \jvs.
 Using isochrone ages for a sample of exoplanet hosts on the long-period pileup suggests that stars slightly hotter than the Sun may populate the pileup until an age of $\sim$6~Gyr.
@@
     \item It is yet unclear whether the Sun resides on the long-period pileup or has already evolved through it.
-\edit1{Offsets of $\approx$110--140~K between the observational and theoretical \teff scales would place the long-period pileup at Rossby numbers consistent with other literature values of the location of the WMB transition.} If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
+Offsets of $\approx$110--140~K between the observational and theoretical \teff scales would place the long-period pileup at Rossby numbers consistent with other literature values of the location of the WMB transition.
+If the Sun has evolved through the pileup, there is some modest tension with the Sun's age and the ages of the oldest stars on the pileup; some of the cooler long-period pileup stars in the CKS sample are both more massive and older than the Sun, which contradicts the expectation from the WMB model that hotter stars spend a shorter period of time on the pileup.
 However, this tension might be simply explained by inaccurate isochrone ages.
 
-    \item We \edit1{tentatively detect} a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
-This overdensity \edit1{appears to be} less prominent than the long-period overdensity, \edit1{possibly indicating that} the short-period pileup is shorter-lived, $\mathcal{O}(10^8 \text{yr})$, relative to the long-period pileup ($>10^9$~yr).
+    \item We tentatively detect a secondary overdensity of stars at the short-period edge of the \teff-\prot plane.
+This overdensity appears to be less prominent than the long-period overdensity, possibly indicating that the short-period pileup is shorter-lived, $\mathcal{O}(10^8 \text{yr})$, relative to the long-period pileup ($>10^9$~yr).
 The short-period pileup appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core–envelope coupling, as proposed by \curtis.
@@
     \label{fig:comparison}
+    \script{comparison.py}
 \end{figure*}
@@
 
-In Figure~\ref{fig:teffscales} we compare temperatures between the \edit1{LAMOST DR5} catalog \citep{Xiang2019} and temperatures from other surveys.
+In Figure~\ref{fig:teffscales} we compare temperatures between the LAMOST DR5 catalog \citep{Xiang2019} and temperatures from other surveys.
 We find that the LAMOST \teff scale is consistently cooler than other surveys by $\sim$20--80~K, with the exception of the \mma study which sourced photometric \teff estimates from the KIC \citep{Brown2011}.
@@
     \label{fig:teffscales}
+    \script{teffscales.py}
 \end{figure}
@@
 \section{The Gaia color–period plane}\label{app:gaia}
-\edit1{Given that the long-period pileup was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
+Given that the long-period pileup was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high-precision photometric colors provided by the Gaia mission \citep{Gaia2016}.
 We cross-matched the \mma and \santos samples with Gaia EDR3 \citep{GaiaEDR3} using a 1 arcsec search radius and the \texttt{astroquery} package \citep{astroquery}.
@@
 While there is no justification for such a large shift, it may indicate the presence of a systematic offset in the \taucz relation.
-We also note that we do not have an explanation for why the solar value Rossby curve does not pass through the Sun.}
-
-\edit1{We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$.
+We also note that we do not have an explanation for why the solar value Rossby curve does not pass through the Sun.
+
+We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$.
 This break is not apparent in the Gaia–McQuillan sample.
 Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}, suggesting the origin of the discontinuity is in the rotation periods rather than the temperatures or colors.
-At present, we do not have a satisfactory explanation for this behavior, though we note that the \santos catalog employed various time series filters which might introduce systematic artifacts.}
+At present, we do not have a satisfactory explanation for this behavior, though we note that the \santos catalog employed various time series filters which might introduce systematic artifacts.
 
@@
     \label{fig:gaia-mcquillan}
+    \script{gaia-mcquillan.py}
 \end{figure}
@@
     \label{fig:gaia-santos}
+    \script{gaia-santos.py}
 \end{figure}
@@
 
-\edit1{Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pileup, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
+Here we consider the possibility that there is only one astrophysical overdensity in the true \teff–\prot distribution, the long-period pileup, and other features in the observed \teff–\prot distribution appear at period harmonics of this feature due to the difficult problem of reliable, automated rotation period measurement for large samples of stars.
 Figure~\ref{fig:harmonic} shows a Gaussian kernel density estimation of the \lamostmcq \teff–\prot distribution for stars with \logg~$>$~4.1.
@@
 Taking this curve and dividing the periods by factors of 2 and 3, we find that neither resulting curve satisfactorily matches the short-period pileup, though they both bracket the feature seen through density estimation.
-We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.}
+We interpret this as evidence that the short-period pileup is not simply due to measurement error, although we have not definitively ruled out that scenario.
 
@@
     \label{fig:harmonic}
+    \script{harmonic.py}    
 \end{figure}

### 2022-05-09 src/scripts/ages.py
cleaning up for arxiv

@@ -39,11 +39,6 @@ mcq_koi = mcq_koi.add_prefix('mcq_')
 
 
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 ######################################################################################
 
@@ -127,8 +122,6 @@ roc['ro'] = roc['period']/(roc['taucz']/86400)
 roc = roc[roc['evo']==1] # limit to main-sequence
 
 
-
-
 fig, ax = plt.subplots()
 sns.kdeplot(
     x=lam["Teff_lam"], 
@@ -136,7 +129,6 @@ sns.kdeplot(
     fill=True, 
     bw_adjust=0.25,
     levels=4,
-    #levels=[0.25,0.5,0.75,1],
     ax=ax
 )
 
@@ -204,14 +196,6 @@ ax = axes[1]
 cks['cks_e_age'] = cks['cks_age'] - (10.**(cks['cks_logAiso']-cks['cks_e_logAiso'])/1.0e9)
 cks['cks_E_age'] = (10.**(cks['cks_logAiso']+cks['cks_E_logAiso'])/1.0e9) - cks['cks_age']
 
-# ax[0].errorbar(cks['p20_cks_steff'], cks['cks_age'], 
-#              xerr=[cks['cks_e_Teff'], cks['cks_E_Teff']],
-#              yerr=[cks['cks_e_age'], cks['cks_E_age']], fmt='o', color='lightgrey',mec='lightgrey', linewidth=0, ecolor='lightgrey', zorder=1, alpha=0.5)
-
-# ax[0].errorbar(cks['p20_cks_steff'][ridge], cks['cks_age'][ridge], 
-#              xerr=[cks['cks_e_Teff'][ridge], cks['cks_E_Teff'][ridge]],
-#              yerr=[cks['cks_e_age'][ridge], cks['cks_E_age'][ridge]], fmt='o', mec='white', linewidth=0, color='k', ecolor='k', zorder=2)
-
 cks_age_err = np.max([np.nanmedian(cks['cks_e_age']), np.nanmedian(cks['cks_E_age'])])
[…]

### 2022-05-09 src/scripts/asteroseismic.py

@@ -57,20 +57,6 @@ lam = pd.read_parquet(paths.data / 'kepler_lamost.parquet')
 #Drop duplicates
 lam = lam.drop_duplicates(subset=['KIC'], keep='first')
 
-#ik = (lam['Teff_lam']>0)
-#ik &= (lam['Gmag']<15)
-
-#ik &= (lam['l20_Rvar']/lam['rrmscdpp03p0'] > 10)
-#ik &= (lam['logg_lam']-3*lam['e_logg_lam']>4)
-#ik &= (lam['duplicated_source']==False)
-#ik &= (abs(lam["Gmag"]-lam["phot_g_mean_mag"])<1)
-#ik &= (abs(lam["Gmag"]-lam["kepmag"])<1)
-#ik &= (abs(lam["Teff_lam"]-lam["teff_val"])<1000)
-#ik &= (lam['parallax_over_error']>50)
-#ik &= (lam["vim_r"]<0.5)
-#ik &= (lam["ruwe"]<1.2)
-
-
 # ### Comparison with Hall et al. 2021 asteroseismic sample
 sns.set(font_scale=1.2, context="paper", style="ticks")
 
@@ -115,7 +101,6 @@ for i in range(2):
                     label="Masuda et al. 2021", ms=4, lw=1, 
                     **ebar_kws)
     ax1[i].set_xlim(6500,5000)
-    #ax1[i].set_ylim(0,60)
 
     ax1[i].plot(sun["teff"], sun["prot"], zorder=np.inf, **sun_kws)
     ax1[i].plot(sun["teff"], sun["prot"], 'k.', zorder=np.inf)

### 2022-05-09 src/scripts/fraction.py

@@ -30,11 +30,6 @@ sun["logg"] = np.log10(c.GM_sun.cgs.value/c.R_sun.cgs.value**2)
 
 ######################################################################################
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 ######################################################################################
 
@@ -51,7 +46,6 @@ lam = lam.drop_duplicates(subset=['KIC'], keep='first')
 
 lam_mask = (lam["Teff_lam"]>3000)
 lam_mask = (lam["Teff_lam"]<8000)
-#lam_mask &= (lam["logg_lam"]>3)
 lam_mask &= (lam["logg_lam"]>4.1)
 lam_mask &= (lam["logg_lam"]<5)
 lam_mask &= (abs(lam["feh_lam"])<2)
@@ -63,8 +57,6 @@ print('Median LAMOST Teff error:', np.median(lam["e_Teff_lam"]))
 ######################################################################################
 
 ######################################################################################
-#bk = pd.read_csv(paths.data / "_kim_2010/-kim-2010.csv")
-
 def convective_turnover_timescale(teff):    
     #Returns convective turnover timescale in days
     #Gunn et al. 1998 relation, from Cranmer & Saar 2011

### 2022-05-09 src/scripts/gaia-santos.py

@@ -72,15 +72,12 @@ def gaia_edr3_vizier_xmatch(vizier_code):
     edr3['BP-RP'] = edr3['bp_rp']
     edr3['Gmag'] = edr3['phot_g_mean_mag']
     
-    # Store in the object named "Amy"
     cat1 = edr3.copy()
 
     cmd_coeff = [-0.0319809, 4.08935, 5.76321, -6.98323, 3.06721, -0.589493, 0.0417076]
     hyades_cmd = np.poly1d(np.flip(cmd_coeff))
 
     #Query the dust map
-
-    # I have McQuillan+2014 loaded in an astropy table called "Amy"
     distance_cat1 = 1e3/cat1['parallax']
     ifix = np.where(distance_cat1 <= 0)[0]
     if len(ifix) > 0:
@@ -218,8 +215,6 @@ for ax in zip(axes,axs):
         ax[i].plot(_bprp, constant_rossby(_bprp, 0.8*0.496), color='orange', ls=':', label=r'$\mathregular{Ro} = 0.8 \mathregular{R}_\odot$', zorder=np.inf)
         ax[i].legend(prop={"size":8})
         
-        #for j,seq in enumerate(gyro_sequences):
-        #    ax[i].plot(_bprp_cluster, curtis_gyrochrone(_bprp_cluster, kind=seq), label=gyro_ages[j], color='grey', lw=3, alpha=0.4)
 
 for i in range(3):
     for j,seq in enumerate(gyro_sequences):

### 2022-05-09 src/scripts/gap.py

@@ -62,9 +62,6 @@ lam_mask = (lam["Teff_lam"]<8000)
 lam_mask &= (lam["logg_lam"]>3)
 lam_mask &= (lam["logg_lam"]<5)
 lam_mask &= (abs(lam["feh_lam"])<2)
-#lam_mask &= (lam["mcq_Rper"]>2)
-#lam_mask &= (lam["phot_g_mean_mag"]<15)
-#lam_mask &= (lam["r_est"]>0.) & (lam["r_est"]<500.)
 lam = lam[lam_mask]
 
 print('LAMOST unique KIC targets:', len(np.unique(lam["KIC"])))
@@ -152,4 +149,3 @@ plt.colorbar(cb2, label=r'log(R$_\mathregular{per}$/ppm) [dex]', ax=ax2)
 plt.tight_layout()
 sns.despine()
 plt.savefig(paths.figures / 'gap.pdf')
-#plt.show()

### 2022-05-09 src/scripts/kde.py

@@ -131,11 +131,6 @@ mcq_koi = mcq_koi.add_prefix('mcq_')
 
 
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 ######################################################################################
 

### 2022-05-09 src/scripts/mcmc.py

@@ -56,8 +56,6 @@ cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
 ######################################################################################
 # LAMOST-Kepler 
 lam = pd.read_parquet(paths.data / 'kepler_lamost.parquet')
-#print('LAMOST unique KIC targets:', len(np.unique(lam["KIC"])))
-#print('LAMOST unique DR2 targets:', len(np.unique(lam["DR2Name"])))
 
 # Drop duplicate sources, keeping the one with the brighter G magnitude
 lam = lam.sort_values(["KIC", "Gmag"], ascending = (True, True))
@@ -243,7 +241,6 @@ def constant_rossby_sampler(x, y, yerr,
 
         for i in range(flat_samples.shape[1]):
             print(np.mean(flat_samples[:,i]), np.std(flat_samples[:,i]))        
-            #print(np.median(flat_samples[:,i]), np.percentile(flat_samples[:,i],[16,84])-np.median(flat_samples[:,i]))        
     
     
     if corner_plot==True:        

### 2022-05-09 src/scripts/models.py

@@ -39,11 +39,6 @@ mcq_koi = mcq_koi.add_prefix('mcq_')
 
 
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 ######################################################################################
 
@@ -56,8 +51,6 @@ cks = pd.read_parquet(paths.data / 'cks_merged.parquet')
 # where N is the number of KOIs detected around that star so we drop duplicates.
 cks = cks.drop_duplicates(subset=['kepid'], keep='first')
 cks = cks.merge(mcq_koi, how='left', left_on='kepid', right_on='mcq_KIC')
-# Filter stars that have discrepant Teff between CKS and SPOCS
-#cks_mask = abs(cks['cks_Teff']-cks['bf18_Teff'])/cks['cks_Teff'] < 0.02
 ######################################################################################
 
 
@@ -102,7 +95,6 @@ sns.set(font_scale=1.2, context="paper", style="ticks")
 sc_kws = {"marker":",", "color":"orange", "s":8, "rasterized":True}
 sun_kws = {"marker":"o", "color":"black", "ms":8, "mfc":"None", "mew":1}
 
-#sns.displot(data=std, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
 
 #We only want to plot dwarf stars
 logg_thresh = 4.1
@@ -123,7 +115,7 @@ plt.ylabel('Rotation period [d]')
 plt.text(1.15,1.05,"a",transform=plt.gca().transAxes,weight="bold",size=14)
 plt.savefig(paths.figures / 'std-model-cks.pdf')
 
-#sns.displot(data=roc, x="Teff(K)", y="Prot(days)", binwidth=(20, 0.5), cbar=True, cbar_kws={'label': r'N$_\mathregular{stars}$'})
+
[…]

### 2022-05-09 src/scripts/ridge.py

@@ -176,6 +176,6 @@ for i,y in enumerate(ys):
 sns.despine()    
 plt.tight_layout()
 plt.savefig(paths.figures / 'ridge.pdf')
-#plt.show()
+
 
 

### 2022-05-09 src/scripts/skumanich.py

@@ -118,28 +118,6 @@ def curtis_teff_bprp(teff):
     return f(teff)
 
 
-# from scipy.interpolate import interp1d
-
-# std = pd.read_hdf(paths.data / 'standard_population.h5', key='sample')
-# #roc = pd.read_hdf('../models/rocrit_population.h5', key='sample')
-
-# idx = (abs(std['age']-5)<0.01) & (std['evo']==1) & (abs(std['[Fe/H]'])<0.05)
-# model_teff = np.array(std['Teff'][idx])
-# model_prot = np.array(std['period'][idx])
-
-# order = np.argsort(model_teff)
-# model_teff = model_teff[order]
-# model_prot = model_prot[order]
-
-# df = pd.DataFrame({"x": model_teff,
-#                    "y": model_prot})
-# df = df.drop_duplicates()
-
-# x = np.array(df.x)
-# y = np.array(df.y)
-# f = interp1d(x, y, kind='linear', fill_value='extrapolate')
-
-
 #McQuillan et al. 2014
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 
@@ -162,14 +140,6 @@ lam_mask &= (lam["logg_lam"]<5)
 lam_mask &= (abs(lam["feh_lam"])<2)
 lam = lam[lam_mask]
 ######################################################################################
-
-# ######################################################################################
-# hall = Table.read("https://cdsarc.cds.unistra.fr/ftp/J/other/NatAs/5.707/table1.dat",
-#                   readme="https://cdsarc.cds.unistra.fr/ftp/J/other/NatAs/5.707/ReadMe",
-#                   format="ascii.cds")
-
-# hall.info()
-# ######################################################################################
 
 sns.set(style='ticks', font_scale=1.5, context='paper')
 

### 2022-05-09 src/scripts/surveys.py

@@ -36,11 +36,6 @@ mcq_koi = mcq_koi.to_pandas()
 mcq_koi = mcq_koi.add_prefix('mcq_')
 
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 ######################################################################################
 

### 2022-05-09 src/scripts/teffscales.py

@@ -52,7 +52,7 @@ table = table.to_pandas()
 unq = np.unique(table['gaia_ra'], return_index=True, return_counts=True)
 
 table['xmatch_count'] = np.zeros(len(table))
-#table['xmatch_count'].iloc[unq[1]] = unq[2]
+
 for i in range(len(table)):
     arg = unq[0] == table['gaia_ra'].iloc[i]
     table['xmatch_count'].iloc[i] = unq[2][arg]
@@ -70,11 +70,6 @@ lam = pd.read_parquet(paths.data / 'kepler_lamost.parquet')
 lam = lam.drop_duplicates(subset=['KIC'], keep='first')
 
 #McQuillan et al. 2014
-# mcq = Table.read(paths.data / 'mcquillan2014/table1.dat',
-#                 readme=paths.data / 'mcquillan2014/ReadMe',
-#                 format='ascii.cds')
-# mcq = mcq.to_pandas()
-# mcq = mcq.add_prefix('mcq_')
 mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 
 lam = lam.merge(mcq, how="left", left_on="KIC", right_on="mcq_KIC")

### 2022-05-09 src/scripts/weights.py

@@ -26,14 +26,12 @@ mcq = pd.read_parquet(paths.data / 'mcquillan2014_table1.parquet')
 gk = pd.read_parquet(paths.data / 'kepler_dr2_1arcsec.parquet')
 
 mcq = mcq.merge(gk, how="left", left_on="mcq_KIC", right_on="kepid")
-#np.shape(mcq)
 mcq = mcq.sort_values(by=["kepid", "kepler_gaia_ang_dist"])
 mcq = mcq.drop_duplicates(subset=['kepid'], keep='first')
 
 def lamost_xmatch(df):
     
     df_pos = df[['ra', 'dec']].copy()
-    #df_pos.head()
     df_pos.to_csv(paths.data / 'coords.csv', index=False)
     
     table = XMatch.query(cat1=open(paths.data / 'coords.csv'),
@@ -44,8 +42,6 @@ def lamost_xmatch(df):
                          colRA2='RAJ2000',
                          colDec2='DEJ2000')
 
-    #type(table)
-    #print(table)
     table = table.to_pandas()
     
     unq = np.unique(table['ra'], return_index=True, return_counts=True)
@@ -56,7 +52,6 @@ def lamost_xmatch(df):
         arg = unq[0] == table['ra'].iloc[i]
         table['xmatch_count'].iloc[i] = unq[2][arg]
 
-    #table = table.sort_values(by=["gaia_ra"])
     table = table.merge(df, how='right', left_on='ra', right_on='ra')
     table = table.sort_values(["ra","angDist"], ascending = (True, True))
     table = table.drop_duplicates(subset="ra", keep="first")

### 2022-05-09 src/scripts/xmatch.py

@@ -33,7 +33,6 @@ mcq = mcq.drop_duplicates(subset=['kepid'], keep='first')
 def lamost_xmatch(df):
     
     df_pos = df[['ra', 'dec']].copy()
-    #df_pos.head()
     df_pos.to_csv(paths.data / 'coords.csv', index=False)
     
     table = XMatch.query(cat1=open(paths.data / 'coords.csv'),
@@ -44,8 +43,6 @@ def lamost_xmatch(df):
                          colRA2='RAJ2000',
                          colDec2='DEJ2000')
 
-    #type(table)
-    #print(table)
     table = table.to_pandas()
     
     unq = np.unique(table['ra'], return_index=True, return_counts=True)
@@ -56,7 +53,6 @@ def lamost_xmatch(df):
         arg = unq[0] == table['ra'].iloc[i]
         table['xmatch_count'].iloc[i] = unq[2][arg]
 
-    #table = table.sort_values(by=["gaia_ra"])
     table = table.merge(df, how='right', left_on='ra', right_on='ra')
     table = table.sort_values(["ra","angDist"], ascending = (True, True))
     table = table.drop_duplicates(subset="ra", keep="first")
